### 2013-11-29 tex/evogp13/evogp.tex
minor tex ..

\documentclass{llncs}

%\usepackage{times}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
%\usepackage{qtree}
\usepackage{xspace}

%\usepackage{makeidx}  % allows for indexgeneration
%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
%\usepackage{hyperref}
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{graphicx}

\usepackage[ampersand]{easylist}
\usepackage[usenames,dvipsnames]{color}

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\la}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\C}{\mathbb{C}\xspace}
\newcommand{\Real}{\mathbb{R}}

\newenvironment{todo}
{~\\ {\color{red}\textbf{TODO}}
  \begin{easylist}[itemize]}
{ \end{easylist}}

\newcommand{\Lpr}{\Lambda^\prime}
\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
\newcommand{\ro}[1]{{\color{blue} #1}}
\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
\newcommand{\red}[1]{{\color{red} #1}}

%\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}

\begin{document}

\title{
Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
\author{Tom\'{a}\v{s} K\v{r}en\inst{1} \and Roman Neruda\inst{2}}
\institute{%
Faculty of Mathematics and Physics\\
Charles University in Prague\\
Malostransk\'e n\'am\v{e}st\'\i~25, 11000 Prague, Czech Republic\\
\email{tomas.kren@mff.cuni.cz}
\and
Institute of Computer Science\\
Academy of Sciences of the Czech Republic\\
Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207, Prague, Czech Republic\\
\email{roman@cs.cas.cz}
}
\maketitle

\begin{abstract}
In this paper, a generalization of genetic programming 
for simply typed lambda calculus is presented. We propose three population 
initialization methods depending on different search strategies. 
The first strategy corresponds to standard ramped half-and-half initialization, 
the second one corresponds to exhaustive systematic search, and the third one 
represents a novel geometric strategy. Three well-known benchmark experiments
support that the geometric strategy outperforms
the standard generating method in success rate, time consumption and average individual size. 
%Other performance enhancements based on theory of lambda calculus are 
%proposed and supported by experiments, including abstraction elimination 
%that enables us to utilize a simple tree-swapping crossover.
\end{abstract}

\section{Introduction}

%\subsection{gp a typy jsou dobrý / motivace}
%...
%\subsection{příběh}

%(...)\\

Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}. Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.   
Use of types naturally opens door to enriching S-expressions,
the traditional GP representation of individuals, with concepts from
lambda calculus, which is simple yet powerful functional mathematical and programming 
language extensively used in type theory. Such attempts has shown to be 
successful \cite{yu01}.

The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation. During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies. Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus. Thus, the natural idea is to employ an exhaustive systematic search. On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.  
These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies. This design goal also differentiate our approach from 
the three state of the art proposals for typed GP known to us that are discussed in the following section. 
Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search. Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.

%~\\(...)
%
%~\\
%\textbf{poznamky naky}\textit{\\
%- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
%jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
%je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
%a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
%- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
%- křížení by eliminace 
%}

%\subsection{obsah kapitol článku v 1 větě}

The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions. Main original results about search strategies in individual generating are described in section~\ref{approach}. Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.

\section{Related work}
\label{related}

In \cite{yu01} Yu presents a GP system utilizing
polymorphic higher-order functions\footnote{Higher-order 
function is a function taking another function as 
input parameter.} and lambda abstractions.
Important point of interest in this work is use of
\texttt{foldr} function as a tool for \textit{implicit recursion},
i.e. recursion without explicit recursive calls. 
The terminal set for constructing lambda abstraction subtrees 
is limited to use only constants and variables of that particular
lambda abstraction, i.e., outer variables are not allowed to be used
as terminals in this work. This is significant difference from our approach 
since we permit all well-typed normalized \lterms. From this difference also
comes different crossover operation. We focus more on term generating process; 
their term generation is performed in a similar way as the standard one, 
whereas our term generation also tries to utilize techniques of systematic enumeration.

In \cite{kes} Briggs and O’Neill present technique 
utilizing typed GP with combinators.
The difference between approach presented in this work
and our approach is that in this work terms are generated
straight from \textit{library} of combinators and no lambda abstractions
are used. They are using more general polymorphic type system then us
-- the Hindley–Milner type system. They also discuss the 
properties of exhaustive enumeration of terms and compare it with GP search.  
They also present interesting concept of \textit{Generalized
genetic operator} based on term generation.

In \cite{binard2008genetic} by Binard and Felty even 
stronger type system (\textit{System F}) is used.  
But with increasing power of the type system comes increasing difficulty of term generation.
For this reason evolution in this work takes interesting and nonstandard shape 
(fitness is associated with \textit{genes} which are evolved together with \textit{species}
which together participate in creation of individuals).
This differs from our approach, which tries to be generalization of
the standard GP\cite{koza92}.

In contrast with above mentioned works our approach uses very simple type system 
(simply typed lambda calculus) and concentrates on process of generation  
able to generate all possible well-typed normalized lambda terms. In order to do
so we use technique based on \textit{inhabitation machines} 
described by Barendregt in~\cite{barendregt10}.

\section{Preliminaries}
\label{preliminaries}
%definice, pojmy
%\subsection{Term}

In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
First, \lets 
%Here we  %formally 
describe a programming language, 
in which the GP algorithm generates individual programs --- the so called \lterms.

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	 	
Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
\begin{align*}
x   \in V \cup C  &\then x     \in \Lambda \\
M,N \in \Lambda   &\then (M~N) \in \Lambda 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

\textit{Function application} and 
\textit{$\lambda$-abstraction} are concepts
well known from common programming languages. 
For example in JavaScript 
$(M~N)$ translates to expression \texttt{$M$($N$)} and
$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
In other words, the function application 
corresponds to the act of supplying a function 
with an argument, and
the $\lambda$-abstraction is equivalent to 
\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
they were recently introduced to C++, and they are expected to be supported in Java 8.}.
$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for $(\dots((M_1~M_2)~M_3)~\dots~M_n)$
and $\lam{x_1 x_2 \dots x_n }{M}$ for $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.

%\subsubsection{Reductions}
%(... zmiňovat ? a pokud ano tak tady nebo radši v sekci která diskutuje 
%normálnost vygenerovanejch termů a jejich zkrácení pomocí $eta$-redukce)
%\subsubsection{Tree representations}
%(... zmiňovat ?)
%\subsection{Type}

A \lterm as described above
corresponds to a program expression with no type information
included. Now we will describe \textit{types} (or \textit{type terms}).

\begin{definition}
Let $A$ be set of {\it atomic type names}. 
Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
\begin{align*}
\alpha      \in A  &\then   \alpha \in \T \\
\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
\end{align*}
\end{definition}

Type $\sigma \ar \tau$ is type for functions taking as input
something of a type $\sigma$ and returning 
as output something of a type $\tau$. 
$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
combining the previously defined \textit{\lterms} and \textit{types} together.

%\subsection{Context}

\begin{definition}\begin{enumerate}
 \item 	Let $\Lambda$ be set of {\it \lterms}. 
	Let $\mathbb{T}$ be set of {\it types}.       
	A {\it statement} $M : \sigma$ is a pair 
	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
	Statement $M : \sigma$ is vocalized as 
	{\it "$M$ has type $\sigma$"}.
	The term $M$ is called the {\it subject} of the 
	statement $M : \sigma$.
 \item A \textit{declaration} is a statement 
 $x : \sigma$ where $x \in V \cup C$.

 \item A \textit{context} 
 %(or \textit{basis}) 
 is set of declarations with distinct variables as subjects.
\end{enumerate}
\end{definition}

Context is a basic type theoretic concept suitable as a typed alternative
for terminal and function set in standard GP. 
Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
such that $\Gamma$ does not contain any declaration with $x$ as subject.
We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.

%\subsection{Simply typed $\lambda$-calculus}

\begin{definition}
A statement $M\colon\sigma$ is \textit{derivable from}
a context $\Gamma$ (notation 
\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
if it can be produced by the following rules.
\begin{align*}
x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
\tur{\Gamma,x:\sigma}{M}{\tau}
&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
\end{align*}
\end{definition}

Our goal in term generation is to produce terms $M$
for a given pair $\ul{\tau}{\Gamma}$
such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.

%\begin{todo}
%    & From this definition we can derive following rules...
%
%	& citovat něco, asi barendrechta 
%	(on to tam ale nějak moc nedokazuje) 
%	případně mojí diplomku (tam to dokazuju)
%\end{todo}

%\subsection{Unfinished term}

\begin{definition}
Let $V$ be infinite countable set of {\it 
variable names}. Let $C$ be set of {\it constant names}, 
$V \cap C = \emptyset$.	
Let $\T$ be set of types.
Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
Then $\Lpr$ is set of 
\textit{unfinished  \lterms} defined inductively as follows.	
\begin{align*}
\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
\textit{~~~~~~~~(Unfinished leaf)}\\
x   \in V \cup C  &\then x     \in \Lpr \\
M,N \in \Lpr   &\then (M~N) \in \Lpr 
\textit{~~~~~~(Function application)} \\
x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
\textit{~~~~($\lambda$-abstraction)} 
\end{align*}
\end{definition}

\textit{Unfinished leaf} $\ul{\tau}{\Gamma}$ 
stands for yet not specified \lterm of the type $\tau$ 
build from symbols of $\Gamma$.

\section{Our approach}
\label{approach}
\subsection{Introduction}

Our approach to \lterm gerating is based on technique 
briefly described in \cite{barendregt10}, which generates
well-typed \lterms in their long normal form. 
We use this technique to perform systematic exhaustive enumeration
of \lterms in their long normal form in order from smallest to largest.
We use well known \textit{A* algorithm} \cite{AIMA} for this task.
A* is used to search in a given state space for a goal state. 
It finds the optimal solution (in our case the smallest term)
and uses "advising" heuristic function.
It maintains a priority queue to organize states yet to be explored.
Initially this queue contains only the initial state.

Our state space to search in is the space of unfinished \lterms. 
The initial state is the unfinished term $\ul{\tau}{\Gamma}$, 
where $\tau$ is the desired type of
terms to be generated and $\Gamma$ is the context
representing the set of building symbols to be used in construction of
terms (it corresponds to the set $T \cup F$ in
standard GP enriched with types). The process of determining 
successors of a state described below is designed so it constructs well-typed 
\lterms and omits no \lterm in its long normal form. 
A state is considered a goal state if it contains no unfinished
leaf, i.e., it is a finished \lterm.

Our generating method is based on simple modification of the
standard A*, which we call \textit{forgetful A*}. This modification consist in 
additional parameter for the A* algorithm -- the \textit{search strategy}. 
It is a simple filtration function (along with initialization procedure)
that is given the set of all successors of the state that is being examined
and returns a subset of this input. This subset is added to the priority queue 
to be further explored. In this way the search space may be reduced as 
the filtration function may \textit{forget} some successors.
If the queue becomes empty before the desired number of \lterms
is generated, then the initial statete is inserted to the queue
and the process continues. For the standard A* this would be meaningless,
but since our A* is forgetful this kind of restart makes sense.

A* keeps a priority queue of states during the generation process,
on the other hand the \textit{ramped half-and-half method}, 
the standard GP algorithm for generating individuals, 
keeps only one individual which is gradually constructed. This 
behavior is easily achieved by use of suitable search strategy 
that returns subset consisting of only one successor.
The systematic search is obtained by search strategy that 
returns whole input set.      
Our novel \textit{geometric strategy} can be understood as
point somewhere between those two extremes.

\subsection{Algorithm}

The inputs for the term generating algorithm are following.
\begin{enumerate}
 \item Desired type $\tau$ of generated terms.
 \item Context $\Gamma$ representing set of building symbols.
 \item Number $n$ of terms to be generated.
 \item Search strategy $S$. 
\end{enumerate}

Essential data structure of our algorithm 
is priority queue of unfinished terms. 
Priority of an unfinished term is given by its size\footnote{
A* heuristic function is hidden in method of computing
size of unfinished leafs $\ul{\tau}{\Gamma}$. Our algorithm uses
trivial estimate $\vert\ul{\tau}{\Gamma}\vert = 1$ which is trivially admissible.
This heuristic is not as silly as it might seem since it is
quite usual to have $x$ such that $x : \tau \in \Gamma$.
Since the true value of $\vert\ul{\tau}{\Gamma}\vert$ depends only on
$\tau$ and \textit{types} in $\Gamma$ (the \textit{signature}), 
no matter how many variables/constants of
each type there are, it is should by pretty effective to compute this
value precisely and store them for later. 
%This topic has other interesting 
%subtopics (such as surprisingly small 
%number of such "signatures" for a given problem), 
%but it is unfortunately beyond the scope of this paper.
}.
At the beginning, the queue contains only one unfinished term; 
$\ul{\tau}{\Gamma}$. The search strategy $S$ also 
initializes its internal state (if it has one).

At each step, the term $M$ with the smallest size
is pulled from the queue.
According to the number of unfinished leafs in $M$ one of
the following actions is performed.
\begin{enumerate}
 \item If the term $M$ has no unfinished leaf (i.e., it is a finished
 term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
 result set
% \footnote{
% \red{TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
% některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
% , případně to vůbec nezminovbat když by to kradlo moc místa
% }} 
 of generated terms.   
 \item Otherwise, \textit{successors} of the unfinished term $M$ are
       filtered out by \textit{search strategy} $S$ and
       those successors that outlast the filtration 
       are inserted into the queue.
\end{enumerate}

\textit{Successors} of an unfinished term $M$ are obtained by 
\textit{expansion} of the \mbox{\textit{DFS-first}} unfinished leaf $L$
(i.e., the leftmost unfinished leaf of $M$).

Expansion of the selected unfinished leaf $L$ leads to creation of 
one or many (possibly zero) successors.
In this process, $L$ is replaced
by a new subterm defined by the following rules\footnote{
For the sake of simplicity it is presented as two separate rules. 
Since the first rule results in exactly one successor it is smarter
to combine those two rules into one resulting in that unfinished leafs
have only atomic types. At the beginning we transform the initial type into atomic by first rule (if necessary). After that
only second rule is applied, but if it results in creation of some unatomic 
$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }.
\begin{enumerate}
 \item If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
 	   where $\alpha$ is atomic type and $n \geq 1$, 
       then $L$ is replaced by 
       $\lamb{x_1 \dots x_n}{\ul{\alpha}
       {\Gamma,x_1 \colon \rho_1,\dots,x_n \colon \rho_n}}$.
       Thus this expansion results in exactly one successor.  
 \item If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type,
       then for each 
       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}
       the unfinished leaf $L$ is replaced by 
       $(~f~\ul{\tau_1}{\Gamma}~\dots~\ul{\tau_m}{\Gamma}~)$.
       Thus this expansion results in many (possibly zero or one) successors.
\end{enumerate}

Now that we have all possible successors of $M$, we are about to apply
the \textit{search strategy} $S$. A search strategy is a procedure
which takes as input a set of unfinished terms and returns a subset
of the input set. Therefore, search strategy acts as a filter reducing 
the search space.

If the queue becomes empty before the desired number $n$ of terms
is generated, then the initial unfinished term $\ul{\tau}{\Gamma}$ 
is inserted to the queue, search strategy $S$
again initializes its internal state and the process continues.

\Lets now discuss three such search strategies.

\subsubsection{Systematic strategy}

If we use trivial strategy that returns all the inputs, 
then the algorithm systematically generates 
first $n$ smallest lambda terms in their
\textit{long normal form}.

\subsubsection{Ramped half-and-half strategy}
is generalization of the standard ramped half-and-half method described 
in \cite{koza92}. If applied to context satisfying closure requirement
(all constants/variables are of the same type, all functions are operations
on this type), it will behave in the same way as the standard method.
The internal state of this strategy consists of two variables.
It is the only one strategy described here that uses an internal state.

\begin{enumerate}
 \item \textit{isFull} - A boolean value, determining whether \textit{full}
                     or \textit{grow} method will be performed.
 \item \textit{d} - An integer value from $\setDots{2}{D_{init}}$, where 
                $D_{init}$ is predefined maximal depth (e.g. 6).                    
\end{enumerate}

This strategy returns precisely one randomly
(uniformly) selected  element from 
the \textit{selection subset} of input set
(or zero elements if the input set is empty). 
The \textit{selection subset} 
to select from is determined by $depth, d$ and $isFull$.
The $depth$ parameter is the depth (in the term tree) 
of the unfinished leaf that was expanded.
Those elements of input set whose newly added subtree contains one ore more 
unfinished leafs are regarded as \textit{non-terminals}, whereas 
those whose newly added subtree contains no unfinished leaf are regarded as 
\textit{terminals}.
If $depth = 0$, then the subset to select from is  
set of all \textit{non-terminals} of the input set.
If $depth = d$, then the subset to select from is
set of all \textit{terminals} of the input set.
In other cases of $depth$ it depends on value of $isFull$.
If $isFull = true$, then the subset to select from is 
set of all \textit{non-terminals} of the input set.
If $isFull = false$, then the subset to select from is 
the whole input set.

\subsubsection{Geometric strategy}

We can see those two previous strategies as two extremes on the spectrum of 
possible strategies. 
\textit{Systematic strategy} filters no successor state thus performing
exhaustive search resulting in discovery of $n$ smallest terms in one run.
On the other hand, \textit{ramped half-and-half strategy} filters 
all but one successor states resulting in degradation of 
the priority queue into "fancy variable".
\textit{Geometric strategy} is simple yet fairly effective term generating 
strategy somewhere in the middle of this spectrum.
It is parameterized by parameter $q \in (0,1)$, its default well-performing 
value is $q = 0.75$.
For each element of the input set 
it is probabilistically decided whether
it will be returned or omitted. A probability $p$ of returning is
same for all elements, but depends on the $depth$, 
which is defined in the same way as in previous strategy. 
It is computed as follows.
$$ p = q^{depth} $$
This formula is motivated by idea that it is important to
explore all possible root symbols, but as the $depth$ 
increases it becomes less "dangerous" to omit 
an exploration branch. 
We can see this by considering that this strategy results in
somehow forgetful A* search.
With each omission we make the search space smaller. But with
increasing depth these omissions have smaller impact on the search space,
i.e., they cut out lesser portion of the search space.
Another slightly esoteric argument supporting this formula is that "root 
parts" of a program usually stand for more crucial parts
with radical impact on global behavior of a program, 
whereas "leaf parts" of a program usually
stand for less important local parts (e.g. constants).  
This strategy also plays nicely with the idea that 
"too big trees should be killed".

~\\
Furthermore, our system utilizes generalization of the standard 
tree-swapping crossover operator. Since it is beyond the scope of this 
paper we mention it only briefly. Two main concerns with swapping typed subtrees 
are types and free variables. Well-typed offspring is obtained 
by swapping only subtrees of the same type. Only subtrees with 
corresponding counterpart in the second parent are randomly chosen from.
More interesting problem lies in free variables, which may cause trouble
if swapped somewhere where it is suddenly not bounded. In order to circumvent this
difficulty we utilize technique called \textit{abstraction elimination}\cite{jones87}
that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions
and no bound variables. After the initial population is generated,  it is transformed
by abstraction elimination. Another possible transformation taking place
after initialization is $\eta$-normalization shortening 
rather long long normal form into $\beta\eta$-normal form.  
Another performence enhancing transformation is use of "applicative" tree 
representation (coming directly from inductive definition of \lterms).

\section{Experiments}
\label{experiments}

We made tree experiments comparing performance of standard 
\textit{ramped half-and-half strategy} with our
\textit{geometric strategy} (with default parameter $q=0.75$). 
Each experiment consisted of 50 independent runs 
of GP algorithm. Each run had maximally 51 generations and population size
was 500 individuals.
In those experiments we analyze the ability of the system to produce 
a correct solution, computational cost estimates and average size 
of an individual throughout generations. 
For the first two metrics we use the popular measurement  
methods within GP field --- the \textit{performance curves}
described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
and $I(M,i,z)$ (the total number of individuals that must be processed 
to yield a correct solution with probability $z =99\%$).

\subsection{Simple symbolic regression}
\textit{Simple Symbolic Regression} is a problem described
in \cite{koza92}. Objective of this problem is to 
find a function $f(x)$ that fits a sample
of twenty given points. The target function is 
function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
Desired type of generated programs $\tau$ and 
building blocks context $\Gamma$ are following.
\begin{align*}
\tau = \Real \ar &\Real\\
\Gamma = \{
  (+)  &: \Real \ar \Real \ar \Real    ,
  (-)   : \Real \ar \Real \ar \Real    ,
  (*)   : \Real \ar \Real \ar \Real    ,
  rdiv  : \Real \ar \Real \ar \Real    ,\\
  sin  &: \Real \ar \Real              ,
  cos   : \Real \ar \Real              ,
  exp   : \Real \ar \Real              , 
  rlog  : \Real \ar \Real              \}
\end{align*}
where
\noindent
\begin{minipage}{.5\linewidth}
\begin{align*}
rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
p/q & \mbox{otherwise } \end{cases}  
\end{align*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
\begin{align*}
rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
log(\vert x\vert) & \mbox{otherwise}. \end{cases}
\end{align*}
\end{minipage}

\begin{figure}[h!]
  \caption{Performance curves for Simple symbolic regression.}
  \centering
    \includegraphics[scale=0.135]{imgs/SSR_FINAL.png}
\end{figure}
Fitness function is computed as follows.
$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
considered as a correct individual.

Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
The average individual size for generation 50 was 76.
The experiment took 46 minutes.
Our \textit{geometric strategy} scored 21/50 (42\%) success rate. 
Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
The average individual size for generation 50 was 47.
The experiment took 26 minutes.
Figure~1 shows the progress of those metrics throughout generations.
So for the \textit{Simple symbolic regression}
our geometric strategy is slightly more successful then 
ramped half-and-half strategy in all observed metrics.

\subsection{Artificial ant}
\textit{Artificial Ant} is another problem described
in \cite{koza92}. Objective of this problem is to 
find a control program for an artificial ant so
that it can find all food located on "Santa Fe" trail.
The Santa Fe trail lies on toroidal square grid.
The ant is in the upper left corner, facing right.
The ant is able to move forward, turn left, and sense if a food 
piece is ahead of him.

\begin{align*}
\tau = An&tAction\\
\Gamma = \{~~
  l    &: AntAction                              ,
  r    : AntAction                               ,
  m    : AntAction                               ,\\
  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
\end{align*}

Action $l$ turns the ant left. 
Action $r$ turns the ant right.
Action $m$ moves the ant forward.
Action $ifa~x~y$ (if-food-ahead) performs action $x$ 
if a food piece is ahead of the ant,
otherwise it performs action $y$.
Action $p2~x~y$ first performs action $x$ and after it action $y$.
Action $p3~x~y~z$ first performs action $x$, 
after that action $y$ and finally $z$.
Actions $l, r$ and $m$ each take one time step to execute.
Ants action is performed over and over again until it reaches predefined
maximal number of steps.

\begin{figure}[h!]
  \caption{Performance curves for Artificial ant problem.}
  \centering
    \includegraphics[scale=0.135]{imgs/ANT_FINAL.png}
\end{figure}

Fitness value is equal to number of eaten food pieces.
An individual such that eats all 89 pieces of food is 
considered as a correct solution.
This limit is set to be 600 time steps\footnote{
In \cite{koza92} this limit is said to be 400 time steps.
But there is also mentioned following solution, 
which is described as correct solution:
\texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
This program needs 545
time steps; if it is given only 400 time steps, then it eats only 79 pieces
of food. Thus we use 600 time steps. 
}.

Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
The average individual size for generation 50 was circa 400.
The experiment took 265 minutes.
Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
The average individual size for generation 50 was circa 90.
The experiment took 107 minutes.
This is a big improvement in all watched factors.
Figure~2 shows it in greater detail.

\subsection{Even parity problem}

Previous two problems were instances of standard GP, i.e. their building symbols
obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
The even-parity function is a boolean function taking as inputs $N$
boolean values and returning \textit{True} if an even number of inputs 
are \textit{True}. For odd number it returns \textit{False}.
This problem has been used by many researchers
as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
where is presented approach to evolve recursive and modular programs
by use of higher-order functions and $\lambda$-abstractions.
We use very similar set of \textit{building symbols} as in \cite{yu01}. 
\begin{align*}
\tau = [Boo&l] \ar Bool\\
\Gamma = \{
  and   &: Bool \ar Bool \ar Bool                              ,
  or    : Bool \ar Bool \ar Bool                              ,\\
  nand  &: Bool \ar Bool \ar Bool                              ,
  nor    : Bool \ar Bool \ar Bool                              ,\\
  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
  head' &: [Bool] \ar Bool                                   ,
  tail' : [Bool] \ar [Bool]                              \}
\end{align*}
\begin{figure}[h!]
  \caption{Performance curves for Even parity problem.}
  \centering
    \includegraphics[scale=0.135]{imgs/EP_FINAL.png}
\end{figure}

The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
this problem is considered atomic.
Unlike in \cite{yu01}, we use specific instance of polymorphic 
function \texttt{foldr}. 
And modifications of functions \texttt{head} 
(returning the first element of the list) 
and \texttt{tail} (returning the list without the first element) are used; 
making them total by returning default value \textit{False}
and \texttt{[]}, respectively.
We use the same fitness function as in \cite{yu01}. 
The fitness function examines the individual by giving
it all possible boolean lists of length 2 and 3.

Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
The average individual size for generation 50 was 39.
The experiment took 28 minutes.
Our \textit{geometric strategy} scored 32/50 (64\%) success rate. 
Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
The average individual size for generation 50 was circa 100.
The experiment took 33 minutes.
Figure~3 shows it in greater detail.
Once again this is a big improvement against standard method. 
One may get the impression that geometric strategy suffers from bloat here,
but considering that it performs significantly better and
that term size around 100 is similar as for the previous two problems, 
it seems to us that ramped half-and-half maybe has difficulties with 
constructing more complex terms, and therefore is "staying small".
Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
the 100\% correct solution was present in the generation 0. This is pretty good
result for uninformed search. However, our results are less successful 
then those presented in \cite{yu01} (with slightly different set of 
building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
17,500 in generation 4. But by observation of typical solution which is often
some modification of \texttt{foldr xor False inputList} one sees that important task
for GP here is to create \texttt{xor} function. This is advantage for the more specialized term representation used in \cite{yu01} 
(restricting use of outer variables). 
The difference in crossover operators may also be involved in the difference, 
but such analysis is beyond the scope of this paper. Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
scored 17/60 (28\%); min $I(M,i,z) = 220,000$. \textit{GP with ADFs} 
scored 10/29 (34\%); min $I(M,i,z) = 1,440,000$.

\section{Conclusions}
\label{conclusions}

In this work, we have proposed a generalization of genetic programming 
for simply typed lambda calculus. The main focus of this paper is on individual generation algorithms. 
The main idea is to explore the spectrum of available approaches bounded on one end by the traditional
ramped half-and-half strategy, while the second bound represents an exhaustive systematic search. 
Thus, three population initialization methods have been designed, 
depending on different search strategies. 
The first strategy corresponds to the above mentioned traditional genetic programming initialization,  
the second one corresponds to exhaustive search, and the third one 
represents a novel geometric strategy which outperforms standard genetic 
programming in success rate, resource consumption and average individual size 
on the presented experiments.

%\red{ [větu dvě o experimentech]}

The relevance of these algorithms does not effect only initialization, but it can be used during mutation
operator in a straightforward way. 
%Other performance enhancements based on theory of lambda calculus are 
%proposed and supported by experiments, including abstraction 
%elimination that enables us to utilize
%a simple tree-swapping crossover.

Our future research will focus on exploring other possible search strategies positioned in the middle ground 
identified in this paper. Since the search strategy can be easily described by a rather simple filtration algorithm, it might be
interesting to allow the typed GP system to evolve the search strategy it 
utilizes in a kind of meta-evolution way.

%\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }

\nocite{*}

\bibliographystyle{splncs}
\bibliography{evogp}

\end{document}

%
%\begin{thebibliography}{1}
%
%
%\bibitem{koza92}
  %John R. Koza,
  %\emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
  %MIT Press, Cambridge, MA,
  %1992. 
%
%\bibitem{koza05}
  %Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
  %\emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
  %Springer, 2005. ISBN 978-0-387-26417-2 
%
%\bibitem{fg}
 %Riccardo Poli, William B. Langdon, Nicholas F. McPhee
 %\emph{A Field Guide to Genetic Programming}.
 %Lulu Enterprises, UK Ltd, 2008.
%
%\bibitem{yu01}
  %T. Yu. 
  %\emph{Hierachical processing for evolving recursive and modular 
        %programs using higher order functions and lambda abstractions}. 
  %Genetic Programming and Evolvable Machines,
  %2(4):345–380, December 2001. ISSN 1389-2576.
%
%
%\bibitem{montana95}
%D. J. Montana. 
%\emph{Strongly typed genetic programming.} 
%Evolutionary Computation, 3(2): 199–230, 1995.
%%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
%
%\bibitem{haynes96}
%T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
%\emph{Type inheritance in strongly typed genetic programming.} 
%In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
%in Genetic Programming 2, chapter 18, pages 359–376.
%MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
%%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
%
%\bibitem{olsson94}
%J. R. Olsson. 
%\emph{Inductive functional programming using incremental program 
%transformation and Execution of logic programs by 
%iterative-deepening A* SLD-tree search.} 
%Dr scient thesis, University of Oslo, Norway, 1994.
%
%\bibitem{kes}
%Forrest Briggs, Melissa O’Neill.
%\emph{Functional Genetic Programming and Exhaustive
%Program Search with Combinator Expressions.}
%International Journal of Knowledge-based and Intelligent Engineering Systems,
%Volume 12 Issue 1, Pages 47-68, January 2008. 
%
%
%\bibitem{barendregt84}
%H. P. Barendregt,
%\emph{The Lambda Calculus: its Syntax and Semantics}, 
%revised ed., North-Holland, 1984.
%
%\bibitem{barendregt92}
%H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
%\emph{Lambda Calculi with Types.} 
%Handbook of Logic in Computer Science, 1992. 
%
%\bibitem{barendregt10}
%
  %Henk Barendregt, Wil Dekkers, Richard Statman,
  %\emph{Lambda Calculus With Types}.
  %Cambridge University Press,
  %2010. 
  %%URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
%
%\bibitem{jones87}
%Simon Peyton Jones. 
%\emph{The Implementation of Functional Programming Languages}. 
%Prentice Hall, 1987.
%
%
%\bibitem{AIAMA}
	%Stuart J. Russell, Peter Norvig,
	%\emph{Artificial Intelligence: A Modern Approach}.
	%Pearson Education,
	%2003. 
%
%
%\end{thebibliography}

### 2013-11-30 js/fishtronDB.js
change apiURL for fishtronDB

@@ -1,9 +1,10 @@
 var FishtronDB = function () {
 
-  var apiURL = 'http://kutil.php5.cz/fishtronDB'; 
+  var apiURL = 'http://146.255.30.53/fishtronDB'; 
+  //var apiURL = 'http://kutil.php5.cz/fishtronDB'; 
   //var apiURL = 'http://localhost/fishtronDB';
+  
   var nextHandlerID = 1;
-
   var handlers = {};
 
   function load (args,fun) {

### 2013-12-01 js/node/mysql-test.js
node test

@@ -1,0 +1,19 @@
+var mysql = require('mysql');
+
+var connection = mysql.createConnection({
+  host     : 'localhost',
+  user     : 'root',//'sekol',
+  password : '',//'tribar42',
+  database : 'test',
+});
+
+connection.connect();
+
+connection.query("SELECT `val` FROM `big_object` WHERE `key` = 'zobrazeni'", 
+  function(err, rows, fields) {
+    if (err) throw err;
+    console.log('Tadaa: ', rows[0].val);
+  }
+);
+
+connection.end();

### 2013-12-01 js/node/test.js

@@ -1,0 +1,57 @@
+var http  = require('http');
+var mysql = require('mysql');
+var fs    = require('fs');
+
+var i  = 0;
+
+function run (config) {
+  http.createServer(function (req, res) {
+    i++;
+
+    var connection = mysql.createConnection(config.mysql);
+    connection.connect();
+
+    var sql = "SELECT `val` FROM `big_object` WHERE `key` = 'zobrazeni'";
+    connection.query(sql, function (err, rows, fields) {
+    
+      res.writeHead(200, {'Content-Type': 'text/plain; charset=utf-8'});
+      res.write('Ahoj pane číslo '+i+'!\n');
+    
+
+      if (err) {
+        console.log(err.toString());
+      } else {
+        res.write('big_object.zobrazeni = '+ rows[0].val+'\n');  
+      }
+
+      res.end();
+      
+    });
+
+    connection.end();
+
+    console.log(i);
+  }).listen(config.port, config.ip);
+  console.log('Server running at http://'+config.ip+':'+config.port );
+}
+
+function loadConfig (onLoad) {
+  var localConfigFile = __dirname + '/config.json';
+  fs.readFile(localConfigFile, 'utf8', function (err, data) {
+    if (!err) {
+      onLoad(JSON.parse(data));
+    } else {
+      var defaultConfigFile = __dirname + '/config.default.json';
+      fs.readFile(defaultConfigFile, 'utf8', function (err, data) {
+        if (!err) {
+          console.log('(using config.default.json)');
+          onLoad(JSON.parse(data));
+        } else {
+          console.log(err.toString());
+        }
+      });
+    }
+  });
+}
+
+loadConfig(run);

### 2013-12-02 tex/evogp13/vyrez-relwork.tex
minor tex

--- 2013-11-29 tex/evogp13/evogp.tex
@@
-\documentclass{llncs}
+\documentclass{article}
 
-%\usepackage{times}
-\usepackage[utf8]{inputenc}
-\usepackage{amsfonts}
-%\usepackage{qtree}
 \usepackage{xspace}
 
-%\usepackage{makeidx}  % allows for indexgeneration
-%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
-%\usepackage{hyperref}
-\usepackage{amsmath}
-%\usepackage{amssymb}
-\usepackage{graphicx}
-
-\usepackage[ampersand]{easylist}
-\usepackage[usenames,dvipsnames]{color}
-
-\newcommand{\Lets}{Let us\xspace}
-\newcommand{\lets}{let us\xspace}
 \newcommand{\lterm}{$\lambda$-term\xspace}
 \newcommand{\lterms}{$\lambda$-terms\xspace}
-\newcommand{\lhead}{$\lambda$-head\xspace}
-\newcommand{\lheads}{$\lambda$-heads\xspace}
-\newcommand{\la}{\leftarrow\xspace}
-\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
-\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
-\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
-\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
-\newcommand{\atTree}{@-tree\xspace}
-\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
-\newcommand{\lh}[1]{\lambda #1}
-\newcommand{\sexprTree}{sexpr-tree\xspace}
-\newcommand{\SexprTree}{Sexpr-tree\xspace}
-\newcommand{\then}{\Rightarrow\xspace}
-\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
-\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
-\newcommand{\ST}{\mathop{\mathrm{ST}}}
-\newcommand{\FV}{\mathop{\mathrm{FV}}}
-\newcommand{\Scomb }{\mathbf{S}}
-\newcommand{\Kcomb }{\mathbf{K}}
-\newcommand{\Icomb }{\mathbf{I}}
-\newcommand{\bbarr}{\twoheadrightarrow_\beta}
-\newcommand{\barr}{\rightarrow_\beta}
-\newcommand{\beq}{=_\beta}
-\newcommand{\eearr}{\twoheadrightarrow_\eta}
-\newcommand{\earr}{\rightarrow_\eta}
-\newcommand{\eeq}{=_\eta}
-\newcommand{\bearr}{\rightarrow_{\beta\eta}}
-\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
-\newcommand{\beeq}{=_{\beta\eta}}
-\newcommand{\etar}{\twoheadrightarrow_\eta}
-\newcommand{\ered}{$\eta$-reduction\xspace}
-\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
-\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
-\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
-\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
-\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
-\newcommand{\bredex}{$\beta$-redex\xspace} 
-\newcommand{\lnf}{\textit{lnf}\xspace}
-\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
-\newcommand{\Bcomb }{\mathbf{B}}   
-\newcommand{\BBcomb }{\mathbf{B*}}
-\newcommand{\Ccomb }{\mathbf{C}}   
-\newcommand{\CCcomb }{\mathbf{C'}}
-\newcommand{\SScomb }{\mathbf{S'}}
-\newcommand{\ar}{\rightarrow\xspace}
-\newcommand{\T}{\mathbb{T}\xspace}
-\newcommand{\C}{\mathbb{C}\xspace}
-\newcommand{\Real}{\mathbb{R}}
-
-\newenvironment{todo}
-{~\\ {\color{red}\textbf{TODO}}
-  \begin{easylist}[itemize]}
-{ \end{easylist}}
-
-\newcommand{\Lpr}{\Lambda^\prime}
-\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
 \newcommand{\ro}[1]{{\color{blue} #1}}
-\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
-\newcommand{\red}[1]{{\color{red} #1}}
-
-%\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}
 
 \begin{document}
-
-\title{
-Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
-\author{Tom\'{a}\v{s} K\v{r}en\inst{1} \and Roman Neruda\inst{2}}
-\institute{%
-Faculty of Mathematics and Physics\\
-Charles University in Prague\\
-Malostransk\'e n\'am\v{e}st\'\i~25, 11000 Prague, Czech Republic\\
-\email{tomas.kren@mff.cuni.cz}
-\and
-Institute of Computer Science\\
-Academy of Sciences of the Czech Republic\\
-Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207, Prague, Czech Republic\\
-\email{roman@cs.cas.cz}
-}
-\maketitle
-
-\begin{abstract}
-In this paper, a generalization of genetic programming 
-for simply typed lambda calculus is presented. We propose three population 
-initialization methods depending on different search strategies. 
-The first strategy corresponds to standard ramped half-and-half initialization, 
-the second one corresponds to exhaustive systematic search, and the third one 
-represents a novel geometric strategy. Three well-known benchmark experiments
-support that the geometric strategy outperforms
-the standard generating method in success rate, time consumption and average individual size. 
-%Other performance enhancements based on theory of lambda calculus are 
-%proposed and supported by experiments, including abstraction elimination 
-%that enables us to utilize a simple tree-swapping crossover.
-\end{abstract}
-
-\section{Introduction}
-
-%\subsection{gp a typy jsou dobrý / motivace}
-%...
-%\subsection{příběh}
-
-%(...)\\
-
-Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
-Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
-Use of types naturally opens door to enriching S-expressions,
-the traditional GP representation of individuals, with concepts from
-lambda calculus, which is simple yet powerful functional mathematical and programming 
-language extensively used in type theory. Such attempts has shown to be 
-successful \cite{yu01}.
-
-The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
-During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
-Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
-Thus, the natural idea is to employ an exhaustive systematic search.
-On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.
-These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
-This design goal also differentiate our approach from
-the three state of the art proposals for typed GP known to us that are discussed in the following section. 
-Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
-Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
-
-%~\\(...)
-%
-%~\\
-%\textbf{poznamky naky}\textit{\\
-%- geometrická strategie generování termů jakožto hybrid systematického a náhodného sytylu --- to že jsme si vytičili ty dva cíle který plníme tou 
-%jednoduchou strategií, tak máme možnost nalízt jednoduchou strategii která 
-%je na pomezí obou cílů - taková strategie je právě ta naše geometrická 
-%a ukazuje se že se bohulibě chová právě k strašáku GP -- \textbf{bloatu}.\\
-%- teoretické LC konstrukty s výhodou použity - @-stromy a eta-redukce\\
-%- křížení by eliminace 
-%}
-
-%\subsection{obsah kapitol článku v 1 větě}
-
-The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
-Main original results about search strategies in individual generating are described in section~\ref{approach}.
-Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.
 
@@
 
+\ro{
 In \cite{yu01} Yu presents a GP system utilizing
@@
 whereas our term generation also tries to utilize techniques of systematic enumeration. 
-
+}
 In \cite{kes} Briggs and O’Neill present technique 
@@
 
-\section{Preliminaries}
-\label{preliminaries}
-%definice, pojmy
-%\subsection{Term}
+\begin{thebibliography}{1}
 
-In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
-First, \lets 
-%Here we  %formally 
-describe a programming language, 
-in which the GP algorithm generates individual programs --- the so called \lterms.
+\bibitem{koza92}
+  John R. Koza,
+  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
+  MIT Press, Cambridge, MA,
+  1992.
 
-\begin{definition}
-Let $V$ be infinite countable set of {\it 
-variable names}. Let $C$ be set of {\it constant names}, 
-$V \cap C = \emptyset$.	 	
-Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
-\begin{align*}
-x   \in V \cup C  &\then x     \in \Lambda \\
-M,N \in \Lambda   &\then (M~N) \in \Lambda 
-\textit{~~~~~~(Function application)} \\
-x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
-\textit{~~~~($\lambda$-abstraction)} 
-\end{align*}
-\end{definition}
+\bibitem{yu01}
+  T. Yu. 
+  \emph{Hierachical processing for evolving recursive and modular 
+        programs using higher order functions and lambda abstractions}. 
+  Genetic Programming and Evolvable Machines,
+  2(4):345–380, December 2001. ISSN 1389-2576.
 
-\textit{Function application} and 
-\textit{$\lambda$-abstraction} are concepts
-well known from common programming languages. 
-For example in JavaScript 
-$(M~N)$ translates to expression \texttt{$M$($N$)} and
-$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
-In other words, the function application 
-corresponds to the act of supplying a function 
-with an argument, and
-the $\lambda$-abstraction is equivalent to 
-\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
-they were recently introduced to C++, and they are expected to be supported in Java 8.}.
-$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for $(\dots((M_1~M_2)~M_3)~\dots~M_n)$
-and $\lam{x_1 x_2 \dots x_n }{M}$ for $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
+\bibitem{binard2008genetic}
+Binard, F., Felty, A.: Genetic programming with polymorphic types and higher-order 
+functions. In: Proceedings of the 10th annual conference on Genetic and
+evolutionary computation, ACM (2008) 1187-1194
 
-%\subsubsection{Reductions}
-%(... zmiňovat ? a pokud ano tak tady nebo radši v sekci která diskutuje 
-%normálnost vygenerovanejch termů a jejich zkrácení pomocí $eta$-redukce)
-%\subsubsection{Tree representations}
-%(... zmiňovat ?)
-%\subsection{Type}
+\bibitem{kes}
+Forrest Briggs, Melissa O’Neill.
+\emph{Functional Genetic Programming and Exhaustive
+Program Search with Combinator Expressions.}
+International Journal of Knowledge-based and Intelligent Engineering Systems,
+Volume 12 Issue 1, Pages 47-68, January 2008.
 
-A \lterm as described above
-corresponds to a program expression with no type information
-included. Now we will describe \textit{types} (or \textit{type terms}).
+\bibitem{barendregt10}
+  Henk Barendregt, Wil Dekkers, Richard Statman,
+  \emph{Lambda Calculus With Types}.
+  Cambridge University Press,
+  2010. 
+  %URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
 
-\begin{definition}
-Let $A$ be set of {\it atomic type names}. 
-Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
-\begin{align*}
-\alpha      \in A  &\then   \alpha \in \T \\
-\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
-\end{align*}
-\end{definition}
-
-Type $\sigma \ar \tau$ is type for functions taking as input
-something of a type $\sigma$ and returning 
-as output something of a type $\tau$. 
-$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
-$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
-The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
-combining the previously defined \textit{\lterms} and \textit{types} together.
-
-%\subsection{Context}
-
-\begin{definition}\begin{enumerate}
- \item 	Let $\Lambda$ be set of {\it \lterms}. 
-	Let $\mathbb{T}$ be set of {\it types}.       
-	A {\it statement} $M : \sigma$ is a pair 
-	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
-	Statement $M : \sigma$ is vocalized as 
-	{\it "$M$ has type $\sigma$"}.
-	The term $M$ is called the {\it subject} of the 
-	statement $M : \sigma$.
- \item A \textit{declaration} is a statement 
- $x : \sigma$ where $x \in V \cup C$.
-
- \item A \textit{context} 
- %(or \textit{basis}) 
- is set of declarations with distinct variables as subjects.
-\end{enumerate}
-\end{definition}
-
-Context is a basic type theoretic concept suitable as a typed alternative
-for terminal and function set in standard GP. 
-Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
-such that $\Gamma$ does not contain any declaration with $x$ as subject.
-We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
-
-%\subsection{Simply typed $\lambda$-calculus}
-
-\begin{definition}
-A statement $M\colon\sigma$ is \textit{derivable from}
-a context $\Gamma$ (notation 
-\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
-if it can be produced by the following rules.
-\begin{align*}
-x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
-\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
-&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
-\tur{\Gamma,x:\sigma}{M}{\tau}
-&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
-\end{align*}
-\end{definition}
-
-Our goal in term generation is to produce terms $M$
-for a given pair $\ul{\tau}{\Gamma}$
-such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
-
-%\begin{todo}
-%    & From this definition we can derive following rules...
-%
-%	& citovat něco, asi barendrechta 
-%	(on to tam ale nějak moc nedokazuje) 
-%	případně mojí diplomku (tam to dokazuju)
-%\end{todo}
-
-%\subsection{Unfinished term}
-
-\begin{definition}
-Let $V$ be infinite countable set of {\it 
-variable names}. Let $C$ be set of {\it constant names}, 
-$V \cap C = \emptyset$.	
-Let $\T$ be set of types.
-Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
-Then $\Lpr$ is set of 
-\textit{unfinished  \lterms} defined inductively as follows.	
-\begin{align*}
-\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
-\textit{~~~~~~~~(Unfinished leaf)}\\
-x   \in V \cup C  &\then x     \in \Lpr \\
-M,N \in \Lpr   &\then (M~N) \in \Lpr 
-\textit{~~~~~~(Function application)} \\
-x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
-\textit{~~~~($\lambda$-abstraction)} 
-\end{align*}
-\end{definition}
-
-\textit{Unfinished leaf} $\ul{\tau}{\Gamma}$ 
-stands for yet not specified \lterm of the type $\tau$ 
-build from symbols of $\Gamma$.
-
-\section{Our approach}
-\label{approach}
-\subsection{Introduction}
-
-Our approach to \lterm gerating is based on technique 
-briefly described in \cite{barendregt10}, which generates
-well-typed \lterms in their long normal form. 
-We use this technique to perform systematic exhaustive enumeration
-of \lterms in their long normal form in order from smallest to largest.
-We use well known \textit{A* algorithm} \cite{AIMA} for this task.
-A* is used to search in a given state space for a goal state. 
-It finds the optimal solution (in our case the smallest term)
-and uses "advising" heuristic function.
-It maintains a priority queue to organize states yet to be explored.
-Initially this queue contains only the initial state.
-
-Our state space to search in is the space of unfinished \lterms. 
-The initial state is the unfinished term $\ul{\tau}{\Gamma}$, 
-where $\tau$ is the desired type of
-terms to be generated and $\Gamma$ is the context
-representing the set of building symbols to be used in construction of
-terms (it corresponds to the set $T \cup F$ in
-standard GP enriched with types). The process of determining 
-successors of a state described below is designed so it constructs well-typed 
-\lterms and omits no \lterm in its long normal form. 
-A state is considered a goal state if it contains no unfinished
-leaf, i.e., it is a finished \lterm.
-
-Our generating method is based on simple modification of the
-standard A*, which we call \textit{forgetful A*}. This modification consist in 
-additional parameter for the A* algorithm -- the \textit{search strategy}. 
-It is a simple filtration function (along with initialization procedure)
-that is given the set of all successors of the state that is being examined
-and returns a subset of this input. This subset is added to the priority queue 
-to be further explored. In this way the search space may be reduced as 
-the filtration function may \textit{forget} some successors.
-If the queue becomes empty before the desired number of \lterms
-is generated, then the initial statete is inserted to the queue
-and the process continues. For the standard A* this would be meaningless,
-but since our A* is forgetful this kind of restart makes sense.
-
-A* keeps a priority queue of states during the generation process,
-on the other hand the \textit{ramped half-and-half method}, 
-the standard GP algorithm for generating individuals, 
-keeps only one individual which is gradually constructed. This 
-behavior is easily achieved by use of suitable search strategy 
-that returns subset consisting of only one successor.
-The systematic search is obtained by search strategy that 
-returns whole input set.      
-Our novel \textit{geometric strategy} can be understood as
-point somewhere between those two extremes.
-
-\subsection{Algorithm}
-
-The inputs for the term generating algorithm are following.
-\begin{enumerate}
- \item Desired type $\tau$ of generated terms.
- \item Context $\Gamma$ representing set of building symbols.
- \item Number $n$ of terms to be generated.
- \item Search strategy $S$. 
-\end{enumerate}
-
-Essential data structure of our algorithm 
-is priority queue of unfinished terms. 
-Priority of an unfinished term is given by its size\footnote{
-A* heuristic function is hidden in method of computing
-size of unfinished leafs $\ul{\tau}{\Gamma}$. Our algorithm uses
-trivial estimate $\vert\ul{\tau}{\Gamma}\vert = 1$ which is trivially admissible.
-This heuristic is not as silly as it might seem since it is
-quite usual to have $x$ such that $x : \tau \in \Gamma$.
-Since the true value of $\vert\ul{\tau}{\Gamma}\vert$ depends only on
-$\tau$ and \textit{types} in $\Gamma$ (the \textit{signature}), 
-no matter how many variables/constants of
-each type there are, it is should by pretty effective to compute this
-value precisely and store them for later. 
-%This topic has other interesting 
-%subtopics (such as surprisingly small 
-%number of such "signatures" for a given problem), 
-%but it is unfortunately beyond the scope of this paper.
-}.
-At the beginning, the queue contains only one unfinished term; 
-$\ul{\tau}{\Gamma}$. The search strategy $S$ also 
-initializes its internal state (if it has one).
-
-At each step, the term $M$ with the smallest size
-is pulled from the queue.
-According to the number of unfinished leafs in $M$ one of
-the following actions is performed.
-\begin{enumerate}
- \item If the term $M$ has no unfinished leaf (i.e., it is a finished
- term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
- result set
-% \footnote{
-% \red{TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
-% některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
-% , případně to vůbec nezminovbat když by to kradlo moc místa
-% }} 
- of generated terms.   
- \item Otherwise, \textit{successors} of the unfinished term $M$ are
-       filtered out by \textit{search strategy} $S$ and
-       those successors that outlast the filtration 
-       are inserted into the queue.
-\end{enumerate}
-
-\textit{Successors} of an unfinished term $M$ are obtained by 
-\textit{expansion} of the \mbox{\textit{DFS-first}} unfinished leaf $L$
-(i.e., the leftmost unfinished leaf of $M$).
-
-Expansion of the selected unfinished leaf $L$ leads to creation of 
-one or many (possibly zero) successors.
-In this process, $L$ is replaced
-by a new subterm defined by the following rules\footnote{
-For the sake of simplicity it is presented as two separate rules. 
-Since the first rule results in exactly one successor it is smarter
-to combine those two rules into one resulting in that unfinished leafs
-have only atomic types. At the beginning we transform the initial type into atomic by first rule (if necessary). After that
-only second rule is applied, but if it results in creation of some unatomic 
-$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }.
-\begin{enumerate}
- \item If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
- 	   where $\alpha$ is atomic type and $n \geq 1$, 
-       then $L$ is replaced by 
-       $\lamb{x_1 \dots x_n}{\ul{\alpha}
-       {\Gamma,x_1 \colon \rho_1,\dots,x_n \colon \rho_n}}$.
-       Thus this expansion results in exactly one successor.  
- \item If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type,
-       then for each 
-       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}
-       the unfinished leaf $L$ is replaced by 
-       $(~f~\ul{\tau_1}{\Gamma}~\dots~\ul{\tau_m}{\Gamma}~)$.
-       Thus this expansion results in many (possibly zero or one) successors.
-\end{enumerate}
-
-Now that we have all possible successors of $M$, we are about to apply
-the \textit{search strategy} $S$. A search strategy is a procedure
-which takes as input a set of unfinished terms and returns a subset
-of the input set. Therefore, search strategy acts as a filter reducing 
-the search space.
-
-If the queue becomes empty before the desired number $n$ of terms
-is generated, then the initial unfinished term $\ul{\tau}{\Gamma}$ 
-is inserted to the queue, search strategy $S$
-again initializes its internal state and the process continues.
-
-\Lets now discuss three such search strategies.
-
-\subsubsection{Systematic strategy}
-
-If we use trivial strategy that returns all the inputs, 
-then the algorithm systematically generates 
-first $n$ smallest lambda terms in their
-\textit{long normal form}.
-
-\subsubsection{Ramped half-and-half strategy}
-is generalization of the standard ramped half-and-half method described 
-in \cite{koza92}. If applied to context satisfying closure requirement
-(all constants/variables are of the same type, all functions are operations
-on this type), it will behave in the same way as the standard method.
-The internal state of this strategy consists of two variables.
-It is the only one strategy described here that uses an internal state.
-
-\begin{enumerate}
- \item \textit{isFull} - A boolean value, determining whether \textit{full}
-                     or \textit{grow} method will be performed.
- \item \textit{d} - An integer value from $\setDots{2}{D_{init}}$, where 
-                $D_{init}$ is predefined maximal depth (e.g. 6).                    
-\end{enumerate}
-
-This strategy returns precisely one randomly
-(uniformly) selected  element from 
-the \textit{selection subset} of input set
-(or zero elements if the input set is empty). 
-The \textit{selection subset} 
-to select from is determined by $depth, d$ and $isFull$.
-The $depth$ parameter is the depth (in the term tree) 
-of the unfinished leaf that was expanded.
-Those elements of input set whose newly added subtree contains one ore more 
-unfinished leafs are regarded as \textit{non-terminals}, whereas 
-those whose newly added subtree contains no unfinished leaf are regarded as 
-\textit{terminals}.
-If $depth = 0$, then the subset to select from is  
-set of all \textit{non-terminals} of the input set.
-If $depth = d$, then the subset to select from is
-set of all \textit{terminals} of the input set.
-In other cases of $depth$ it depends on value of $isFull$.
-If $isFull = true$, then the subset to select from is 
-set of all \textit{non-terminals} of the input set.
-If $isFull = false$, then the subset to select from is 
-the whole input set.
-
-\subsubsection{Geometric strategy}
-
-We can see those two previous strategies as two extremes on the spectrum of 
-possible strategies. 
-\textit{Systematic strategy} filters no successor state thus performing
-exhaustive search resulting in discovery of $n$ smallest terms in one run.
-On the other hand, \textit{ramped half-and-half strategy} filters 
-all but one successor states resulting in degradation of 
-the priority queue into "fancy variable".
-\textit{Geometric strategy} is simple yet fairly effective term generating 
-strategy somewhere in the middle of this spectrum.
-It is parameterized by parameter $q \in (0,1)$, its default well-performing 
-value is $q = 0.75$.
-For each element of the input set 
-it is probabilistically decided whether
-it will be returned or omitted. A probability $p$ of returning is
-same for all elements, but depends on the $depth$, 
-which is defined in the same way as in previous strategy. 
-It is computed as follows.
-$$ p = q^{depth} $$
-This formula is motivated by idea that it is important to
-explore all possible root symbols, but as the $depth$ 
-increases it becomes less "dangerous" to omit 
-an exploration branch. 
-We can see this by considering that this strategy results in
-somehow forgetful A* search.
-With each omission we make the search space smaller. But with
-increasing depth these omissions have smaller impact on the search space,
-i.e., they cut out lesser portion of the search space.
-Another slightly esoteric argument supporting this formula is that "root 
-parts" of a program usually stand for more crucial parts
-with radical impact on global behavior of a program, 
-whereas "leaf parts" of a program usually
-stand for less important local parts (e.g. constants).  
-This strategy also plays nicely with the idea that 
-"too big trees should be killed".
-
-~\\
-Furthermore, our system utilizes generalization of the standard 
-tree-swapping crossover operator. Since it is beyond the scope of this 
-paper we mention it only briefly. Two main concerns with swapping typed subtrees 
-are types and free variables. Well-typed offspring is obtained 
-by swapping only subtrees of the same type. Only subtrees with 
-corresponding counterpart in the second parent are randomly chosen from.
-More interesting problem lies in free variables, which may cause trouble
-if swapped somewhere where it is suddenly not bounded. In order to circumvent this
-difficulty we utilize technique called \textit{abstraction elimination}\cite{jones87}
-that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions
-and no bound variables. After the initial population is generated,  it is transformed
-by abstraction elimination. Another possible transformation taking place
-after initialization is $\eta$-normalization shortening 
-rather long long normal form into $\beta\eta$-normal form.  
-Another performence enhancing transformation is use of "applicative" tree 
-representation (coming directly from inductive definition of \lterms).
-
-\section{Experiments}
-\label{experiments}
-
-We made tree experiments comparing performance of standard 
-\textit{ramped half-and-half strategy} with our
-\textit{geometric strategy} (with default parameter $q=0.75$). 
-Each experiment consisted of 50 independent runs 
-of GP algorithm. Each run had maximally 51 generations and population size
-was 500 individuals.
-In those experiments we analyze the ability of the system to produce 
-a correct solution, computational cost estimates and average size 
-of an individual throughout generations. 
-For the first two metrics we use the popular measurement  
-methods within GP field --- the \textit{performance curves}
-described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
-and $I(M,i,z)$ (the total number of individuals that must be processed 
-to yield a correct solution with probability $z =99\%$).
-
-\subsection{Simple symbolic regression}
-\textit{Simple Symbolic Regression} is a problem described
-in \cite{koza92}. Objective of this problem is to 
-find a function $f(x)$ that fits a sample
-of twenty given points. The target function is 
-function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
-Desired type of generated programs $\tau$ and 
-building blocks context $\Gamma$ are following.
-\begin{align*}
-\tau = \Real \ar &\Real\\
-\Gamma = \{
-  (+)  &: \Real \ar \Real \ar \Real    ,
-  (-)   : \Real \ar \Real \ar \Real    ,
-  (*)   : \Real \ar \Real \ar \Real    ,
-  rdiv  : \Real \ar \Real \ar \Real    ,\\
-  sin  &: \Real \ar \Real              ,
-  cos   : \Real \ar \Real              ,
-  exp   : \Real \ar \Real              , 
-  rlog  : \Real \ar \Real              \}
-\end{align*}
-where
-\noindent
-\begin{minipage}{.5\linewidth}
-\begin{align*}
-rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
-p/q & \mbox{otherwise } \end{cases}  
-\end{align*}
-\end{minipage}%
-\begin{minipage}{.5\linewidth}
-\begin{align*}
-rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
-log(\vert x\vert) & \mbox{otherwise}. \end{cases}
-\end{align*}
-\end{minipage}
-
-\begin{figure}[h!]
-  \caption{Performance curves for Simple symbolic regression.}
-  \centering
-    \includegraphics[scale=0.135]{imgs/SSR_FINAL.png}
-\end{figure}
-Fitness function is computed as follows.
-$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
-where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
-An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
-considered as a correct individual.
-
-Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
-The average individual size for generation 50 was 76.
-The experiment took 46 minutes.
-Our \textit{geometric strategy} scored 21/50 (42\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
-The average individual size for generation 50 was 47.
-The experiment took 26 minutes.
-Figure~1 shows the progress of those metrics throughout generations.
-So for the \textit{Simple symbolic regression}
-our geometric strategy is slightly more successful then 
-ramped half-and-half strategy in all observed metrics.
-
-\subsection{Artificial ant}
-\textit{Artificial Ant} is another problem described
-in \cite{koza92}. Objective of this problem is to 
-find a control program for an artificial ant so
-that it can find all food located on "Santa Fe" trail.
-The Santa Fe trail lies on toroidal square grid.
-The ant is in the upper left corner, facing right.
-The ant is able to move forward, turn left, and sense if a food 
-piece is ahead of him.
-
-\begin{align*}
-\tau = An&tAction\\
-\Gamma = \{~~
-  l    &: AntAction                              ,
-  r    : AntAction                               ,
-  m    : AntAction                               ,\\
-  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
-  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
-  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
-\end{align*}
-
-Action $l$ turns the ant left. 
-Action $r$ turns the ant right.
-Action $m$ moves the ant forward.
-Action $ifa~x~y$ (if-food-ahead) performs action $x$ 
-if a food piece is ahead of the ant,
-otherwise it performs action $y$.
-Action $p2~x~y$ first performs action $x$ and after it action $y$.
-Action $p3~x~y~z$ first performs action $x$, 
-after that action $y$ and finally $z$.
-Actions $l, r$ and $m$ each take one time step to execute.
-Ants action is performed over and over again until it reaches predefined
-maximal number of steps.
-
-\begin{figure}[h!]
-  \caption{Performance curves for Artificial ant problem.}
-  \centering
-    \includegraphics[scale=0.135]{imgs/ANT_FINAL.png}
-\end{figure}
-
-Fitness value is equal to number of eaten food pieces.
-An individual such that eats all 89 pieces of food is 
-considered as a correct solution.
-This limit is set to be 600 time steps\footnote{
-In \cite{koza92} this limit is said to be 400 time steps.
-But there is also mentioned following solution, 
-which is described as correct solution:
-\texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
-This program needs 545
-time steps; if it is given only 400 time steps, then it eats only 79 pieces
-of food. Thus we use 600 time steps. 
-}.
-
-Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
-The average individual size for generation 50 was circa 400.
-The experiment took 265 minutes.
-Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
-The average individual size for generation 50 was circa 90.
-The experiment took 107 minutes.
-This is a big improvement in all watched factors.
-Figure~2 shows it in greater detail.
-
-\subsection{Even parity problem}
-
-Previous two problems were instances of standard GP, i.e. their building symbols
-obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
-The even-parity function is a boolean function taking as inputs $N$
-boolean values and returning \textit{True} if an even number of inputs 
-are \textit{True}. For odd number it returns \textit{False}.
-This problem has been used by many researchers
-as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
-where is presented approach to evolve recursive and modular programs
-by use of higher-order functions and $\lambda$-abstractions.
-We use very similar set of \textit{building symbols} as in \cite{yu01}. 
-\begin{align*}
-\tau = [Boo&l] \ar Bool\\
-\Gamma = \{
-  and   &: Bool \ar Bool \ar Bool                              ,
-  or    : Bool \ar Bool \ar Bool                              ,\\
-  nand  &: Bool \ar Bool \ar Bool                              ,
-  nor    : Bool \ar Bool \ar Bool                              ,\\
-  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
-  head' &: [Bool] \ar Bool                                   ,
-  tail' : [Bool] \ar [Bool]                              \}
-\end{align*}
-\begin{figure}[h!]
-  \caption{Performance curves for Even parity problem.}
-  \centering
-    \includegraphics[scale=0.135]{imgs/EP_FINAL.png}
-\end{figure}
-
-The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
-this problem is considered atomic.
-Unlike in \cite{yu01}, we use specific instance of polymorphic 
-function \texttt{foldr}. 
-And modifications of functions \texttt{head} 
-(returning the first element of the list) 
-and \texttt{tail} (returning the list without the first element) are used; 
-making them total by returning default value \textit{False}
-and \texttt{[]}, respectively.
-We use the same fitness function as in \cite{yu01}. 
-The fitness function examines the individual by giving
-it all possible boolean lists of length 2 and 3.
-
-Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
-The average individual size for generation 50 was 39.
-The experiment took 28 minutes.
-Our \textit{geometric strategy} scored 32/50 (64\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
-The average individual size for generation 50 was circa 100.
-The experiment took 33 minutes.
-Figure~3 shows it in greater detail.
-Once again this is a big improvement against standard method. 
-One may get the impression that geometric strategy suffers from bloat here,
-but considering that it performs significantly better and
-that term size around 100 is similar as for the previous two problems, 
-it seems to us that ramped half-and-half maybe has difficulties with 
-constructing more complex terms, and therefore is "staying small".
-Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
-the 100\% correct solution was present in the generation 0. This is pretty good
-result for uninformed search. However, our results are less successful 
-then those presented in \cite{yu01} (with slightly different set of 
-building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
-17,500 in generation 4. But by observation of typical solution which is often
-some modification of \texttt{foldr xor False inputList} one sees that important task
-for GP here is to create \texttt{xor} function. This is advantage for the more specialized term representation used in \cite{yu01} 
-(restricting use of outer variables). 
-The difference in crossover operators may also be involved in the difference, 
-but such analysis is beyond the scope of this paper. Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
-scored 17/60 (28\%); min $I(M,i,z) = 220,000$. \textit{GP with ADFs} 
-scored 10/29 (34\%); min $I(M,i,z) = 1,440,000$.
-
-\section{Conclusions}
-\label{conclusions}
-
-In this work, we have proposed a generalization of genetic programming 
-for simply typed lambda calculus. The main focus of this paper is on individual generation algorithms. 
-The main idea is to explore the spectrum of available approaches bounded on one end by the traditional
-ramped half-and-half strategy, while the second bound represents an exhaustive systematic search. 
-Thus, three population initialization methods have been designed, 
-depending on different search strategies. 
-The first strategy corresponds to the above mentioned traditional genetic programming initialization,  
-the second one corresponds to exhaustive search, and the third one 
-represents a novel geometric strategy which outperforms standard genetic 
-programming in success rate, resource consumption and average individual size 
-on the presented experiments.
-
-%\red{ [větu dvě o experimentech]}
-
-The relevance of these algorithms does not effect only initialization, but it can be used during mutation
-operator in a straightforward way. 
-%Other performance enhancements based on theory of lambda calculus are 
-%proposed and supported by experiments, including abstraction 
-%elimination that enables us to utilize
-%a simple tree-swapping crossover.
-
-Our future research will focus on exploring other possible search strategies positioned in the middle ground 
-identified in this paper. Since the search strategy can be easily described by a rather simple filtration algorithm, it might be
-interesting to allow the typed GP system to evolve the search strategy it 
-utilizes in a kind of meta-evolution way.
-
-%\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }
-
-\nocite{*}
-
-\bibliographystyle{splncs}
-\bibliography{evogp}
+\end{thebibliography}
 
 \end{document}
-
-%
-%\begin{thebibliography}{1}
-%
-%
-%\bibitem{koza92}
-  %John R. Koza,
-  %\emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
-  %MIT Press, Cambridge, MA,
-  %1992. 
-%
-%\bibitem{koza05}
-  %Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
-  %\emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
-  %Springer, 2005. ISBN 978-0-387-26417-2 
-%
-%\bibitem{fg}
- %Riccardo Poli, William B. Langdon, Nicholas F. McPhee
- %\emph{A Field Guide to Genetic Programming}.
- %Lulu Enterprises, UK Ltd, 2008.
-%
-%\bibitem{yu01}
-  %T. Yu. 
-  %\emph{Hierachical processing for evolving recursive and modular 
-        %programs using higher order functions and lambda abstractions}. 
-  %Genetic Programming and Evolvable Machines,
-  %2(4):345–380, December 2001. ISSN 1389-2576.
-%
-%
-%\bibitem{montana95}
-%D. J. Montana. 
-%\emph{Strongly typed genetic programming.} 
-%Evolutionary Computation, 3(2): 199–230, 1995.
-%%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
-%
-%\bibitem{haynes96}
-%T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
-%\emph{Type inheritance in strongly typed genetic programming.} 
-%In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
-%in Genetic Programming 2, chapter 18, pages 359–376.
-%MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
-%%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
-%
-%\bibitem{olsson94}
-%J. R. Olsson. 
-%\emph{Inductive functional programming using incremental program 
-%transformation and Execution of logic programs by 
-%iterative-deepening A* SLD-tree search.} 
-%Dr scient thesis, University of Oslo, Norway, 1994.
-%
-%\bibitem{kes}
-%Forrest Briggs, Melissa O’Neill.
-%\emph{Functional Genetic Programming and Exhaustive
-%Program Search with Combinator Expressions.}
-%International Journal of Knowledge-based and Intelligent Engineering Systems,
-%Volume 12 Issue 1, Pages 47-68, January 2008. 
-%
-%
-%\bibitem{barendregt84}
-%H. P. Barendregt,
-%\emph{The Lambda Calculus: its Syntax and Semantics}, 
-%revised ed., North-Holland, 1984.
-%
-%\bibitem{barendregt92}
-%H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
-%\emph{Lambda Calculi with Types.} 
-%Handbook of Logic in Computer Science, 1992. 
-%
-%\bibitem{barendregt10}
-%
-  %Henk Barendregt, Wil Dekkers, Richard Statman,
-  %\emph{Lambda Calculus With Types}.
-  %Cambridge University Press,
-  %2010. 
-  %%URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
-%
-%\bibitem{jones87}
-%Simon Peyton Jones. 
-%\emph{The Implementation of Functional Programming Languages}. 
-%Prentice Hall, 1987.
-%
-%
-%\bibitem{AIAMA}
-	%Stuart J. Russell, Peter Norvig,
-	%\emph{Artificial Intelligence: A Modern Approach}.
-	%Pearson Education,
-	%2003. 
-%
-%
-%\end{thebibliography}

### 2013-12-03 kit/pid.sh
Pridan kit unixovych prikazu pro praci se servrem.

@@ -1,0 +1,2 @@
+#!/bin/bash
+ps x | grep "node test.js$" | awk '{print $1}' | cat

### 2013-12-03 kit/pull.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+echo
+echo Pulling tomkren.github.io from github...
+echo
+cd /tom
+git pull

### 2013-12-03 kit/start.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+echo
+echo Starting test.js server by script using nohup...
+echo
+cd /tom/js/node
+nohup node test.js &
+sleep 1
+PID=`ps x | grep "node test.js$" | awk '{print $1}'`
+echo $PID > pid.txt
+echo 'pid' $PID

### 2013-12-03 kit/stop.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+echo
+echo Stoping test.js server by script...
+echo
+cd /tom/js/node
+PID=`cat pid.txt`
+rm pid.txt
+kill -15 $PID

### 2013-12-03 kit/tkreset.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+echo
+echo Reseting contents of the tomkren.cz WWW folder..
+echo
+cd /tkwww
+rm -R *
+cp -R /tom/* /tkwww

### 2013-12-03 kit/bigpull.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+echo
+echo BIG PULL
+echo
+cd /kit
+./pull.sh
+./tkreset.sh

### 2013-12-04 kit/_tenkrat_poprve/porod.sh
snad zaklad porodniho skriptu

@@ -1,0 +1,13 @@
+#!/bin/bash
+echo TODO : BACHA JE TO rozdelany
+echo
+echo Rodim server ...
+echo
+ln -s /home/domeny/tomkren.cz/web/subdomeny/git/ /git-repos
+cd /git-repos
+git clone https://github.com/tomkren/tomkren/tomkren.github.io.git
+git clone https://github.com/tomkren/kutil.git
+ln -s /git-repos/tomkren.github.io/ /tom
+ln -s /home/domeny/tomkren.cz/web/subdomeny/www/ /tkwww
+ln -s /home/domeny/fishtron.net/web/subdomeny/www/ /ftwww
+echo 'Zdravicko sefe!' > /etc/motd

### 2013-12-04 kit/bigpull.sh

@@ -5,3 +5,4 @@ echo
 cd /kit
 ./pull.sh
 ./tkreset.sh
+./ftreset.sh

### 2013-12-04 kit/ftreset.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+echo
+echo Reseting contents of the fishtron.net WWW folder..
+echo
+cd /ftwww
+rm -R *
+cp -R /tom/* /ftwww
+mv fishtron.html index.html

### 2014-01-13 tex/cec14/article.tex
[tex] cec start + thesis dany do slozky

\documentclass[conference]{IEEEtran}
%\usepackage{stmaryrd}
\usepackage{amsfonts}

%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
%\usepackage{hyperref}
\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{qtree}
\usepackage{xspace}

\newcommand{\Lets}{Let us\xspace}
\newcommand{\lets}{let us\xspace}
\newcommand{\lterm}{$\lambda$-term\xspace}
\newcommand{\lterms}{$\lambda$-terms\xspace}
\newcommand{\lhead}{$\lambda$-head\xspace}
\newcommand{\lheads}{$\lambda$-heads\xspace}
\newcommand{\la}{\leftarrow\xspace}
\newcommand{\Lp}  {\Lambda^{\prime}\xspace}
\newcommand{\tur}[3]{#1\vdash{}#2:#3}
\newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
\newcommand{\GMS}{\turst{\Gamma}{M}{\sigma}}
\newcommand{\atTree}{@-tree\xspace}
\newcommand{\setDots}[2]{ \lbrace #1 , \dots , #2 \rbrace}
\newcommand{\lh}[1]{\lambda #1}
\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\SexprTree}{Sexpr-tree\xspace}
\newcommand{\then}{\Rightarrow\xspace}
\newcommand{\lamb}[2]{( \lambda \, #1 \, . \, #2 )}
\newcommand{\lam}[2]{\lambda \, #1 \, . \, #2}
\newcommand{\ST}{\mathop{\mathrm{ST}}}
\newcommand{\FV}{\mathop{\mathrm{FV}}}
\newcommand{\Scomb }{\mathbf{S}}
\newcommand{\Kcomb }{\mathbf{K}}
\newcommand{\Icomb }{\mathbf{I}}
\newcommand{\bbarr}{\twoheadrightarrow_\beta}
\newcommand{\barr}{\rightarrow_\beta}
\newcommand{\beq}{=_\beta}
\newcommand{\eearr}{\twoheadrightarrow_\eta}
\newcommand{\earr}{\rightarrow_\eta}
\newcommand{\eeq}{=_\eta}
\newcommand{\bearr}{\rightarrow_{\beta\eta}}
\newcommand{\bbeearr}{\twoheadrightarrow_{\beta\eta}}
\newcommand{\beeq}{=_{\beta\eta}}
\newcommand{\etar}{\twoheadrightarrow_\eta}
\newcommand{\ered}{$\eta$-reduction\xspace}
\newcommand{\bnf}{$\beta$-\textit{nf}\xspace}
\newcommand{\enf}{$\eta$-\textit{nf}\xspace}
\newcommand{\eenf}{$\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\beenf}{$\beta\eta^{-1}$-\textit{nf}\xspace}
\newcommand{\benf}{$\beta\eta$-\textit{nf}\xspace}
\newcommand{\bredex}{$\beta$-redex\xspace} 
\newcommand{\lnf}{\textit{lnf}\xspace}
\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
\newcommand{\Bcomb }{\mathbf{B}}   
\newcommand{\BBcomb }{\mathbf{B*}}
\newcommand{\Ccomb }{\mathbf{C}}   
\newcommand{\CCcomb }{\mathbf{C'}}
\newcommand{\SScomb }{\mathbf{S'}}
\newcommand{\ar}{\rightarrow\xspace}
\newcommand{\T}{\mathbb{T}\xspace}
\newcommand{\Real}{\mathbb{R}}

\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}

\IEEEoverridecommandlockouts    % to create the author's affliation portion

\textwidth 178mm    % <------ These are the adjustments we made 10/18/2005
\textheight 239mm   % You may or may not need to adjust these numbers again
\oddsidemargin -7mm
\evensidemargin -7mm
\topmargin -6mm
\columnsep 5mm

\begin{document}

% paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
\title{\ \\ \LARGE\bf Dodělat název (...)}

\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}

\maketitle

\begin{abstract}
Dodělat abstrakt (...)
\end{abstract}
% no key words

\section{Introduction}
\PARstart{D}{odělat} úvod (...) aaa bbb ccc aaa bbb ccc 
aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc 
aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc.

Jak to celý pojmout?

Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama. Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).

Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)

Po vygenerování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce). To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.

Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy. Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů). Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (x y . x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál. Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho. Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.

\section{Related work}
\label{related}

Vzit asi dost stejny jako při minulym članku, vyzvednout 
kombinátoři a barendrechta.

\section{Preliminaries}
\label{preliminaries}

Definice lambda kalkulu tentokrát bez nutnosti do toho extra podrobně zavádět typy.

Redukce

Eliminaci abstrakcí

---

Jak reprezentovat stromy programů pro GP?

Reprezentace v klasickym kozovy je S-expression.

Nebo můžem používat kombinátory jako Briggs a O’Neil (to si myslim je jakoby hlavní 
konkurence, vůči který by to chtělo obhájit)

Lambda termy a jejich awesomeness

Povídaní o redukcích

Povídání o lnf

že lnf je přirozený rozšíření s-exprešnů do lambda kalkulu

vlastnosti termů v lnf

proč je eta redukovat (eta redukcí lnf dostanem beta-eta-nf a že nemusíme beta redukovat 
pač se to tim nerozbyje)

Generování

Generování gramatikou

Gramatika pro lnf

Inhabitation trees jako intuitivní model takovýhodle lnf generování

Problémy s proměnýma a jak je řešit.

Křížit lambda termy i s proměnejma a abstrakcema, problémy řešit když nastanou (pomluvit 
a odsoudit)

Zmenšit prostor termů (tim že někerý nejsme schopný vygenerovat) s kterým operujeme tak 
aby křížení už nebyl problém (to dělá Yu - v těle lambda fce dovoluje jen použití 
proměnných z její hlavy)

Převod hned po vygenerování

Převod až při křížení

eliminace abstrakcí

skřížim

vložený kombinátory nahradim odpovídajícím termem

celý to redukuju

\section{Our approach}
\label{approach}

\subsection{Introduction}

\subsection{Algorithm}

\section{Experiments}
\label{experiments}

\section{Conclusions}

Dodělat závěr (...)

\begin{thebibliography}{22}

\bibitem{koza92}
  TODO. Todo,
  \emph{Dát sem místo toho ten bibtex}.
  MIT Press, Cambridge, MA,
  2014.

\bibitem{koza92}
  John R. Koza,
  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
  MIT Press, Cambridge, MA,
  1992.

\bibitem{koza05}
  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
  Springer, 2005. ISBN 978-0-387-26417-2

\bibitem{fg}
 Riccardo Poli, William B. Langdon, Nicholas F. McPhee
 \emph{A Field Guide to Genetic Programming}.
 Lulu Enterprises, UK Ltd, 2008.

\bibitem{yu01}
  T. Yu. 
  \emph{Hierachical processing for evolving recursive and modular 
        programs using higher order functions and lambda abstractions}. 
  Genetic Programming and Evolvable Machines,
  2(4):345–380, December 2001. ISSN 1389-2576.

\bibitem{montana95}
D. J. Montana. 
\emph{Strongly typed genetic programming.} 
Evolutionary Computation, 3(2): 199–230, 1995.
%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha

\bibitem{haynes96}
T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
\emph{Type inheritance in strongly typed genetic programming.} 
In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
in Genetic Programming 2, chapter 18, pages 359–376.
MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.

\bibitem{olsson94}
J. R. Olsson. 
\emph{Inductive functional programming using incremental program 
transformation and Execution of logic programs by 
iterative-deepening A* SLD-tree search.} 
Dr scient thesis, University of Oslo, Norway, 1994.

\bibitem{kes}
Forrest Briggs, Melissa O’Neill.
\emph{Functional Genetic Programming and Exhaustive
Program Search with Combinator Expressions.}
International Journal of Knowledge-based and Intelligent Engineering Systems,
Volume 12 Issue 1, Pages 47-68, January 2008.

\bibitem{barendregt84}
H. P. Barendregt,
\emph{The Lambda Calculus: its Syntax and Semantics}, 
revised ed., North-Holland, 1984.

\bibitem{barendregt92}
H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
\emph{Lambda Calculi with Types.} 
Handbook of Logic in Computer Science, 1992.

\bibitem{barendregt10}

  Henk Barendregt, Wil Dekkers, Richard Statman,
  \emph{Lambda Calculus With Types}.
  Cambridge University Press,
  2010. 
  %URL \url{http://www.cs.ru.nl/~henk/book.pdf}.

\bibitem{jones87}
Simon Peyton Jones. 
\emph{The Implementation of Functional Programming Languages}. 
Prentice Hall, 1987.

\bibitem{AIAMA}
	Stuart J. Russell, Peter Norvig,
	\emph{Artificial Intelligence: A Modern Approach}.
	Pearson Education,
	2003.

\end{thebibliography}

% that's all folks
\end{document}

### 2014-01-15 tex/cec14/article.tex
[tex] minor

@@
 \usepackage{xspace}
+
+\usepackage[ampersand]{easylist}
+
+\newenvironment{lizt}
+{\begin{easylist}[itemize]}
+{\end{easylist}}
 
@@
 
-Jak reprezentovat stromy programů pro GP?
-
-Reprezentace v klasickym kozovy je S-expression.
-
-Nebo můžem používat kombinátory jako Briggs a O’Neil (to si myslim je jakoby hlavní 
-konkurence, vůči který by to chtělo obhájit)
-
-Lambda termy a jejich awesomeness
-
-Povídaní o redukcích
-
-Povídání o lnf
+\begin{lizt}
+& Jak reprezentovat stromy programu pro GP?
+  && Reprezentace v klasickym kozovy je S-expression.
+  && Nebo mužem používat kombinátory jako Briggs a O’Neil (to si 
+     myslim je jakoby hlavní konkurence, vuči který by to chtělo 
+     obhájit)
+& Lambda termy a jejich awesomeness
+& Povídaní o redukcích
+& Povídání o lnf
+\end{lizt}
 

### 2014-01-16 tex/cec14/article.tex
[tex]

@@
 % paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
-\title{\ \\ \LARGE\bf Dodělat název (...)}
+\title{\ \\ \LARGE\bf 
+Utilization of Reductions and Abstraction Elimination in Typed Genetic Programming}
 
@@
 \begin{abstract}
-Dodělat abstrakt (...)
+Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions.
+In this paper we discuss advantages of this representation coming from use of reductions (beta and eta) and how to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm.
+We discuss the role of those reductions in the process of generating initial population and compare several crossover approaches including novel approach to crossover operator based both on reductions and abstraction elimination.
+The design goal of this operator is to turn the disadvantage of abstraction elimination  - possibly quadratic increase of program size - into a virtue; our approach leads to more crossover points.
+At the same time, utilization of reductions provides offspring of small sizes.
 \end{abstract}

### 2014-01-17 tex/cec14/article.tex
[tex]

--- 2014-01-15 tex/cec14/article.tex
@@
 \documentclass[conference]{IEEEtran}
+
 %\usepackage{stmaryrd}
+%\usepackage{amsfonts}
+%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
+%\usepackage{hyperref}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%\usepackage[utf8]{inputenc}
+%\usepackage{qtree}
+%\usepackage{xspace}
+%\usepackage[ampersand]{easylist}
+
+%\usepackage{times}
+\usepackage[utf8]{inputenc}
 \usepackage{amsfonts}
+%\usepackage{qtree}
+\usepackage{xspace}
 
+%\usepackage{makeidx}  % allows for indexgeneration
 %\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
@@
 %\usepackage{amssymb}
-\usepackage[utf8]{inputenc}
-\usepackage{qtree}
-\usepackage{xspace}
+\usepackage{graphicx}
 
 \usepackage[ampersand]{easylist}
+\usepackage[usenames,dvipsnames]{color}
+
+%\usepackage{amsthm}
+\newtheorem{definition}{Definition}
 
@@
 \newcommand{\Lp}  {\Lambda^{\prime}\xspace}
-\newcommand{\tur}[3]{#1\vdash{}#2:#3}
+\newcommand{\tur}[3]{#1\vdash{}#2 \colon #3}
 \newcommand{\turst}[3]{$#1\vdash{}#2:#3$\xspace}
@@
 \newcommand{\T}{\mathbb{T}\xspace}
+\newcommand{\C}{\mathbb{C}\xspace}
 \newcommand{\Real}{\mathbb{R}}
+
+\newenvironment{todo}
+{~\\ {\color{red}\textbf{TODO}}
+  \begin{easylist}[itemize]}
+{ \end{easylist}}
+
+\newcommand{\Lpr}{\Lambda^\prime}
+\newcommand{\ul}[2]{\langle #1 ; #2 \rangle}
+\newcommand{\ro}[1]{{\color{blue} #1}}
+\newcommand{\tom}[1]{{\color{ForestGreen} #1}}
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 % paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
-\title{\ \\ \LARGE\bf Dodělat název (...)}
+\title{\ \\ \LARGE\bf 
+Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
 
@@
 \begin{abstract}
-Dodělat abstrakt (...)
+In this paper, a generalization of genetic programming 
+for simply typed lambda calculus is presented. We propose three population 
+initialization methods depending on different search strategies. 
+The first strategy corresponds to standard ramped half-and-half initialization, 
+the second one corresponds to exhaustive systematic search, and the third one 
+represents a novel geometric strategy. Three well-known benchmark experiments
+support that the geometric strategy outperforms
+the standard generating method in success rate, time consumption and average individual size. 
 \end{abstract}
@@
 \section{Introduction}
-\PARstart{D}{odělat} úvod (...) aaa bbb ccc aaa bbb ccc 
-aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc 
-aaa bbb ccc aaa bbb ccc aaa bbb ccc aaa bbb ccc.
+\PARstart{G}{enetic} programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
+Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
+Use of types naturally opens door to enriching S-expressions,
+the traditional GP representation of individuals, with concepts from
+lambda calculus, which is simple yet powerful functional mathematical and programming 
+language extensively used in type theory. Such attempts has shown to be 
+successful \cite{yu01}.
 
-Jak to celý pojmout?
+The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
+During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
+Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
+Thus, the natural idea is to employ an exhaustive systematic search.
+On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.
+These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
+This design goal also differentiate our approach from
+the three state of the art proposals for typed GP known to us that are discussed in the following section. 
+Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
+Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
 
-Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
-Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).
-
-Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)
-
-Po vygenerování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce).
-To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.
-
-Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
-Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
-Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (x y .
-x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
-Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
-Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.
-Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
+The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
+Main original results about search strategies in individual generating are described in section~\ref{approach}.
+Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.
 
@@
 
-Vzit asi dost stejny jako při minulym članku, vyzvednout 
-kombinátoři a barendrechta.
+In \cite{yu01} Yu presents a GP system utilizing
+polymorphic higher-order functions\footnote{Higher-order 
+function is a function taking another function as 
+input parameter.} and lambda abstractions.
+Important point of interest in this work is use of
+\texttt{foldr} function as a tool for \textit{implicit recursion},
+i.e. recursion without explicit recursive calls. 
+The terminal set for constructing lambda abstraction subtrees 
+is limited to use only constants and variables of that particular
+lambda abstraction, i.e., outer variables are not allowed to be used
+as terminals in this work. This is significant difference from our approach 
+since we permit all well-typed normalized \lterms. From this difference also
+comes different crossover operation. We focus more on term generating process; 
+their term generation is performed in a similar way as the standard one, 
+whereas our term generation also tries to utilize techniques of systematic enumeration.
+
+In \cite{kes} Briggs and O’Neill present technique 
+utilizing typed GP with combinators.
+The difference between approach presented in this work
+and our approach is that in this work terms are generated
+straight from \textit{library} of combinators and no lambda abstractions
+are used. They are using more general polymorphic type system then us
+-- the Hindley–Milner type system. They also discuss the 
+properties of exhaustive enumeration of terms and compare it with GP search.  
+They also present interesting concept of \textit{Generalized
+genetic operator} based on term generation.
+
+In \cite{binard2008genetic} by Binard and Felty even 
+stronger type system (\textit{System F}) is used.  
+But with increasing power of the type system comes increasing difficulty of term generation.
+For this reason evolution in this work takes interesting and nonstandard shape 
+(fitness is associated with \textit{genes} which are evolved together with \textit{species}
+which together participate in creation of individuals).
+This differs from our approach, which tries to be generalization of
+the standard GP\cite{koza92}.
+
+In contrast with above mentioned works our approach uses very simple type system 
+(simply typed lambda calculus) and concentrates on process of generation  
+able to generate all possible well-typed normalized lambda terms. In order to do
+so we use technique based on \textit{inhabitation machines} 
+described by Barendregt in~\cite{barendregt10}.
 
@@
 
-Definice lambda kalkulu tentokrát bez nutnosti do toho extra podrobně zavádět typy.
+In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
+First, \lets 
+describe a programming language, 
+in which the GP algorithm generates individual programs --- the so called \lterms.
 
-Redukce
+\begin{definition}
+Let $V$ be infinite countable set of {\it 
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	 	
+Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
+\begin{align*}
+x   \in V \cup C  &\then x     \in \Lambda \\
+M,N \in \Lambda   &\then (M~N) \in \Lambda 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
+\textit{~~~~($\lambda$-abstraction)} 
+\end{align*}
+\end{definition}
 
-Eliminaci abstrakcí
+\textit{Function application} and 
+\textit{$\lambda$-abstraction} are concepts
+well known from common programming languages. 
+For example in JavaScript 
+$(M~N)$ translates to expression \texttt{$M$($N$)} and
+$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
+In other words, the function application 
+corresponds to the act of supplying a function 
+with an argument, and
+the $\lambda$-abstraction is equivalent to 
+\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
+they were recently introduced to C++, and they are expected to be supported in Java 8.}.
+$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for $(\dots((M_1~M_2)~M_3)~\dots~M_n)$
+and $\lam{x_1 x_2 \dots x_n }{M}$ for $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
 
----
+A \lterm as described above
+corresponds to a program expression with no type information
+included. Now we will describe \textit{types} (or \textit{type terms}).
 
-\begin{lizt}
-& Jak reprezentovat stromy programu pro GP?
-  && Reprezentace v klasickym kozovy je S-expression.
-  && Nebo mužem používat kombinátory jako Briggs a O’Neil (to si 
-     myslim je jakoby hlavní konkurence, vuči který by to chtělo 
-     obhájit)
-& Lambda termy a jejich awesomeness
-& Povídaní o redukcích
-& Povídání o lnf
-\end{lizt}
+\begin{definition}
+Let $A$ be set of {\it atomic type names}. 
+Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
+\begin{align*}
+\alpha      \in A  &\then   \alpha \in \T \\
+\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
+\end{align*}
+\end{definition}
 
-že lnf je přirozený rozšíření s-exprešnů do lambda kalkulu
+Type $\sigma \ar \tau$ is type for functions taking as input
+something of a type $\sigma$ and returning 
+as output something of a type $\tau$. 
+$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
+$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
+The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
+combining the previously defined \textit{\lterms} and \textit{types} together.
 
-vlastnosti termů v lnf
+\begin{definition}\begin{enumerate}
+ \item 	Let $\Lambda$ be set of {\it \lterms}. 
+	Let $\mathbb{T}$ be set of {\it types}.       
+	A {\it statement} $M : \sigma$ is a pair 
+	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
+	Statement $M : \sigma$ is vocalized as 
+	{\it "$M$ has type $\sigma$"}.
+	The term $M$ is called the {\it subject} of the 
+	statement $M : \sigma$.
+ \item A \textit{declaration} is a statement 
+ $x : \sigma$ where $x \in V \cup C$.
 
-proč je eta redukovat (eta redukcí lnf dostanem beta-eta-nf a že nemusíme beta redukovat 
-pač se to tim nerozbyje)
+ \item A \textit{context} 
+ is set of declarations with distinct variables as subjects.
+\end{enumerate}
+\end{definition}
 
-Generování
+Context is a basic type theoretic concept suitable as a typed alternative
+for terminal and function set in standard GP. 
+Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
+such that $\Gamma$ does not contain any declaration with $x$ as subject.
+We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
 
-Generování gramatikou
+\begin{definition}
+A statement $M\colon\sigma$ is \textit{derivable from}
+a context $\Gamma$ (notation 
+\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
+if it can be produced by the following rules.
+\begin{align*}
+x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
+\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
+&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
+\tur{\Gamma,x:\sigma}{M}{\tau}
+&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
+\end{align*}
+\end{definition}
 
-Gramatika pro lnf
+Our goal in term generation is to produce terms $M$
+for a given pair $\ul{\tau}{\Gamma}$
+such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
 
-Inhabitation trees jako intuitivní model takovýhodle lnf generování
+\begin{definition}
+Let $V$ be infinite countable set of {\it 
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	
+Let $\T$ be set of types.
+Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
+Then $\Lpr$ is set of 
+\textit{unfinished  \lterms} defined inductively as follows.	
+\begin{align*}
+\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
+\textit{~~~~~~~~(Unfinished leaf)}\\
+x   \in V \cup C  &\then x     \in \Lpr \\
+M,N \in \Lpr   &\then (M~N) \in \Lpr 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
+\textit{~~~~($\lambda$-abstraction)} 
+\end{align*}
+\end{definition}
 
-Problémy s proměnýma a jak je řešit.
-
-Křížit lambda termy i s proměnejma a abstrakcema, problémy řešit když nastanou (pomluvit 
-a odsoudit)
-
-Zmenšit prostor termů (tim že někerý nejsme schopný vygenerovat) s kterým operujeme tak 
-aby křížení už nebyl problém (to dělá Yu - v těle lambda fce dovoluje jen použití 
-proměnných z její hlavy)
-
-Převod hned po vygenerování
-
-Převod až při křížení
-
-eliminace abstrakcí
-
-skřížim
-
-vložený kombinátory nahradim odpovídajícím termem
-
-celý to redukuju
+\textit{Unfinished leaf} $\ul{\tau}{\Gamma}$ 
+stands for yet not specified \lterm of the type $\tau$ 
+build from symbols of $\Gamma$.
 
@@
 
+Our approach to \lterm gerating is based on technique 
+briefly described in \cite{barendregt10}, which generates
+well-typed \lterms in their long normal form. 
+We use this technique to perform systematic exhaustive enumeration
+of \lterms in their long normal form in order from smallest to largest.
+We use well known \textit{A* algorithm} \cite{AIMA} for this task.
+A* is used to search in a given state space for a goal state. 
+It finds the optimal solution (in our case the smallest term)
+and uses "advising" heuristic function.
+It maintains a priority queue to organize states yet to be explored.
+Initially this queue contains only the initial state.
+
+Our state space to search in is the space of unfinished \lterms. 
+The initial state is the unfinished term $\ul{\tau}{\Gamma}$, 
+where $\tau$ is the desired type of
+terms to be generated and $\Gamma$ is the context
+representing the set of building symbols to be used in construction of
+terms (it corresponds to the set $T \cup F$ in
+standard GP enriched with types). The process of determining 
+successors of a state described below is designed so it constructs well-typed 
+\lterms and omits no \lterm in its long normal form. 
+A state is considered a goal state if it contains no unfinished
+leaf, i.e., it is a finished \lterm.
+
+Our generating method is based on simple modification of the
+standard A*, which we call \textit{forgetful A*}. This modification consist in 
+additional parameter for the A* algorithm -- the \textit{search strategy}. 
+It is a simple filtration function (along with initialization procedure)
+that is given the set of all successors of the state that is being examined
+and returns a subset of this input. This subset is added to the priority queue 
+to be further explored. In this way the search space may be reduced as 
+the filtration function may \textit{forget} some successors.
+If the queue becomes empty before the desired number of \lterms
+is generated, then the initial statete is inserted to the queue
+and the process continues. For the standard A* this would be meaningless,
+but since our A* is forgetful this kind of restart makes sense.
+
+A* keeps a priority queue of states during the generation process,
+on the other hand the \textit{ramped half-and-half method}, 
+the standard GP algorithm for generating individuals, 
+keeps only one individual which is gradually constructed. This 
+behavior is easily achieved by use of suitable search strategy 
+that returns subset consisting of only one successor.
+The systematic search is obtained by search strategy that 
+returns whole input set.      
+Our novel \textit{geometric strategy} can be understood as
+point somewhere between those two extremes.
+
 \subsection{Algorithm}
+
+The inputs for the term generating algorithm are following.
+\begin{enumerate}
+ \item Desired type $\tau$ of generated terms.
+ \item Context $\Gamma$ representing set of building symbols.
+ \item Number $n$ of terms to be generated.
+ \item Search strategy $S$. 
+\end{enumerate}
+
+Essential data structure of our algorithm 
+is priority queue of unfinished terms. 
+Priority of an unfinished term is given by its size\footnote{
+A* heuristic function is hidden in method of computing
+size of unfinished leafs $\ul{\tau}{\Gamma}$. Our algorithm uses
+trivial estimate $\vert\ul{\tau}{\Gamma}\vert = 1$ which is trivially admissible.
+This heuristic is not as silly as it might seem since it is
+quite usual to have $x$ such that $x : \tau \in \Gamma$.
+Since the true value of $\vert\ul{\tau}{\Gamma}\vert$ depends only on
+$\tau$ and \textit{types} in $\Gamma$ (the \textit{signature}), 
+no matter how many variables/constants of
+each type there are, it is should by pretty effective to compute this
+value precisely and store them for later. 
+%This topic has other interesting 
+%subtopics (such as surprisingly small 
+%number of such "signatures" for a given problem), 
+%but it is unfortunately beyond the scope of this paper.
+}.
+At the beginning, the queue contains only one unfinished term; 
+$\ul{\tau}{\Gamma}$. The search strategy $S$ also 
+initializes its internal state (if it has one).
+
+At each step, the term $M$ with the smallest size
+is pulled from the queue.
+According to the number of unfinished leafs in $M$ one of
+the following actions is performed.
+\begin{enumerate}
+ \item If the term $M$ has no unfinished leaf (i.e., it is a finished
+ term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
+ result set
+% \footnote{
+% \red{TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
+% některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
+% , případně to vůbec nezminovbat když by to kradlo moc místa
+% }} 
+ of generated terms.   
+ \item Otherwise, \textit{successors} of the unfinished term $M$ are
+       filtered out by \textit{search strategy} $S$ and
+       those successors that outlast the filtration 
+       are inserted into the queue.
+\end{enumerate}
+
+\textit{Successors} of an unfinished term $M$ are obtained by 
+\textit{expansion} of the \mbox{\textit{DFS-first}} unfinished leaf $L$
+(i.e., the leftmost unfinished leaf of $M$).
+
+Expansion of the selected unfinished leaf $L$ leads to creation of 
+one or many (possibly zero) successors.
+In this process, $L$ is replaced
+by a new subterm defined by the following rules\footnote{
+For the sake of simplicity it is presented as two separate rules. 
+Since the first rule results in exactly one successor it is smarter
+to combine those two rules into one resulting in that unfinished leafs
+have only atomic types. At the beginning we transform the initial type into atomic by first rule (if necessary). After that
+only second rule is applied, but if it results in creation of some unatomic 
+$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }.
+\begin{enumerate}
+ \item If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
+ 	   where $\alpha$ is atomic type and $n \geq 1$, 
+       then $L$ is replaced by 
+       $\lamb{x_1 \dots x_n}{\ul{\alpha}
+       {\Gamma,x_1 \colon \rho_1,\dots,x_n \colon \rho_n}}$.
+       Thus this expansion results in exactly one successor.  
+ \item If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type,
+       then for each 
+       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}
+       the unfinished leaf $L$ is replaced by 
+       $(~f~\ul{\tau_1}{\Gamma}~\dots~\ul{\tau_m}{\Gamma}~)$.
+       Thus this expansion results in many (possibly zero or one) successors.
+\end{enumerate}
+
+Now that we have all possible successors of $M$, we are about to apply
+the \textit{search strategy} $S$. A search strategy is a procedure
+which takes as input a set of unfinished terms and returns a subset
+of the input set. Therefore, search strategy acts as a filter reducing 
+the search space.
+
+If the queue becomes empty before the desired number $n$ of terms
+is generated, then the initial unfinished term $\ul{\tau}{\Gamma}$ 
+is inserted to the queue, search strategy $S$
+again initializes its internal state and the process continues.
+
+\Lets now discuss three such search strategies.
+
+\subsubsection{Systematic strategy}
+
+If we use trivial strategy that returns all the inputs, 
+then the algorithm systematically generates 
+first $n$ smallest lambda terms in their
+\textit{long normal form}.
+
+\subsubsection{Ramped half-and-half strategy}
+is generalization of the standard ramped half-and-half method described 
+in \cite{koza92}. If applied to context satisfying closure requirement
+(all constants/variables are of the same type, all functions are operations
+on this type), it will behave in the same way as the standard method.
+The internal state of this strategy consists of two variables.
+It is the only one strategy described here that uses an internal state.
+
+\begin{enumerate}
+ \item \textit{isFull} - A boolean value, determining whether \textit{full}
+                     or \textit{grow} method will be performed.
+ \item \textit{d} - An integer value from $\setDots{2}{D_{init}}$, where 
+                $D_{init}$ is predefined maximal depth (e.g. 6).                    
+\end{enumerate}
+
+This strategy returns precisely one randomly
+(uniformly) selected  element from 
+the \textit{selection subset} of input set
+(or zero elements if the input set is empty). 
+The \textit{selection subset} 
+to select from is determined by $depth, d$ and $isFull$.
+The $depth$ parameter is the depth (in the term tree) 
+of the unfinished leaf that was expanded.
+Those elements of input set whose newly added subtree contains one ore more 
+unfinished leafs are regarded as \textit{non-terminals}, whereas 
+those whose newly added subtree contains no unfinished leaf are regarded as 
+\textit{terminals}.
+If $depth = 0$, then the subset to select from is  
+set of all \textit{non-terminals} of the input set.
+If $depth = d$, then the subset to select from is
+set of all \textit{terminals} of the input set.
+In other cases of $depth$ it depends on value of $isFull$.
+If $isFull = true$, then the subset to select from is 
+set of all \textit{non-terminals} of the input set.
+If $isFull = false$, then the subset to select from is 
+the whole input set.
+
+\subsubsection{Geometric strategy}
+
+We can see those two previous strategies as two extremes on the spectrum of 
+possible strategies. 
+\textit{Systematic strategy} filters no successor state thus performing
+exhaustive search resulting in discovery of $n$ smallest terms in one run.
+On the other hand, \textit{ramped half-and-half strategy} filters 
+all but one successor states resulting in degradation of 
+the priority queue into "fancy variable".
+\textit{Geometric strategy} is simple yet fairly effective term generating 
+strategy somewhere in the middle of this spectrum.
+It is parameterized by parameter $q \in (0,1)$, its default well-performing 
+value is $q = 0.75$.
+For each element of the input set 
+it is probabilistically decided whether
+it will be returned or omitted. A probability $p$ of returning is
+same for all elements, but depends on the $depth$, 
+which is defined in the same way as in previous strategy. 
+It is computed as follows.
+$$ p = q^{depth} $$
+This formula is motivated by idea that it is important to
+explore all possible root symbols, but as the $depth$ 
+increases it becomes less "dangerous" to omit 
+an exploration branch. 
+We can see this by considering that this strategy results in
+somehow forgetful A* search.
+With each omission we make the search space smaller. But with
+increasing depth these omissions have smaller impact on the search space,
+i.e., they cut out lesser portion of the search space.
+Another slightly esoteric argument supporting this formula is that "root 
+parts" of a program usually stand for more crucial parts
+with radical impact on global behavior of a program, 
+whereas "leaf parts" of a program usually
+stand for less important local parts (e.g. constants).  
+This strategy also plays nicely with the idea that 
+"too big trees should be killed".
+
+~\\
+Furthermore, our system utilizes generalization of the standard 
+tree-swapping crossover operator. Since it is beyond the scope of this 
+paper we mention it only briefly. Two main concerns with swapping typed subtrees 
+are types and free variables. Well-typed offspring is obtained 
+by swapping only subtrees of the same type. Only subtrees with 
+corresponding counterpart in the second parent are randomly chosen from.
+More interesting problem lies in free variables, which may cause trouble
+if swapped somewhere where it is suddenly not bounded. In order to circumvent this
+difficulty we utilize technique called \textit{abstraction elimination}\cite{jones87}
+that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables. After the initial population is generated,  it is transformed
+by abstraction elimination. Another possible transformation taking place
+after initialization is $\eta$-normalization shortening 
+rather long long normal form into $\beta\eta$-normal form.  
+Another performence enhancing transformation is use of "applicative" tree 
+representation (coming directly from inductive definition of \lterms).
 

### 2014-01-17 tex/cec14/article.tex
[tex]

@@
 
+We made tree experiments comparing performance of standard 
+\textit{ramped half-and-half strategy} with our
+\textit{geometric strategy} (with default parameter $q=0.75$). 
+Each experiment consisted of 50 independent runs 
+of GP algorithm. Each run had maximally 51 generations and population size
+was 500 individuals.
+In those experiments we analyze the ability of the system to produce 
+a correct solution, computational cost estimates and average size 
+of an individual throughout generations. 
+For the first two metrics we use the popular measurement  
+methods within GP field --- the \textit{performance curves}
+described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
+and $I(M,i,z)$ (the total number of individuals that must be processed 
+to yield a correct solution with probability $z =99\%$).
+
+\subsection{Simple symbolic regression}
+\textit{Simple Symbolic Regression} is a problem described
+in \cite{koza92}. Objective of this problem is to 
+find a function $f(x)$ that fits a sample
+of twenty given points. The target function is 
+function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
+Desired type of generated programs $\tau$ and 
+building blocks context $\Gamma$ are following.
+\begin{align*}
+\tau = \Real \ar &\Real\\
+\Gamma = \{
+  (+)  &: \Real \ar \Real \ar \Real    ,
+  (-)   : \Real \ar \Real \ar \Real    ,
+  (*)   : \Real \ar \Real \ar \Real    ,
+  rdiv  : \Real \ar \Real \ar \Real    ,\\
+  sin  &: \Real \ar \Real              ,
+  cos   : \Real \ar \Real              ,
+  exp   : \Real \ar \Real              , 
+  rlog  : \Real \ar \Real              \}
+\end{align*}
+where
+\noindent
+\begin{minipage}{.5\linewidth}
+\begin{align*}
+rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
+p/q & \mbox{otherwise } \end{cases}  
+\end{align*}
+\end{minipage}%
+\begin{minipage}{.5\linewidth}
+\begin{align*}
+rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
+log(\vert x\vert) & \mbox{otherwise}. \end{cases}
+\end{align*}
+\end{minipage}
+
+\begin{figure}[h!]
+  \caption{Performance curves for Simple symbolic regression.}
+  \centering
+    \includegraphics[scale=0.135]{imgs/SSR_FINAL.png}
+\end{figure}
+Fitness function is computed as follows.
+$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
+where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
+An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
+considered as a correct individual.
+
+Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
+The average individual size for generation 50 was 76.
+The experiment took 46 minutes.
+Our \textit{geometric strategy} scored 21/50 (42\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
+The average individual size for generation 50 was 47.
+The experiment took 26 minutes.
+Figure~1 shows the progress of those metrics throughout generations.
+So for the \textit{Simple symbolic regression}
+our geometric strategy is slightly more successful then 
+ramped half-and-half strategy in all observed metrics.
+
+\subsection{Artificial ant}
+\textit{Artificial Ant} is another problem described
+in \cite{koza92}. Objective of this problem is to 
+find a control program for an artificial ant so
+that it can find all food located on "Santa Fe" trail.
+The Santa Fe trail lies on toroidal square grid.
+The ant is in the upper left corner, facing right.
+The ant is able to move forward, turn left, and sense if a food 
+piece is ahead of him.
+
+\begin{align*}
+\tau = An&tAction\\
+\Gamma = \{~~
+  l    &: AntAction                              ,
+  r    : AntAction                               ,
+  m    : AntAction                               ,\\
+  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
+  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
+  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
+\end{align*}
+
+Action $l$ turns the ant left. 
+Action $r$ turns the ant right.
+Action $m$ moves the ant forward.
+Action $ifa~x~y$ (if-food-ahead) performs action $x$ 
+if a food piece is ahead of the ant,
+otherwise it performs action $y$.
+Action $p2~x~y$ first performs action $x$ and after it action $y$.
+Action $p3~x~y~z$ first performs action $x$, 
+after that action $y$ and finally $z$.
+Actions $l, r$ and $m$ each take one time step to execute.
+Ants action is performed over and over again until it reaches predefined
+maximal number of steps.
+
+\begin{figure}[h!]
+  \caption{Performance curves for Artificial ant problem.}
+  \centering
+    \includegraphics[scale=0.135]{imgs/ANT_FINAL.png}
+\end{figure}
+
+Fitness value is equal to number of eaten food pieces.
+An individual such that eats all 89 pieces of food is 
+considered as a correct solution.
+This limit is set to be 600 time steps\footnote{
+In \cite{koza92} this limit is said to be 400 time steps.
+But there is also mentioned following solution, 
+which is described as correct solution:
+\texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
+This program needs 545
+time steps; if it is given only 400 time steps, then it eats only 79 pieces
+of food. Thus we use 600 time steps. 
+}.
+
+Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
+The average individual size for generation 50 was circa 400.
+The experiment took 265 minutes.
+Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
+The average individual size for generation 50 was circa 90.
+The experiment took 107 minutes.
+This is a big improvement in all watched factors.
+Figure~2 shows it in greater detail.
+
+\subsection{Even parity problem}
+
+Previous two problems were instances of standard GP, i.e. their building symbols
+obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
+The even-parity function is a boolean function taking as inputs $N$
+boolean values and returning \textit{True} if an even number of inputs 
+are \textit{True}. For odd number it returns \textit{False}.
+This problem has been used by many researchers
+as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
+where is presented approach to evolve recursive and modular programs
+by use of higher-order functions and $\lambda$-abstractions.
+We use very similar set of \textit{building symbols} as in \cite{yu01}. 
+\begin{align*}
+\tau = [Boo&l] \ar Bool\\
+\Gamma = \{
+  and   &: Bool \ar Bool \ar Bool                              ,
+  or    : Bool \ar Bool \ar Bool                              ,\\
+  nand  &: Bool \ar Bool \ar Bool                              ,
+  nor    : Bool \ar Bool \ar Bool                              ,\\
+  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
+  head' &: [Bool] \ar Bool                                   ,
+  tail' : [Bool] \ar [Bool]                              \}
+\end{align*}
+\begin{figure}[h!]
+  \caption{Performance curves for Even parity problem.}
+  \centering
+    \includegraphics[scale=0.135]{imgs/EP_FINAL.png}
+\end{figure}
+
+The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
+this problem is considered atomic.
+Unlike in \cite{yu01}, we use specific instance of polymorphic 
+function \texttt{foldr}. 
+And modifications of functions \texttt{head} 
+(returning the first element of the list) 
+and \texttt{tail} (returning the list without the first element) are used; 
+making them total by returning default value \textit{False}
+and \texttt{[]}, respectively.
+We use the same fitness function as in \cite{yu01}. 
+The fitness function examines the individual by giving
+it all possible boolean lists of length 2 and 3.
+
+Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
+The average individual size for generation 50 was 39.
+The experiment took 28 minutes.
+Our \textit{geometric strategy} scored 32/50 (64\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
+The average individual size for generation 50 was circa 100.
+The experiment took 33 minutes.
+Figure~3 shows it in greater detail.
+Once again this is a big improvement against standard method. 
+One may get the impression that geometric strategy suffers from bloat here,
+but considering that it performs significantly better and
+that term size around 100 is similar as for the previous two problems, 
+it seems to us that ramped half-and-half maybe has difficulties with 
+constructing more complex terms, and therefore is "staying small".
+Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
+the 100\% correct solution was present in the generation 0. This is pretty good
+result for uninformed search. However, our results are less successful 
+then those presented in \cite{yu01} (with slightly different set of 
+building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
+17,500 in generation 4. But by observation of typical solution which is often
+some modification of \texttt{foldr xor False inputList} one sees that important task
+for GP here is to create \texttt{xor} function. This is advantage for the more specialized term representation used in \cite{yu01} 
+(restricting use of outer variables). 
+The difference in crossover operators may also be involved in the difference, 
+but such analysis is beyond the scope of this paper. Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
+scored 17/60 (28\%); min $I(M,i,z) = 220,000$. \textit{GP with ADFs} 
+scored 10/29 (34\%); min $I(M,i,z) = 1,440,000$.
+
 \section{Conclusions}
+\label{conclusions}
 
-Dodělat závěr (...)
+In this work, we have proposed a generalization of genetic programming 
+for simply typed lambda calculus. The main focus of this paper is on individual generation algorithms. 
+The main idea is to explore the spectrum of available approaches bounded on one end by the traditional
+ramped half-and-half strategy, while the second bound represents an exhaustive systematic search. 
+Thus, three population initialization methods have been designed, 
+depending on different search strategies. 
+The first strategy corresponds to the above mentioned traditional genetic programming initialization,  
+the second one corresponds to exhaustive search, and the third one 
+represents a novel geometric strategy which outperforms standard genetic 
+programming in success rate, resource consumption and average individual size 
+on the presented experiments.
+
+%\red{ [větu dvě o experimentech]}
+
+The relevance of these algorithms does not effect only initialization, but it can be used during mutation
+operator in a straightforward way. 
+%Other performance enhancements based on theory of lambda calculus are 
+%proposed and supported by experiments, including abstraction 
+%elimination that enables us to utilize
+%a simple tree-swapping crossover.
+
+Our future research will focus on exploring other possible search strategies positioned in the middle ground 
+identified in this paper. Since the search strategy can be easily described by a rather simple filtration algorithm, it might be
+interesting to allow the typed GP system to evolve the search strategy it 
+utilizes in a kind of meta-evolution way.
+
+%\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }
 

### 2014-01-17 tex/cec14/article.tex
[tex]

@@
 In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
-First, \lets 
-describe a programming language, 
-in which the GP algorithm generates individual programs --- the so called \lterms.
+First, \lets describe a programming language, 
+in which the GP algorithm generates individual programs --- the so called \lterms.\\
 
@@
 \end{definition}
+~\\
 
@@
 corresponds to a program expression with no type information
-included. Now we will describe \textit{types} (or \textit{type terms}).
+included. Now we will describe \textit{types} (or \textit{type terms}).\\
 
@@
 The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
-combining the previously defined \textit{\lterms} and \textit{types} together.
+combining the previously defined \textit{\lterms} and \textit{types} together. \\
 
-\begin{definition}\begin{enumerate}
+\begin{definition}~
+
+\begin{enumerate}
  \item 	Let $\Lambda$ be set of {\it \lterms}. 
@@
 \end{definition}
+%~\\
 
@@
 such that $\Gamma$ does not contain any declaration with $x$ as subject.
-We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
+We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.\\
 
@@
 for a given pair $\ul{\tau}{\Gamma}$
-such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
+such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.\\
 
@@
 
-\Lets now discuss three such search strategies.
+\Lets now discuss three such search strategies.\\
 
-\subsubsection{Systematic strategy}
-
+%\subsubsection{Systematic strategy}
+\textit{Systematic strategy}:~~ 
 If we use trivial strategy that returns all the inputs, 
@@
 first $n$ smallest lambda terms in their
-\textit{long normal form}.
+\textit{long normal form}.\\
 
-\subsubsection{Ramped half-and-half strategy}
-is generalization of the standard ramped half-and-half method described 
+%\subsubsection{Ramped half-and-half strategy}
+\textit{Ramped half-and-half strategy}:~~
+This is generalization of the standard ramped half-and-half method described 
 in \cite{koza92}. If applied to context satisfying closure requirement
@@
 If $isFull = false$, then the subset to select from is 
-the whole input set.
+the whole input set.\\
 
-\subsubsection{Geometric strategy}
-
+%\subsubsection{Geometric strategy}
+\textit{Geometric strategy}:~~
 We can see those two previous strategies as two extremes on the spectrum of 
@@
   (-)   : \Real \ar \Real \ar \Real    ,
-  (*)   : \Real \ar \Real \ar \Real    ,
-  rdiv  : \Real \ar \Real \ar \Real    ,\\
-  sin  &: \Real \ar \Real              ,
-  cos   : \Real \ar \Real              ,
-  exp   : \Real \ar \Real              , 
+  (*)   : \Real \ar \Real \ar \Real    ,\\
+  rdiv &: \Real \ar \Real \ar \Real    ,
+  sin   : \Real \ar \Real              ,
+  cos   : \Real \ar \Real              ,\\
+  exp  &: \Real \ar \Real              , 
   rlog  : \Real \ar \Real              \}
@@
 where
+
 \noindent
@@
 \end{minipage}%
+
 \begin{minipage}{.5\linewidth}
@@
   \centering
-    \includegraphics[scale=0.135]{imgs/SSR_FINAL.png}
+    \includegraphics[scale=0.1]{imgs/SSR_FINAL.png}
 \end{figure}
@@
 \begin{align*}
-\tau = An&tAction\\
+\tau = An&tAct\\
 \Gamma = \{~~
-  l    &: AntAction                              ,
-  r    : AntAction                               ,
-  m    : AntAction                               ,\\
-  ifa  &: AntAction \ar AntAction \ar AntAction  ,\\
-  p2   &: AntAction \ar AntAction \ar AntAction  ,\\
-  p3   &: AntAction \ar AntAction \ar AntAction \ar AntAction  \}
+  l    &: AntAct                              ,
+  r    : AntAct                               ,
+  m    : AntAct                               ,\\
+  ifa  &: AntAct \ar AntAct \ar AntAct  ,\\
+  p2   &: AntAct \ar AntAct \ar AntAct  ,\\
+  p3   &: AntAct \ar AntAct \ar AntAct \ar AntAct  \}
 \end{align*}
@@
   \centering
-    \includegraphics[scale=0.135]{imgs/ANT_FINAL.png}
+    \includegraphics[scale=0.1]{imgs/ANT_FINAL.png}
 \end{figure}
@@
   \centering
-    \includegraphics[scale=0.135]{imgs/EP_FINAL.png}
+    \includegraphics[scale=0.1]{imgs/EP_FINAL.png}
 \end{figure}
@@
 
-\begin{thebibliography}{22}
+%\nocite{*}
 
-\bibitem{koza92}
-  TODO. Todo,
-  \emph{Dát sem místo toho ten bibtex}.
-  MIT Press, Cambridge, MA,
-  2014.
-
-\bibitem{koza92}
-  John R. Koza,
-  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
-  MIT Press, Cambridge, MA,
-  1992.
-
-\bibitem{koza05}
-  Koza, J.R., Keane, M., Streeter, M., Mydlowec, W.,Yu, J., Lanza, G. 
-  \emph{Genetic Programming IV: Routine Human-Competitive Machine Intelligence.} 
-  Springer, 2005. ISBN 978-0-387-26417-2
-
-\bibitem{fg}
- Riccardo Poli, William B. Langdon, Nicholas F. McPhee
- \emph{A Field Guide to Genetic Programming}.
- Lulu Enterprises, UK Ltd, 2008.
-
-\bibitem{yu01}
-  T. Yu. 
-  \emph{Hierachical processing for evolving recursive and modular 
-        programs using higher order functions and lambda abstractions}. 
-  Genetic Programming and Evolvable Machines,
-  2(4):345–380, December 2001. ISSN 1389-2576.
-
-\bibitem{montana95}
-D. J. Montana. 
-\emph{Strongly typed genetic programming.} 
-Evolutionary Computation, 3(2): 199–230, 1995.
-%URL \url{ http://vishnu.bbn.com/papers/stgp.pdf }. nefacha
-
-\bibitem{haynes96}
-T. D. Haynes, D. A. Schoenefeld, and R. L. Wainwright. 
-\emph{Type inheritance in strongly typed genetic programming.} 
-In P. J. Angeline and K. E. Kinnear, Jr., editors, Advances
-in Genetic Programming 2, chapter 18, pages 359–376.
-MIT Press, Cambridge, MA, USA, 1996. ISBN 0-262-01158-1. 
-%URL \url{http://www.mcs.utulsa.edu/~rogerw/papers/Haynes-hier.pdf}.
-
-\bibitem{olsson94}
-J. R. Olsson. 
-\emph{Inductive functional programming using incremental program 
-transformation and Execution of logic programs by 
-iterative-deepening A* SLD-tree search.} 
-Dr scient thesis, University of Oslo, Norway, 1994.
-
-\bibitem{kes}
-Forrest Briggs, Melissa O’Neill.
-\emph{Functional Genetic Programming and Exhaustive
-Program Search with Combinator Expressions.}
-International Journal of Knowledge-based and Intelligent Engineering Systems,
-Volume 12 Issue 1, Pages 47-68, January 2008.
-
-\bibitem{barendregt84}
-H. P. Barendregt,
-\emph{The Lambda Calculus: its Syntax and Semantics}, 
-revised ed., North-Holland, 1984.
-
-\bibitem{barendregt92}
-H. Barendregt , S. Abramsky , D. M. Gabbay , T. S. E. Maibaum.
-\emph{Lambda Calculi with Types.} 
-Handbook of Logic in Computer Science, 1992.
-
-\bibitem{barendregt10}
-
-  Henk Barendregt, Wil Dekkers, Richard Statman,
-  \emph{Lambda Calculus With Types}.
-  Cambridge University Press,
-  2010. 
-  %URL \url{http://www.cs.ru.nl/~henk/book.pdf}.
-
-\bibitem{jones87}
-Simon Peyton Jones. 
-\emph{The Implementation of Functional Programming Languages}. 
-Prentice Hall, 1987.
-
-\bibitem{AIAMA}
-	Stuart J. Russell, Peter Norvig,
-	\emph{Artificial Intelligence: A Modern Approach}.
-	Pearson Education,
-	2003.
-
-\end{thebibliography}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,evogp}
 

### 2014-01-19 tex/cec14/article.tex
cecky

@@
 was 500 individuals.
-In those experiments we analyze the ability of the system to produce 
-a correct solution, computational cost estimates and average size 
-of an individual throughout generations. 
-For the first two metrics we use the popular measurement  
+In those experiments we analyze the fitness of the best individual,
+the average size of an individual,
+the ability of the system to produce a correct solution 
+and computational cost estimates throughout generations. 
+For the last two metrics we use the popular measurement  
 methods within GP field --- the \textit{performance curves}
@@
 to yield a correct solution with probability $z =99\%$).
+Difference of those two generating methods 
+is statistically analyzed by the Welch t-test
+by comparing the fitness values of the best individuals of the run. 
+Graphs showing mean values of the fitness of the best individual and
+the average size of an individual throughout generations
+contain error bars representing the standard error of the mean (SEM).
 
@@
 \begin{figure}[h!]
-  \caption{Performance curves for Simple symbolic regression.}
+  \caption{Graphs for Simple symbolic regression.}
   \centering
-    \includegraphics[scale=0.1]{imgs/SSR_FINAL.png}
+    \includegraphics[scale=0.15]{imgs/cec/SSR.PNG}
 \end{figure}
-Fitness function is computed as follows.
-$$ fitness(f) =  \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert   $$
+
+~\\Fitness function is computed as follows.
+
+$$ fitness(f) =  \frac{1}{1+ \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert }  $$
 where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
@@
 \begin{figure}[h!]
-  \caption{Performance curves for Artificial ant problem.}
+  \caption{Graphs for Artificial ant problem.}
   \centering
-    \includegraphics[scale=0.1]{imgs/ANT_FINAL.png}
+    \includegraphics[scale=0.15]{imgs/cec/ANT.PNG}
 \end{figure}
@@
 }.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|cc|}
+\hline
+& Ramped half-and-half & Geometric \\
+\hline
+Mean & 0.7999030182 & 0.8174067512 \\
+SD	 & 0.1521369287	& 0.1826109800 \\
+SEM	 & 0.0215154108	& 0.0258250925 \\
+N	 & 50 & 50  \\
+\hline
+t-value & 0.5207 & \\
+p-value & 0.6038 & \\
+\hline
+\end{tabular}
+\caption{Statistical analysis of Simple symbolic regression.
+By conventional criteria, this difference is considered to be not statistically significant.}
+\end{table}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|cc|}
+\hline
+& Ramped half-and-half & Geometric \\
+\hline
+Mean & 66.20 & 79.78 \\
+SD	 & 12.33 & 8.77  \\
+SEM	 & 1.74	 & 1.24  \\
+N	 & 50    & 50    \\
+\hline
+t-value &  6.3479           & \\
+p-value &  less than 0.0001 &  \\
+\hline
+\end{tabular}
+\caption{Statistical analysis of Artificial ant problem.
+ By conventional criteria, this difference is considered to be extremely statistically significant. }
+\end{table}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|cc|}
+\hline
+& Ramped half-and-half & Geometric \\
+\hline
+Mean & 0.895000	& 0.930000 \\
+SD	 & 0.075170	& 0.078490 \\
+SEM	 & 0.010631	& 0.011100 \\
+N	 & 50    & 50    \\
+\hline
+t-value &  2.2772 & \\
+p-value &  0.0250 &  \\
+\hline
+\end{tabular}
+\caption{Statistical analysis of Even parity problem. 
+By conventional criteria, this difference is considered to be statistically significant.}
+\end{table}
 
@@
 \Gamma = \{
-  and   &: Bool \ar Bool \ar Bool                              ,
-  or    : Bool \ar Bool \ar Bool                              ,\\
-  nand  &: Bool \ar Bool \ar Bool                              ,
-  nor    : Bool \ar Bool \ar Bool                              ,\\
-  foldr &: (Bool \ar Bool \ar Bool) \ar Bool \ar [Bool] \ar Bool ,\\
-  head' &: [Bool] \ar Bool                                   ,
-  tail' : [Bool] \ar [Bool]                              \}
+  and   &: Bool \ar Bool \ar Bool                              ,\\
+  or    &: Bool \ar Bool \ar Bool                              ,\\
+  nand  &: Bool \ar Bool \ar Bool                              ,\\
+  nor   &: Bool \ar Bool \ar Bool                              ,\\
+  foldr &: (Bool \ar Bool \ar Bool) \\
+        &~~~\ar Bool \ar [Bool] \ar Bool ,\\
+  head' &: [Bool] \ar Bool                                   ,\\
+  tail' &: [Bool] \ar [Bool]                              \}
 \end{align*}
 \begin{figure}[h!]
-  \caption{Performance curves for Even parity problem.}
+  \caption{Graphs for Even parity problem.}
   \centering
-    \includegraphics[scale=0.1]{imgs/EP_FINAL.png}
+    \includegraphics[scale=0.15]{imgs/cec/EP.PNG}
 \end{figure}

### 2014-01-20 tex/cec14/article.tex
cecky..

@@
 \begin{abstract}
+Tree based genetic programming traditionally uses simple S-expressions to represent programs, however more expressive representations can exhibit better results and are well suited for typed genetic programming.
+One of such representations is lambda calculus.
 In this paper, a generalization of genetic programming 
-for simply typed lambda calculus is presented. We propose three population 
-initialization methods depending on different search strategies. 
+for simply typed lambda calculus is presented. We mainly focus on
+population initialization methods. 
+Three such initialization methods are proposed 
+depending on different search strategies. 
 The first strategy corresponds to standard ramped half-and-half initialization, 
 the second one corresponds to exhaustive systematic search, and the third one 
-represents a novel geometric strategy. Three well-known benchmark experiments
+represents a novel geometric strategy. 
+Three well-known benchmark experiments (simple symbolic regression,
+artificial ant problem and even parity problem)
 support that the geometric strategy outperforms
-the standard generating method in success rate, time consumption and average individual size. 
+the standard generating method in success rate, best fitness value,
+time consumption and average individual size. 
+Despite the fact that the presented system is designed for typed genetic 
+programming, when given standard genetic programming problem satisfying
+standard closure requirement it produces simple S-expressions and
+the novel geometric strategy outperforms the standard system. 
+This claim is supported by that first two experiments both satisfying 
+the closure requirement.
+The fact that a search strategy can be easily described by a rather simple filtration algorithm opens opportunity to easy exploration of more 
+population initialization methods, possibly in a kind of meta-evolution way.
 \end{abstract}
@@
 
-In \cite{yu01} Yu presents a GP system utilizing
+Yu presents a GP system utilizing
 polymorphic higher-order functions\footnote{Higher-order 
 function is a function taking another function as 
-input parameter.} and lambda abstractions.
+input parameter.} and lambda abstractions  \cite{yu01}.
 Important point of interest in this work is use of
@@
 
-In \cite{kes} Briggs and O’Neill present technique 
-utilizing typed GP with combinators.
+Briggs and O’Neill present technique 
+utilizing typed GP with combinators \cite{kes}.
 The difference between approach presented in this work
@@
 
-In \cite{binard2008genetic} by Binard and Felty even 
-stronger type system (\textit{System F}) is used.  
+Binard and Felty use even 
+stronger type system (\textit{System F}) \cite{binard2008genetic}.  
 But with increasing power of the type system comes increasing difficulty of term generation.
@@
 so we use technique based on \textit{inhabitation machines} 
-described by Barendregt in~\cite{barendregt10}.
+described by Barendregt \cite{barendregt10}.
 
@@
 
+\red{Přestože je naše generujicí metoda navržena obecně pro typovanou verzi,
+tak pracuje dobře i pro problémy standardního GP splňující closure podmínku.
+Pro taklovéto problémy strategie generuje standardní S-výrazy. Proto jsou
+první dva ze tří experimentů právě takového charakteru. Dalším důvodem pro volbu 
+takovýchto problémů je fakt, že výsledky pokusů výrazně ovlivňuje použitý křížící 
+operátor. Návrh obecného křížícího operatoru pro GP v symply typed lc je otevřený problém. Protože křížící operátor použitý v našem systému je poměrně komplikovaný
+a tím pádem beyond the scope of this paper, ale pro problémy splnující
+clousure podmínku se chová přesně stejně jako standardní křížící operator,
+tak se to zdá jako další důvod proč 2/3 experimentů udělat na closed problémech.}
+
 \subsection{Simple symbolic regression}
@@
 
-\begin{figure}[h!]
-  \caption{Graphs for Simple symbolic regression.}
-  \centering
-    \includegraphics[scale=0.15]{imgs/cec/SSR.PNG}
-\end{figure}
-
 ~\\Fitness function is computed as follows.
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.14]{imgs/cec/SSR.PNG}
+  \caption{Graphs for Simple symbolic regression.}
+\end{figure}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|cc|}
+\hline
+& Ramped half-and-half & Geometric \\
+\hline
+Mean & 0.7999030182 & 0.8174067512 \\
+SD	 & 0.1521369287	& 0.1826109800 \\
+SEM	 & 0.0215154108	& 0.0258250925 \\
+N	 & 50 & 50  \\
+\hline
+t-value & 0.5207 & $\alpha = 0.05$\\
+p-value & 0.6038 & \textbf{Not statistically significant.}\\
+\hline
+\end{tabular}
+\caption{Statistical analysis - Simple symbolic regression.}
+\end{table}
+
+~\\
+
+Figure~1 shows results of this experiment. 
+Table~I summarizes statistical analysis
+of the fitness values of the best individuals of the run.  
+By conventional criteria, this difference is considered to be not statistically significant.
 Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
 Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
-The average individual size for generation 50 was 76.
+The average individual size for generation 50 was 68.8.
 The experiment took 46 minutes.
@@
 Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
-The average individual size for generation 50 was 47.
+The average individual size for generation 50 was 44.1.
 The experiment took 26 minutes.
-Figure~1 shows the progress of those metrics throughout generations.
 So for the \textit{Simple symbolic regression}
 our geometric strategy is slightly more successful then 
-ramped half-and-half strategy in all observed metrics.
+ramped half-and-half strategy in these observed metrics.
 
@@
 
-\begin{figure}[h!]
-  \caption{Graphs for Artificial ant problem.}
-  \centering
-    \includegraphics[scale=0.15]{imgs/cec/ANT.PNG}
-\end{figure}
-
 Fitness value is equal to number of eaten food pieces.
@@
 
-\begin{table}[t]
+\begin{figure}[h!]
   \centering
-\begin{tabular}{|l|cc|}
-\hline
-& Ramped half-and-half & Geometric \\
-\hline
-Mean & 0.7999030182 & 0.8174067512 \\
-SD	 & 0.1521369287	& 0.1826109800 \\
-SEM	 & 0.0215154108	& 0.0258250925 \\
-N	 & 50 & 50  \\
-\hline
-t-value & 0.5207 & \\
-p-value & 0.6038 & \\
-\hline
-\end{tabular}
-\caption{Statistical analysis of Simple symbolic regression.
-By conventional criteria, this difference is considered to be not statistically significant.}
-\end{table}
+  \includegraphics[scale=0.14]{imgs/cec/ANT.PNG}
+  \caption{Graphs for Artificial ant problem.}
+\end{figure}
 
@@
 \hline
-t-value &  6.3479           & \\
-p-value &  less than 0.0001 &  \\
+t-value &  6.3479           & $\alpha = 0.05$\\
+p-value &  \textless 0.0001 &  \textbf{Extremely statistically significant.}\\
 \hline
 \end{tabular}
-\caption{Statistical analysis of Artificial ant problem.
- By conventional criteria, this difference is considered to be extremely statistically significant. }
+\caption{Statistical analysis - Artificial ant problem.}
 \end{table}
 
-\begin{table}[t]
-\centering
-\begin{tabular}{|l|cc|}
-\hline
-& Ramped half-and-half & Geometric \\
-\hline
-Mean & 0.895000	& 0.930000 \\
-SD	 & 0.075170	& 0.078490 \\
-SEM	 & 0.010631	& 0.011100 \\
-N	 & 50    & 50    \\
-\hline
-t-value &  2.2772 & \\
-p-value &  0.0250 &  \\
-\hline
-\end{tabular}
-\caption{Statistical analysis of Even parity problem. 
-By conventional criteria, this difference is considered to be statistically significant.}
-\end{table}
+Figure~2 shows results of this experiment. 
+Table~II summarizes statistical analysis
+of the fitness values of the best individuals of the run.  
+According to this analysis our strategy performed better;
+by conventional criteria, this difference is considered to be extremely statistically significant.
 
@@
 Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
-The average individual size for generation 50 was circa 400.
+The average individual size for generation 50 was 279.5.
 The experiment took 265 minutes.
+
 Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
 Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
-The average individual size for generation 50 was circa 90.
+The average individual size for generation 50 was 94.1.
 The experiment took 107 minutes.
+
 This is a big improvement in all watched factors.
-Figure~2 shows it in greater detail.
 
@@
 
-Previous two problems were instances of standard GP, i.e. their building symbols
-obeyed the closure requirement. This third experiment will break the closure requirement, thus ideal for testing typed GP techniques. 
+Previous two problems were instances of standard GP, i.e. their building 
+symbols obeyed the closure requirement. This third experiment will 
+break the closure requirement, thus ideal for testing typed GP techniques. 
 The even-parity function is a boolean function taking as inputs $N$
@@
 This problem has been used by many researchers
-as a benchmark for GP. We compare our results with that obtained by Yu in \cite{yu01},
+as a benchmark for GP. We compare our results with 
+that obtained by Yu in \cite{yu01},
 where is presented approach to evolve recursive and modular programs
@@
 \end{align*}
-\begin{figure}[h!]
-  \caption{Graphs for Even parity problem.}
-  \centering
-    \includegraphics[scale=0.15]{imgs/cec/EP.PNG}
-\end{figure}
 
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.14]{imgs/cec/EP.PNG}
+  \caption{Graphs for Even parity problem.}
+\end{figure}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|cc|}
+\hline
+& Ramped half-and-half & Geometric \\
+\hline
+Mean & 0.895000	& 0.930000 \\
+SD	 & 0.075170	& 0.078490 \\
+SEM	 & 0.010631	& 0.011100 \\
+N	 & 50    & 50    \\
+\hline
+t-value &  2.2772 & $\alpha = 0.05$\\
+p-value &  0.0250 &  \textbf{Statistically significant.}\\
+\hline
+\end{tabular}
+\caption{Statistical analysis - Even parity problem.}
+\end{table}
+
+Figure~3 shows results of this experiment. 
+Table~III summarizes statistical analysis
+of the fitness values of the best individuals of the run.  
+According to this analysis our strategy performed better;
+by conventional criteria, this difference is considered to be 
+statistically significant.
+
 Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
 Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
-The average individual size for generation 50 was 39.
+The average individual size for generation 50 was 47.7.
 The experiment took 28 minutes.
@@
 Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
-The average individual size for generation 50 was circa 100.
+The average individual size for generation 50 was 74.3.
 The experiment took 33 minutes.
-Figure~3 shows it in greater detail.
+
 Once again this is a big improvement against standard method.
+
 One may get the impression that geometric strategy suffers from bloat here,
@@
 17,500 in generation 4. But by observation of typical solution which is often
-some modification of \texttt{foldr xor False inputList} one sees that important task
-for GP here is to create \texttt{xor} function. This is advantage for the more specialized term representation used in \cite{yu01} 
+some modification of \texttt{foldr xor False inputList} one sees that important 
+task for GP here is to create \texttt{xor} function. This is advantage for the 
+more specialized term representation used in \cite{yu01} 
 (restricting use of outer variables). 
@@
 
+\red{Ještě nějak rozšířit.}
+
 %\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }

### 2014-01-20 tex/cec14/article.tex
cecky... a odeslano

@@
 
-\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
+\author{Tom\'{a}\v{s} K\v{r}en\thanks{Tomáš Křen (tomkren@gmail.cz) is with Charles Univetsity in Prague, Faculty of Mathematics and Physics; Department of Theoretical Computer Science and Mathematical Logic.} \and Roman Neruda\thanks{Roman Neruda (roman@cs.cas.cz) is with Institute of Computer Science, Academy of Sciences of the Czech Republic and has been partially supported by the The Ministry of Education of the Czech Republic project COST LD 13002}}
 
@@
 represents a novel geometric strategy.
-Three well-known benchmark experiments (simple symbolic regression,
-artificial ant problem and even parity problem)
-support that the geometric strategy outperforms
+Experiments on three well known benchmark problems show
+that the geometric strategy outperforms
 the standard generating method in success rate, best fitness value,
 time consumption and average individual size.
-Despite the fact that the presented system is designed for typed genetic 
-programming, when given standard genetic programming problem satisfying
-standard closure requirement it produces simple S-expressions and
-the novel geometric strategy outperforms the standard system. 
-This claim is supported by that first two experiments both satisfying 
-the closure requirement.
 The fact that a search strategy can be easily described by a rather simple filtration algorithm opens opportunity to easy exploration of more 
@@
 straight from \textit{library} of combinators and no lambda abstractions
-are used. They are using more general polymorphic type system then us
+are used. They are using more general polymorphic type system than us
 -- the Hindley–Milner type system. They also discuss the 
@@
 
-Our approach to \lterm gerating is based on technique 
+Our approach to \lterm generating is based on technique 
 briefly described in \cite{barendregt10}, which generates
@@
 If the queue becomes empty before the desired number of \lterms
-is generated, then the initial statete is inserted to the queue
+is generated, then the initial state is inserted to the queue
 and the process continues. For the standard A* this would be meaningless,
@@
 no matter how many variables/constants of
-each type there are, it is should by pretty effective to compute this
+each type there are, it should by pretty effective to compute this
 value precisely and store them for later. 
@@
 rather long long normal form into $\beta\eta$-normal form.  
-Another performence enhancing transformation is use of "applicative" tree 
+Another performance enhancing transformation is use of "applicative" tree 
 representation (coming directly from inductive definition of \lterms).
@@
 
-We made tree experiments comparing performance of standard 
+We made three experiments comparing performance of standard 
 \textit{ramped half-and-half strategy} with our
@@
 
-\red{Přestože je naše generujicí metoda navržena obecně pro typovanou verzi,
-tak pracuje dobře i pro problémy standardního GP splňující closure podmínku.
-Pro taklovéto problémy strategie generuje standardní S-výrazy. Proto jsou
-první dva ze tří experimentů právě takového charakteru. Dalším důvodem pro volbu 
-takovýchto problémů je fakt, že výsledky pokusů výrazně ovlivňuje použitý křížící 
-operátor. Návrh obecného křížícího operatoru pro GP v symply typed lc je otevřený problém. Protože křížící operátor použitý v našem systému je poměrně komplikovaný
-a tím pádem beyond the scope of this paper, ale pro problémy splnující
-clousure podmínku se chová přesně stejně jako standardní křížící operator,
-tak se to zdá jako další důvod proč 2/3 experimentů udělat na closed problémech.}
+Despite the fact that the presented system is designed
+for typed genetic programming it works well for
+standard GP problems satisfying the closure requirement. For such problems our approach 
+generates standard S-expressions only. 
+Another reason for including standard GP
+benchmarks in our experiments is that fact that
+the results are seriously affected by a crossover 
+operator. Design of a good crossover for typed GP
+remains an open problem. 
+Our crossover is quite complicated and beyond
+the scope of this paper, but for problems satisfying
+the closure requirement it behaves exactly the same
+as the standard GP crossover. 
+Three well known benchmark problems
+(Simple symbolic regression, Artificial ant and
+Even parity problem) were chosen for our experiments.
+First two satisfy the closure requirement, 
+whereas the third one does not.
 
@@
 So for the \textit{Simple symbolic regression}
-our geometric strategy is slightly more successful then 
+our geometric strategy is slightly more successful than 
 ramped half-and-half strategy in these observed metrics.
@@
 
-One may get the impression that geometric strategy suffers from bloat here,
-but considering that it performs significantly better and
-that term size around 100 is similar as for the previous two problems, 
+One may get the impression that geometric strategy 
+suffers from bloat here, but as is stated in \cite{fg}: 
+\textit{"We should not equate growth with bloat
+and we should define bloat as program growth without 
+(significant) return in terms of fitness."}
+Considering that it performs significantly better and
+that term size around 75 is similar as for the previous two problems, 
 it seems to us that ramped half-and-half maybe has difficulties with 
 constructing more complex terms, and therefore is "staying small".
+
 Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
@@
 result for uninformed search. However, our results are less successful 
-then those presented in \cite{yu01} (with slightly different set of 
+than those presented in \cite{yu01} (with slightly different set of 
 building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
@@
 
+Experiments on three well 
+known benchmark problems show
+that the geometric strategy outperforms
+the standard ramped half-and-half 
+generating method in success rate, best fitness value,
+time consumption and average individual size.
+
 The relevance of these algorithms does not effect only initialization, but it can be used during mutation
@@
 
-\red{Ještě nějak rozšířit.}
+%\red{Ještě nějak rozšířit.}
 

### 2014-01-21 tex/cec14b/article.tex
cecek ... resubmitly-finale

--- 2014-01-20 tex/cec14/article.tex
@@
 tree-swapping crossover operator. Since it is beyond the scope of this 
-paper we mention it only briefly. Two main concerns with swapping typed subtrees 
-are types and free variables. Well-typed offspring is obtained 
+paper we mention it only briefly. Two main concerns with swapping typed 
+subtrees are types and free variables. Well-typed offspring is obtained 
 by swapping only subtrees of the same type. Only subtrees with 
@@
 after initialization is $\eta$-normalization shortening 
-rather long long normal form into $\beta\eta$-normal form.  
-Another performance enhancing transformation is use of "applicative" tree 
-representation (coming directly from inductive definition of \lterms).
+rather long long normal form into shorter $\beta\eta$-normal form.  
+Another performance enhancing transformation is option of using "applicative" tree representation (coming directly from inductive definition of \lterms) instead of more traditional S-expression representation.
+
+A remarkable benefit of parameterizing the generating method 
+by a search strategy taking the form of a simple filtration function
+is that such a function can be expressed by functional language in a very 
+economical way. Take for instance a search strategy that behaves same
+as the \textit{geometric} except that when it comes to 
+situation where all the elements are filtered out it
+acts as the \textit{ramped half-and-half}. 
+Such \textit{strategy composition}
+is easily describable as a higher order function. Thus comes 
+to mind the possibility of using our system to come up with new
+search strategies on its own.
 
@@
 \Gamma = \{
-  (+)  &: \Real \ar \Real \ar \Real    ,
-  (-)   : \Real \ar \Real \ar \Real    ,
-  (*)   : \Real \ar \Real \ar \Real    ,\\
-  rdiv &: \Real \ar \Real \ar \Real    ,
-  sin   : \Real \ar \Real              ,
-  cos   : \Real \ar \Real              ,\\
-  exp  &: \Real \ar \Real              , 
+  (+)  &: \Real \ar \Real \ar \Real    ,~ 
+  (-)   : \Real \ar \Real \ar \Real    ,~\\
+  (*)  &: \Real \ar \Real \ar \Real    ,~ 
+  rdiv  : \Real \ar \Real \ar \Real    ,~\\
+  sin  &: \Real \ar \Real              ,~ 
+  cos   : \Real \ar \Real              ,~\\
+  exp  &: \Real \ar \Real              ,~ 
   rlog  : \Real \ar \Real              \}
@@
 
-~\\Fitness function is computed as follows.
+~\\
+
+Fitness function is computed as follows.
 
@@
 An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
-considered as a correct individual.
+considered as a correct individual.\\
 
@@
   \centering
-  \includegraphics[scale=0.14]{imgs/cec/SSR.PNG}
   \caption{Graphs for Simple symbolic regression.}
+  \includegraphics[scale=0.241]{imgs/cec/SSRb.PNG}
 \end{figure}
@@
 
-~\\
-
 Figure~1 shows results of this experiment. 
 Table~I summarizes statistical analysis
-of the fitness values of the best individuals of the run.  
+of the fitness values of the best individuals of the run 
+(\textit{p-value} $ = 0.6038$).  
 By conventional criteria, this difference is considered to be not statistically significant.
@@
 piece is ahead of him.
-
 \begin{align*}
-\tau = An&tAct\\
+\tau~ = An&tAct\\
 \Gamma = \{~~
@@
 
-Action $l$ turns the ant left. 
-Action $r$ turns the ant right.
+Actions $l$ and $r$ turn the ant left and right, respectively. 
 Action $m$ moves the ant forward.
-Action $ifa~x~y$ (if-food-ahead) performs action $x$ 
-if a food piece is ahead of the ant,
+Action $ifa~x~y$ \mbox{(\textit{if-food-ahead})} 
+performs action $x$ if a food piece is ahead of the ant,
 otherwise it performs action $y$.
-Action $p2~x~y$ first performs action $x$ and after it action $y$.
-Action $p3~x~y~z$ first performs action $x$, 
-after that action $y$ and finally $z$.
+Actions $p2$ and $p3$ perform two and three consecutive actions, respectively.
 Actions $l, r$ and $m$ each take one time step to execute.
@@
 maximal number of steps. 
-
 Fitness value is equal to number of eaten food pieces.
@@
   \centering
-  \includegraphics[scale=0.14]{imgs/cec/ANT.PNG}
   \caption{Graphs for Artificial ant problem.}
+  \includegraphics[scale=0.26]{imgs/cec/ANTb.PNG}
 \end{figure}
@@
 Table~II summarizes statistical analysis
-of the fitness values of the best individuals of the run.  
+of the fitness values of the best individuals of the run 
+(\textit{p-value} $ < 0.0001$).  
 According to this analysis our strategy performed better;
 by conventional criteria, this difference is considered to be extremely statistically significant.
-
 Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
@@
 The experiment took 265 minutes.
-
 Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
@@
 The experiment took 107 minutes.
-
 This is a big improvement in all watched factors.
@@
   \centering
-  \includegraphics[scale=0.14]{imgs/cec/EP.PNG}
   \caption{Graphs for Even parity problem.}
+  \includegraphics[scale=0.26]{imgs/cec/EPb.PNG}
 \end{figure}
@@
 Table~III summarizes statistical analysis
-of the fitness values of the best individuals of the run.  
+of the fitness values of the best individuals of the run 
+(\textit{p-value} $ = 0.0250$).  
 According to this analysis our strategy performed better;

### 2014-01-21 js/gui.js
PSO tak nak vachrlate funguje

@@ -147,7 +147,7 @@ function mkGUI (containerId) {
 
   var $editorContainer = $('<div>').append([$editorButts, $elRyba, $editor]);
 
-  var sTabsNames = ['mxr','editor','results','stats','log'];
+  var sTabsNames = ['editor','mxr','results','stats','log'];
   var sTabs = mkTabs(sTabsNames);
 
   var $clearButt = mkAButt('clear','remove',false,false,true);
@@ -310,7 +310,7 @@ function mkGUI (containerId) {
   function resize () {
     var newHeight = window.innerHeight-106;
     $log   .css('height', (newHeight-30)+'px');
-    $editor.css('height', (newHeight)+'px');
+    $editor.css('height', (newHeight-2)+'px');
     $graphContainer.css('height',(newHeight-20)+'px');
     var $bars = Phenotype.get$bars();
     if ($bars) {
@@ -373,6 +373,7 @@ function mkGUI (containerId) {
       var content = msg.content;
       switch (msg.subject) {
         case 'log'     : guiLog(content); break;
+        case 'consoleLog': log(content); break;
         case 'stats'   : 
           var stats = content;
           graphs.handleStats(stats);

### 2014-01-21 js/problems/RB.js

@@ -12,74 +12,61 @@
 opts = {
   name: 'RB',
 
-  numRuns: 50,
-  numGens: 50,
-  popSize: 1000,
+  minimization : true,
 
-  saveBest  : false,
+  numRuns: 10,
+  numGens: 100,
+  popSize: 30,
 
-  statsOpts: StatsOpts.default,
+  saveBest  : true,
+
   logOpts  : LogOpts.default,
 
   init: function () {
     
-    var data = TSPutils.tsp225; //Ulysses22; //16; //.EU4,
+    var benchmark = Real.Sphere; //Real.Rosenbrock;
 
-    this.solver = ACO.mkSolver(
-      TSPutils.mkTspProblem({
-        Q: data.optVal,//6792.12,
-        initTauVal: 1,
-        from: '1', // 'Praha'
-        data: data, 
-        //dists: TSPutils.EU4_dists,
-      }, ACO.defaultAntOpts)
-    );
+    this.solver = PSO.mkSolver({
+      n:     3,
+      omega: 0.5,
+      phi_p: 0.5,
+      phi_g: 0.5,
+      b_lo:  benchmark.interval[0],
+      b_up:  benchmark.interval[1]
+    });
 
-    this.fitness   = this.solver.antProblem.fitness;
-    this.operators = [[this.solver.Operators.generatePop, 1.0]];
+    this.fitness   = benchmark.f;
+    this.operators = [[this.solver.Operators.movePopOperator, 1.0]];
 
-    var h = 320;
+    var h = 20;
 
     this.phenotype = {
       height: h,
       init: function ($el){
-        var $panel = $('<div id="__TSPpanel">');
+        var $panel = $('<div>');
         $el.append($panel);
-
-        var w = $(window).width(); 
-        var w_small = (w-15)/2; 
-                
-        //$panel.html(w_small);
-        
- 
-        $panel.html(
-          '<table style="margin-bottom: 3px;"><tr>'+
-          '<td><canvas id="TSP_canvas1" width="'+w_small+'" height="'+h+'"></canvas>'+
-          '<td><canvas id="TSP_canvas2" width="'+w_small+'" height="'+h+'"></canvas>'+
-          '</table>'
-        );
-
-
-
-        $('#__TSPpanel td').css({border: '1px dotted gray'});
-        
-        
-
         return $panel;
       },
       update: function (el, indivFun, indiv, runKnowledge) {
 
-        TSPutils.draw($('#TSP_canvas1'), data , {
[…]

### 2014-01-21 js/problems/Real/Real.js

@@ -1,0 +1,38 @@
+var Real = (function () {
+
+
+  function repFun (x) {
+    return function (n) {
+      return Utils.replicate(n, x);
+    };
+  }
+
+  var Sphere = {
+    f : function (xs) {
+      return sum(_.map(xs, function(x){return x*x;}));
+    },
+    interval : [-5.12, 5.12],
+    optimum  : repFun(0) 
+  };
+
+  var Rosenbrock = {
+    f : function (xs) {
+      var ret = 0;
+      for (var i=0; i<xs.length-1; i++) {
+        var x = xs[i];
+        var a = xs[i+1] - x*x; 
+        var b = (x-1)*(x-1);
+        ret  += 100*a*a + b*b;  
+      }
+      return ret;
+    },
+    interval : [-2.048,2.048],
+    optimum  : repFun(1)
+  };
+
+  return {
+    Sphere: Sphere,
+    Rosenbrock: Rosenbrock,
+  };
+
+})();

### 2014-01-21 js/problems/_register.js

@@ -1,6 +1,6 @@
 var ProblemRegister = [
+  'TSP.js',
+  'RB.js',
   'SSR.js',
-  'TSP.js'
-  //'ant.js',
-  //'SSR_old.js'
+
 ];

### 2014-01-21 js/solver.js

@@ -90,14 +90,15 @@ var Solver = (function () {
     return function (pop) {
       return evalPop_core( evalPopOpts.mkIndivArr(pop), 
                            evalPopOpts.mkIndivFitness(opts),
-                           evalPopOpts.onEvalIndivUpdate );
+                           evalPopOpts.onEvalIndivUpdate,
+                           opts.minimization );
     }
   }
 
-  function evalPop_core (indivArr, indivFitness, onEvalIndivUpdate) {
+  function evalPop_core (indivArr, indivFitness, onEvalIndivUpdate, minimization) {
 
     var terminate = false;
-    var best      = {fitVal: 0};
+    var best      = {fitVal: (minimization ? Number.MAX_VALUE : 0) }  ;
     var popDist   = mkDist(_.map(indivArr, function (indiv) {
       
       var fitResult = indivFitness(indiv);
@@ -124,10 +125,14 @@ var Solver = (function () {
 
       fitResult.indiv = indiv;
 
-      best = updateBest(best, fitResult);
+      best = updateBest(best, fitResult, minimization);
 
       return [indiv,fitResult.fitVal];
     }));
+
+    //log(best);
+    //log(popDist.bestVal(true) );
+    //log('---');
 
     var evaledPop = {
       popDist:   popDist,
@@ -164,9 +169,9 @@ var Solver = (function () {
       run:        run,
       gen:        gen,
       terminate:  evaledPop.terminate,
-      best:       popDist.bestVal(),
+      best:       opts.minimization ? popDist.worstVal() : popDist.bestVal(),
       avg:        popDist.avgVal(),
-      worst:      popDist.worstVal(),
+      worst:      opts.minimization ? popDist.bestVal() : popDist.worstVal(),
       bestSize:   solver.commOpts.size(bestTerm),  //    <-----------
       avgSize:    avgTermSize,
       maxSize:    maxTermSize,

### 2014-01-21 js/solvers/PSO.js

@@ -7,9 +7,16 @@ var PSO = (function () {
       omega: 1,
       phi_p: 1,
       phi_g: 1,
-      b_lo:  [-1],
-      b_up:  [ 1]
+      b_lo:  -1,
+      b_up:  1
     }, PSOopts);
+
+    if(_.isNumber(PSOopts.b_lo)) {
+      PSOopts.b_lo = Utils.replicate(PSOopts.n, PSOopts.b_lo);
+    }
+    if(_.isNumber(PSOopts.b_up)) {
+      PSOopts.b_up = Utils.replicate(PSOopts.n, PSOopts.b_up);
+    }
 
     function mkParticle () {
       var n = PSOopts.n;
@@ -77,7 +84,7 @@ var PSO = (function () {
     }
 
     var Operators = {
-      generatePop : {
+      movePopOperator : {
         in : function (opts) {
           return opts.popSize;
         },
@@ -92,7 +99,7 @@ var PSO = (function () {
       return {
         globalBest: {
           x: null,
-          fitVal: 0
+          fitVal: Number.MAX_VALUE
         }
       };
     }
@@ -121,7 +128,7 @@ var PSO = (function () {
       },
       onEvalIndivUpdate: function (indiv, fitVal) {
 
-        if (indiv.term.p_fitVal < fitVal) {
+        if (indiv.term.p_fitVal > fitVal) { //minimize
       
           var indiv_new = _.clone(indiv);
           indiv_new.term.p = indiv.term.x;
@@ -135,7 +142,7 @@ var PSO = (function () {
     };
 
     function updateRunKnowledge (runKnowledge, evaledPop, _opts) {
-      if (evaledPop.best.fitVal > runKnowledge.globalBest.fitVal) {
+      if (evaledPop.best.fitVal < runKnowledge.globalBest.fitVal) { //minimize
         return {
           globalBest: {
             x:      evaledPop.best.indiv.term.x,
@@ -166,9 +173,17 @@ var PSO = (function () {
       commOpts: commOpts,
       ctx: ctx,
 
-      Operators: Operators
+      Operators: Operators,
+
+      //for testing, could be private
+      mkParticle:   mkParticle,
+      moveParticle: moveParticle
     };
   }
+
+  //function mkFiness_minimization (realBenchmark) {
+  //  return 
+  //}
 
   return {
     mkSolver : mkSolver

### 2014-01-21 js/utils.js

@@ -4,9 +4,16 @@ var App = {
 
 // TODO : postupně všecko z tohoto souboru přendat do tohodle namespacu
 var Utils = {
-  uniformRand : function (a, b) {
+  uniformRand: function (a, b) {
     return (b-a)*Math.random()+a;
-  }
+  },
+  replicate: function (n, x) {
+    var ret = [];
+    for (var i=0; i<n; i++) {
+      ret[i] = x;
+    }
+    return ret;
+  } 
 };
 
 function evalOptsStr (optsStr) {
@@ -121,11 +128,11 @@ var partition = function(mustBeTrue,array){
            notSatisfy : notSatisfy };
 };
 
-function updateBest( oldRes , newRes ){
+function updateBest (oldRes , newRes, minimization) {
   if( oldRes.fitVal < newRes.fitVal ){
-    return newRes;
+    return (minimization ? oldRes : newRes) ;
   }
-  return oldRes;
+  return (minimization ? newRes : oldRes) ;
 }
 
 function mkDist( distArr ){

### 2014-01-21 js/worker.js

@@ -18,14 +18,19 @@ importScripts(
  'solver.js',
  'solvers/GP.js',
  'solvers/ACO.js',
+ 'solvers/PSO.js',
  'problem-opts.js',
- 'problems/TSP/TSPutils.js'
+ 'problems/TSP/TSPutils.js',
+ 'problems/Real/Real.js'
 );
 
 var log = function (logStr) {
-  send('log', logStr); 
+  send('consoleLog', logStr); 
 };
 
+var guiLog = function (logStr) {
+  send('log', logStr); 
+};
 
 
 function send (subject, content) {
@@ -43,7 +48,7 @@ self.addEventListener('message', function(e) {
   var opts = evalOptsStr( msg.optsStr );
 
   var communicator = {
-    log: log, 
+    log: guiLog, 
     sendStats: function (stats) {
       send('stats', stats);
     },

### 2014-01-21 js/problems/TSP.js
ACO-TSP už zbytečně neposílá ty kopy dat ale je to teď udelany prasacky, nutno předelat
poznamykv komentech v TSP.js a TSPutils.js

@@ -37,6 +37,44 @@ opts = {
     this.fitness   = this.solver.antProblem.fitness;
     this.operators = [[this.solver.Operators.generatePop, 1.0]];
 
+
+    var maxEdges = 500;
+
+    //TODO: drawStyledEdges v TSPutils udělany prasacky jak sviň, předělat !
+    // tady to odešle spíš než objekt pole a tam se to už je vytiskne, dvaktrat se to zbytečně třídí
+    this.preprocessRK = function (tauGraph) {
+
+      var arr = [];
+
+      for (var i in tauGraph) {
+        for (var j in tauGraph[i]) {
+          arr.push({i:i, j:j, tau:tauGraph[i][j]});
+        }
+      }
+
+      arr = _.sortBy(arr, function (o) {
+        return o.tau;
+      });
+
+      maxEdges = maxEdges || arr.length;
+      var from = arr.length - maxEdges;
+      from = from >=0 ? from : 0;
+
+      var ret = {};
+
+      for (var k = from; k<arr.length; k++) {
+        var o = arr[k];
+        if (ret[o.i] === undefined) {
+          ret[o.i] = {};
+        }
+        ret[o.i][o.j] = o.tau;
+      }
+
+      log(ret);
+
+      return ret;
+    };
+
     var h = 320;
 
     this.phenotype = {

### 2014-01-21 js/problems/TSP/TSPutils.js

@@ -169,6 +169,8 @@ var TSPutils = (function () {
     return graph;
   }
 
+  //TODO !!!!!!!!!!!! udelany prasacky jak sviň, lepší aby se tady jen tisklo co příde z 
+  // preprocesu
   function drawStyledEdges (ctx, coords, pxPos, graph) {
     var keys = _.keys(coords);
     var len  = keys.length;
@@ -179,9 +181,12 @@ var TSPutils = (function () {
       for (var j = i+1; j<len; j++) {
         var I = keys[i];
         var J = keys[j];
-        var x = graph[I][J];
-        if (x < min) {min = x;}
-        if (x > max) {max = x;}
+
+        if (graph[I] !== undefined && graph[I][J] !== undefined){
+          var x = graph[I][J];
+          if (x < min) {min = x;}
+          if (x > max) {max = x;}
+        }
       }
     }
 
@@ -200,17 +205,20 @@ var TSPutils = (function () {
         var I = keys[i];
         var J = keys[j];
 
+        if (graph[I] !== undefined && graph[I][J] !== undefined) {
 
-        var x = Math.round( 250 * ( 1-normalize(graph[I][J]) ) ) ;
+          var x = Math.round( 250 * ( 1-normalize(graph[I][J]) ) ) ;
 
-        var colorStr = 'rgb('+x+','+x+','+x+')';
+          var colorStr = 'rgb('+x+','+x+','+x+')';
 
-        edgeArr.push({ 
-          colorStr: colorStr, 
-          tau: graph[I][J],
-          iPos: pxPos(coords[I]),
-          jPos: pxPos(coords[J]) 
-        });
+          edgeArr.push({ 
+            colorStr: colorStr, 
+            tau: graph[I][J],
+            iPos: pxPos(coords[I]),
+            jPos: pxPos(coords[J]) 
+          });
+
+        }
                 
       }
     }
@@ -218,6 +226,7 @@ var TSPutils = (function () {
     edgeArr = _.sortBy( edgeArr, function (x) {
       return x.tau;
     });
+
 
     for (var k = 0; k<edgeArr.length; k++) {
       var e = edgeArr[k];

### 2014-01-21 js/solver.js

@@ -177,7 +177,7 @@ var Solver = (function () {
       maxSize:    maxTermSize,
       minSize:    minTermSize,
       best_jsStr: solver.commOpts.indivStr(bestIndiv), //    <-----------
-      runKnowledge: runKnowledge
+      runKnowledge: opts.preprocessRK ?  opts.preprocessRK(runKnowledge) : runKnowledge
     });
 
     var logOpts = opts.logOpts; 

### 2014-01-22 js/problems/TSP.js
přípravana mmas

@@ -15,23 +15,28 @@ opts = {
   numGens: 201,
   popSize: 10,
 
-  saveBest : true,
+  saveBest : !true,
   
   statsOpts: StatsOpts.default,
   logOpts  : LogOpts.default,
 
   init: function () {
     
-    var data = TSPutils.tsp225; //Ulysses22; //16; //.EU4,
+    var data = TSPutils.tsp225; //TSPutils.Ulysses22; //16; //.EU4,
 
     this.solver = ACO.mkSolver(
       TSPutils.mkTspProblem({
+        method: 'AS',
         Q: data.optVal,//6792.12,
         initTauVal: 1,
         from: '1', // 'Praha'
         data: data, 
         //dists: TSPutils.EU4_dists,
-      }, ACO.defaultAntOpts)
+      },{
+        rho  : 0.15 ,
+        alpha: 1.2  ,
+        beta : 1.2   
+      })
     );
 
     this.fitness   = this.solver.antProblem.fitness;
@@ -69,8 +74,6 @@ opts = {
         }
         ret[o.i][o.j] = o.tau;
       }
-
-      log(ret);
 
       return ret;
     };

### 2014-01-22 js/problems/TSP/TSPutils.js

@@ -16,6 +16,7 @@ var TSPutils = (function () {
     }
 
     return {
+      method   : tspOpts.method,
       from     : tspOpts.from,
       fitness  : mkTSPFitness(tsp, tspOpts.Q),
       heur     : mkTspHeur(tsp),

### 2014-01-22 js/solvers/ACO.js

@@ -3,9 +3,7 @@ var ACO = (function () {
   var defaultAntOpts = {
     rho          : 0.15 ,
     alpha        : 1.2  ,
-    beta         : 1.2  ,
-    antsPerRound : 100  ,
-    numRounds    : 10   ,
+    beta         : 1.2   
   };
 
   function selectSucc (from, path, tau, succsFun, heur, antOpts) {
@@ -84,7 +82,7 @@ var ACO = (function () {
     }
 
 
-    function updateRunKnowledge (runKnowledge, evaledPop, opts) {
+    function update_AS (runKnowledge, evaledPop) {
       //TODO !!!
       
       var antOpts = antProblem.opts;
@@ -124,6 +122,14 @@ var ACO = (function () {
       });
 
       return newTau;
+    }
+
+    function updateRunKnowledge (runKnowledge, evaledPop, opts) {
+      switch (antProblem.method) {
+        case 'AS'   : return update_AS(runKnowledge, evaledPop); 
+        case 'MMAS' : throw 'TODO';
+        default     : throw 'ACO.updateRunKnowledge : unsupported method';
+      } 
     }
 
     var evalPopOpts = {

### 2014-01-22 js/gui.js
mmas věci základ

@@ -377,7 +377,7 @@ function mkGUI (containerId) {
         case 'stats'   : 
           var stats = content;
           graphs.handleStats(stats);
-          Phenotype.update(pheno, $phenoUpdateEl, stats.best_jsStr, stats.best, stats.runKnowledge);
+          Phenotype.update(pheno, $phenoUpdateEl, stats.best_jsStr, stats.best, stats.runKnowledge, opts);
           resize();
           break;
         case 'result' : 

### 2014-01-22 js/phenotype.js

@@ -39,13 +39,15 @@ var Phenotype = function () {
 
   var theGreen = '#5CB85C';
 
-  function update (pheno_, $el, best_jsStr, fitVal, runKnowledge) {
+  function update (pheno_, $el, best_jsStr, fitVal, runKnowledge, opts) {
 
     var __result;
     with(ctxToWithobj(ctx)){
       eval('__result=' + best_jsStr );
     }
 
+    var greenBar = opts.greenBar || 1;
+    
     pheno = pheno_;
     var $link = $('<div>').attr({
       //'data-toggle':"tooltip",
@@ -60,7 +62,7 @@ var Phenotype = function () {
         
     }).html($('<div>').css({
       'background-color': '#F5F5F5',
-      'height': (100-fitVal*100).toFixed(0) + '%'
+      'height': ( (100-greenBar*fitVal*100) ).toFixed(0) + '%' 
     }))
     .click(function(){
       pheno.update($el, __result, best_jsStr, runKnowledge);

### 2014-01-22 js/problems/TSP.js

@@ -15,29 +15,33 @@ opts = {
   numGens: 201,
   popSize: 10,
 
-  saveBest : !true,
+  saveBest : true,
   
-  statsOpts: StatsOpts.default,
+  //statsOpts: StatsOpts.default,
   logOpts  : LogOpts.default,
 
   init: function () {
     
-    var data = TSPutils.tsp225; //TSPutils.Ulysses22; //16; //.EU4,
+    var data = TSPutils.tsp225; //TSPutils.tsp225; //TSPutils.Ulysses22; //16; //.EU4,
 
     this.solver = ACO.mkSolver(
       TSPutils.mkTspProblem({
-        method: 'AS',
-        Q: data.optVal,//6792.12,
-        initTauVal: 1,
+        method: 'MMAS',
+        Q: 1, //data.optVal,//6792.12,
+        initTauVal: 0.001,
         from: '1', // 'Praha'
         data: data, 
         //dists: TSPutils.EU4_dists,
       },{
         rho  : 0.15 ,
-        alpha: 1.2  ,
-        beta : 1.2   
+        alpha: 1  ,
+        beta : 5  ,
+        maxTau: 0.03,
+        minTau: 0.0001
       })
     );
+
+    this.greenBar = data.optVal;
 
     this.fitness   = this.solver.antProblem.fitness;
     this.operators = [[this.solver.Operators.generatePop, 1.0]];
@@ -118,8 +122,25 @@ opts = {
           tauGraph: runKnowledge
         });
 
-      },  
+      },
+
+
+
     };
+  },
+
+  statsOpts: {
+    drawStep: 1,
+    graphs: {
+      fitness: {
+        //yMax: 1.0,
+        vars: {
+          best:  { color: 'green', avg: true, minmax: true }, 
+          avg:   { color: 'blue',  avg: true, minmax: true },
+          worst: { color: 'red',   avg: true, minmax: true }
+        }
+      }
+    }
   },
 
 }

### 2014-01-22 js/solvers/ACO.js

@@ -40,12 +40,14 @@ var ACO = (function () {
 
   function mkPath (antProblem, tau) {
 
+    //log(tau);
+
     var succsFun = antProblem.succsFun;     
     var heur     = antProblem.heur; 
     var antOpts  = antProblem.opts;   
 
     var isGoal = antProblem.isGoal;   
-    var next   = antProblem.from;
+    var next   = _.sample(_.keys(tau)); //antProblem.from;
     var path   = [];
 
     while( !isGoal(path) && next !== null ){
@@ -97,7 +99,7 @@ var ACO = (function () {
       // vypařování
       for (i in oldTau) {
         if (newTau[i] === undefined) {newTau[i] = {};}
-        for (j in oldTau) {
+        for (j in oldTau[i]) {
           newTau[i][j] = oldTau[i][j] * (1-rho); 
         }
       }
@@ -124,11 +126,55 @@ var ACO = (function () {
       return newTau;
     }
 
+    function update_MMAS (runKnowledge, evaledPop) {
+      //TODO !!!
+      
+      var antOpts = antProblem.opts;
+      var rho     = antOpts.rho; 
+
+      var maxTau  = antOpts.maxTau;
+      var minTau  = antOpts.minTau;
+
+      function checkTau (tau) {
+        if (tau < minTau) {return minTau;}
+        if (tau > maxTau) {return maxTau;}
+        return tau;
+      }
+      
+      var oldTau = runKnowledge;
+      var newTau = {};
+
+      var i,j;
+
+      for (i in oldTau) {
+        if (newTau[i] === undefined) {newTau[i] = {};}
+        for (j in oldTau[i]) {
+          newTau[i][j] = checkTau( oldTau[i][j] * (1-rho) ); 
+        }
+      }
+
+
+      var bestPath   = evaledPop.best.indiv.term;
+      var bestFitVal = evaledPop.best.fitVal;
+      var pathLen    = bestPath.length;
+
+      for (var s = 0; s < pathLen-1; s++) {
+
+        i = bestPath[s];
+        j = bestPath[s+1];
+                                                                // TODO : (!!!)
+        newTau[i][j] = checkTau( newTau[i][j] + bestFitVal );   // viz otázky z update_AS
+      }      
+
+      return newTau;
+    }
+
+
     function updateRunKnowledge (runKnowledge, evaledPop, opts) {
       switch (antProblem.method) {
[…]

### 2014-01-22 js/worker.js

@@ -1,7 +1,7 @@
 var $ = function (x){}
 
 importScripts(
- 'libs/underscore-1.5.1.js',
+ 'libs/underscore-1.5.2.js',
  'libs/peg-0.7.0.js',
  'libs/wu-0.1.8.js', 
  'libs/geo.js',

### 2014-01-24 tex/gecco14/gecco14.tex
gecco

--- 2014-01-20 tex/cec14/article.tex
@@
-\documentclass[conference]{IEEEtran}
+\documentclass{sig-alternate}
 
-%\usepackage{stmaryrd}
-%\usepackage{amsfonts}
-%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
-%\usepackage{hyperref}
-%\usepackage{amsmath}
-%\usepackage{amssymb}
-%\usepackage[utf8]{inputenc}
-%\usepackage{qtree}
-%\usepackage{xspace}
-%\usepackage[ampersand]{easylist}
-
-%\usepackage{times}
 \usepackage[utf8]{inputenc}
-\usepackage{amsfonts}
-%\usepackage{qtree}
 \usepackage{xspace}
 
-%\usepackage{makeidx}  % allows for indexgeneration
-%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
-%\usepackage{hyperref}
-\usepackage{amsmath}
-%\usepackage{amssymb}
-\usepackage{graphicx}
-
-\usepackage[ampersand]{easylist}
-\usepackage[usenames,dvipsnames]{color}
-
-%\usepackage{amsthm}
 \newtheorem{definition}{Definition}
@@
 
-\hyphenation{op-tical net-works semi-conduc-tor IEEEtran}
-
-\IEEEoverridecommandlockouts    % to create the author's affliation portion
-
-\textwidth 178mm    % <------ These are the adjustments we made 10/18/2005
-\textheight 239mm   % You may or may not need to adjust these numbers again
-\oddsidemargin -7mm
-\evensidemargin -7mm
-\topmargin -6mm
-\columnsep 5mm
-
 \begin{document}
 
-% paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
-\title{\ \\ \LARGE\bf 
-Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
+\conferenceinfo{GECCO'14,} {July 12-16, 2014, Vancouver, BC, Canada.}
+\CopyrightYear{2014}
+\crdata{TBA}
+\clubpenalty=10000
+\widowpenalty = 10000
 
-\author{Tom\'{a}\v{s} K\v{r}en \and Roman Neruda}
+\title{Utilization of Reductions and Abstraction Elimination in Typed Genetic Programming}
+
+\numberofauthors{2}
+
+\author{
+% 1st. author
+\alignauthor
+Tom\'{a}\v{s} K\v{r}en\\
+  \affaddr{Faculty of Mathematics and Physics}\\
+  \affaddr{Charles University in Prague}\\
+  \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
+  \affaddr{Prague, Czech Republic}\\
+  \email{tomkren@gmail.com}
+% 2nd. author
+\alignauthor
+Roman Neruda\\
+  \affaddr{Institute of Computer Science}\\
+  \affaddr{Academy of Sciences of the Czech Republic}\\
+  \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
+  \affaddr{Prague, Czech Republic}\\
+  \email{roman@cs.cas.cz}
+}
+
+%\date{24 January 2014}
 
 \maketitle
+\begin{abstract}
+Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions.
+In this paper we discuss advantages of this representation coming from use of reductions (beta and eta) and how to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm.
+We discuss the role of those reductions in the process of generating initial population and compare several crossover approaches including novel approach to crossover operator based both on reductions and abstraction elimination.
+The design goal of this operator is to turn the disadvantage of abstraction elimination  - possibly quadratic increase of program size - into a virtue; our approach leads to more crossover points.
+At the same time, utilization of reductions provides offspring of small sizes.
+\end{abstract}
 
-\begin{abstract}
-Tree based genetic programming traditionally uses simple S-expressions to represent programs, however more expressive representations can exhibit better results and are well suited for typed genetic programming.
-One of such representations is lambda calculus.
-In this paper, a generalization of genetic programming 
-for simply typed lambda calculus is presented. We mainly focus on
-population initialization methods. 
-Three such initialization methods are proposed 
-depending on different search strategies. 
-The first strategy corresponds to standard ramped half-and-half initialization, 
-the second one corresponds to exhaustive systematic search, and the third one 
-represents a novel geometric strategy. 
-Three well-known benchmark experiments (simple symbolic regression,
-artificial ant problem and even parity problem)
-support that the geometric strategy outperforms
-the standard generating method in success rate, best fitness value,
-time consumption and average individual size. 
-Despite the fact that the presented system is designed for typed genetic 
-programming, when given standard genetic programming problem satisfying
-standard closure requirement it produces simple S-expressions and
-the novel geometric strategy outperforms the standard system. 
-This claim is supported by that first two experiments both satisfying 
-the closure requirement.
-The fact that a search strategy can be easily described by a rather simple filtration algorithm opens opportunity to easy exploration of more 
-population initialization methods, possibly in a kind of meta-evolution way.
-\end{abstract}
-% no key words
+% A category with the (minimum) three required fields
+\category{H.4}{TODO}{Miscellaneous}
+%A category including the fourth, optional field follows...
+\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
+
+\terms{TODO}
+
+\keywords{TODO, ToDo, todo}
 
 \section{Introduction}
-\PARstart{G}{enetic} programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
+
+Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
 Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
@@
 First, \lets describe a programming language, 
-in which the GP algorithm generates individual programs --- the so called \lterms.\\
+in which the GP algorithm generates individual programs --- the so called \lterms.
 
@@
 \end{definition}
-~\\
 
@@
 they were recently introduced to C++, and they are expected to be supported in Java 8.}.
-$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for $(\dots((M_1~M_2)~M_3)~\dots~M_n)$
-and $\lam{x_1 x_2 \dots x_n }{M}$ for $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
+
+For better readability, 
+$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for
+$(\dots((M_1~M_2)~M_3)~\dots~M_n)$
+and $\lam{x_1 x_2 \dots x_n }{M}$ for 
+$\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
 
@@
 corresponds to a program expression with no type information
-included. Now we will describe \textit{types} (or \textit{type terms}).\\
+included. Now we will describe \textit{types} (or \textit{type terms}).
 
@@
 The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
-combining the previously defined \textit{\lterms} and \textit{types} together. \\
+combining the previously defined \textit{\lterms} and \textit{types} together.
 
@@
 such that $\Gamma$ does not contain any declaration with $x$ as subject.
-We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.\\
+We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
 
@@
 for a given pair $\ul{\tau}{\Gamma}$
-such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.\\
+such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
 
@@
 
-\subsection{Introduction}
-
-Our approach to \lterm gerating is based on technique 
-briefly described in \cite{barendregt10}, which generates
-well-typed \lterms in their long normal form. 
-We use this technique to perform systematic exhaustive enumeration
-of \lterms in their long normal form in order from smallest to largest.
-We use well known \textit{A* algorithm} \cite{AIMA} for this task.
-A* is used to search in a given state space for a goal state. 
-It finds the optimal solution (in our case the smallest term)
-and uses "advising" heuristic function.
-It maintains a priority queue to organize states yet to be explored.
-Initially this queue contains only the initial state.
-
-Our state space to search in is the space of unfinished \lterms. 
-The initial state is the unfinished term $\ul{\tau}{\Gamma}$, 
-where $\tau$ is the desired type of
-terms to be generated and $\Gamma$ is the context
-representing the set of building symbols to be used in construction of
-terms (it corresponds to the set $T \cup F$ in
-standard GP enriched with types). The process of determining 
-successors of a state described below is designed so it constructs well-typed 
-\lterms and omits no \lterm in its long normal form. 
-A state is considered a goal state if it contains no unfinished
-leaf, i.e., it is a finished \lterm.
-
-Our generating method is based on simple modification of the
-standard A*, which we call \textit{forgetful A*}. This modification consist in 
-additional parameter for the A* algorithm -- the \textit{search strategy}. 
-It is a simple filtration function (along with initialization procedure)
-that is given the set of all successors of the state that is being examined
-and returns a subset of this input. This subset is added to the priority queue 
-to be further explored. In this way the search space may be reduced as 
-the filtration function may \textit{forget} some successors.
-If the queue becomes empty before the desired number of \lterms
-is generated, then the initial statete is inserted to the queue
-and the process continues. For the standard A* this would be meaningless,
-but since our A* is forgetful this kind of restart makes sense.
-
-A* keeps a priority queue of states during the generation process,
-on the other hand the \textit{ramped half-and-half method}, 
-the standard GP algorithm for generating individuals, 
-keeps only one individual which is gradually constructed. This 
-behavior is easily achieved by use of suitable search strategy 
-that returns subset consisting of only one successor.
-The systematic search is obtained by search strategy that 
-returns whole input set.      
-Our novel \textit{geometric strategy} can be understood as
-point somewhere between those two extremes.
-
-\subsection{Algorithm}
-
-The inputs for the term generating algorithm are following.
-\begin{enumerate}
- \item Desired type $\tau$ of generated terms.
- \item Context $\Gamma$ representing set of building symbols.
- \item Number $n$ of terms to be generated.
- \item Search strategy $S$. 
-\end{enumerate}
-
-Essential data structure of our algorithm 
-is priority queue of unfinished terms. 
-Priority of an unfinished term is given by its size\footnote{
-A* heuristic function is hidden in method of computing
-size of unfinished leafs $\ul{\tau}{\Gamma}$. Our algorithm uses
-trivial estimate $\vert\ul{\tau}{\Gamma}\vert = 1$ which is trivially admissible.
-This heuristic is not as silly as it might seem since it is
-quite usual to have $x$ such that $x : \tau \in \Gamma$.
-Since the true value of $\vert\ul{\tau}{\Gamma}\vert$ depends only on
-$\tau$ and \textit{types} in $\Gamma$ (the \textit{signature}), 
-no matter how many variables/constants of
-each type there are, it is should by pretty effective to compute this
-value precisely and store them for later. 
-%This topic has other interesting 
-%subtopics (such as surprisingly small 
-%number of such "signatures" for a given problem), 
-%but it is unfortunately beyond the scope of this paper.
-}.
-At the beginning, the queue contains only one unfinished term; 
-$\ul{\tau}{\Gamma}$. The search strategy $S$ also 
-initializes its internal state (if it has one).
-
-At each step, the term $M$ with the smallest size
-is pulled from the queue.
-According to the number of unfinished leafs in $M$ one of
-the following actions is performed.
-\begin{enumerate}
- \item If the term $M$ has no unfinished leaf (i.e., it is a finished
- term satisfying \mbox{$\tur{\Gamma}{M}{\tau}$}), then it is added to the
- result set
-% \footnote{
-% \red{TODO: říct že tim že je to set se řeší problém toho, aby se tam nevyskytovali
-% některý vygenerovaný víckrát, pokud však toto nechcem můžem použít jinou collection
-% , případně to vůbec nezminovbat když by to kradlo moc místa
-% }} 
- of generated terms.   
- \item Otherwise, \textit{successors} of the unfinished term $M$ are
-       filtered out by \textit{search strategy} $S$ and
-       those successors that outlast the filtration 
-       are inserted into the queue.
-\end{enumerate}
-
-\textit{Successors} of an unfinished term $M$ are obtained by 
-\textit{expansion} of the \mbox{\textit{DFS-first}} unfinished leaf $L$
-(i.e., the leftmost unfinished leaf of $M$).
-
-Expansion of the selected unfinished leaf $L$ leads to creation of 
-one or many (possibly zero) successors.
-In this process, $L$ is replaced
-by a new subterm defined by the following rules\footnote{
-For the sake of simplicity it is presented as two separate rules. 
-Since the first rule results in exactly one successor it is smarter
-to combine those two rules into one resulting in that unfinished leafs
-have only atomic types. At the beginning we transform the initial type into atomic by first rule (if necessary). After that
-only second rule is applied, but if it results in creation of some unatomic 
-$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }.
-\begin{enumerate}
- \item If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
- 	   where $\alpha$ is atomic type and $n \geq 1$, 
-       then $L$ is replaced by 
-       $\lamb{x_1 \dots x_n}{\ul{\alpha}
-       {\Gamma,x_1 \colon \rho_1,\dots,x_n \colon \rho_n}}$.
-       Thus this expansion results in exactly one successor.  
- \item If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type,
-       then for each 
-       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}
-       the unfinished leaf $L$ is replaced by 
-       $(~f~\ul{\tau_1}{\Gamma}~\dots~\ul{\tau_m}{\Gamma}~)$.
-       Thus this expansion results in many (possibly zero or one) successors.
-\end{enumerate}
-
-Now that we have all possible successors of $M$, we are about to apply
-the \textit{search strategy} $S$. A search strategy is a procedure
-which takes as input a set of unfinished terms and returns a subset
-of the input set. Therefore, search strategy acts as a filter reducing 
-the search space.
-
-If the queue becomes empty before the desired number $n$ of terms
-is generated, then the initial unfinished term $\ul{\tau}{\Gamma}$ 
-is inserted to the queue, search strategy $S$
-again initializes its internal state and the process continues.
-
-\Lets now discuss three such search strategies.\\
-
-%\subsubsection{Systematic strategy}
-\textit{Systematic strategy}:~~ 
-If we use trivial strategy that returns all the inputs, 
-then the algorithm systematically generates 
-first $n$ smallest lambda terms in their
-\textit{long normal form}.\\
-
-%\subsubsection{Ramped half-and-half strategy}
-\textit{Ramped half-and-half strategy}:~~
-This is generalization of the standard ramped half-and-half method described 
-in \cite{koza92}. If applied to context satisfying closure requirement
-(all constants/variables are of the same type, all functions are operations
-on this type), it will behave in the same way as the standard method.
-The internal state of this strategy consists of two variables.
-It is the only one strategy described here that uses an internal state.
-
-\begin{enumerate}
- \item \textit{isFull} - A boolean value, determining whether \textit{full}
-                     or \textit{grow} method will be performed.
- \item \textit{d} - An integer value from $\setDots{2}{D_{init}}$, where 
-                $D_{init}$ is predefined maximal depth (e.g. 6).                    
-\end{enumerate}
-
-This strategy returns precisely one randomly
-(uniformly) selected  element from 
-the \textit{selection subset} of input set
-(or zero elements if the input set is empty). 
-The \textit{selection subset} 
-to select from is determined by $depth, d$ and $isFull$.
-The $depth$ parameter is the depth (in the term tree) 
-of the unfinished leaf that was expanded.
-Those elements of input set whose newly added subtree contains one ore more 
-unfinished leafs are regarded as \textit{non-terminals}, whereas 
-those whose newly added subtree contains no unfinished leaf are regarded as 
-\textit{terminals}.
-If $depth = 0$, then the subset to select from is  
-set of all \textit{non-terminals} of the input set.
-If $depth = d$, then the subset to select from is
-set of all \textit{terminals} of the input set.
-In other cases of $depth$ it depends on value of $isFull$.
-If $isFull = true$, then the subset to select from is 
-set of all \textit{non-terminals} of the input set.
-If $isFull = false$, then the subset to select from is 
-the whole input set.\\
-
-%\subsubsection{Geometric strategy}
-\textit{Geometric strategy}:~~
-We can see those two previous strategies as two extremes on the spectrum of 
-possible strategies. 
-\textit{Systematic strategy} filters no successor state thus performing
-exhaustive search resulting in discovery of $n$ smallest terms in one run.
-On the other hand, \textit{ramped half-and-half strategy} filters 
-all but one successor states resulting in degradation of 
-the priority queue into "fancy variable".
-\textit{Geometric strategy} is simple yet fairly effective term generating 
-strategy somewhere in the middle of this spectrum.
-It is parameterized by parameter $q \in (0,1)$, its default well-performing 
-value is $q = 0.75$.
-For each element of the input set 
-it is probabilistically decided whether
-it will be returned or omitted. A probability $p$ of returning is
-same for all elements, but depends on the $depth$, 
-which is defined in the same way as in previous strategy. 
-It is computed as follows.
-$$ p = q^{depth} $$
-This formula is motivated by idea that it is important to
-explore all possible root symbols, but as the $depth$ 
-increases it becomes less "dangerous" to omit 
-an exploration branch. 
-We can see this by considering that this strategy results in
-somehow forgetful A* search.
-With each omission we make the search space smaller. But with
-increasing depth these omissions have smaller impact on the search space,
-i.e., they cut out lesser portion of the search space.
-Another slightly esoteric argument supporting this formula is that "root 
-parts" of a program usually stand for more crucial parts
-with radical impact on global behavior of a program, 
-whereas "leaf parts" of a program usually
-stand for less important local parts (e.g. constants).  
-This strategy also plays nicely with the idea that 
-"too big trees should be killed".
-
-~\\
-Furthermore, our system utilizes generalization of the standard 
-tree-swapping crossover operator. Since it is beyond the scope of this 
-paper we mention it only briefly. Two main concerns with swapping typed subtrees 
-are types and free variables. Well-typed offspring is obtained 
-by swapping only subtrees of the same type. Only subtrees with 
-corresponding counterpart in the second parent are randomly chosen from.
-More interesting problem lies in free variables, which may cause trouble
-if swapped somewhere where it is suddenly not bounded. In order to circumvent this
-difficulty we utilize technique called \textit{abstraction elimination}\cite{jones87}
-that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions
-and no bound variables. After the initial population is generated,  it is transformed
-by abstraction elimination. Another possible transformation taking place
-after initialization is $\eta$-normalization shortening 
-rather long long normal form into $\beta\eta$-normal form.  
-Another performence enhancing transformation is use of "applicative" tree 
-representation (coming directly from inductive definition of \lterms).
-
 \section{Experiments}
 \label{experiments}
-
-We made tree experiments comparing performance of standard 
-\textit{ramped half-and-half strategy} with our
-\textit{geometric strategy} (with default parameter $q=0.75$). 
-Each experiment consisted of 50 independent runs 
-of GP algorithm. Each run had maximally 51 generations and population size
-was 500 individuals.
-In those experiments we analyze the fitness of the best individual,
-the average size of an individual,
-the ability of the system to produce a correct solution 
-and computational cost estimates throughout generations. 
-For the last two metrics we use the popular measurement  
-methods within GP field --- the \textit{performance curves}
-described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
-and $I(M,i,z)$ (the total number of individuals that must be processed 
-to yield a correct solution with probability $z =99\%$).
-Difference of those two generating methods 
-is statistically analyzed by the Welch t-test
-by comparing the fitness values of the best individuals of the run. 
-Graphs showing mean values of the fitness of the best individual and
-the average size of an individual throughout generations
-contain error bars representing the standard error of the mean (SEM).
-
-\red{Přestože je naše generujicí metoda navržena obecně pro typovanou verzi,
-tak pracuje dobře i pro problémy standardního GP splňující closure podmínku.
-Pro taklovéto problémy strategie generuje standardní S-výrazy. Proto jsou
-první dva ze tří experimentů právě takového charakteru. Dalším důvodem pro volbu 
-takovýchto problémů je fakt, že výsledky pokusů výrazně ovlivňuje použitý křížící 
-operátor. Návrh obecného křížícího operatoru pro GP v symply typed lc je otevřený problém. Protože křížící operátor použitý v našem systému je poměrně komplikovaný
-a tím pádem beyond the scope of this paper, ale pro problémy splnující
-clousure podmínku se chová přesně stejně jako standardní křížící operator,
-tak se to zdá jako další důvod proč 2/3 experimentů udělat na closed problémech.}
-
-\subsection{Simple symbolic regression}
-\textit{Simple Symbolic Regression} is a problem described
-in \cite{koza92}. Objective of this problem is to 
-find a function $f(x)$ that fits a sample
-of twenty given points. The target function is 
-function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
-Desired type of generated programs $\tau$ and 
-building blocks context $\Gamma$ are following.
-\begin{align*}
-\tau = \Real \ar &\Real\\
-\Gamma = \{
-  (+)  &: \Real \ar \Real \ar \Real    ,
-  (-)   : \Real \ar \Real \ar \Real    ,
-  (*)   : \Real \ar \Real \ar \Real    ,\\
-  rdiv &: \Real \ar \Real \ar \Real    ,
-  sin   : \Real \ar \Real              ,
-  cos   : \Real \ar \Real              ,\\
-  exp  &: \Real \ar \Real              , 
-  rlog  : \Real \ar \Real              \}
-\end{align*}
-where
-
-\noindent
-\begin{minipage}{.5\linewidth}
-\begin{align*}
-rdiv(p,q) &= \begin{cases} 1 &\mbox{if } q = 0 \\
-p/q & \mbox{otherwise } \end{cases}  
-\end{align*}
-\end{minipage}%
-
-\begin{minipage}{.5\linewidth}
-\begin{align*}
-rlog(x) &= \begin{cases} 0 &\mbox{if } x = 0 \\
-log(\vert x\vert) & \mbox{otherwise}. \end{cases}
-\end{align*}
-\end{minipage}
-
-~\\Fitness function is computed as follows.
-
-$$ fitness(f) =  \frac{1}{1+ \sum\limits_{i=1}^{20}{ \vert f(x_i)-y_i }\vert }  $$
-where $(x_i,y_i)$ are 20 data samples from $[-1,1]$, such that $y_i = f_t(x_i)$.
-An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
-considered as a correct individual.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.14]{imgs/cec/SSR.PNG}
-  \caption{Graphs for Simple symbolic regression.}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\begin{tabular}{|l|cc|}
-\hline
-& Ramped half-and-half & Geometric \\
-\hline
-Mean & 0.7999030182 & 0.8174067512 \\
-SD	 & 0.1521369287	& 0.1826109800 \\
-SEM	 & 0.0215154108	& 0.0258250925 \\
-N	 & 50 & 50  \\
-\hline
-t-value & 0.5207 & $\alpha = 0.05$\\
-p-value & 0.6038 & \textbf{Not statistically significant.}\\
-\hline
-\end{tabular}
-\caption{Statistical analysis - Simple symbolic regression.}
-\end{table}
-
-~\\
-
-Figure~1 shows results of this experiment. 
-Table~I summarizes statistical analysis
-of the fitness values of the best individuals of the run.  
-By conventional criteria, this difference is considered to be not statistically significant.
-Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
-The average individual size for generation 50 was 68.8.
-The experiment took 46 minutes.
-Our \textit{geometric strategy} scored 21/50 (42\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
-The average individual size for generation 50 was 44.1.
-The experiment took 26 minutes.
-So for the \textit{Simple symbolic regression}
-our geometric strategy is slightly more successful then 
-ramped half-and-half strategy in these observed metrics.
-
-\subsection{Artificial ant}
-\textit{Artificial Ant} is another problem described
-in \cite{koza92}. Objective of this problem is to 
-find a control program for an artificial ant so
-that it can find all food located on "Santa Fe" trail.
-The Santa Fe trail lies on toroidal square grid.
-The ant is in the upper left corner, facing right.
-The ant is able to move forward, turn left, and sense if a food 
-piece is ahead of him.
-
-\begin{align*}
-\tau = An&tAct\\
-\Gamma = \{~~
-  l    &: AntAct                              ,
-  r    : AntAct                               ,
-  m    : AntAct                               ,\\
-  ifa  &: AntAct \ar AntAct \ar AntAct  ,\\
-  p2   &: AntAct \ar AntAct \ar AntAct  ,\\
-  p3   &: AntAct \ar AntAct \ar AntAct \ar AntAct  \}
-\end{align*}
-
-Action $l$ turns the ant left. 
-Action $r$ turns the ant right.
-Action $m$ moves the ant forward.
-Action $ifa~x~y$ (if-food-ahead) performs action $x$ 
-if a food piece is ahead of the ant,
-otherwise it performs action $y$.
-Action $p2~x~y$ first performs action $x$ and after it action $y$.
-Action $p3~x~y~z$ first performs action $x$, 
-after that action $y$ and finally $z$.
-Actions $l, r$ and $m$ each take one time step to execute.
-Ants action is performed over and over again until it reaches predefined
-maximal number of steps.
-
-Fitness value is equal to number of eaten food pieces.
-An individual such that eats all 89 pieces of food is 
-considered as a correct solution.
-This limit is set to be 600 time steps\footnote{
-In \cite{koza92} this limit is said to be 400 time steps.
-But there is also mentioned following solution, 
-which is described as correct solution:
-\texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
-This program needs 545
-time steps; if it is given only 400 time steps, then it eats only 79 pieces
-of food. Thus we use 600 time steps. 
-}.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.14]{imgs/cec/ANT.PNG}
-  \caption{Graphs for Artificial ant problem.}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\begin{tabular}{|l|cc|}
-\hline
-& Ramped half-and-half & Geometric \\
-\hline
-Mean & 66.20 & 79.78 \\
-SD	 & 12.33 & 8.77  \\
-SEM	 & 1.74	 & 1.24  \\
-N	 & 50    & 50    \\
-\hline
-t-value &  6.3479           & $\alpha = 0.05$\\
-p-value &  \textless 0.0001 &  \textbf{Extremely statistically significant.}\\
-\hline
-\end{tabular}
-\caption{Statistical analysis - Artificial ant problem.}
-\end{table}
-
-Figure~2 shows results of this experiment. 
-Table~II summarizes statistical analysis
-of the fitness values of the best individuals of the run.  
-According to this analysis our strategy performed better;
-by conventional criteria, this difference is considered to be extremely statistically significant.
-
-Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
-The average individual size for generation 50 was 279.5.
-The experiment took 265 minutes.
-
-Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
-The average individual size for generation 50 was 94.1.
-The experiment took 107 minutes.
-
-This is a big improvement in all watched factors.
-
-\subsection{Even parity problem}
-
-Previous two problems were instances of standard GP, i.e. their building 
-symbols obeyed the closure requirement. This third experiment will 
-break the closure requirement, thus ideal for testing typed GP techniques. 
-The even-parity function is a boolean function taking as inputs $N$
-boolean values and returning \textit{True} if an even number of inputs 
-are \textit{True}. For odd number it returns \textit{False}.
-This problem has been used by many researchers
-as a benchmark for GP. We compare our results with 
-that obtained by Yu in \cite{yu01},
-where is presented approach to evolve recursive and modular programs
-by use of higher-order functions and $\lambda$-abstractions.
-We use very similar set of \textit{building symbols} as in \cite{yu01}. 
-\begin{align*}
-\tau = [Boo&l] \ar Bool\\
-\Gamma = \{
-  and   &: Bool \ar Bool \ar Bool                              ,\\
-  or    &: Bool \ar Bool \ar Bool                              ,\\
-  nand  &: Bool \ar Bool \ar Bool                              ,\\
-  nor   &: Bool \ar Bool \ar Bool                              ,\\
-  foldr &: (Bool \ar Bool \ar Bool) \\
-        &~~~\ar Bool \ar [Bool] \ar Bool ,\\
-  head' &: [Bool] \ar Bool                                   ,\\
-  tail' &: [Bool] \ar [Bool]                              \}
-\end{align*}
-
-The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
-this problem is considered atomic.
-Unlike in \cite{yu01}, we use specific instance of polymorphic 
-function \texttt{foldr}. 
-And modifications of functions \texttt{head} 
-(returning the first element of the list) 
-and \texttt{tail} (returning the list without the first element) are used; 
-making them total by returning default value \textit{False}
-and \texttt{[]}, respectively.
-We use the same fitness function as in \cite{yu01}. 
-The fitness function examines the individual by giving
-it all possible boolean lists of length 2 and 3.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.14]{imgs/cec/EP.PNG}
-  \caption{Graphs for Even parity problem.}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\begin{tabular}{|l|cc|}
-\hline
-& Ramped half-and-half & Geometric \\
-\hline
-Mean & 0.895000	& 0.930000 \\
-SD	 & 0.075170	& 0.078490 \\
-SEM	 & 0.010631	& 0.011100 \\
-N	 & 50    & 50    \\
-\hline
-t-value &  2.2772 & $\alpha = 0.05$\\
-p-value &  0.0250 &  \textbf{Statistically significant.}\\
-\hline
-\end{tabular}
-\caption{Statistical analysis - Even parity problem.}
-\end{table}
-
-Figure~3 shows results of this experiment. 
-Table~III summarizes statistical analysis
-of the fitness values of the best individuals of the run.  
-According to this analysis our strategy performed better;
-by conventional criteria, this difference is considered to be 
-statistically significant.
-
-Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
-The average individual size for generation 50 was 47.7.
-The experiment took 28 minutes.
-Our \textit{geometric strategy} scored 32/50 (64\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
-The average individual size for generation 50 was 74.3.
-The experiment took 33 minutes.
-
-Once again this is a big improvement against standard method.
-
-One may get the impression that geometric strategy suffers from bloat here,
-but considering that it performs significantly better and
-that term size around 100 is similar as for the previous two problems, 
-it seems to us that ramped half-and-half maybe has difficulties with 
-constructing more complex terms, and therefore is "staying small".
-Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
-the 100\% correct solution was present in the generation 0. This is pretty good
-result for uninformed search. However, our results are less successful 
-then those presented in \cite{yu01} (with slightly different set of 
-building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
-17,500 in generation 4. But by observation of typical solution which is often
-some modification of \texttt{foldr xor False inputList} one sees that important 
-task for GP here is to create \texttt{xor} function. This is advantage for the 
-more specialized term representation used in \cite{yu01} 
-(restricting use of outer variables). 
-The difference in crossover operators may also be involved in the difference, 
-but such analysis is beyond the scope of this paper. Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
-scored 17/60 (28\%); min $I(M,i,z) = 220,000$. \textit{GP with ADFs} 
-scored 10/29 (34\%); min $I(M,i,z) = 1,440,000$.
 
@@
 
-In this work, we have proposed a generalization of genetic programming 
-for simply typed lambda calculus. The main focus of this paper is on individual generation algorithms. 
-The main idea is to explore the spectrum of available approaches bounded on one end by the traditional
-ramped half-and-half strategy, while the second bound represents an exhaustive systematic search. 
-Thus, three population initialization methods have been designed, 
-depending on different search strategies. 
-The first strategy corresponds to the above mentioned traditional genetic programming initialization,  
-the second one corresponds to exhaustive search, and the third one 
-represents a novel geometric strategy which outperforms standard genetic 
-programming in success rate, resource consumption and average individual size 
-on the presented experiments.
+%ACKNOWLEDGMENTS are optional
+\section{Acknowledgments}
+This section is optional; it is a location for you
+to acknowledge grants, funding, editing assistance and
+what have you.  In the present case, for example, the
+authors would like to thank Gerald Murray of ACM for
+his help in codifying this \textit{Author's Guide}
+and the \textbf{.cls} and \textbf{.tex} files that it describes.
 
-%\red{ [větu dvě o experimentech]}
+%
+% The following two commands are all you need in the
+% initial runs of your .tex file to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+\bibliography{evogp}  % sigproc.bib is the name of the Bibliography in this case
+% You must have a proper ".bib" file
+%  and remember to run:
+% latex bibtex latex latex
+% to resolve all references
+%
+% ACM needs 'a single self-contained file'!
+%
 
-The relevance of these algorithms does not effect only initialization, but it can be used during mutation
-operator in a straightforward way. 
-%Other performance enhancements based on theory of lambda calculus are 
-%proposed and supported by experiments, including abstraction 
-%elimination that enables us to utilize
-%a simple tree-swapping crossover.
-
-Our future research will focus on exploring other possible search strategies positioned in the middle ground 
-identified in this paper. Since the search strategy can be easily described by a rather simple filtration algorithm, it might be
-interesting to allow the typed GP system to evolve the search strategy it 
-utilizes in a kind of meta-evolution way.
-
-\red{Ještě nějak rozšířit.}
-
-%\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }
-
-%\nocite{*}
-
-\bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,evogp}
-
-% that's all folks
+% That's all folks!
 \end{document}

### 2014-01-26 tex/gecco14/gecco14.tex
gecco atd

@@
 \usepackage{xspace}
+\usepackage[ampersand]{easylist}
 
 \newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
 
@@
 \newcommand{\Real}{\mathbb{R}}
+\newcommand{\gar}{\longmapsto}
 
@@
 \keywords{TODO, ToDo, todo}
+
+\section{Plán boje}
+
+\subsection{Jak to celý pojmout?}
+
+Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
+Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).
+
+Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)
+
+\begin{itemize}
+\item Po vy generování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce).
+To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.
+
+\item Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
+Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
+Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y .
+x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
+Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
+Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.
+Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
+\end{itemize}
+
+\subsection{Osnova}
+
+\begin{lizt}
+& Jak reprezentovat stromy programu pro GP?
+  && Reprezentace v klasickym kozovy je S-expression.
+  && Nebo mužem používat kombinátory jako 
+     Briggs a O’Neil (to si myslim je jakoby 
+     hlavní konkurence, vuči který by to chtělo obhájit)
+  && Lambda termy a jejich awesomeness
+& \textbf{Povídaní o redukcích}
+& Povídání o lnf
+  && že lnf je přirozený rozšíření s-exprešnu do lambda 
+     kalkulu vlastnosti termu v lnf 
+     &&& proč je eta redukovat (eta redukcí lnf dostanem 
+         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
+& Generování
+  && Generování gramatikou
+  && Gramatika pro lnf
+  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
+& Problémy s proměnýma a jak je řešit.
+  && Křížit lambda termy i s proměnejma a abstrakcema, 
+     problémy řešit když nastanou (pomluvit a odsoudit)
+  && Zmenšit prostor termu (tim že někerý nejsme schopný 
+     vygenerovat) s kterým operujeme tak aby křížení už 
+     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
+     jen použití proměnných z její hlavy)
+  && Převod hned po vygenerování
+  && Převod až při křížení
+     &&& eliminace abstrakcí
+     &&& skřížim
+     &&& vložený kombinátory nahradim odpovídajícím termem
+     &&& celý to redukuju
+ & Pokusy (?!)
+ & Závěr
+\end{lizt}
 
@@
 
+\subsection{$\beta$-reduction}
+
+In order to perform computation there must be some
+mechanism for term evaluation. In $\lambda$-calculus there
+is $\beta$-reduction for this reason.\\
+
+\newcommand{\bRedex}{$\beta$-redex\xspace}
+\newcommand{\bRedexes}{$\beta$-redexes\xspace}
+\newcommand{\bArrow}{\rightarrow_\beta\xspace}
+\newcommand{\eArrow}{\rightarrow_\eta\xspace}
+\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
+
+A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
+A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
+This fact is written as \textit{relation} $\bArrow$ 
+of those two terms:
+\begin{equation} \label{eq:bRed}
+\lamb{x}{M}N \bArrow M[x:=N]
+\end{equation}
+It is also possible to reduce \textit{subterm \bRedexes} 
+which can be formally stated as:
+\begin{align*}
+P \bArrow Q &\then (R~P)      \bArrow (R~Q) \\
+P \bArrow Q &\then (P~R)      \bArrow (Q~R) \\
+P \bArrow Q &\then \lam{x}{P} \bArrow \lam{x}{Q}  
+\end{align*}
+
+In other words, $\beta$-reduction is the process 
+of insertion of arguments supplied to a function into 
+its body. \\
+
+Another useful relations are $\bbarr$ and $\beq$ defined as follows.
+
+\begin{enumerate}
+ \item \begin{enumerate}
+ 	\item $M \bbarr M$
+ 	\item $M \barr N \then M \bbarr N$
+ 	\item $M \bbarr N , N \bbarr L \then M \bbarr L$ 	
+ \end{enumerate}
+ \item \begin{enumerate}
+ 	\item $M \bbarr N \then M \beq N$
+ 	\item $M \beq N \then N \beq M$
+ 	\item $M \beq N , N \beq L \then M \beq L$
+ \end{enumerate}
+
+\end{enumerate}
+
+We read those relations as follows.
+\begin{enumerate}
+ 	\item $M \bbarr N$ --- "$M$ $\beta$-reduces to $N$."  
+ 	\item $M \barr N$  --- "$M$ $\beta$-reduces to $N$
+ 	      in one step."
+ 	\item $M \beq N$ --- "$M$ is $\beta$-convertible to $N$."	
+ \end{enumerate}
+
+\subsection{$\eta$-reduction}
+
+Similarly as for $\beta$-reduction we can define $\eta$-reduction 
+except that instead of \ref{eq:bRed} we use:
+
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$~
+
+Analogically, a term of a form $\lamb{x}{(M~x)}$ is called 
+\textit{$\eta$-redex}.
+
+Relation $\bearr\;=\;\barr \cup \earr$. 
+(Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)
+
+Similarly as for $\bbarr$ and $\beq$ we can define relations 
+$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.
+
+\subsection{$\eta^{-1}$-reduction}
+
+$\eta^{-1}$-reduction (also called $\eta$-expansion) is 
+the reduction converse to $\eta$-reduction.
+Again it may be obtained by replacing \ref{eq:bRed}, now with:
+
+$$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$
+
+\subsection{Normal forms}
+
+~\begin{easylist}[enumerate]
+& A \lterm is a \textit{$\beta$-normal form} (\bnf) 
+if it does not have a $\beta$-redex as subterm.
+& A \lterm M \textit{has} a \bnf if $M \beq N$
+and $N$ is a \bnf.\\
+\end{easylist}
+A normal form may be thought of as a result of a term evaluation.
+
+Similarly we can define \enf and \benf.
+
+\subsection{Types etc}
+
 A \lterm as described above
@@
 
+Přeformulovat aby se to hodilo do tohodle kontextu....
 Our goal in term generation is to produce terms $M$
@@
 
+\subsection{Long normal form}
+\label{lnf}
+
+%v barendrechtovi je to v definici lnf
+%$\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo $M_m$ 
+%ale to musí bejt určitě překlep...
+
 \begin{definition}
-Let $V$ be infinite countable set of {\it 
-variable names}. Let $C$ be set of {\it constant names}, 
-$V \cap C = \emptyset$.	
-Let $\T$ be set of types.
-Let $\C$ be set of all contexts on ($V \cup C$, $\T$).
-Then $\Lpr$ is set of 
-\textit{unfinished  \lterms} defined inductively as follows.	
+Let \GMS where 
+$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha, n \geq 0$.
+	\begin{enumerate}
+	  \item	
+		Then $M$ is in \textit{long normal form} (\lnf) if following 
+		conditions are satisfied.
+		\begin{enumerate}
+		 \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
+		  (specially for $n = 0$, $M$ is term of the form $f$).
+		 \item Each $M_i$ is in \lnf.
+		\end{enumerate}	
+	  \item 
+	    $M$ has a \lnf if $M =_{\beta\eta} N$ and $N$ is in \lnf.
+	\end{enumerate}
+\end{definition}~
+
+As is shown in \cite{barendregt10}, \lnf has following nice properties.
+
+\begin{proposition}
+If $M$ has a \bnf, 
+%which according to Theorem 2B.4 is always the case, 
+then it also has a unique \lnf, 
+which is also its unique \beenf.
+\end{proposition}
+
+\begin{proposition}
+Every $B$ in \bnf has a \lnf 
+$L$ such that $L \twoheadrightarrow_{\eta} B$.
+\end{proposition}
+
+\subsection{Grammar producing \lterms in \lnf}
+
+In \cite{barendregt10} is shown term generating grammar with 
+following rules (our notation is used, but we will not 
+highlighted terminals anymore).
+
+%\footnote{
+%I was using term generating grammars for term generation
+%on my own before I encountered with \cite{barendregt10},
+%where i happily discovered that Barendregt is using
+%almost identical notation. 
+%But his \lnf grammar was more clever then my grammar
+%that was using grammar rules corresponding to the three basic
+%inference rules. 
+%}
+
 \begin{align*}
-\tau \in \T , \Gamma \in \C &\then \ul{\tau}{\Gamma} \in \Lpr
-\textit{~~~~~~~~(Unfinished leaf)}\\
-x   \in V \cup C  &\then x     \in \Lpr \\
-M,N \in \Lpr   &\then (M~N) \in \Lpr 
-\textit{~~~~~~(Function application)} \\
-x   \in V , M \in \Lpr &\then \lamb{x}{M} \in \Lpr
-\textit{~~~~($\lambda$-abstraction)} 
+( \alpha , \Gamma )  
+&\gar
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+& \textbf{if } \alpha \in A,
+(f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
+\\ 
+( \sigma \rightarrow \tau , \Gamma )  
+&\gar
+(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
+&   
 \end{align*}
-\end{definition}
 
-\textit{Unfinished leaf} $\ul{\tau}{\Gamma}$ 
-stands for yet not specified \lterm of the type $\tau$ 
-build from symbols of $\Gamma$.
+The second rule can be replaced by more effective one.
+\[ 
+( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
+\gar
+(~\lambda~x_1~\dots~x_n~.~
+( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
+~~~~ \textbf{if } n > 0
+\]
+
+This rule packs consecutive uses of the second rule into one use.
+This is valid since the use of the second rule is deterministic;
+it is used if and only if the non-terminal's type is not atomic.
 

### 2014-01-26 tex/gecco14_b/gecco14.tex
gecco14_b

--- 2014-01-26 tex/gecco14/gecco14.tex
@@
 \usepackage[ampersand]{easylist}
+\usepackage{qtree}
+\usepackage{color}
 
@@
 \newcommand{\red}[1]{{\color{red} #1}}
+\newcommand{\nehodi}[1]{{\color{yellow} #1}}
 
@@
 
-\section{Plán boje}
+\section{Introduction}
 
-\subsection{Jak to celý pojmout?}
+Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
+Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
+Use of types naturally opens door to enriching S-expressions,
+the traditional GP representation of individuals, with concepts from
+lambda calculus, which is simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts has shown to be successful \cite{yu01}.
 
+~\\
+
+\begin{lizt}
+& Jak reprezentovat stromy programu pro GP?
+  && Reprezentace v klasickym kozovy je S-expression.
+  && Nebo mužem používat kombinátory jako 
+     Briggs a O’Neil (to si myslim je jakoby 
+     hlavní konkurence, vuči který by to chtělo obhájit)
+  && Lambda termy a jejich awesomeness
+\end{lizt}
+
+~\\
+
+\red{
 Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
@@
 \end{itemize}
+}
 
-\subsection{Osnova}
-
-\begin{lizt}
-& Jak reprezentovat stromy programu pro GP?
-  && Reprezentace v klasickym kozovy je S-expression.
-  && Nebo mužem používat kombinátory jako 
-     Briggs a O’Neil (to si myslim je jakoby 
-     hlavní konkurence, vuči který by to chtělo obhájit)
-  && Lambda termy a jejich awesomeness
-& \textbf{Povídaní o redukcích}
-& Povídání o lnf
-  && že lnf je přirozený rozšíření s-exprešnu do lambda 
-     kalkulu vlastnosti termu v lnf 
-     &&& proč je eta redukovat (eta redukcí lnf dostanem 
-         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
-& Generování
-  && Generování gramatikou
-  && Gramatika pro lnf
-  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
-& Problémy s proměnýma a jak je řešit.
-  && Křížit lambda termy i s proměnejma a abstrakcema, 
-     problémy řešit když nastanou (pomluvit a odsoudit)
-  && Zmenšit prostor termu (tim že někerý nejsme schopný 
-     vygenerovat) s kterým operujeme tak aby křížení už 
-     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
-     jen použití proměnných z její hlavy)
-  && Převod hned po vygenerování
-  && Převod až při křížení
-     &&& eliminace abstrakcí
-     &&& skřížim
-     &&& vložený kombinátory nahradim odpovídajícím termem
-     &&& celý to redukuju
- & Pokusy (?!)
- & Závěr
-\end{lizt}
-
-\section{Introduction}
-
-Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
-Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
-Use of types naturally opens door to enriching S-expressions,
-the traditional GP representation of individuals, with concepts from
-lambda calculus, which is simple yet powerful functional mathematical and programming 
-language extensively used in type theory. Such attempts has shown to be 
-successful \cite{yu01}.
-
-The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
+%\nehodi{The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
 During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
@@
 These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
-This design goal also differentiate our approach from
-the three state of the art proposals for typed GP known to us that are discussed in the following section. 
+This design goal also differentiate our approach from the three state of the art proposals for typed GP known to us that are discussed in the following section.
 Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
-Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
+Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.}
 
@@
 Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.
+
+\subsection{Randombits}
+\red{
+
+- někam zduraznit že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
+
+~\\- pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, protže stačí brát v potaz třídy ekvivalence a ne zamotný termy - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy.
+
+~\\- Dost podstatný ospravedlnění lambda reprezentace je fakt, že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.
+
+~\\- nekam se může dát - since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
+Pro more soffisticated type systems would be situation similar.
+
+~\\- (promaklejší formulaci na velkym kompu zabednena) Komb jak je představuje Briggs a spol jsou husty, přinašej výhody jako 
+jednoduchost atd. ale je husty taky to skusit druhou metodou a to je lambdy pač tyhle argumenty:
+
+~\\- stručnější
+
+~\\- do defaultu vyžadujou jednoduší typovej system 
+(SKI musej bejt polymorfní)
+
+~\\- lidštější
+
+}
 
@@
 
+\subsection{Abstraction elimination}
+
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\Lets describe transformation $\Ae$ performing this 
+process.\footnote{In our implementation we must also deal with types, because 
+our individuals are annotated with types. But since this process
+is fairly straightforward but cumbersome to describe, we will 
+skip explaining it here. Reader interested in seeing it can see 
+the source code of our implementation. }
+\begin{align*}
+\Ae[x]           &= x &\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb &\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
+&\textbf{if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
+&\textbf{if } x \in \FV(M)\\
+&&\vee~ x \in \FV(N)\\
+\end{align*}
+
+This is simple version of this process. More optimized version,
+in the means of the size of resulting term and its performance
+is following one, presented in \cite{jones87}.\\
+
+As is stated in \cite{jones87},
+the biggest disadvantage of this technique is that the translated
+term is often much larger than in its lambda form --- the size of
+the translated term can be proportional to the
+square of the size of the original term.
+
+But the advantage is also tempting --- no need to deal with variables
+and lambda heads.
+
+\section{Term generating and reductions}
+
+\begin{lizt}
+& Povídání o lnf
+  && že lnf je přirozený rozšíření s-exprešnu do lambda 
+     kalkulu vlastnosti termu v lnf 
+     &&& proč je eta redukovat (eta redukcí lnf dostanem 
+         beta-eta-nf a že nemusíme beta redukovat pač se to tim 
+         nerozbyje)
+\end{lizt}
+
 \subsection{Grammar producing \lterms in \lnf}
@@
 
-%\footnote{
-%I was using term generating grammars for term generation
-%on my own before I encountered with \cite{barendregt10},
-%where i happily discovered that Barendregt is using
-%almost identical notation. 
-%But his \lnf grammar was more clever then my grammar
-%that was using grammar rules corresponding to the three basic
-%inference rules. 
-%}
-
 \begin{align*}
@@
 (~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
-& \textbf{if } \alpha \in A,
+\\& \textbf{if } \alpha \in A,
 (f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
@@
 The second rule can be replaced by more effective one.
-\[ 
+$ 
 ( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
-\gar
+\gar \\
 (~\lambda~x_1~\dots~x_n~.~
@@
 ~~~~ \textbf{if } n > 0
-\]
+$
 
@@
 
-\section{Our approach}
-\label{approach}
+\subsection{Benefits of generating \lterms in \lnf}
+\label{benefits}
+
+By generating \lterms in \textit{lnf} we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$, but $M =_{\beta\eta} N$.
+In other words, we avoid generating two programs with different 
+source codes, but performing the same computation.
+
+Every \lterm $M$ such that \GMS has its unique \lnf $L$, 
+for which $L =_{\beta\eta} M$.
+Therefore the computation performed by \lterm $M$ 
+is not omitted, because it is the same computation
+as the computation performed by \lterm $L$. \\
+
+Generating \lterms in \lnf is even better than generating 
+\lterms in \bnf. Since \lnf is the same thing as \beenf,
+every \lterm in \lnf is also in \bnf.
+
+This comes straight from the definition of \beenf, 
+but one can also see it from observing method for generating
+terms in \bnf. As is shown in \cite{barendregt10}, 
+this method is obtained simply by replacing  
+the first grammar rule by slightly modified rule,
+resulting in the following grammar.
+\begin{align*}
+( \pi , \Gamma )  
+&\gar
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+\\& \textbf{if } (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
+\\ 
+( \sigma \rightarrow \tau , \Gamma )  
+&\gar
+(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
+&   
+\end{align*}
+
+The difference lies in that $f$'s type is no longer needed to be fully expanded
+($\pi \in \T$ instead of $\alpha \in A$). This makes the grammar less deterministic,
+resulting in a bigger search space. The new rule is generalization of the old one,
+thus all terms in \lnf will be generated, along with many new terms in \bnf that 
+are not in \lnf.
+
+By generating \lterms in \lnf we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$ and $M =_{\eta} N$; 
+but by generating in \bnf we do not avoid it.\\
+
+The disadvantage of the \lnf, as the name suggests, is that it is long.
+Terms in \lnf are said to be \textit{fully $\eta$-expanded} \cite{barendregt10}. 
+Relevant property of $\eta$-reduction is that it always shortens the term
+that is being reduced be it. And conversely, $\eta$-expansion prolongs.
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$
+
+Now we show that for every \lterm $M$ 
+every sequence of $\eta$-reductions is finite and
+leads to unique \enf $N$.
+
+\begin{enumerate}
+ \item Every application of \ered shortens the term.
+       Since every term has finite size, this process must 
+       end at some point. Thus every \lterm has \enf.
+ \item Since \ered is \textit{Church–Rosser}, \enf is unique (see \cite{barendregt84}). 
+\end{enumerate}
+
+So we can take every generated \lterm $M$ in 
+\lnf and transform it to shorter term in \enf. 
+The question is whether it remains in \bnf, thus being in \benf.
+The answer is yes; it can be proven by showing that no 
+new \bredex is created by \ered.
+
+\begin{proposition}
+Let $P$ be in \bnf and $P \eArrow Q$. Then $Q$ is in \bnf.    
+\end{proposition}
+\begin{proof}
+
+For better clarity \lets show the \ered and \bredex using trees.\\
+\ered: \\
+
+\Tree [.$\lh{x}$ [.@ $M$ $x$ ] ] 
+~~~~$\eArrow$~~~~
+$M$
+
+And \bredex:
+
+\Tree [.@ [.$\lh{x}$ $M$ ] $N$ ] \\
+
+\Lets assume that $P \eArrow Q$ creates a new \bredex $B$ in $Q$.
+
+Since \ered only destroys and never creates \textit{function applications} (i.e. @),
+the root @ of $B$ must be present in $P$.  
+But since $P$ contains no \bredex, the left subterm $L$ of this root @
+is not $\lambda$-abstraction.
+Only possible way for $L$ to be changed by $\eArrow$ into 
+a $\lambda$-abstraction is that $L$ is the reduced subterm (so that
+$L$ is changed for its subterm).
+But that is in contradiction with $P$ not containing any \bredex,
+because it would cause $L$ be a $\lambda$-abstraction.
+
+\end{proof}
+
+Notable property of \lnf and \benf is that there is \textit{bijection} 
+(i.e. one-to-one correspondence) of 
+the set of simply typed \lterms in \lnf and 
+the set of simply typed \lterms in \benf.
+
+\begin{proposition}
+
+Reduction to \enf is bijection between  
+the set of simply typed \lterms in \lnf and 
+the set of simply typed \lterms in \benf.
+\end{proposition}
+\begin{proof}
+
+Since reduction to \enf always leads to an unique term, it is a function.
+In previous proposition is shown that \ered of \lnf
+leads to a term in \benf.\\
+
+In order to show that a function is bijection it is sufficient to show that it is
+both \textit{injection} and \textit{surjection}.\\
+
+Suppose it is not injection.
+
+So there must be $M_1,M_2$ in \lnf such that $M_1 \not= M_2$
+and $N$ in \benf such that $M_1 \etar N$, $M_2 \etar N$.
+Therefore $M_1 =_\eta M_2$, 
+so $M_1 =_{\beta\eta^{-1}} M_2$.
+This contradicts with $M_1,M_2$ being distinct \lnf{}s.\\
+
+Every $M$ in \bnf has a \lnf $N$ such that 
+$N \twoheadrightarrow_{\eta} M$ (proposition from \ref{lnf}).
+Term $M$ in \benf is in \bnf, thus it has desired \lnf $N$
+which reduces to it.
+
+Therefore it is surjection. 
+\end{proof}
+
+Suppose we have systematic (i.e. gradually generating all terms, 
+but no term twice) method for generating terms in \lnf,
+we may transform it to systematic method for generating terms in \benf
+by simply reducing each generated term to its \enf.
+
+\section{Crossover operator}
+
+\red{Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.
+
+Další metody jak zacházet s lambdama:
+
+- použít méně obecnou reprezentaci, takovym
+  příkladem je Yu, která nepovoluje vnější proměný uvnitř
+  lambda abstrakcí a další restrikce. Touto cestou také nechceme jít,
+  myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí
+  normální formu žádného well-typed termu.
+
+- křížit tak jako by se nic nedělo a problémy řešit až když nastanou:
+  tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu. 
+  Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá
+  prvek mutace. K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.
+
+Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných.
+Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I)
+které zastanou puvodní roli proměnných. We idetified two methods using this approach:
+
+- Na konci generovací fáze všechny vygenerované termy převést do kombinatorové
+  podoby pomocí elmin abstr.
+Tento přístup můžeme chápat jako kompromis (najit lepší slovo, hibrid...) mezi lambda reprezentací a komb reprezentací.
+Nastává však otázka zda má pak smysl generovat termy v lc, když je pak stejně převedeme.
+Převod má totiž až kvadratický nárůst velikosti a dá se očekávat, že termy jsou při přímém generování v komb metodě generovany usporněji.
+
+- Jedince držet po celou dobu v be-nf (té nejúspornější co se týče 
+  beta,eta-redukcí) a pouze při křížení je převést na kombinátorovou formu
+  eliminací abstrakcí. Po zkřížení je zase uvést do be-nf neobsahující přidané kombinátory tak, že jejich výskyty nahradíme jejich definicemi 
+  (e.g. K nahradime termem (\ x y . x) ) a tento term redukujeme 
+  do beta eta nf.
+Takový přístup využívá dříve nelibou vlastnost eliminace totiž kvadratický narust velikosti ku svému prospěchu, zde totiž k tomuto nárůstu dochází pouze dočasně a tím zvyšuje počet křížících bodů, což se ukazuje jako výhoda (citovat Yu kde při použití @-reprezentace místo s-výrazů dosahuje lepších výsledků).
+Díky redukcím jsou pak výsledné termy zase nejmenčí co v dáné "třídě ekvivalence" mohou být.
+Nevýhodou je, že
+  takový operace žerou nějakej čas. Je však otázkou, do jaké míry je takový overhead (kouknout co toslovo znamena:D) podstatný, jelikož na druhou stranu 
+  z keeping trees small přichází že to bude celý rychlejší a je otazka zda bude takový overhead podstatný v porovnání s overheadem pro výhodnocení termu, který je typicky nejdražší.
+
+}
+
+\subsection{Pokračujou zkopčenosti z théze}
+
+Design goal behind our approach to crossover operation is
+to try to generalize standard tree swapping crossover.
+This is not because of dislike of more specialized crossovers;
+it comes from systematic approach for exploring the domain of
+typed functional GP --- thus I would like to start with the most common one.\\
+
+%In the future, after finding sufficiently nice generalization of the standard tree %swapping crossover I plan to move my attention to more 
+%specialized crossovers, like the one presented in \cite{yu01}.  \\
+
+The crossover operation in standard GP is performed 
+by swapping randomly selected subtrees in each parent 
+S-expression (see \ref{GPxover}).
+
+For typed lambda terms two difficulties arise: Types and variables.\\
+
+We will show how to crossover \textit{typed \lterm trees} in both 
+\atTree and \sexprTree notation.\\
+
+As in standard GP our crossover will be performed by swapping
+two subtrees. But now with constraint that both subtrees have
+the same type. This is discussed in grater detail in \ref{typed-swapping}.\\
+
+Variables bring more difficulties then types do.
+This problem arises from variables that are free in subterms corresponding 
+to swapped subtrees.
+
+Following example illustrates the problem. \Lets have these two
+parent trees with selected nodes in bold.\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
+
+~\\The swap of subtrees results in following trees:\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
+
+~\\The problem is that variable $x_2$ in second tree
+is not bound by any $\lambda$-head and since
+it is not element of $\Gamma$, the second tree is not well-typed \lterm.
+
+% - This is only one example of a problem caused by swapping subtrees with variables,
+% - Resolve problems with free variables or avoid variables completely.
+
+%In \ref{repairing-method} is presented more detailed discussion of problems with
+%variables and method which repairs such "broken" terms.
+
+First approach tried at this problem was to repair the broken term,
+but since this method is not currently implemented we will describe it
+in just a few words. If the moved free variable ends in the scope 
+where is defined variable with the same name and type, then it
+is OK. If it is not this case, we may rename this variable to some 
+other variable "visible" in its new scope if such a variable has the same 
+type. If there is no such a suitable variable we may generate new value 
+to replace the variable by term generating method.
+
+As you can see, this is not much elegant solution.
+
+But we can avoid dealing with this problem by avoiding use of variables.
+This can be achieved by process called abstraction elimination.
+
+\subsection{Typed subtree swapping}
+\label{typed-swapping}
+
+First thing to do in standard subtree swapping is to select random node
+in the first parent.
+
+We modify this procedure so that we allow
+selection only of those nodes with such a type that there exists  
+a node in the second parent with the same type.
+
+Standard subtree swapping crossover as a first thing selects 
+whether the selected node will be inner node (usually with probability 
+$p_{ip} = 90\%$) or leaf node (with probability 10\%).
+
+We are in a more complicated situation, because one of those 
+sets may be empty, because of allowing only nodes with possible "partner"
+in the second parent. Thus we do this step only if both sets are
+nonempty.
+
+After selecting a node in the first parent we select node in the
+second parent such that type of that node must by the same as the type 
+of the first node. Again, this may eliminate the "90-10" step of
+first deciding whether the selected node will be internal node 
+or leaf node.
+
+When both nodes are selected we may swap the trees. \\
+
+If the abstraction elimination was performed, then 
+since the trees are of the same type and there are no variables to be 
+moved from their scope, the offspring trees are well typed.\\
+
+Both \sexprTree and \atTree are able to be crossed by this 
+mechanism. But \atTree has more possibilities then \atTree.
+This comes from the fact that every subtree of the \sexprTree
+corresponds to a subtree of \atTree, but there are subtrees
+of \atTree that do not correspond to a subtree of a \sexprTree.\\
+
+Following example should clarify this.
+
+\Tree[.@	
+   [.@ \textbf{f} x ]
+   [.y ]  		 			
+]
+\Tree[.\textbf{f} x y ]~\\
+
+In \atTree, \textbf{f} is leaf thus subtree, 
+whereas in \sexprTree it is internal node thus not a subtree.\\
+
+Another nice property of \sexprTree{}s with no lambdas 
+is that they are the same representation as S-expressions
+used by standard GP.\\
+
+Again, similarly as for standard version, 
+a maximum permissible depth $D_{created}$ 
+for offspring individuals is defined (e.g. $D_{created} = 17$).
+If one of the offspring has greater depth than this limit, then 
+this offspring is replaced by the first parent in the result of 
+the crossover operator. If both offspring exceeds this limit, than 
+both are replaced by both parents.
+
+For \atTree the $D_{created}$ must be larger since
+\atTree (without lambdas) is a binary tree. This 
+enlargement is approximately proportionate to average number of 
+function arguments. We use generous $D_{created} = 17\times3$.
 
@@
 and the \textbf{.cls} and \textbf{.tex} files that it describes.
+
+\section{Osnova}
+
+\begin{lizt}
+& Jak reprezentovat stromy programu pro GP?
+  && Reprezentace v klasickym kozovy je S-expression.
+  && Nebo mužem používat kombinátory jako 
+     Briggs a O’Neil (to si myslim je jakoby 
+     hlavní konkurence, vuči který by to chtělo obhájit)
+  && Lambda termy a jejich awesomeness
+& \textbf{Povídaní o redukcích}
+& Povídání o lnf
+  && že lnf je přirozený rozšíření s-exprešnu do lambda 
+     kalkulu vlastnosti termu v lnf 
+     &&& proč je eta redukovat (eta redukcí lnf dostanem 
+         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
+& Generování
+  && Generování gramatikou
+  && Gramatika pro lnf
+  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
+& Problémy s proměnýma a jak je řešit.
+  && Křížit lambda termy i s proměnejma a abstrakcema, 
+     problémy řešit když nastanou (pomluvit a odsoudit)
+  && Zmenšit prostor termu (tim že někerý nejsme schopný 
+     vygenerovat) s kterým operujeme tak aby křížení už 
+     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
+     jen použití proměnných z její hlavy)
+  && Převod hned po vygenerování
+  && Převod až při křížení
+     &&& eliminace abstrakcí
+     &&& skřížim
+     &&& vložený kombinátory nahradim odpovídajícím termem
+     &&& celý to redukuju
+ & Pokusy (?!)
+ & Závěr
+\end{lizt}
 

### 2014-01-26 tex/gecco14/gecco14.tex
gecco stary zmeny pred viromgem

@@
 \usepackage[ampersand]{easylist}
+\usepackage{color}
+\usepackage{qtree}
+%\usepackage{graphicx}
 
@@
 
+\red{možná dát do těla, v preliminarýs by mohlo stačit def lnf}
+
 In \cite{barendregt10} is shown term generating grammar with 
@@
 &\gar
-(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)\\
 & \textbf{if } \alpha \in A,
@@
 The second rule can be replaced by more effective one.
-\[ 
+$
 ( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
-\gar
+\gar \\
 (~\lambda~x_1~\dots~x_n~.~
@@
 
-\section{Our approach}
+\subsection{Abstraction elimination}
+\label{toSKI}
+%... Conversion to SKI combinators
+
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\Lets describe transformation $\Ae$ performing this 
+process.
+\begin{align*}
+\Ae[x]           &= x &\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb &\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
+&\textbf{if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
+&\textbf{if } x \in \FV(M) \\
+& &\vee~ x \in \FV(N)\\
+\end{align*}
+
+This is simple version of this process. More optimized version,
+in the means of the size of resulting term and its performance
+is following one, presented in \cite{jones87}.\\
+
+As is stated in \cite{jones87},
+the biggest disadvantage of this technique is that the translated
+term is often much larger than in its lambda form --- the size of
+the translated term can be proportional to the
+square of the size of the original term.
+
+But the advantage is also tempting --- no need to deal with variables
+and lambda heads.
+
+\section{\red{Tělo článku, jakoby Our approach} }
 \label{approach}
+
+\subsection{Introduction}
+\red{
+(Tématem je utilization redukcí a eliminace abstrakcí.)
+
+Nejprve si ukážerme jak využít eta redukci při konstrukci 
+počáteční populace.
+
+Hlavním tématem článku je operátor křížení.
+Jeho hlavním problémem je výskyt proměnných v individualt trees.
+To jak to můžem řešit je několik možností:
+
+- Vůbec nepoužívat lambda abstrakce, (asi v tomdle seznamu vubec neuvádět,
+  ale tohle přesunout někam hodně na začátek, tuhle kapitolu už začít s tim
+  že chceme navrhnout co nejlepší reprezentaci programů jakožto jedinců pro GP 
+  tak že involvuje lambda abstrakce)
+  generovat už rovnou kombinatorový termy (dělá Briggs).
+  [Argumenty] proč má smysl zabejvat se programama s lam abstrakcema
+  a ne čístě jen v kombinátorech (pak přesunout možná někam jinam) -
+  Možná to podat takle, "zdá se že použití kombinátorů má řadu výhod (
+  vymenovat naKY co jmenuje briggs), ale konzistentní prozkoumání 
+  obecnýho řešení lambdovýho chgarakteru určitě stojí taky za zkoužku,
+  což podporují nasledující argumenty":
+  (a) úspornost - abstrakce eliminací produkuje 
+  termy až kvadraticky větší než je původní 
+  term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
+  míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
+  podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
+  takto vnořeného podtermu složitě dopravit). 
+  (b) použití kombů vyžaduje silnější typovej sytém - 
+  kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
+  lambdy v pohodě běžej na simply typed lc. 
+  (c) použití proměnných je typické pro lidi, proto by se mohlo sát že
+  i programy generovaný systémem kterej s proměnejma počitá budou o něco 
+  "lidštější"
+
+- Použít omezenější strukturu která nepovoluje všechny možné lambda termy 
+  (vyfootnoutit nebo někam vejš říct, že všma možnejma termama myslíme i
+  redukovaný termy)
+
+}
+
+\subsection{Benefits of generating \lterms in \lnf}
+\label{benefits}
+
+By generating \lterms in \textit{lnf} we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$, but $M =_{\beta\eta} N$.
+In other words, we avoid generating two programs with different 
+source codes, but performing the same computation.
+
+Every \lterm $M$ such that \GMS has its unique \lnf $L$, 
+for which $L =_{\beta\eta} M$.
+Therefore the computation performed by \lterm $M$ 
+is not omitted, because it is the same computation
+as the computation performed by \lterm $L$. \\
+
+Generating \lterms in \lnf is even better than generating 
+\lterms in \bnf. Since \lnf is the same thing as \beenf,
+every \lterm in \lnf is also in \bnf.
+
+This comes straight from the definition of \beenf, 
+but one can also see it from observing method for generating
+terms in \bnf. As is shown in \cite{barendregt10}, 
+this method is obtained simply by replacing  
+the first grammar rule by slightly modified rule,
+resulting in the following grammar.
+\begin{align*}
+( \pi , \Gamma )  
+&\gar
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)\\
+& \textbf{if } (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
+\\ 
+( \sigma \rightarrow \tau , \Gamma )  
+&\gar
+(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
+&   
+\end{align*}
+
+The difference lies in that $f$'s type is no longer needed to be fully expanded
+($\pi \in \T$ instead of $\alpha \in A$). This makes the grammar less deterministic,
+resulting in a bigger search space. The new rule is generalization of the old one,
+thus all terms in \lnf will be generated, along with many new terms in \bnf that 
+are not in \lnf.
+
+By generating \lterms in \lnf we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$ and $M =_{\eta} N$; 
+but by generating in \bnf we do not avoid it.\\
+
+The disadvantage of the \lnf, as the name suggests, is that it is long.
+Terms in \lnf are said to be \textit{fully $\eta$-expanded} \cite{barendregt10}. 
+Relevant property of $\eta$-reduction is that it always shortens the term
+that is being reduced be it. And conversely, $\eta$-expansion prolongs.
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$
+
+Now we show that for every \lterm $M$ 
+every sequence of $\eta$-reductions is finite and
+leads to unique \enf $N$.
+
+\begin{enumerate}
+ \item Every application of \ered shortens the term.
+       Since every term has finite size, this process must 
+       end at some point. Thus every \lterm has \enf.
+ \item Since \ered is \textit{Church–Rosser}, \enf is unique (see \cite{barendregt84}). 
+\end{enumerate}
+
+So we can take every generated \lterm $M$ in 
+\lnf and transform it to shorter term in \enf. 
+The question is whether it remains in \bnf, thus being in \benf.
+The answer is yes; it can be proven by showing that no 
+new \bredex is created by \ered.
+
+\begin{proposition}
+Let $P$ be in \bnf and $P \eArrow Q$. Then $Q$ is in \bnf.    
+\end{proposition}
+\begin{proof}
+
+For better clarity \lets show the \ered and \bredex using trees.\\
+\ered: \\
+
+\Tree [.$\lh{x}$ [.@ $M$ $x$ ] ] 
+~~~~$\eArrow$~~~~
+$M$
+
+And \bredex:
+
+\Tree [.@ [.$\lh{x}$ $M$ ] $N$ ] \\
+
+\Lets assume that $P \eArrow Q$ creates a new \bredex $B$ in $Q$.
+
+Since \ered only destroys and never creates \textit{function applications} (i.e. @),
+the root @ of $B$ must be present in $P$.  
+But since $P$ contains no \bredex, the left subterm $L$ of this root @
+is not $\lambda$-abstraction.
+Only possible way for $L$ to be changed by $\eArrow$ into 
+a $\lambda$-abstraction is that $L$ is the reduced subterm (so that
+$L$ is changed for its subterm).
+But that is in contradiction with $P$ not containing any \bredex,
+because it would cause $L$ be a $\lambda$-abstraction.
+
+\end{proof}
+
+Notable property of \lnf and \benf is that there is \textit{bijection} 
+(i.e. one-to-one correspondence) of 
+the set of simply typed \lterms in \lnf and 
+the set of simply typed \lterms in \benf.
+
+\begin{proposition}
+
+Reduction to \enf is bijection between  
+the set of simply typed \lterms in \lnf and 
+the set of simply typed \lterms in \benf.
+\end{proposition}
+\begin{proof}
+
+Since reduction to \enf always leads to an unique term, it is a function.
+In previous proposition is shown that \ered of \lnf
+leads to a term in \benf.\\
+
+In order to show that a function is bijection it is sufficient to show that it is
+both \textit{injection} and \textit{surjection}.\\
+
+Suppose it is not injection.
+
+So there must be $M_1,M_2$ in \lnf such that $M_1 \not= M_2$
+and $N$ in \benf such that $M_1 \etar N$, $M_2 \etar N$.
+Therefore $M_1 =_\eta M_2$, 
+so $M_1 =_{\beta\eta^{-1}} M_2$.
+This contradicts with $M_1,M_2$ being distinct \lnf{}s.\\
+
+Every $M$ in \bnf has a \lnf $N$ such that 
+$N \twoheadrightarrow_{\eta} M$ (proposition from \ref{lnf}).
+Term $M$ in \benf is in \bnf, thus it has desired \lnf $N$
+which reduces to it.
+
+Therefore it is surjection. 
+\end{proof}
+
+Suppose we have systematic (i.e. gradually generating all terms, 
+but no term twice) method for generating terms in \lnf,
+we may transform it to systematic method for generating terms in \benf
+by simply reducing each generated term to its \enf.
+
+\subsection{Crossover}
+
+Design goal behind our approach to crossover operation is
+to try to generalize standard tree swapping crossover.
+This is not because of dislike of more specialized crossovers;
+it comes from systematic approach for exploring the domain of
+typed functional GP --- thus I would like to start with the most common one.\\
+
+The crossover operation in standard GP is performed 
+by swapping randomly selected subtrees in each parent 
+S-expression (see \ref{GPxover}).
+
+For typed lambda terms two difficulties arise: Types and variables.\\
+
+We will show how to crossover \textit{typed \lterm trees} in both 
+\atTree and \sexprTree notation.\\
+
+As in standard GP our crossover will be performed by swapping
+two subtrees. But now with constraint that both subtrees have
+the same type. This is discussed in grater detail in \ref{typed-swapping}.\\
+
+Variables bring more difficulties then types do.
+This problem arises from variables that are free in subterms corresponding 
+to swapped subtrees.
+
+Following example illustrates the problem. \Lets have these two
+parent trees with selected nodes in bold.\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
+
+~\\The swap of subtrees results in following trees:\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
+
+~\\The problem is that variable $x_2$ in second tree
+is not bound by any $\lambda$-head and since
+it is not element of $\Gamma$, the second tree is not well-typed \lterm.
+
+% - This is only one example of a problem caused by swapping subtrees with variables,
+% - Resolve problems with free variables or avoid variables completely.
+
+%In \ref{repairing-method} is presented more detailed discussion of problems with
+%variables and method which repairs such "broken" terms.
+
+First approach tried at this problem was to repair the broken term,
+but since this method is not currently implemented we will describe it
+in just a few words. If the moved free variable ends in the scope 
+where is defined variable with the same name and type, then it
+is OK. If it is not this case, we may rename this variable to some 
+other variable "visible" in its new scope if such a variable has the same 
+type. If there is no such a suitable variable we may generate new value 
+to replace the variable by term generating method.
+
+As you can see, this is not much elegant solution.
+
+But we can avoid dealing with this problem by avoiding use of variables.
+This can be achieved by process called abstraction elimination.
+
+\subsubsection{Typed subtree swapping}
+\label{typed-swapping}
+
+First thing to do in standard subtree swapping is to select random node
+in the first parent.
+
+We modify this procedure so that we allow
+selection only of those nodes with such a type that there exists  
+a node in the second parent with the same type.
+
+Standard subtree swapping crossover as a first thing selects 
+whether the selected node will be inner node (usually with probability 
+$p_{ip} = 90\%$) or leaf node (with probability 10\%).
+
+We are in a more complicated situation, because one of those 
+sets may be empty, because of allowing only nodes with possible "partner"
+in the second parent. Thus we do this step only if both sets are
+nonempty.
+
+After selecting a node in the first parent we select node in the
+second parent such that type of that node must by the same as the type 
+of the first node. Again, this may eliminate the "90-10" step of
+first deciding whether the selected node will be internal node 
+or leaf node.
+
+When both nodes are selected we may swap the trees. \\
+
+If the abstraction elimination was performed, then 
+since the trees are of the same type and there are no variables to be 
+moved from their scope, the offspring trees are well typed.\\
+
+Both \sexprTree and \atTree are able to be crossed by this 
+mechanism. But \atTree has more possibilities then \atTree.
+This comes from the fact that every subtree of the \sexprTree
+corresponds to a subtree of \atTree, but there are subtrees
+of \atTree that do not correspond to a subtree of a \sexprTree.\\
+
+Following example should clarify this.
+
+\Tree[.@	
+   [.@ \textbf{f} x ]
+   [.y ]  		 			
+]
+\Tree[.\textbf{f} x y ]~\\
+
+In \atTree, \textbf{f} is leaf thus subtree, 
+whereas in \sexprTree it is internal node thus not a subtree.\\
+
+Another nice property of \sexprTree{}s with no lambdas 
+is that they are the same representation as S-expressions
+used by standard GP.\\
+
+Again, similarly as for standard version, 
+a maximum permissible depth $D_{created}$ 
+for offspring individuals is defined (e.g. $D_{created} = 17$).
+If one of the offspring has greater depth than this limit, then 
+this offspring is replaced by the first parent in the result of 
+the crossover operator. If both offspring exceeds this limit, than 
+both are replaced by both parents.
+
+For \atTree the $D_{created}$ must be larger since
+\atTree (without lambdas) is a binary tree. This 
+enlargement is approximately proportionate to average number of 
+function arguments. We use generous $D_{created} = 17\times3$.
+
+\subsection{Unpacking crossover}
 

### 2014-01-27 tex/gecco14_b/gecco14.tex
gecco

@@
 
-~\\
+\red{ 
+Reprezentace v klasickym kozovy je S-expression.
+Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
 
-\begin{lizt}
-& Jak reprezentovat stromy programu pro GP?
-  && Reprezentace v klasickym kozovy je S-expression.
-  && Nebo mužem používat kombinátory jako 
-     Briggs a O’Neil (to si myslim je jakoby 
-     hlavní konkurence, vuči který by to chtělo obhájit)
-  && Lambda termy a jejich awesomeness
-\end{lizt}
+Zdá se že použití kombinátorů má řadu výhod (vymenovat nnějaký co jmenuje briggs), 
+ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
+taky za zkoužku, což podporují nasledující argumenty:
 
-~\\
+  (a) motivace pro práci s lambda termama je, že za nima košatá teorie, 
+      která například popisuje redukce lambda termů (ty zajišťujou jak 
+      zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, 
+      díky tomu, že se různé stromy redukují na ten samej).
+      (ALT znění: 
+      pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, 
+      protže stačí brát v potaz třídy ekvivalence a ne zamotný termy 
+      - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy. )
+      (ALT2: 
+      Dost podstatný ospravedlnění lambda reprezentace je fakt, 
+      že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.)
 
-\red{
-Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
-Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).
+  (b) úspornost - abstrakce eliminací produkuje 
+      termy až kvadraticky větší než je původní 
+      term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
+      míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
+      podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
+      takto vnořeného podtermu složitě dopravit).
 
-Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)
+  (c) použití kombů vyžaduje silnější typovej sytém - 
+      kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
+      lambdy v pohodě běžej na simply typed lc.
 
-\begin{itemize}
-\item Po vy generování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce).
-To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.
+  (d) použití proměnných je typické pro lidi, proto by se mohlo sát že
+      i programy generovaný systémem kterej s proměnejma počitá budou o něco 
+      "lidštější"
 
-\item Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
-Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
-Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y .
-x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
-Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
-Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.
-Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
-\end{itemize}
+Since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
+Pro more soffisticated type systems would be situation similar.
+
+Hlavním problémem lambda reprezentac proti kombinátorový je 
+jak se vypořádat s problémy zpusobenými použitím 
+localních proměných (aka jak křížit).
+
 }
@@
 
+\red{ve shrnutí paperu říct, že jeho tělo je tvořeno dvěma částma,
+první se zabejvá generovánim a jeho vztahem k redukcím, druhá řeší křížení
+s velkym důrazem na vyřešení problémů s lokálníma proměnejma.}
+
 The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
@@
 
-\subsection{Randombits}
+\subsection{\red{Randombits}}
 \red{
 
-- někam zduraznit že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
-
-~\\- pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, protže stačí brát v potaz třídy ekvivalence a ne zamotný termy - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy.
-
-~\\- Dost podstatný ospravedlnění lambda reprezentace je fakt, že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.
-
-~\\- nekam se může dát - since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
-Pro more soffisticated type systems would be situation similar.
-
-~\\- (promaklejší formulaci na velkym kompu zabednena) Komb jak je představuje Briggs a spol jsou husty, přinašej výhody jako 
-jednoduchost atd. ale je husty taky to skusit druhou metodou a to je lambdy pač tyhle argumenty:
-
-~\\- stručnější
-
-~\\- do defaultu vyžadujou jednoduší typovej system 
-(SKI musej bejt polymorfní)
-
-~\\- lidštější
+- někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
 
@@
 \label{related}
+
+\red{TODO zatim jen zkopčený z minula, opravit podle novýho kontextu,
+možná přesunout velkou obhajobu z úvodu sem..}
 
@@
 
-\begin{lizt}
-& Povídání o lnf
-  && že lnf je přirozený rozšíření s-exprešnu do lambda 
-     kalkulu vlastnosti termu v lnf 
-     &&& proč je eta redukovat (eta redukcí lnf dostanem 
-         beta-eta-nf a že nemusíme beta redukovat pač se to tim 
-         nerozbyje)
-\end{lizt}
+\red{
+V této sekci prodiskutujeme generování termův lnf, kteroužto můžeme chápat jako
+příme zobecnění S-expressions do lambda termů. Povíme si jaký vztah takové termy mají
+k beta a eta redukci, jaká je jejich výhoda a jak je lze jednoduše převést
+do úspornější beta eta nf. 
+}
 
@@
 In \cite{barendregt10} is shown term generating grammar with 
-following rules (our notation is used, but we will not 
-highlighted terminals anymore).
+following rules.
+
+\red{předělat ten zápis, víc to okecat, říct že to je jakoby gramatika, 
+formálně 2-level-gramatika do footnoty, pač má nekonečně neterminálu  
+i terminálů. že neterminaly jsou dvojice (typ, kontext) a že terminaly jsou
+symboly kterejma se zapisujou lambda termy}
 
@@
 it is used if and only if the non-terminal's type is not atomic.
+
+\red{řict že lnf je přirozený rozšíření s-exprešnu do lambda 
+kalkulu: pokud vezmeme gama že splnuje closure podmínku, bude 
+se používat jen pravidlo jedna, one can see that f stands for 
+root node and (ty závorky) for direct subtrees. For context containing 
+at least one higer ord function we get special kind of root node containing
+lambda head with one direct subtree standing for body of 
+the lambda function, v tomto tele se navíc mohou objevit proměnné 
+definované v této hlavě (asi doplnit obrázkama).
+
+Diskuze toho jakou přesnou strategií je nejvhodnější generovat
+termy v lnf je beyond the scope of this paper, mužeme například použít
+klasickou raped half-and half strategii. Footnotenout kraťoučkou
+zmínku o gemetrický. 
+}
 
@@
 
-Další metody jak zacházet s lambdama:
+Metody jak zacházet s lambdama:
 
@@
   příkladem je Yu, která nepovoluje vnější proměný uvnitř
-  lambda abstrakcí a další restrikce. Touto cestou také nechceme jít,
+  lambda abstrakcí a další restrikce. Touto cestou nechceme jít,
   myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí
@@
 Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných.
-Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I)
-které zastanou puvodní roli proměnných. We idetified two methods using this approach:
+Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I), které zastanou puvodní roli proměnných.
+We idetified two methods using this approach:
 
@@
 Nastává však otázka zda má pak smysl generovat termy v lc, když je pak stejně převedeme.
-Převod má totiž až kvadratický nárůst velikosti a dá se očekávat, že termy jsou při přímém generování v komb metodě generovany usporněji.
+(Výhodou je, že
+při generování v kombinátorech nemáme zajištěno že (pokud bychom nahradili S,K,I 
+za odpovídající termy a redukovlaito,tak že nevygenerujeme dvakrát různej term co se redukuje do toho samýho - což je ale takova divná výhoda kerou bych asi ani nepsal nebo nevim) Převod má totiž až kvadratický nárůst velikosti a dá se očekávat, že termy jsou při přímém generování v komb metodě generovany usporněji (čili to co sme nahnali na redukcích stratíme eliminací).
 
 - Jedince držet po celou dobu v be-nf (té nejúspornější co se týče 
-  beta,eta-redukcí) a pouze při křížení je převést na kombinátorovou formu
+  beta,eta-redukcí) a pouze při křížení oba rodiče převést na kombinátorovou formu
   eliminací abstrakcí. Po zkřížení je zase uvést do be-nf neobsahující přidané kombinátory tak, že jejich výskyty nahradíme jejich definicemi 
@@
 Díky redukcím jsou pak výsledné termy zase nejmenčí co v dáné "třídě ekvivalence" mohou být.
-Nevýhodou je, že
-  takový operace žerou nějakej čas. Je však otázkou, do jaké míry je takový overhead (kouknout co toslovo znamena:D) podstatný, jelikož na druhou stranu 
+Nevýhodou je, že takový operace žerou nějakej čas.
+Je však otázkou, do jaké míry je takový overhead (kouknout co toslovo znamena:D) podstatný, jelikož na druhou stranu
   z keeping trees small přichází že to bude celý rychlejší a je otazka zda bude takový overhead podstatný v porovnání s overheadem pro výhodnocení termu, který je typicky nejdražší.
+Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
+
+%(ALTernativní znění ze začátku: Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
+Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměný a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
+Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y .
+x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
+Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.)
 
@@
 
-\subsection{Pokračujou zkopčenosti z théze}
+\subsection{Pokračujou zkopčenosti z théze až dokonce }
+
+\red{todle se dá dát asi na začátek crossover operation spíš, nebo nevim}
 
@@
 
-%In the future, after finding sufficiently nice generalization of the standard tree %swapping crossover I plan to move my attention to more 
-%specialized crossovers, like the one presented in \cite{yu01}.  \\
+\red{to furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje}
 
@@
 by swapping randomly selected subtrees in each parent 
-S-expression (see \ref{GPxover}).
+S-expression.
 
@@
 
-We will show how to crossover \textit{typed \lterm trees} in both 
-\atTree and \sexprTree notation.\\
+\red{CO S @-stromama řešit je explicitně tady, případně někde jinde?
+Pro generování nemá reprezentace vliv, ta má vliv až vpři křížení, kde at-tree zajišťuje víc míst.
+We will show how to crossover \textit{typed \lterm trees} in both \atTree and \sexprTree notation.}
 
@@
 
-\subsection{Typed subtree swapping}
+\subsubsection{Typed subtree swapping}
 \label{typed-swapping}
@@
 
+TODO
+
+\red{mimojiné zmínit, že v budouznu bysme chtěli porovnat kombinátory 
+s tou naší nějak extenzivně (což by bylo hezký už ted ale nestíhá se to..)}
+
 %ACKNOWLEDGMENTS are optional
@@
 
-\section{Osnova}
-
-\begin{lizt}
-& Jak reprezentovat stromy programu pro GP?
-  && Reprezentace v klasickym kozovy je S-expression.
-  && Nebo mužem používat kombinátory jako 
-     Briggs a O’Neil (to si myslim je jakoby 
-     hlavní konkurence, vuči který by to chtělo obhájit)
-  && Lambda termy a jejich awesomeness
-& \textbf{Povídaní o redukcích}
-& Povídání o lnf
-  && že lnf je přirozený rozšíření s-exprešnu do lambda 
-     kalkulu vlastnosti termu v lnf 
-     &&& proč je eta redukovat (eta redukcí lnf dostanem 
-         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
-& Generování
-  && Generování gramatikou
-  && Gramatika pro lnf
-  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
-& Problémy s proměnýma a jak je řešit.
-  && Křížit lambda termy i s proměnejma a abstrakcema, 
-     problémy řešit když nastanou (pomluvit a odsoudit)
-  && Zmenšit prostor termu (tim že někerý nejsme schopný 
-     vygenerovat) s kterým operujeme tak aby křížení už 
-     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
-     jen použití proměnných z její hlavy)
-  && Převod hned po vygenerování
-  && Převod až při křížení
-     &&& eliminace abstrakcí
-     &&& skřížim
-     &&& vložený kombinátory nahradim odpovídajícím termem
-     &&& celý to redukuju
- & Pokusy (?!)
- & Závěr
-\end{lizt}
+%\section{Osnova}
+%\begin{lizt}
+%& Jak reprezentovat stromy programu pro GP?
+%  && Reprezentace v klasickym kozovy je S-expression.
+%  && Nebo mužem používat kombinátory jako 
+%     Briggs a O’Neil (to si myslim je jakoby 
+%     hlavní konkurence, vuči který by to chtělo obhájit)
+%  && Lambda termy a jejich awesomeness
+%& \textbf{Povídaní o redukcích}
+%& Povídání o lnf
+%  && že lnf je přirozený rozšíření s-exprešnu do lambda 
+%     kalkulu vlastnosti termu v lnf 
+%     &&& proč je eta redukovat (eta redukcí lnf dostanem 
+%         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
+%& Generování
+%  && Generování gramatikou
+%  && Gramatika pro lnf
+%  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
+%& Problémy s proměnýma a jak je řešit.
+%  && Křížit lambda termy i s proměnejma a abstrakcema, 
+%     problémy řešit když nastanou (pomluvit a odsoudit)
+%  && Zmenšit prostor termu (tim že někerý nejsme schopný 
+%     vygenerovat) s kterým operujeme tak aby křížení už 
+%     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
+%     jen použití proměnných z její hlavy)
+%  && Převod hned po vygenerování
+%  && Převod až při křížení
+%     &&& eliminace abstrakcí
+%     &&& skřížim
+%     &&& vložený kombinátory nahradim odpovídajícím termem
+%     &&& celý to redukuju
+% & Pokusy (?!)
+% & Závěr
+%\end{lizt}
 

### 2014-01-27 tex/gecco14/gecco14.tex
gecco

@@
 \usepackage[ampersand]{easylist}
+\usepackage{qtree}
 \usepackage{color}
-\usepackage{qtree}
-%\usepackage{graphicx}
 
@@
 
-\section{Plán boje}
-
-\subsection{Jak to celý pojmout?}
-
-Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
-Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).
-
-Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)
-
-\begin{itemize}
-\item Po vy generování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce).
-To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.
-
-\item Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
-Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
-Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y .
-x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
-Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
-Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.
-Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
-\end{itemize}
-
-\subsection{Osnova}
-
-\begin{lizt}
-& Jak reprezentovat stromy programu pro GP?
-  && Reprezentace v klasickym kozovy je S-expression.
-  && Nebo mužem používat kombinátory jako 
-     Briggs a O’Neil (to si myslim je jakoby 
-     hlavní konkurence, vuči který by to chtělo obhájit)
-  && Lambda termy a jejich awesomeness
-& \textbf{Povídaní o redukcích}
-& Povídání o lnf
-  && že lnf je přirozený rozšíření s-exprešnu do lambda 
-     kalkulu vlastnosti termu v lnf 
-     &&& proč je eta redukovat (eta redukcí lnf dostanem 
-         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
-& Generování
-  && Generování gramatikou
-  && Gramatika pro lnf
-  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
-& Problémy s proměnýma a jak je řešit.
-  && Křížit lambda termy i s proměnejma a abstrakcema, 
-     problémy řešit když nastanou (pomluvit a odsoudit)
-  && Zmenšit prostor termu (tim že někerý nejsme schopný 
-     vygenerovat) s kterým operujeme tak aby křížení už 
-     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
-     jen použití proměnných z její hlavy)
-  && Převod hned po vygenerování
-  && Převod až při křížení
-     &&& eliminace abstrakcí
-     &&& skřížim
-     &&& vložený kombinátory nahradim odpovídajícím termem
-     &&& celý to redukuju
- & Pokusy (?!)
- & Závěr
-\end{lizt}
-
 \section{Introduction}
@@
 the traditional GP representation of individuals, with concepts from
-lambda calculus, which is simple yet powerful functional mathematical and programming 
-language extensively used in type theory. Such attempts has shown to be 
-successful \cite{yu01}.
+lambda calculus, which is simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts has shown to be successful \cite{yu01}.
 
-The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
+%Other ...
+\red{
+
+%Reprezentace v klasickym kozovy je S-expression.
+
+The individual representation used in standard GP is S-expression. 
+Basically there are two general approaches of generalizing S-expressions
+into trees of better expressive power suitable for typed GP with higher 
+order functions: We can enrich the method with the polymorphic combinators \cite{kes}
+or we can enrich the method with the lambda abstractions and local variables
+\cite{yu01}.
+
+Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
+
+Zdá se že použití kombinátorů má řadu výhod (vymenovat nnějaký co jmenuje briggs), 
+ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
+taky za zkoužku, což podporují nasledující argumenty:
+
+  (a) motivace pro práci s lambda termama je, že za nima košatá teorie, 
+      která například popisuje redukce lambda termů (ty zajišťujou jak 
+      zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, 
+      díky tomu, že se různé stromy redukují na ten samej).
+      (ALT znění: 
+      pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, 
+      protže stačí brát v potaz třídy ekvivalence a ne zamotný termy 
+      - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy. )
+      (ALT2: 
+      Dost podstatný ospravedlnění lambda reprezentace je fakt, 
+      že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.)
+
+  (b) úspornost - abstrakce eliminací produkuje 
+      termy až kvadraticky větší než je původní 
+      term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
+      míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
+      podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
+      takto vnořeného podtermu složitě dopravit).
+
+  (c) použití kombů vyžaduje silnější typovej sytém - 
+      kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
+      lambdy v pohodě běžej na simply typed lc.
+
+  (d) použití proměnných je typické pro lidi, proto by se mohlo sát že
+      i programy generovaný systémem kterej s proměnejma počitá budou o něco 
+      "lidštější"
+
+Since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
+Pro more soffisticated type systems would be situation similar.
+
+Hlavním problémem lambda reprezentac proti kombinátorový je 
+jak se vypořádat s problémy zpusobenými použitím 
+localních proměných (aka jak křížit).
+
+}
+
+%\nehodi{The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
 During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
@@
 These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
-This design goal also differentiate our approach from
-the three state of the art proposals for typed GP known to us that are discussed in the following section. 
+This design goal also differentiate our approach from the three state of the art proposals for typed GP known to us that are discussed in the following section.
 Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
-Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
+Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.}
+
+\red{ve shrnutí paperu říct, že jeho tělo je tvořeno dvěma částma,
+první se zabejvá generovánim a jeho vztahem k redukcím, druhá řeší křížení
+s velkym důrazem na vyřešení problémů s lokálníma proměnejma.}
 
@@
 
+\subsection{\red{Randombits}}
+\red{
+
+- někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
+
+}
+
 \section{Related work}
 \label{related}
+
+\red{TODO zatim jen zkopčený z minula, opravit podle novýho kontextu,
+možná přesunout velkou obhajobu z úvodu sem..}
 
@@
 
+\subsection{Abstraction elimination}
+
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\Lets describe transformation $\Ae$ performing this 
+process.\footnote{In our implementation we must also deal with types, because 
+our individuals are annotated with types. But since this process
+is fairly straightforward but cumbersome to describe, we will 
+skip explaining it here. Reader interested in seeing it can see 
+the source code of our implementation. }
+\begin{align*}
+\Ae[x]           &= x &\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb &\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
+&\textbf{if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
+&\textbf{if } x \in \FV(M)\\
+&&\vee~ x \in \FV(N)\\
+\end{align*}
+
+This is simple version of this process. More optimized version,
+in the means of the size of resulting term and its performance
+is following one, presented in \cite{jones87}.\\
+
+As is stated in \cite{jones87},
+the biggest disadvantage of this technique is that the translated
+term is often much larger than in its lambda form --- the size of
+the translated term can be proportional to the
+square of the size of the original term.
+
+But the advantage is also tempting --- no need to deal with variables
+and lambda heads.
+
+\section{Term generating and reductions}
+
+\red{
+V této sekci prodiskutujeme generování termův lnf, kteroužto můžeme chápat jako
+příme zobecnění S-expressions do lambda termů. Povíme si jaký vztah takové termy mají
+k beta a eta redukci, jaká je jejich výhoda a jak je lze jednoduše převést
+do úspornější beta eta nf. 
+}
+
 \subsection{Grammar producing \lterms in \lnf}
 
-\red{možná dát do těla, v preliminarýs by mohlo stačit def lnf}
+In \cite{barendregt10} is shown term generating grammar with 
+following rules.
 
-In \cite{barendregt10} is shown term generating grammar with 
-following rules (our notation is used, but we will not 
-highlighted terminals anymore).
-
-%\footnote{
-%I was using term generating grammars for term generation
-%on my own before I encountered with \cite{barendregt10},
-%where i happily discovered that Barendregt is using
-%almost identical notation. 
-%But his \lnf grammar was more clever then my grammar
-%that was using grammar rules corresponding to the three basic
-%inference rules. 
-%}
+\red{předělat ten zápis, víc to okecat, říct že to je jakoby gramatika, 
+formálně 2-level-gramatika do footnoty, pač má nekonečně neterminálu  
+i terminálů. že neterminaly jsou dvojice (typ, kontext) a že terminaly jsou
+symboly kterejma se zapisujou lambda termy}
 
@@
 &\gar
-(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)\\
-& \textbf{if } \alpha \in A,
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+\\& \textbf{if } \alpha \in A,
 (f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
@@
 
-\subsection{Abstraction elimination}
-\label{toSKI}
-%... Conversion to SKI combinators
+\red{řict že lnf je přirozený rozšíření s-exprešnu do lambda 
+kalkulu: pokud vezmeme gama že splnuje closure podmínku, bude 
+se používat jen pravidlo jedna, one can see that f stands for 
+root node and (ty závorky) for direct subtrees. For context containing 
+at least one higer ord function we get special kind of root node containing
+lambda head with one direct subtree standing for body of 
+the lambda function, v tomto tele se navíc mohou objevit proměnné 
+definované v této hlavě (asi doplnit obrázkama).
 
-\textit{Abstraction elimination} is a process of transforming 
-an arbitrary \lterm into \lterm that contains no lambda abstractions
-and no bound variables.
-The newly produced \lterm may contain function applications, 
-free symbols from former \lterm and some new symbols standing for 
-combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
-
-Those combinators are defined as:
-\begin{align*}
-\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
-\Kcomb &= \lam{x\,y}{x} \\
-\Icomb &= \lam{x}{x} 
-\end{align*}
-
-\Lets describe transformation $\Ae$ performing this 
-process.
-\begin{align*}
-\Ae[x]           &= x &\\[0.4em]
-\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
-\Ae[\lam{x}{x}]  &= \Icomb &\\
-\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
-\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
-&\textbf{if } x \in \FV(M)\\
-\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
-&\textbf{if } x \in \FV(M) \\
-& &\vee~ x \in \FV(N)\\
-\end{align*}
-
-This is simple version of this process. More optimized version,
-in the means of the size of resulting term and its performance
-is following one, presented in \cite{jones87}.\\
-
-As is stated in \cite{jones87},
-the biggest disadvantage of this technique is that the translated
-term is often much larger than in its lambda form --- the size of
-the translated term can be proportional to the
-square of the size of the original term.
-
-But the advantage is also tempting --- no need to deal with variables
-and lambda heads.
-
-\section{\red{Tělo článku, jakoby Our approach} }
-\label{approach}
-
-\subsection{Introduction}
-\red{
-(Tématem je utilization redukcí a eliminace abstrakcí.)
-
-Nejprve si ukážerme jak využít eta redukci při konstrukci 
-počáteční populace.
-
-Hlavním tématem článku je operátor křížení.
-Jeho hlavním problémem je výskyt proměnných v individualt trees.
-To jak to můžem řešit je několik možností:
-
-- Vůbec nepoužívat lambda abstrakce, (asi v tomdle seznamu vubec neuvádět,
-  ale tohle přesunout někam hodně na začátek, tuhle kapitolu už začít s tim
-  že chceme navrhnout co nejlepší reprezentaci programů jakožto jedinců pro GP 
-  tak že involvuje lambda abstrakce)
-  generovat už rovnou kombinatorový termy (dělá Briggs).
-  [Argumenty] proč má smysl zabejvat se programama s lam abstrakcema
-  a ne čístě jen v kombinátorech (pak přesunout možná někam jinam) -
-  Možná to podat takle, "zdá se že použití kombinátorů má řadu výhod (
-  vymenovat naKY co jmenuje briggs), ale konzistentní prozkoumání 
-  obecnýho řešení lambdovýho chgarakteru určitě stojí taky za zkoužku,
-  což podporují nasledující argumenty":
-  (a) úspornost - abstrakce eliminací produkuje 
-  termy až kvadraticky větší než je původní 
-  term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
-  míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
-  podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
-  takto vnořeného podtermu složitě dopravit). 
-  (b) použití kombů vyžaduje silnější typovej sytém - 
-  kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
-  lambdy v pohodě běžej na simply typed lc. 
-  (c) použití proměnných je typické pro lidi, proto by se mohlo sát že
-  i programy generovaný systémem kterej s proměnejma počitá budou o něco 
-  "lidštější"
-
-- Použít omezenější strukturu která nepovoluje všechny možné lambda termy 
-  (vyfootnoutit nebo někam vejš říct, že všma možnejma termama myslíme i
-  redukovaný termy)
-
+Diskuze toho jakou přesnou strategií je nejvhodnější generovat
+termy v lnf je beyond the scope of this paper, mužeme například použít
+klasickou raped half-and half strategii. Footnotenout kraťoučkou
+zmínku o gemetrický. 
 }
@@
 &\gar
-(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)\\
-& \textbf{if } (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+\\& \textbf{if } (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
 \\ 
@@
 
-\subsection{Crossover}
+\section{Crossover operator}
+
+\red{Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.
+
+Metody jak zacházet s lambdama:
+
+- použít méně obecnou reprezentaci, takovym
+  příkladem je Yu, která nepovoluje vnější proměný uvnitř
+  lambda abstrakcí a další restrikce. Touto cestou nechceme jít,
+  myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí
+  normální formu žádného well-typed termu.
+
+- křížit tak jako by se nic nedělo a problémy řešit až když nastanou:
+  tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu. 
+  Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá
+  prvek mutace. K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.
+
+Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných.
+Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I), které zastanou puvodní roli proměnných.
+We idetified two methods using this approach:
+
+- Na konci generovací fáze všechny vygenerované termy převést do kombinatorové
+  podoby pomocí elmin abstr.
+Tento přístup můžeme chápat jako kompromis (najit lepší slovo, hibrid...) mezi lambda reprezentací a komb reprezentací.
+Nastává však otázka zda má pak smysl generovat termy v lc, když je pak stejně převedeme.
+(Výhodou je, že
+při generování v kombinátorech nemáme zajištěno že (pokud bychom nahradili S,K,I 
+za odpovídající termy a redukovlaito,tak že nevygenerujeme dvakrát různej term co se redukuje do toho samýho - což je ale takova divná výhoda kerou bych asi ani nepsal nebo nevim) Převod má totiž až kvadratický nárůst velikosti a dá se očekávat, že termy jsou při přímém generování v komb metodě generovany usporněji (čili to co sme nahnali na redukcích stratíme eliminací).
+
+- Jedince držet po celou dobu v be-nf (té nejúspornější co se týče 
+  beta,eta-redukcí) a pouze při křížení oba rodiče převést na kombinátorovou formu
+  eliminací abstrakcí. Po zkřížení je zase uvést do be-nf neobsahující přidané kombinátory tak, že jejich výskyty nahradíme jejich definicemi 
+  (e.g. K nahradime termem (\ x y . x) ) a tento term redukujeme 
+  do beta eta nf.
+Takový přístup využívá dříve nelibou vlastnost eliminace totiž kvadratický narust velikosti ku svému prospěchu, zde totiž k tomuto nárůstu dochází pouze dočasně a tím zvyšuje počet křížících bodů, což se ukazuje jako výhoda (citovat Yu kde při použití @-reprezentace místo s-výrazů dosahuje lepších výsledků).
+Díky redukcím jsou pak výsledné termy zase nejmenčí co v dáné "třídě ekvivalence" mohou být.
+Nevýhodou je, že takový operace žerou nějakej čas.
+Je však otázkou, do jaké míry je takový overhead (kouknout co toslovo znamena:D) podstatný, jelikož na druhou stranu
+  z keeping trees small přichází že to bude celý rychlejší a je otazka zda bude takový overhead podstatný v porovnání s overheadem pro výhodnocení termu, který je typicky nejdražší.
+Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
+
+%(ALTernativní znění ze začátku: Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
+Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměný a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
+Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y .
+x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
+Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.)
+
+}
+
+\subsection{Pokračujou zkopčenosti z théze až dokonce }
+
+\red{todle se dá dát asi na začátek crossover operation spíš, nebo nevim}
 
@@
 
+\red{to furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje}
+
 The crossover operation in standard GP is performed 
 by swapping randomly selected subtrees in each parent 
-S-expression (see \ref{GPxover}).
+S-expression.
 
@@
 
-\subsection{Unpacking crossover}
-
 \section{Experiments}
@@
 \label{conclusions}
+
+TODO
+
+\red{mimojiné zmínit, že v budouznu bysme chtěli porovnat kombinátory 
+s tou naší nějak extenzivně (což by bylo hezký už ted ale nestíhá se to..)}
 
@@
 and the \textbf{.cls} and \textbf{.tex} files that it describes.
+
+%\section{Osnova}
+%\begin{lizt}
+%& Jak reprezentovat stromy programu pro GP?
+%  && Reprezentace v klasickym kozovy je S-expression.
+%  && Nebo mužem používat kombinátory jako 
+%     Briggs a O’Neil (to si myslim je jakoby 
+%     hlavní konkurence, vuči který by to chtělo obhájit)
+%  && Lambda termy a jejich awesomeness
+%& \textbf{Povídaní o redukcích}
+%& Povídání o lnf
+%  && že lnf je přirozený rozšíření s-exprešnu do lambda 
+%     kalkulu vlastnosti termu v lnf 
+%     &&& proč je eta redukovat (eta redukcí lnf dostanem 
+%         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
+%& Generování
+%  && Generování gramatikou
+%  && Gramatika pro lnf
+%  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
+%& Problémy s proměnýma a jak je řešit.
+%  && Křížit lambda termy i s proměnejma a abstrakcema, 
+%     problémy řešit když nastanou (pomluvit a odsoudit)
+%  && Zmenšit prostor termu (tim že někerý nejsme schopný 
+%     vygenerovat) s kterým operujeme tak aby křížení už 
+%     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
+%     jen použití proměnných z její hlavy)
+%  && Převod hned po vygenerování
+%  && Převod až při křížení
+%     &&& eliminace abstrakcí
+%     &&& skřížim
+%     &&& vložený kombinátory nahradim odpovídajícím termem
+%     &&& celý to redukuju
+% & Pokusy (?!)
+% & Závěr
+%\end{lizt}
 

### 2014-01-27 tex/gecco14/gecco14.tex
gecco

@@
 
+Todele asi šoupnout do related work.
+
 %Reprezentace v klasickym kozovy je S-expression.
@@
 
-Zdá se že použití kombinátorů má řadu výhod (vymenovat nnějaký co jmenuje briggs), 
+Zdá se že použití kombinátorů má řadu výhod (vymenovat nějaký co jmenuje briggs), 
 ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
@@
 \Lets describe transformation $\Ae$ performing this 
-process.\footnote{In our implementation we must also deal with types, because 
-our individuals are annotated with types. But since this process
-is fairly straightforward but cumbersome to describe, we will 
-skip explaining it here. Reader interested in seeing it can see 
-the source code of our implementation. }
+process.
+%\footnote{In our implementation we must also deal with types, because our individuals are annotated with types.
+But since this process is fairly straightforward but cumbersome to describe, we will skip explaining it here.
+Reader interested in seeing it can see the source code of our implementation.
+}
 \begin{align*}
@@
 
+\red{tady je to v původnim kontextu, odebrat slova co se tam nehodí}
+
 \section{Term generating and reductions}
 
-\red{
-V této sekci prodiskutujeme generování termův lnf, kteroužto můžeme chápat jako
-příme zobecnění S-expressions do lambda termů. Povíme si jaký vztah takové termy mají
-k beta a eta redukci, jaká je jejich výhoda a jak je lze jednoduše převést
-do úspornější beta eta nf. 
-}
+This section is focused on the individual generating method producing
+terms in their long normal form, which can be understood as a straight 
+generalization of S-expressions into lambda calculus. 
+We discuss the relation between terms in \lnf with beta and eta reductions,
+the advantages of such representation and we also show how to easily 
+transform such terms into short \benf.
 
@@
 
-\red{řict že lnf je přirozený rozšíření s-exprešnu do lambda 
-kalkulu: pokud vezmeme gama že splnuje closure podmínku, bude 
-se používat jen pravidlo jedna, one can see that f stands for 
-root node and (ty závorky) for direct subtrees. For context containing 
-at least one higer ord function we get special kind of root node containing
-lambda head with one direct subtree standing for body of 
+Long normal form is a generalization of S-expressions into lambda calculus:
+Let $\Gamma$ be a context satisfying the closure requirement, then only
+the first rule will be applicable. 
+$(\alpha, \Gamma)$ rewrites to $(~f~( \alpha, \Gamma )~\dots~( \alpha, \Gamma )~)$.
+One can see that $f$ stands for 
+the root node and that each $( \rho_i , \Gamma )$ will produce a direct subtree.
+
+\red{
+For context containing at least one higher order function we get special kind of root node containing lambda head with one direct subtree standing for body of 
 the lambda function, v tomto tele se navíc mohou objevit proměnné 
@@
 because it would cause $L$ be a $\lambda$-abstraction.
-
 \end{proof}
@@
 
-\red{Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.
-
-Metody jak zacházet s lambdama:
-
-- použít méně obecnou reprezentaci, takovym
-  příkladem je Yu, která nepovoluje vnější proměný uvnitř
-  lambda abstrakcí a další restrikce. Touto cestou nechceme jít,
-  myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí
-  normální formu žádného well-typed termu.
-
-- křížit tak jako by se nic nedělo a problémy řešit až když nastanou:
-  tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu. 
-  Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá
-  prvek mutace. K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.
-
-Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných.
+The design goal behind our approach to the crossover operation is
+to try to generalize the standard tree swapping crossover.
+
+The crossover operation in standard GP is performed 
+by swapping randomly selected subtrees in each parent 
+S-expression.
+For typed lambda terms two difficulties arise: Types and variables.
+
+As in standard GP our crossover will be performed by swapping
+two subtrees. But now with constraint that both subtrees have
+the same type.
+
+Variables bring more difficulties then types do.
+This problem arises from variables that are free in subterms corresponding 
+to swapped subtrees.
+
+Following example illustrates the problem. \Lets have these two
+parent trees with selected nodes in bold.\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
+
+~\\The swap of subtrees results in following trees:\\
+
+\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
+\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
+
+~\\The problem is that variable $x_2$ in second tree
+is not bound by any $\lambda$-head and since
+it is not element of $\Gamma$, the second tree is not well-typed \lterm.
+
+Generally speaking, the variable problem is caused by that the local variable is not defined in the new place or the variable is defined in the new place, but has
+some another type.
+
+Let us list some possible approaches to solving this problem:
+
+%Už přeloženo: 
+%Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.
+%Metody jak zacházet s lambdama:
+%- použít méně obecnou reprezentaci, takovym příkladem je Yu, která nepovoluje vnější proměný uvnitř lambda abstrakcí a další restrikce.
+Touto cestou nechceme jít, myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí normální formu žádného well-typed termu.
+%- křížit tak jako by se nic nedělo a problémy řešit až když nastanou: tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu.
+Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá prvek mutace.
+K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.
+%Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných.
 Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I), které zastanou puvodní roli proměnných.
 We idetified two methods using this approach:
+
+One option is to choose a less general representation of lambda terms.
+Such approach is successfully used in \cite{yu01} where the 
+terminal set for constructing lambda abstraction subtrees 
+is limited to use only constants and variables of that particular
+lambda abstraction and an appropriate variable naming convention is
+used in order to prevent troubles with variables.
+We choose not to follow such approach because we do not want to
+lose the opportunity to represent every possible well-typed 
+term in its normal form.
+
+Another option is to overlook the variable problems and to correct the
+defects when they occur. Such correction can be performed by 
+renaming the problematic variable to some another variable defined in that place 
+or, if there is no suitable candidate, to replace its occurrence with 
+a newly generated term of required type. This approach seems to us as a very
+unwieldy one -- there are clearly problems for which such collisions would
+occur very often and when there is no suitable rename candidate the 
+generation of new term brings the unnecessary element of a mutation.
+Such solution seems to be useful only as a last resort.
+
+There is an elegant way to overcome the problem with variables by getting rid of them.
+This is possible thanks to the abstraction elimination algorithm described above.
+It turns a \lterm into a \lterm that contains no lambda abstractions
+and no variables, instead, it contains additional function symbols standing for 
+polymorphic combinators (i.e. $\Scomb$, $\Kcomb$ and $\Icomb$ 
+in the simple case of the algorithm).
+
+We identified two methods using this approach:
+
+\red{
 
@@
 
-\subsection{Pokračujou zkopčenosti z théze až dokonce }
-
-\red{todle se dá dát asi na začátek crossover operation spíš, nebo nevim}
-
-Design goal behind our approach to crossover operation is
-to try to generalize standard tree swapping crossover.
-This is not because of dislike of more specialized crossovers;
-it comes from systematic approach for exploring the domain of
-typed functional GP --- thus I would like to start with the most common one.\\
-
-\red{to furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje}
-
-The crossover operation in standard GP is performed 
-by swapping randomly selected subtrees in each parent 
-S-expression.
-
-For typed lambda terms two difficulties arise: Types and variables.\\
-
-\red{CO S @-stromama řešit je explicitně tady, případně někde jinde?
+\subsection{Poznámky k uvodu ke křížení}
+\red{
+POZNAMKY:
+
+CO S @-stromama řešit je explicitně tady, případně někde jinde?
 Pro generování nemá reprezentace vliv, ta má vliv až vpři křížení, kde at-tree zajišťuje víc míst.
-We will show how to crossover \textit{typed \lterm trees} in both \atTree and \sexprTree notation.}
-
-As in standard GP our crossover will be performed by swapping
-two subtrees. But now with constraint that both subtrees have
-the same type. This is discussed in grater detail in \ref{typed-swapping}.\\
-
-Variables bring more difficulties then types do.
-This problem arises from variables that are free in subterms corresponding 
-to swapped subtrees.
-
-Following example illustrates the problem. \Lets have these two
-parent trees with selected nodes in bold.\\
-
-\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ [.\textbf{g} $x_2$ c ] ] $x_1$ ] ]
-\Tree [.$\lh{x_1}$ [.h $x_1$ $\mathbf{x_1}$ ] ]
-
-~\\The swap of subtrees results in following trees:\\
-
-\Tree [.$\lh{x_1}$ [.f [.$\lh{x_2}$ $\mathbf{x_1}$ ] $x_1$ ] ]
-\Tree [.$\lh{x_1}$ [.h $x_1$ [.\textbf{g} $\mathbf{x_2}$ \textbf{c} ] ] ]
-
-~\\The problem is that variable $x_2$ in second tree
-is not bound by any $\lambda$-head and since
-it is not element of $\Gamma$, the second tree is not well-typed \lterm.
-
-% - This is only one example of a problem caused by swapping subtrees with variables,
-% - Resolve problems with free variables or avoid variables completely.
-
-%In \ref{repairing-method} is presented more detailed discussion of problems with
-%variables and method which repairs such "broken" terms.
-
-First approach tried at this problem was to repair the broken term,
-but since this method is not currently implemented we will describe it
-in just a few words. If the moved free variable ends in the scope 
-where is defined variable with the same name and type, then it
-is OK. If it is not this case, we may rename this variable to some 
-other variable "visible" in its new scope if such a variable has the same 
-type. If there is no such a suitable variable we may generate new value 
-to replace the variable by term generating method.
-
-As you can see, this is not much elegant solution.
-
-But we can avoid dealing with this problem by avoiding use of variables.
-This can be achieved by process called abstraction elimination.
-
-\subsubsection{Typed subtree swapping}
+We will show how to crossover \textit{typed \lterm trees} in both \atTree and \sexprTree notation.
+
+to o tom že chem zobecnit tu standardní furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje, někde to zmínit že se to na ně nevztahuje
+}
+
+\subsection{Typed subtree swapping in greater detail}
 \label{typed-swapping}
@@
 \red{mimojiné zmínit, že v budouznu bysme chtěli porovnat kombinátory 
-s tou naší nějak extenzivně (což by bylo hezký už ted ale nestíhá se to..)}
+s tou naší nějak extenzivně, ikdyž možná to tam vubec nepsat
+(což by bylo hezký už ted ale nestíhá se to..)}
 

### 2014-01-28 tex/gecco14/gecco14.tex
gecco

@@
 % A category with the (minimum) three required fields
-\category{H.4}{TODO}{Miscellaneous}
+\category{H.4}{\red{TODO}}{Miscellaneous}
 %A category including the fourth, optional field follows...
@@
 
-\terms{TODO}
-
-\keywords{TODO, ToDo, todo}
+\terms{\red{TODO}}
+
+\keywords{\red{TODO, ToDo, todo}}
 
@@
 
-%Other ...
-\red{
-
-Todele asi šoupnout do related work.
-
-%Reprezentace v klasickym kozovy je S-expression.
-
-The individual representation used in standard GP is S-expression. 
-Basically there are two general approaches of generalizing S-expressions
-into trees of better expressive power suitable for typed GP with higher 
-order functions: We can enrich the method with the polymorphic combinators \cite{kes}
-or we can enrich the method with the lambda abstractions and local variables
-\cite{yu01}.
-
-Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
-
-Zdá se že použití kombinátorů má řadu výhod (vymenovat nějaký co jmenuje briggs), 
-ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
-taky za zkoužku, což podporují nasledující argumenty:
-
-  (a) motivace pro práci s lambda termama je, že za nima košatá teorie, 
-      která například popisuje redukce lambda termů (ty zajišťujou jak 
-      zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, 
-      díky tomu, že se různé stromy redukují na ten samej).
-      (ALT znění: 
-      pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, 
-      protže stačí brát v potaz třídy ekvivalence a ne zamotný termy 
-      - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy. )
-      (ALT2: 
-      Dost podstatný ospravedlnění lambda reprezentace je fakt, 
-      že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.)
-
-  (b) úspornost - abstrakce eliminací produkuje 
-      termy až kvadraticky větší než je původní 
-      term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
-      míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
-      podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
-      takto vnořeného podtermu složitě dopravit).
-
-  (c) použití kombů vyžaduje silnější typovej sytém - 
-      kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
-      lambdy v pohodě běžej na simply typed lc.
-
-  (d) použití proměnných je typické pro lidi, proto by se mohlo sát že
-      i programy generovaný systémem kterej s proměnejma počitá budou o něco 
-      "lidštější"
-
-Since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
-Pro more soffisticated type systems would be situation similar.
-
-Hlavním problémem lambda reprezentac proti kombinátorový je 
-jak se vypořádat s problémy zpusobenými použitím 
-localních proměných (aka jak křížit).
-
-}
-
+\red{TODO....}
+
+%Pro inspiraci :
 %\nehodi{The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
@@
 
-- někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
+- Možná někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
 
@@
 
-\red{TODO zatim jen zkopčený z minula, opravit podle novýho kontextu,
-možná přesunout velkou obhajobu z úvodu sem..}
+\red{TODO zatim jen zkopčený z minula a dodaná diskuze 
+zatim česky porovnávající nás a kombinatoři , 
+opravit podle novýho kontextu}
 
@@
 
+\red{
+
+POZNAMKY PRO DISKUSI POROVNAVAJICI NÁS A BRIGGSE:
+
+%Reprezentace v klasickym kozovy je S-expression.
+
+The individual representation used in standard GP is S-expression. 
+Basically there are two general approaches of generalizing S-expressions
+into trees of better expressive power suitable for typed GP with higher 
+order functions: We can enrich the method with the polymorphic combinators \cite{kes}
+or we can enrich the method with the lambda abstractions and local variables
+\cite{yu01}.
+
+Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
+
+Zdá se že použití kombinátorů má řadu výhod (vymenovat nějaký co jmenuje briggs), 
+ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
+taky za zkoužku, což podporují nasledující argumenty:
+
+  (a) motivace pro práci s lambda termama je, že za nima košatá teorie, 
+      která například popisuje redukce lambda termů (ty zajišťujou jak 
+      zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, 
+      díky tomu, že se různé stromy redukují na ten samej).
+      (ALT znění: 
+      pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, 
+      protže stačí brát v potaz třídy ekvivalence a ne zamotný termy 
+      - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy. )
+      (ALT2: 
+      Dost podstatný ospravedlnění lambda reprezentace je fakt, 
+      že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.)
+
+  (b) úspornost - abstrakce eliminací produkuje 
+      termy až kvadraticky větší než je původní 
+      term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
+      míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
+      podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
+      takto vnořeného podtermu složitě dopravit).
+
+  (c) použití kombů vyžaduje silnější typovej sytém - 
+      kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
+      lambdy v pohodě běžej na simply typed lc.
+
+  (d) použití proměnných je typické pro lidi, proto by se mohlo sát že
+      i programy generovaný systémem kterej s proměnejma počitá budou o něco 
+      "lidštější"
+
+Since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
+Pro more soffisticated type systems would be situation similar.
+
+Hlavním problémem lambda reprezentac proti kombinátorový je 
+jak se vypořádat s problémy zpusobenými použitím 
+localních proměných (aka jak křížit).
+
+}
+
 Binard and Felty use even 
@@
 \label{preliminaries}
+
+\red{TODO: ještě určitě projít a vylepšit; někde zestručnit někde rozšířit}
 
@@
 \subsection{Normal forms}
+
+\red{asi řít tady v ty subsekci co že to je normalizace termu jednou větou}
 
@@
 Generally speaking, the variable problem is caused by that the local variable is not defined in the new place or the variable is defined in the new place, but has
-some another type.
-
-Let us list some possible approaches to solving this problem:
-
-%Už přeloženo: 
-%Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.
-%Metody jak zacházet s lambdama:
-%- použít méně obecnou reprezentaci, takovym příkladem je Yu, která nepovoluje vnější proměný uvnitř lambda abstrakcí a další restrikce.
-Touto cestou nechceme jít, myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí normální formu žádného well-typed termu.
-%- křížit tak jako by se nic nedělo a problémy řešit až když nastanou: tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu.
-Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá prvek mutace.
-K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.
-%Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných.
-Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I), které zastanou puvodní roli proměnných.
-We idetified two methods using this approach:
+some another type. Let us list some possible approaches to solving this problem.
 
@@
 polymorphic combinators (i.e. $\Scomb$, $\Kcomb$ and $\Icomb$ 
-in the simple case of the algorithm).
-
-We identified two methods using this approach:
-
-\red{
-
-- Na konci generovací fáze všechny vygenerované termy převést do kombinatorové
-  podoby pomocí elmin abstr.
-Tento přístup můžeme chápat jako kompromis (najit lepší slovo, hibrid...) mezi lambda reprezentací a komb reprezentací.
-Nastává však otázka zda má pak smysl generovat termy v lc, když je pak stejně převedeme.
-(Výhodou je, že
-při generování v kombinátorech nemáme zajištěno že (pokud bychom nahradili S,K,I 
-za odpovídající termy a redukovlaito,tak že nevygenerujeme dvakrát různej term co se redukuje do toho samýho - což je ale takova divná výhoda kerou bych asi ani nepsal nebo nevim) Převod má totiž až kvadratický nárůst velikosti a dá se očekávat, že termy jsou při přímém generování v komb metodě generovany usporněji (čili to co sme nahnali na redukcích stratíme eliminací).
-
-- Jedince držet po celou dobu v be-nf (té nejúspornější co se týče 
-  beta,eta-redukcí) a pouze při křížení oba rodiče převést na kombinátorovou formu
-  eliminací abstrakcí. Po zkřížení je zase uvést do be-nf neobsahující přidané kombinátory tak, že jejich výskyty nahradíme jejich definicemi 
-  (e.g. K nahradime termem (\ x y . x) ) a tento term redukujeme 
-  do beta eta nf.
-Takový přístup využívá dříve nelibou vlastnost eliminace totiž kvadratický narust velikosti ku svému prospěchu, zde totiž k tomuto nárůstu dochází pouze dočasně a tím zvyšuje počet křížících bodů, což se ukazuje jako výhoda (citovat Yu kde při použití @-reprezentace místo s-výrazů dosahuje lepších výsledků).
+in the simple case of the algorithm). When there are no variables
+we can freely swap the subtrees of the same type.
+
+We identified two approaches that utilize abstraction elimination:
+
+The first one converts all the generated terms at the end of the population initialization phase. We call this approach to lambda term crossover 
+the \textit{hybrid}, because it can be understood as hybrid of the
+lambda calculus representation and the purely combinator representation
+used in \cite{kes}. Advantage of generating terms as \lnf lambda terms
+instead of generating them directly as combinator terms is that it reduces
+the search space, because there is more than one way to represent a \lnf 
+lambda term as combinator term (this can be seen considering that 
+\lnf term $\lamb{x}{x}$ is equivalent to both $\Icomb$ and 
+$\Scomb \Kcomb \Kcomb$), but every combinator term has only 
+one unique \lnf form. Unfortunately the disadvantage is also important;
+the abstraction elimination can cause up to a quadratic 
+increase in the term size. It is reasonable to suspect that combinator
+terms generated directly are probably smaller than those mechanically
+translated by the abstraction elimination. 
+The act of producing a small normalized term and translating it
+immediately to a bulky combinator form seems as a wasteful behavior.
+
+Thus we propose following crossover operator, which we call 
+the \textit{unpacking} crossover. Unlike the \textit{hybrid} it operates
+over \lterms and produces \lterm offspring in \benf.
+When two parent \lterms are about to be crossed, they are both
+converted to combinator terms using the elimination abstraction algorithm. 
+After the trees are swapped, the offspring terms are again 
+normalized to the \benf \lterms containing no temporary  
+combinators added during the abstraction elimination. 
+This normalization is achieved by replacement of all 
+occurrences of temporary combinators by their
+respective definitions (e.g. $\Kcomb$ is replaced by $\lamb{x y}{x}$) 
+followed by beta and eta normalization. 
+We believe that the property of increasing the term size is 
+advantageous for the \textit{unpacking} crossover.
+This belief is based on the fact that larger terms have more crossover
+points and on the observation that higher number of crossover points
+is beneficial for crossover operation, this observation will be discussed
+in more detail in the subsequent subsection. 
+\red{(TODO: opravdu jí tam diskutovat)}  
+The difference between the \textit{unpacking} and \textit{hybrid}
+crossover is that in the case of the \textit{unpacking}  
+this increase is only temporary stage, which is reversed
+by the normalization. 
+The most notable disadvantage of this crossover approach is
+that it is more time consuming than the simple ones. 
+\red{TODO: Až bude známo jak moc to žere čas, tak to sem připsat.
+a taky někde zmínit asi výslovně, že pro closed problémy je to 
+stejný jako kozovská pač vnich žadný abstrakce nejsou.
+Taky nějak vhodně zmínit analogii s přírodou.}
+
+% zbytky poznámek k unpacking crossoveru, 
+% ještě projít a dyštak připsat co se bude hodit :
+
+%-  Takový přístup využívá dříve nelibou vlastnost eliminace totiž kvadratický narust velikosti ku svému prospěchu, zde totiž k tomuto nárůstu dochází pouze dočasně a tím zvyšuje počet křížících bodů, což se ukazuje jako výhoda (citovat Yu kde při použití @-reprezentace místo s-výrazů dosahuje lepších výsledků).
 Díky redukcím jsou pak výsledné termy zase nejmenčí co v dáné "třídě ekvivalence" mohou být.
 Nevýhodou je, že takový operace žerou nějakej čas.
-Je však otázkou, do jaké míry je takový overhead (kouknout co toslovo znamena:D) podstatný, jelikož na druhou stranu
-  z keeping trees small přichází že to bude celý rychlejší a je otazka zda bude takový overhead podstatný v porovnání s overheadem pro výhodnocení termu, který je typicky nejdražší.
+Je však otázkou, do jaké míry je takový overhead (kouknout co toslovo znamena:D) podstatný, jelikož na druhou stranu z keeping trees small přichází že to bude celý rychlejší a je otazka zda bude takový overhead podstatný v porovnání s overheadem pro výhodnocení termu, který je typicky nejdražší.
 Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
-
 %(ALTernativní znění ze začátku: Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
@@
 
-}
-
-\subsection{Poznámky k uvodu ke křížení}
+\subsection{Typed subtree swapping in greater detail}
+\label{typed-swapping}
+
 \red{
@@
 
-CO S @-stromama řešit je explicitně tady, případně někde jinde?
-Pro generování nemá reprezentace vliv, ta má vliv až vpři křížení, kde at-tree zajišťuje víc míst.
-We will show how to crossover \textit{typed \lterm trees} in both \atTree and \sexprTree notation.
-
-to o tom že chem zobecnit tu standardní furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje, někde to zmínit že se to na ně nevztahuje
+Tuhle subsekci ještě pořádně projít, zatim je to tu skopčený z diplomky.
+
+Na tuhle subsekci se odkazuje předchozí text o unpacking s tim že je tu rozebranej vliv
+počtu křižicích bodů na kvalitu křížení, takže to tu rozebrat, tzn. musi se odiskutovat
+rozdíl mezi @-tree a S-expresion.
+
 }
-
-\subsection{Typed subtree swapping in greater detail}
-\label{typed-swapping}
 
@@
 \label{experiments}
+\red{TODO!!}
 

### 2014-01-30 tex/gecco14/gecco14.tex
gecco finiš-him!

@@
 \newcommand{\gar}{\longmapsto}
+\newcommand{\bRedex}{$\beta$-redex\xspace}
+\newcommand{\bRedexes}{$\beta$-redexes\xspace}
+\newcommand{\bArrow}{\rightarrow_\beta\xspace}
+\newcommand{\eArrow}{\rightarrow_\eta\xspace}
+\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
 
@@
 
-\numberofauthors{2}
-
-\author{
-% 1st. author
-\alignauthor
-Tom\'{a}\v{s} K\v{r}en\\
-  \affaddr{Faculty of Mathematics and Physics}\\
-  \affaddr{Charles University in Prague}\\
-  \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
-  \affaddr{Prague, Czech Republic}\\
-  \email{tomkren@gmail.com}
-% 2nd. author
-\alignauthor
-Roman Neruda\\
-  \affaddr{Institute of Computer Science}\\
-  \affaddr{Academy of Sciences of the Czech Republic}\\
-  \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
-  \affaddr{Prague, Czech Republic}\\
-  \email{roman@cs.cas.cz}
-}
+% \numberofauthors{2} 
+%
+%  \author{
+%  % 1st. author
+%  \alignauthor
+%  Tom\'{a}\v{s} K\v{r}en\\
+%    \affaddr{Faculty of Mathematics and Physics}\\
+%    \affaddr{Charles University in Prague}\\
+%    \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
+%    \affaddr{Prague, Czech Republic}\\
+%    \email{tomkren@gmail.com}
+%  % 2nd. author
+%  \alignauthor
+%  Roman Neruda\\
+%    \affaddr{Institute of Computer Science}\\
+%    \affaddr{Academy of Sciences of the Czech Republic}\\
+%    \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
+%    \affaddr{Prague, Czech Republic}\\
+%    \email{roman@cs.cas.cz}
+%  }
 
@@
 Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions.
+In this paper we discuss advantages of this representation coming from the use of reductions (beta and eta) and a way to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm.
+We discuss the role of those reductions in the process of generating initial population and propose several crossover approaches, two of which are based on abstraction elimination and compared using the even parity benchmark problem.
+%including novel approach to crossover operator based both on reductions and abstraction elimination. 
+%The design goal of this operator is to turn the disadvantage of abstraction elimination  - possibly quadratic increase of program size - into a virtue; our approach leads to more crossover points.
+At the same time, utilization of reductions provides offspring of small sizes.
+\end{abstract}
+
+%Starej abstrakt než se zjistilo že unpacking neni tak super:
+%Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions.
 In this paper we discuss advantages of this representation coming from use of reductions (beta and eta) and how to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm.
@@
 At the same time, utilization of reductions provides offspring of small sizes.
-\end{abstract}
 
 % A category with the (minimum) three required fields
-\category{H.4}{\red{TODO}}{Miscellaneous}
+\category{I.2.2}{Artificial Intelligence}{Automatic Programming}[Program synthesis]
 %A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{\red{TODO}}
-
-\keywords{\red{TODO, ToDo, todo}}
+\category{F.4.1}{Mathematical Logic and Formal Languages}{Mathematical Logic}[Lambda calculus and related systems]
+
+\terms{Experimentation, Languages, Theory}
+
+\keywords{genetic programming, lambda calculus}
 
@@
 Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
-Use of types naturally opens door to enriching S-expressions,
+Usage of types naturally opens door to enriching S-expressions,
 the traditional GP representation of individuals, with concepts from
-lambda calculus, which is simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts has shown to be successful \cite{yu01}.
-
-\red{TODO....}
+lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts have shown to be successful \cite{yu01}.
+
+One of the motivations for using lambda calculus is that it is backed by a considerable theoretical background.
+From this background we can utilize several useful concepts such as reductions, normalization or abstraction elimination, and use them to enhance the standard GP algorithm.
+We use normalization to reduce the size of the search space of program trees and the abstraction elimination to prevent
+the difficulties caused by the presence of variables in program trees.
+
+In our system we use simply typed lambda calculus.
+Since abstraction elimination and reductions do not involve many type related issues it seems to us as a sufficient choice of type system.
+Techniques described here would be analogous for a more sophisticated type system and can be directly applied to such systems as the Hindley–Milner type system.
+
+%\red{Možná někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP esli to neni blbý..}
 
@@
 
-\red{ve shrnutí paperu říct, že jeho tělo je tvořeno dvěma částma,
-první se zabejvá generovánim a jeho vztahem k redukcím, druhá řeší křížení
-s velkym důrazem na vyřešení problémů s lokálníma proměnejma.}
-
 The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
-Main original results about search strategies in individual generating are described in section~\ref{approach}.
-Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.
-
-\subsection{\red{Randombits}}
-\red{
-
-- Možná někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP..
-
-}
+The body of the paper is presented in sections \ref{generating} and \ref{crossover}; the first discusses the term generating 
+procedure and its relations
+with reductions, the second describes several approaches to lambda term
+crossover and how they solve the variable problem. Section~\ref{experiments} examines those ideas on the even parity benchmark problem, while the paper is concluded by section~\ref{conclusions}.
 
@@
 \label{related}
-
-\red{TODO zatim jen zkopčený z minula a dodaná diskuze 
-zatim česky porovnávající nás a kombinatoři , 
-opravit podle novýho kontextu}
 
@@
 input parameter.} and lambda abstractions  \cite{yu01}.
-Important point of interest in this work is use of
-\texttt{foldr} function as a tool for \textit{implicit recursion},
-i.e. recursion without explicit recursive calls. 
-The terminal set for constructing lambda abstraction subtrees 
-is limited to use only constants and variables of that particular
-lambda abstraction, i.e., outer variables are not allowed to be used
-as terminals in this work. This is significant difference from our approach 
-since we permit all well-typed normalized \lterms. From this difference also
-comes different crossover operation. We focus more on term generating process; 
-their term generation is performed in a similar way as the standard one, 
-whereas our term generation also tries to utilize techniques of systematic enumeration.
-
-Briggs and O’Neill present technique 
+An important point of interest in the work is use of
+\texttt{foldr} function as a tool for \textit{implicit recursion}, i.e.
+recursion without explicit recursive calls.
+The terminal set for constructing lambda abstraction subtrees  is limited to use only constants and variables of that particular lambda abstraction, i.e., outer variables are not allowed to be used as terminals in this work.
+This is a significant difference from our approach since we permit all well-typed normalized \lterms.
+This difference also causes the need for different crossover operators.
+
+Briggs and O’Neill present a technique 
 utilizing typed GP with combinators \cite{kes}.
-The difference between approach presented in this work
-and our approach is that in this work terms are generated
-straight from \textit{library} of combinators and no lambda abstractions
-are used. They are using more general polymorphic type system than us
--- the Hindley–Milner type system. They also discuss the 
-properties of exhaustive enumeration of terms and compare it with GP search.  
-They also present interesting concept of \textit{Generalized
-genetic operator} based on term generation.
-
-\red{
-
-POZNAMKY PRO DISKUSI POROVNAVAJICI NÁS A BRIGGSE:
-
-%Reprezentace v klasickym kozovy je S-expression.
-
-The individual representation used in standard GP is S-expression. 
-Basically there are two general approaches of generalizing S-expressions
-into trees of better expressive power suitable for typed GP with higher 
-order functions: We can enrich the method with the polymorphic combinators \cite{kes}
-or we can enrich the method with the lambda abstractions and local variables
-\cite{yu01}.
-
-Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
-
-Zdá se že použití kombinátorů má řadu výhod (vymenovat nějaký co jmenuje briggs), 
-ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
-taky za zkoužku, což podporují nasledující argumenty:
-
-  (a) motivace pro práci s lambda termama je, že za nima košatá teorie, 
-      která například popisuje redukce lambda termů (ty zajišťujou jak 
-      zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, 
-      díky tomu, že se různé stromy redukují na ten samej).
-      (ALT znění: 
-      pohovořit o tom, že normální formy zmenšujou prohledávaný prostor, 
-      protže stačí brát v potaz třídy ekvivalence a ne zamotný termy 
-      - reprezentanti těchto tříd ekvivalence jsou pak právě normální formy. )
-      (ALT2: 
-      Dost podstatný ospravedlnění lambda reprezentace je fakt, 
-      že redukce (beta, eta) šlapou právě jen při tom, když máme lambdy.)
-
-  (b) úspornost - abstrakce eliminací produkuje 
-      termy až kvadraticky větší než je původní 
-      term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
-      míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
-      podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí do
-      takto vnořeného podtermu složitě dopravit).
-
-  (c) použití kombů vyžaduje silnější typovej sytém - 
-      kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
-      lambdy v pohodě běžej na simply typed lc.
-
-  (d) použití proměnných je typické pro lidi, proto by se mohlo sát že
-      i programy generovaný systémem kterej s proměnejma počitá budou o něco 
-      "lidštější"
-
-Since abstr-elim a redukce neinvolvujou moc typově rilejtlý problémy, je ok celou věc demostrovat na jednoduchém typovém systému jakám je simply typed lc.
-Pro more soffisticated type systems would be situation similar.
-
-Hlavním problémem lambda reprezentac proti kombinátorový je 
-jak se vypořádat s problémy zpusobenými použitím 
-localních proměných (aka jak křížit).
-
-}
-
-Binard and Felty use even 
-stronger type system (\textit{System F}) \cite{binard2008genetic}.  
-But with increasing power of the type system comes increasing difficulty of term generation.
-For this reason evolution in this work takes interesting and nonstandard shape 
-(fitness is associated with \textit{genes} which are evolved together with \textit{species}
-which together participate in creation of individuals).
-This differs from our approach, which tries to be generalization of
-the standard GP\cite{koza92}.
-
-In contrast with above mentioned works our approach uses very simple type system 
-(simply typed lambda calculus) and concentrates on process of generation  
-able to generate all possible well-typed normalized lambda terms. In order to do
-so we use technique based on \textit{inhabitation machines} 
-described by Barendregt \cite{barendregt10}.
+The difference between approach presented in their work
+and our approach is that they generate terms in a straightforward way directly from the library of combinators, without any use of lambda abstractions.
+They use the Hindley–Milner type system.
+They also present interesting concept of \textit{Generalized genetic operator} based on term generation.
+The combinator based genetic algorithm has the advantage of
+avoiding the need of dealing with variables.
+
+Binard and Felty use even stronger type system (\textit{System F}) \cite{binard2008genetic}.
+But with the increasing power of the type system also comes an increasing difficulty of term generation.
+For this reason, the evolution in this work takes interesting and nonstandard shape (fitness is associated with \textit{genes} which are evolved together with \textit{species} which together participate in creation of individuals).
+This differs from our approach, which tries to be a generalization of the standard GP\cite{koza92}.
+
+In contrast with the above mentioned works, our approach uses very simple type system (simply typed lambda calculus) and concentrates on process of generation  
+able to generate all possible well-typed normalized lambda terms. In order to do so we use technique based on \textit{inhabitation machines} described by Barendregt \cite{barendregt10}.
 
@@
 \label{preliminaries}
-
-\red{TODO: ještě určitě projít a vylepšit; někde zestručnit někde rozšířit}
 
@@
 well known from common programming languages. 
-For example in JavaScript 
+For example, in JavaScript, 
 $(M~N)$ translates to expression \texttt{$M$($N$)} and
@@
 For better readability, 
-$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for
-$(\dots((M_1~M_2)~M_3)~\dots~M_n)$
-and $\lam{x_1 x_2 \dots x_n }{M}$ for 
+$M_1~M_2~M_3~\dots~M_n$ will be an abbreviation for
+$(\dots((M_1~M_2)~M_3)~\dots~M_n)$,
+and $\lam{x_1 x_2 \dots x_n }{M}$ will be an abbreviation for 
 $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
 
-\subsection{$\beta$-reduction}
-
-In order to perform computation there must be some
+\subsection{Reductions}
+
+In order to perform computation, there must be some
 mechanism for term evaluation. In $\lambda$-calculus there
-is $\beta$-reduction for this reason.\\
-
-\newcommand{\bRedex}{$\beta$-redex\xspace}
-\newcommand{\bRedexes}{$\beta$-redexes\xspace}
-\newcommand{\bArrow}{\rightarrow_\beta\xspace}
-\newcommand{\eArrow}{\rightarrow_\eta\xspace}
-\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
+is a \mbox{$\beta$-reduction} procedure for this reason.\\
 
@@
 
-\subsection{$\eta$-reduction}
-
-Similarly as for $\beta$-reduction we can define $\eta$-reduction 
-except that instead of \ref{eq:bRed} we use:
-
-$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$~
+Similarly as for $\beta$-reduction we can define $\eta$-reduction except that now it is defined as follows.  
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$
 
@@
 (Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)
-
 Similarly as for $\bbarr$ and $\beq$ we can define relations 
-$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.
-
-\subsection{$\eta^{-1}$-reduction}
+$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.\\
 
@@
 the reduction converse to $\eta$-reduction.
-Again it may be obtained by replacing \ref{eq:bRed}, now with:
-
+It is defined as follows.  
 $$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$
@@
 \subsection{Normal forms}
-
-\red{asi řít tady v ty subsekci co že to je normalizace termu jednou větou}
-
-~\begin{easylist}[enumerate]
+\begin{definition}~
+
+\begin{easylist}[enumerate]
 & A \lterm is a \textit{$\beta$-normal form} (\bnf) 
@@
 \end{easylist}
+\end{definition}
 A normal form may be thought of as a result of a term evaluation. 
+Normalization of the term $M$ is the process of finding normal form of $M$.
 
@@
 
-\subsection{Types etc}
+\subsection{Types}
 
@@
 Context is a basic type theoretic concept suitable as a typed alternative
-for terminal and function set in standard GP. 
+for terminal, and function set in standard GP. 
 Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
@@
 
-Přeformulovat aby se to hodilo do tohodle kontextu....
 Our goal in term generation is to produce terms $M$
@@
 
-This is simple version of this process. More optimized version,
-in the means of the size of resulting term and its performance
-is following one, presented in \cite{jones87}.\\
-
 As is stated in \cite{jones87},
@@
 term is often much larger than in its lambda form --- the size of
-the translated term can be proportional to the
-square of the size of the original term.
-
-But the advantage is also tempting --- no need to deal with variables
-and lambda heads.
-
-\red{tady je to v původnim kontextu, odebrat slova co se tam nehodí}
+the translated term can be proportional to the square of the size 
+of the original term. But the advantage is also tempting --- no need to deal with variables and lambda abstractions.
+
+The algorithm presented here is a simple version of this process. 
+More optimized version (used in our \textit{hybrid} crossover described below), by means of the size of resulting term and its time performance is presented in \cite{jones87}.
 
 \section{Term generating and reductions}
+\label{generating}
 
@@
 
-In \cite{barendregt10} is shown term generating grammar with 
-following rules.
-
-\red{předělat ten zápis, víc to okecat, říct že to je jakoby gramatika, 
-formálně 2-level-gramatika do footnoty, pač má nekonečně neterminálu  
-i terminálů. že neterminaly jsou dvojice (typ, kontext) a že terminaly jsou
-symboly kterejma se zapisujou lambda termy}
-
+In \cite{barendregt10}, the following term generating 2-level grammar\footnote{2-level grammar (unlike a standard grammar) can have infinitely many non-terminals and its rule is better described as \textit{rule schema}.
+The original grammar described in \cite{barendregt10} has $( \sigma \rightarrow \tau , \Gamma ) \gar (~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)$ as second rule, but our grammar is equivalent -- it packs consecutive uses of this rule into our rule.} for generating all possible
+terms in their \textit{long normal form} is described.
+
+If we want to generate a term $M$ satisfying $\tur{\Gamma}{M}{\tau}$,
+then the initial non-terminal is $(\tau, \Gamma)$.
+
+For every $\alpha \in A$ and context $\Gamma$ and $f$ such that \\
+$(f :$ $\rho_1 \ar \dots \ar \rho_m \ar \alpha$ $) \in \Gamma$ 
+we have a rule of the first form:
+$$
+( \alpha , \Gamma ) \gar 
+(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
+$$
+
+In other words, one possible way to construct a term of 
+the atomic type $\alpha$ from the building symbols $\Gamma$
+is to take a function symbol $f$ from $\Gamma$ with 
+the return type $\alpha$ (or if $m=0$ a constant of the type $\alpha$),
+put it into the root of the term and to generate 
+a subtrees of appropriate types for each of $f$'s arguments
+from the building symbols $\Gamma$.
+
+For every atomic type $\alpha$, context $\Gamma$ and types
+$\tau_1 , \dots , \tau_n$ (where $n > 0$) 
+we have a rule of the second form:\\
+
+$
+( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
+ \\ ~~~~~~~~\gar
+(~\lambda~x_1~\dots~x_n~.~
+( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
+$~\\
+
+In other words, the way to construct a term of 
+the function type $\tau_1 \ar \dots \ar \tau_n \ar \alpha$ 
+from the building symbols $\Gamma$
+is to create an anonymous function
+with fresh variables $x_1, \dots , x_n$ in its head 
+whose body is constructed as a term that has type $\alpha$
+and is constructed using symbols from $\Gamma$ 
+enriched with those new variables $x_1, \dots , x_n$,
+which are associated with appropriate types.
+
+% TODO dyštak zmínit že sme druhý pravidlo nahradili fikanou zkratkou
+%The second rule can be replaced by more effective one.
+%This rule packs consecutive uses of the second rule into one use. This is valid since the use of the second rule is deterministic; it is used if and only if the non-terminal's type is not atomic.
+
+Long normal form is a generalization of S-expressions into lambda calculus:
+Let $\Gamma$ be a context satisfying the closure requirement, then only
+the first rule will be applicable. 
+$(\alpha, \Gamma)$ rewrites to $(~f~( \alpha, \Gamma )~\dots~( \alpha, \Gamma )~)$.
+One can see that $f$ stands for 
+the root node and that each $( \alpha , \Gamma )$ will produce a direct subtree.
+
+For a context containing at least one higher order function 
+(or if the desired type of generated terms is a function type) we get special kind of root node containing lambda head with one direct subtree standing for body of the lambda function.
+Local variables defined in the lambda head may also appear in this body.
+
+Discussion of what is the best strategy for browsing the search space given by above described grammar in order to generate \lnf terms is beyond the scope of this paper.
+Standard \textit{ramped half-and-half} strategy can be used for this purpose, or some other,
+which we discuss in \cite{nasecec}. In our experiments we use the \textit{geometric} strategy proposed in 
+\cite{nasecec}.
+
+\subsection{Benefits of generating \lterms in \lnf}
+\label{benefits}
+
+By generating \lterms in \textit{lnf} we avoid generating 
+\lterms $M$,$N$ such that $M \not= N$, but $M =_{\beta\eta} N$.
+In other words, we avoid generating two programs with different 
+source codes, but performing the same computation.
+
+Every \lterm $M$ such that \GMS has its unique \lnf $L$, 
+for which $L =_{\beta\eta} M$.
+Therefore, the computation performed by \lterm $M$ 
+is not omitted, because it is the same computation
+as the computation performed by \lterm $L$. \\
+
+Generating \lterms in \lnf is even better than generating 
+\lterms in \bnf. Since \lnf is the same thing as \beenf,
+every \lterm in \lnf is also in \bnf.
+
+This comes straight from the definition of \beenf, 
+but one can also see it by observing the method for generating
+terms in \bnf. As shown in \cite{barendregt10}, 
+this method is obtained by the following 2-level grammar.
 \begin{align*}
-( \alpha , \Gamma )  
+( \pi , \Gamma )  
 &\gar
 (~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
-\\& \textbf{if } \alpha \in A,
-(f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
 \\ 
@@
 &   
+\\& \textit{where~~~} (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
 \end{align*}
 
-The second rule can be replaced by more effective one.
-$ 
-( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
-\gar \\
-(~\lambda~x_1~\dots~x_n~.~
-( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
-~~~~ \textbf{if } n > 0
-$
-
-This rule packs consecutive uses of the second rule into one use.
-This is valid since the use of the second rule is deterministic;
-it is used if and only if the non-terminal's type is not atomic.
-
-Long normal form is a generalization of S-expressions into lambda calculus:
-Let $\Gamma$ be a context satisfying the closure requirement, then only
-the first rule will be applicable. 
-$(\alpha, \Gamma)$ rewrites to $(~f~( \alpha, \Gamma )~\dots~( \alpha, \Gamma )~)$.
-One can see that $f$ stands for 
-the root node and that each $( \rho_i , \Gamma )$ will produce a direct subtree.
-
-\red{
-For context containing at least one higher order function we get special kind of root node containing lambda head with one direct subtree standing for body of 
-the lambda function, v tomto tele se navíc mohou objevit proměnné 
-definované v této hlavě (asi doplnit obrázkama).
-
-Diskuze toho jakou přesnou strategií je nejvhodnější generovat
-termy v lnf je beyond the scope of this paper, mužeme například použít
-klasickou raped half-and half strategii. Footnotenout kraťoučkou
-zmínku o gemetrický. 
-}
-
-\subsection{Benefits of generating \lterms in \lnf}
-\label{benefits}
-
-By generating \lterms in \textit{lnf} we avoid generating 
-\lterms $M$,$N$ such that $M \not= N$, but $M =_{\beta\eta} N$.
-In other words, we avoid generating two programs with different 
-source codes, but performing the same computation.
-
-Every \lterm $M$ such that \GMS has its unique \lnf $L$, 
-for which $L =_{\beta\eta} M$.
-Therefore the computation performed by \lterm $M$ 
-is not omitted, because it is the same computation
-as the computation performed by \lterm $L$. \\
-
-Generating \lterms in \lnf is even better than generating 
-\lterms in \bnf. Since \lnf is the same thing as \beenf,
-every \lterm in \lnf is also in \bnf.
-
-This comes straight from the definition of \beenf, 
-but one can also see it from observing method for generating
-terms in \bnf. As is shown in \cite{barendregt10}, 
-this method is obtained simply by replacing  
-the first grammar rule by slightly modified rule,
-resulting in the following grammar.
-\begin{align*}
-( \pi , \Gamma )  
-&\gar
-(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
-\\& \textbf{if } (f : \rho_1 \ar \dots \ar \rho_m \ar \pi) \in \Gamma
-\\ 
-( \sigma \rightarrow \tau , \Gamma )  
-&\gar
-(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
-&   
-\end{align*}
-
-The difference lies in that $f$'s type is no longer needed to be fully expanded
-($\pi \in \T$ instead of $\alpha \in A$). This makes the grammar less deterministic,
+This grammar differs from the grammar for \lnf only in the first rule\footnote{We use a different second rule for \lnf grammar, but as is mentioned in footnote above, it also works with this second rule.}.
+Specifically, it differs in that $f$'s type is no longer needed to be fully expanded ($\pi \in \T$ instead of $\alpha \in A$).
+This makes the grammar less deterministic,
 resulting in a bigger search space. The new rule is generalization of the old one,
@@
 \lterms $M$,$N$ such that $M \not= N$ and $M =_{\eta} N$; 
-but by generating in \bnf we do not avoid it.\\
+but generating in \bnf does not have such a property..\\
 
@@
 Terms in \lnf are said to be \textit{fully $\eta$-expanded} \cite{barendregt10}. 
-Relevant property of $\eta$-reduction is that it always shortens the term
-that is being reduced be it. And conversely, $\eta$-expansion prolongs.
+A relevant property of $\eta$-reduction is that it always shortens the term
+that is being reduced by it. And conversely, $\eta$-expansion prolongs.
 $$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$
 
-Now we show that for every \lterm $M$ 
+Now we show that for every \lterm $M$, 
 every sequence of $\eta$-reductions is finite and
-leads to unique \enf $N$.
+it leads to a unique \enf $N$.
 
@@
        end at some point. Thus every \lterm has \enf.
- \item Since \ered is \textit{Church–Rosser}, \enf is unique (see \cite{barendregt84}). 
+ \item Since \ered is \textit{Church–Rosser}, \enf is unique (see~\cite{barendregt84}). 
 \end{enumerate}
@@
 
-For better clarity \lets show the \ered and \bredex using trees.\\
-\ered: \\
-
-\Tree [.$\lh{x}$ [.@ $M$ $x$ ] ] 
-~~~~$\eArrow$~~~~
+For better clarity \lets show the \ered and \bredex using \atTree{}s 
+(a \lterm representation derived from the definition of \lterms 
+where @ stands for \textit{function application}).\\
+\textit{\ered}: \Tree [.$\lh{x}$ [.@ $M$ $x$ ] ] 
+$\eArrow$
 $M$ 
-
-And \bredex:
-
-\Tree [.@ [.$\lh{x}$ $M$ ] $N$ ] \\
+~~~~~~~~~~~
+\textit{\bredex}: \Tree [.@ [.$\lh{x}$ $M$ ] $N$ ] \\
 
 \Lets assume that $P \eArrow Q$ creates a new \bredex $B$ in $Q$.
-
 Since \ered only destroys and never creates \textit{function applications} (i.e. @),
@@
 
-Suppose we have systematic (i.e. gradually generating all terms, 
-but no term twice) method for generating terms in \lnf,
-we may transform it to systematic method for generating terms in \benf
-by simply reducing each generated term to its \enf.
+Suppose we have a systematic method (i.e. gradually generating all terms, but no term is generated twice) for generating terms in \lnf,
+we may transform it to a systematic method for generating terms in \benf by simply reducing each generated term to its \enf.
 
 \section{Crossover operator}
+\label{crossover}
 
@@
 
-As in standard GP our crossover will be performed by swapping
-two subtrees. But now with constraint that both subtrees have
-the same type.
+As in standard GP, our crossover will be performed by swapping
+two subtress of the same type.
 
 Variables bring more difficulties then types do.
-This problem arises from variables that are free in subterms corresponding 
-to swapped subtrees.
-
-Following example illustrates the problem. \Lets have these two
+The problem arises from variables that are free in subterms corresponding to swapped subtrees.
+
+The following example illustrates the problem. \Lets have these two
 parent trees with selected nodes in bold.\\
@@
 
-Generally speaking, the variable problem is caused by that the local variable is not defined in the new place or the variable is defined in the new place, but has
-some another type. Let us list some possible approaches to solving this problem.
+Generally speaking, the variable problem is caused by the fact the local variable is not defined in the new place, or the variable is defined in the new place, but has
+some a different type. Let us list some possible approaches to solving this problem.
 
@@
 is limited to use only constants and variables of that particular
-lambda abstraction and an appropriate variable naming convention is
-used in order to prevent troubles with variables.
+lambda abstraction, and an appropriate variable naming convention is used in order to prevent troubles with variables.
 We choose not to follow such approach because we do not want to
@@
 defects when they occur. Such correction can be performed by 
-renaming the problematic variable to some another variable defined in that place 
-or, if there is no suitable candidate, to replace its occurrence with 
-a newly generated term of required type. This approach seems to us as a very
-unwieldy one -- there are clearly problems for which such collisions would
+renaming the problematic variable to some other variable defined in the new local scope of swapped subtree.
+Or, if there is no suitable candidate, to replace its occurrence with a newly generated term of the required type.
+This approach seems to us as a very unwieldy one -- there are clearly problems for which such collisions would
 occur very often and when there is no suitable rename candidate the 
@@
 
-We identified two approaches that utilize abstraction elimination:
+We identified two approaches that utilize abstraction elimination.
+
+\subsection{Hybrid crossover}
 
 The first one converts all the generated terms at the end of the population initialization phase. We call this approach to lambda term crossover 
-the \textit{hybrid}, because it can be understood as hybrid of the
+\textit{hybrid}, because it can be understood as hybrid of the
 lambda calculus representation and the purely combinator representation
-used in \cite{kes}. Advantage of generating terms as \lnf lambda terms
+used in \cite{kes}. The advantage of generating terms as \lnf lambda terms
 instead of generating them directly as combinator terms is that it reduces
 the search space, because there is more than one way to represent a \lnf 
-lambda term as combinator term (this can be seen considering that 
+lambda term as a combinator term (this can be seen considering that 
 \lnf term $\lamb{x}{x}$ is equivalent to both $\Icomb$ and 
 $\Scomb \Kcomb \Kcomb$), but every combinator term has only 
-one unique \lnf form. Unfortunately the disadvantage is also important;
+one unique \lnf form. Unfortunately, one disadvantage is also important;
 the abstraction elimination can cause up to a quadratic 
 increase in the term size. It is reasonable to suspect that combinator
-terms generated directly are probably smaller than those mechanically
+terms generated directly may be smaller than those mechanically
 translated by the abstraction elimination. 
 The act of producing a small normalized term and translating it
-immediately to a bulky combinator form seems as a wasteful behavior.
-
-Thus we propose following crossover operator, which we call 
-the \textit{unpacking} crossover. Unlike the \textit{hybrid} it operates
-over \lterms and produces \lterm offspring in \benf.
+immediately to a bulky combinator form can be seen as a wasteful behavior.
+
+\subsection{Unpacking crossover}
+
+Thus, we propose the following crossover operator in order to be able to analyze the above mentioned issues, which we call the \textit{unpacking} crossover.
+Unlike the \textit{hybrid} it operates over \lterms and produces \lterm offspring in \benf.
 When two parent \lterms are about to be crossed, they are both
@@
 in more detail in the subsequent subsection. 
-\red{(TODO: opravdu jí tam diskutovat)}  
 The difference between the \textit{unpacking} and \textit{hybrid}
 crossover is that in the case of the \textit{unpacking}  
-this increase is only temporary stage, which is reversed
+this increase is only in the temporary stage, which is reversed
 by the normalization. 
 The most notable disadvantage of this crossover approach is
-that it is more time consuming than the simple ones. 
-\red{TODO: Až bude známo jak moc to žere čas, tak to sem připsat.
-a taky někde zmínit asi výslovně, že pro closed problémy je to 
-stejný jako kozovská pač vnich žadný abstrakce nejsou.
+that it is more time consuming than the simple one.
+
+%\red{TODO: Až bude známo jak moc to žere čas, tak to sem připsat.
+a taky někde zmínit asi výslovně, že pro closed problémy je to stejný jako kozovská pač vnich žadný abstrakce nejsou.
 Taky nějak vhodně zmínit analogii s přírodou.}
@@
 
-\red{
-POZNAMKY:
-
-Tuhle subsekci ještě pořádně projít, zatim je to tu skopčený z diplomky.
-
-Na tuhle subsekci se odkazuje předchozí text o unpacking s tim že je tu rozebranej vliv
-počtu křižicích bodů na kvalitu křížení, takže to tu rozebrat, tzn. musi se odiskutovat
-rozdíl mezi @-tree a S-expresion.
-
-}
-
-First thing to do in standard subtree swapping is to select random node
-in the first parent.
-
-We modify this procedure so that we allow
-selection only of those nodes with such a type that there exists  
-a node in the second parent with the same type.
-
-Standard subtree swapping crossover as a first thing selects 
-whether the selected node will be inner node (usually with probability 
-$p_{ip} = 90\%$) or leaf node (with probability 10\%).
-
-We are in a more complicated situation, because one of those 
+The first thing to do in a standard subtree swapping crossover is to select random node in the first parent.
+
+We modify this procedure so that we allow selection only of nodes with such a type that there exists a node in the second parent with the same type.
+
+The standard subtree swapping crossover selects whether the selected node will be inner node (usually with probability $p_{ip} = 90\%$) or leaf node (with probability 10\%).
+
+We are in a more complicated situation, since one of those 
 sets may be empty, because of allowing only nodes with possible "partner"
@@
 After selecting a node in the first parent we select node in the
-second parent such that type of that node must by the same as the type 
+second parent such that the type of that node must correspond to the type of
 of the first node. Again, this may eliminate the "90-10" step of
@@
 If the abstraction elimination was performed, then 
-since the trees are of the same type and there are no variables to be 
+since the trees are of the same type, and there are no variables to be 
 moved from their scope, the offspring trees are well typed.\\
@@
 Both \sexprTree and \atTree are able to be crossed by this 
-mechanism. But \atTree has more possibilities then \atTree.
+mechanism. But \atTree has more possibilities than \sexprTree.
 This comes from the fact that every subtree of the \sexprTree
@@
 
-Following example should clarify this.
+The following example should clarify this.
 
@@
 
-In \atTree, \textbf{f} is leaf thus subtree, 
-whereas in \sexprTree it is internal node thus not a subtree.\\
+In the \atTree, \textbf{f} is leaf, thus it is a subtree, 
+whereas in \sexprTree it is an internal node and thus not a subtree.
+Those favorable properties of \atTree{} are also reported in \cite{yu1998polygp}.\\
 
@@
 
-Again, similarly as for standard version, 
+Again, similarly to standard GP, 
 a maximum permissible depth $D_{created}$ 
@@
 this offspring is replaced by the first parent in the result of 
-the crossover operator. If both offspring exceeds this limit, than 
-both are replaced by both parents.
+the crossover operator. If both offspring exceed this limit, then 
+both are replaced by parents.
 
 For \atTree the $D_{created}$ must be larger since
-\atTree (without lambdas) is a binary tree. This 
+\atTree (without lambdas) is a binary tree. Then this 
 enlargement is approximately proportionate to average number of 
@@
 
+~\\
+
 \section{Experiments}
 \label{experiments}
-\red{TODO!!}
+
+We have tested our approach on the well-known even parity problem.
+
+We made two experiments, in the first we have used the \textit{hybrid} crossover, while in the second we have used the \textit{unpacking} crossover.
+
+Each experiment consisted of 50 independent runs 
+of GP algorithm. Each run had maximally 51 generations and population size
+was 500 individuals.
+In the experiments we analyze the fitness of the best individual,
+the average size of an individual,
+the ability of the system to produce a correct solution 
+and computational cost estimates throughout generations. 
+For the last two metrics we use the popular measurement  
+methods within the GP field --- the \textit{performance curves}
+described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
+and $I(M,i,z)$ (the total number of individuals that must be processed 
+to yield a correct solution with probability $z =99\%$).
+Difference of those two crossover methods is statistically analyzed by the Welch t-test by comparing the fitness values of the best individuals of the run. 
+Graphs showing mean values of the fitness of the best individual and
+the average size of an individual throughout generations contain error bars representing the standard error of the mean (SEM).
+
+The even parity function is a boolean function taking as inputs $N$
+boolean values and returning \textit{True} if an even number of inputs 
+are \textit{True}. For odd number it returns \textit{False}.
+This problem has been used by many researchers
+as a benchmark for GP. 
+We compare our results with that presented in \cite{yu01, kes}.
+We use very similar set of \textit{building symbols} as in \cite{yu01}. 
+\begin{align*}
+\tau = [Boo&l] \ar Bool\\
+\Gamma = \{
+  and   &: Bool \ar Bool \ar Bool                              ,\\
+  or    &: Bool \ar Bool \ar Bool                              ,\\
+  nand  &: Bool \ar Bool \ar Bool                              ,\\
+  nor   &: Bool \ar Bool \ar Bool                              ,\\
+  foldr &: (Bool \ar Bool \ar Bool) \\
+        &~~~\ar Bool \ar [Bool] \ar Bool ,\\
+  head' &: [Bool] \ar Bool                                   ,\\
+  tail' &: [Bool] \ar [Bool]                              \}
+\end{align*}
+
+The type $[Bool]$ stands for \textit{list of $Bool$s} and for purpose of
+this problem is considered atomic.
+Unlike in \cite{yu01}, we use specific instance of polymorphic 
+function \texttt{foldr}. 
+And modifications of functions \texttt{head} 
+(returning the first element of the list) 
+and \texttt{tail} (returning the list without the first element) are used; 
+making them total by returning default value \textit{False}
+and \texttt{[]}, respectively.
+We use the same fitness function as in \cite{yu01}. 
+The fitness function examines the individual by giving
+it all possible boolean lists of length 2 and 3.
+
+\begin{figure}[h!]
+  \centering
+  \caption{Graphs for Even parity problem.}
+  \includegraphics[scale=0.251]{grafy/EP.PNG}
+\end{figure}
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|cc|}
+\hline
+& Unpacking & Hybrid \\
+\hline
+Mean & 0.935000	& 0.930000 \\
+SD	 & 0.066768	& 0.078490 \\
+SEM	 & 0.009442	& 0.011100 \\
+N	 & 50    & 50    \\
+\hline
+t-value &  0.3431 & $\alpha = 0.05$\\
+p-value &  0.7323 &  \textbf{Not statistically significant}\\
+\hline
+\end{tabular}
+\caption{Statistical analysis.}
+\end{table}~\\
+
+Figure~1 shows results of this experiments. 
+Table~1 summarizes statistical analysis
+of the fitness values of the best individuals of the run 
+(\textit{p-value} $ = 0.7323$).  
+According to this analysis both crossovers performed about the same;
+by conventional criteria, this difference is considered to be not 
+statistically significant.
+
+The \textit{unpacking} crossover scored 23/50 (46\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 0 with 114,000 individuals to be processed.
+The average individual size for generation 50 was 83.1.
+The experiment took 388 minutes.
+
+The \textit{hybrid} crossover scored 32/50 (64\%) success rate. 
+Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
+The average individual size for generation 50 was 74.3.
+The experiment took 33 minutes.
+
+Comparison with other results is summarized in table 2.
+
+\begin{table}[t]
+\centering
+\begin{tabular}{|l|c|}
+\hline
+GP approach &  $I(M,i,z)$ \\
+\hline
+PolyGP                 &  14,000       \\
+\textbf{Our approach (hybrid)}&  \textbf{28,000} \\
+GP with Combinators	   &  58,616        \\
+GP with Iteration      &  60,000       \\
+\textbf{Our approach (unpacking)}&  \textbf{114,000} \\
+Generic GP	           &  220,000      \\
+OOGP	               &  680,000      \\
+GP with ADFs           &1,440,000      \\
+\hline
+\end{tabular}
+\caption{Results comparison.}
+\end{table}~\\
 
@@
 
-TODO
-
-\red{mimojiné zmínit, že v budouznu bysme chtěli porovnat kombinátory 
-s tou naší nějak extenzivně, ikdyž možná to tam vubec nepsat
-(což by bylo hezký už ted ale nestíhá se to..)}
+We have presented two approaches to crossover for typed GP over lambda terms.
+The first one (\textit{hybrid}) transforms all generated individuals into combinator terms at the end of generation phase.
+The second crossover approach (\textit{unpacking}) uses the lambda term representation during the whole run, except for the temporary phases during the crossover.
+We have compared their relative performance on even parity problem.
+It shows that the first one is performing the same or better in all observed metrics.
+This was surprising result for us.
+
+According to comparison of "individuals to be processed" metric our GP approach is performing better than the GP system presented in \cite{kes} and worse than the one presented in \cite{yu01}.
+By observation of typical solution which is often some modification of \texttt{foldr xor False inputList} one sees that an important task for GP here is to create \texttt{xor} function.
+This is the advantage for the more specialized term representation used in \cite{yu01}, which
+restricts the use of outer variables. 
+It is also fair to say, that this popular metric (in which the comparison data are available) is somehow problematic, since it is prone to have large variation across various instances of the same experiment \cite{luke2002perfect}.
+As is also visible from our results; both our approaches scored best in generation 0 thus the result depended only on the generation method which those two approaches shared, yet the individuals to be processed estimates are quite different.
+
+In the future work, we would like to accompany our theoretical results concerning relations of generating procedure and reductions described in this paper with experimental results.
+
+In this paper we touched varied aspects of the two lambda calculus concepts -- reductions and abstraction elimination -- in the hope that it will shed some more
+light on the promising intersection of lambda calculus and genetic programming.
+
+%\red{
+%možná zde zmínit tohle   
+%(b) úspornost - abstrakce eliminací produkuje 
+%    termy až kvadraticky větší než je původní 
+%    term. Dá se očekávat, že i při přímém generování je použití kombinátorů 
+%    míň úsporný (proměná se může bez velký námahy objevit v hluboko vnořeném
+%    podtermu velmi jednoduše, zatímco pomocí kombinátorů se její hodnota musí        
+%    do takto vnořeného podtermu složitě dopravit).
+%(c) použití kombů vyžaduje silnější typovej sytém - 
+%    kombinátory musej bejt polymorfní a přidáváme je do stavební sady,
+%    lambdy v pohodě běžej na simply typed lc. 
+%(d) použití proměnných je typické pro lidi, proto by se mohlo sát že
+%    i programy generovaný systémem kterej s proměnejma počitá budou o něco 
+%    "lidštější" 
+%}
 
 %ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
+%\section{Acknowledgments}
+%This section is optional; it is a location for you to acknowledge grants, funding, editing assistance and what have you.
+In the present case, for example, the authors would like to thank Gerald Murray of ACM for his help in codifying this \textit{Author's Guide} and the \textbf{.cls} and \textbf{.tex} files that it describes.
 

### 2014-04-11 tex/cec14/article.tex
članky stuff

@@
 
-\author{Tom\'{a}\v{s} K\v{r}en\thanks{Tomáš Křen (tomkren@gmail.cz) is with Charles Univetsity in Prague, Faculty of Mathematics and Physics; Department of Theoretical Computer Science and Mathematical Logic.} \and Roman Neruda\thanks{Roman Neruda (roman@cs.cas.cz) is with Institute of Computer Science, Academy of Sciences of the Czech Republic and has been partially supported by the The Ministry of Education of the Czech Republic project COST LD 13002}}
+\author{Tom\'{a}\v{s} K\v{r}en\thanks{Tomáš Křen (tomkren@gmail.cz) is with Charles Univetsity in Prague, Faculty of Mathematics and Physics; Department of Theoretical Computer Science and Mathematical Logic \red{and with TODO} and has been partially supported by the project GA ČR P202/10/1333 and by the SVV project number 260~104.} \and Roman Neruda\thanks{Roman Neruda (roman@cs.cas.cz) is with Institute of Computer Science, Academy of Sciences of the Czech Republic and has been partially supported by the The Ministry of Education of the Czech Republic project COST LD 13002.}}
 

### 2014-04-11 tex/gecco14/gecco14.tex
članky stuff

@@
 {\end{easylist}}
+
+\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
+\def\sharedaffiliation{\end{tabular}\newline\begin{tabular}{c}}
+
+\def\wu{\superscript{*}}
+\def\wg{\superscript{\dag}}
 
@@
 
+\numberofauthors{2}
+\author{
+  \alignauthor \ Tom\'{a}\v{s} K\v{r}en\wu\wg\\
+  \email{tomkren@gmail.com}
+  %
+  \alignauthor \ Roman Neruda\wu\\
+  \email{roman@cs.cas.cz}
+  %
+  \sharedaffiliation
+  \begin{tabular}{ccc}
+    \affaddr{{\wu}Institute of Computer Science{\ }}      & & \affaddr{{\wg}Faculty of Mathematics and Physics{\ }} \\
+    \affaddr{Academy of Sciences of the Czech Republic}   & & \affaddr{Charles University in Prague} \\
+    \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,} & & \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,} \\
+    \affaddr{Prague, Czech Republic}                      & & \affaddr{Prague, Czech Republic} \\
+  \end{tabular}
+}
+
 %\date{24 January 2014}
@@
 We discuss the role of those reductions in the process of generating initial population and propose several crossover approaches, two of which are based on abstraction elimination and compared using the even parity benchmark problem.
+\red{Přidat víc o kříženích tak aby davál insight to the project.}
 %including novel approach to crossover operator based both on reductions and abstraction elimination. 
@@
 Techniques described here would be analogous for a more sophisticated type system and can be directly applied to such systems as the Hindley–Milner type system.
+
+\red{Zdůraznit že main contribution of this paper je popis crossover operations který [rekapitulovat dobrý vlastnosti našeho křížení z 2ky (a dalších sekcí), tj že to nezjednodušuje na méně obecnou notaci a přesto to zůstává efektivní i v problémech kde to zjednodušení hraje do karet tý metodě.]}
 
@@
 \label{preliminaries}
+
+\red{Buď přidat neformální komentář a/nebo odebrat nadbytečný věci.}
 
@@
 source codes, but performing the same computation.
+\red{Zde potřeba doplnit vysvětlení, že to řeší jen čast tohoto
+fenomenu}
 
@@
 %ACKNOWLEDGMENTS are optional
-%\section{Acknowledgments}
+\section{Acknowledgments}
+
+Tomáš Křen has been partially supported by the project GA ČR P202/10/1333 and by the SVV project number 260~104.
+Roman Roman Neruda has been partially supported by the The Ministry of Education of the Czech Republic project COST LD 13002.
+
 %This section is optional; it is a location for you to acknowledge grants, funding, editing assistance and what have you.

### 2014-04-11 tex/cec14/article.tex
članky..

@@
 
-The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
-During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
+%The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation. 
+\red{In this paper we present a tree generating method suitable both for the standard GP\cite{koza92} and  typed GP over lambda calculus.
+We chose to describe the algorithm in the means of the typed lambda calculus since the standard S-expression representation can be understood as a special case of this representation.
+However, for evaluation of our approach we focus more on problems from standard GP.
+We chose this evaluation approach because the result of the GP system depends profoundly on the crossover operation, which remains an open problem for the typed GP over lambda calculus.
+  }
+
+During the expansion phase \red{of the term generating procedure} the \red{space} 
+%set 
+of unfinished terms can be browsed with respect to various search strategies.
 Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
@@
 the three state of the art proposals for typed GP known to us that are discussed in the following section. 
-Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
+Our proposed \emph{geometric
+%al 
+search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
 Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.

### 2014-04-15 tex/cec14/article.tex
cla

@@
 
-%\usepackage{stmaryrd}
-%\usepackage{amsfonts}
-%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
-%\usepackage{hyperref}
-%\usepackage{amsmath}
-%\usepackage{amssymb}
-%\usepackage[utf8]{inputenc}
-%\usepackage{qtree}
-%\usepackage{xspace}
-%\usepackage[ampersand]{easylist}
-
-%\usepackage{times}
 \usepackage[utf8]{inputenc}
 \usepackage{amsfonts}
-%\usepackage{qtree}
 \usepackage{xspace}
-
-%\usepackage{makeidx}  % allows for indexgeneration
-%\usepackage{graphicx,times,psfig,amsmath} % Add all your packages here
-%\usepackage{hyperref}
 \usepackage{amsmath}
-%\usepackage{amssymb}
+
 \usepackage{graphicx}
+\usepackage{epstopdf}
 
@@
 
-%\usepackage{amsthm}
+%\usepackage[ruled,norelsize]{algorithm2e}
+\usepackage[vlined,ruled,norelsize]{algorithm2e}
+
+\newcommand{\Pseudokod}[3]{
+	\begin{figure}[!t]
+	\removelatexerror
+	\begin{algorithm}[H]
+		\caption{#1}
+		\DontPrintSemicolon
+		\SetKwProg{Fn}{function}{}{}
+		\Fn{#2}{#3}
+	\end{algorithm}
+	\end{figure}
+}
+
+\makeatletter
+\newcommand{\removelatexerror}{\let\@latex@error\@gobble}
+\makeatother
+
 \newtheorem{definition}{Definition}
@@
 
-\IEEEoverridecommandlockouts    % to create the author's affliation portion
-
-\textwidth 178mm    % <------ These are the adjustments we made 10/18/2005
-\textheight 239mm   % You may or may not need to adjust these numbers again
-\oddsidemargin -7mm
-\evensidemargin -7mm
-\topmargin -6mm
-\columnsep 5mm
+%\IEEEoverridecommandlockouts    % to create the author's affliation portion
 
@@
 
-% paper title: Must keep \ \\ \LARGE\bf in it to leave enough margin.
-\title{\ \\ \LARGE\bf 
-Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
-
-\author{Tom\'{a}\v{s} K\v{r}en\thanks{Tomáš Křen (tomkren@gmail.cz) is with Charles Univetsity in Prague, Faculty of Mathematics and Physics; Department of Theoretical Computer Science and Mathematical Logic \red{and with TODO} and has been partially supported by the project GA ČR P202/10/1333 and by the SVV project number 260~104.} \and Roman Neruda\thanks{Roman Neruda (roman@cs.cas.cz) is with Institute of Computer Science, Academy of Sciences of the Czech Republic and has been partially supported by the The Ministry of Education of the Czech Republic project COST LD 13002.}}
+\title{Generating Lambda Term Individuals in Typed Genetic Programming Using Forgetful A*}
+
+\author{\IEEEauthorblockN{Tom\'{a}\v{s} K\v{r}en$^{*\dag}$}
+\IEEEauthorblockA{
+tomkren@{}gmail.com\\
+$^{*}$Charles University\\
+Faculty  of  Mathematics  and  Physics\\
+Malostransk\'e n\'am\v{e}st\'\i \ 25, Prague\\
+Czech Republic}
+\and
+\IEEEauthorblockN{Roman Neruda$^{\dag}$}
+\IEEEauthorblockA{roman@{}cs.cas.cz\\
+$^{\dag}$Institute of Computer Science\\
+Academy of Sciences of the Czech Republic\\
+Pod Vod\'arenskou v\v{e}\v{z}\'\i \ 2, Prague\\
+Czech Republic}
+}
 
@@
 \begin{abstract}
-Tree based genetic programming traditionally uses simple S-expressions to represent programs, however more expressive representations can exhibit better results and are well suited for typed genetic programming.
-One of such representations is lambda calculus.
-In this paper, a generalization of genetic programming 
-for simply typed lambda calculus is presented. We mainly focus on
-population initialization methods. 
-Three such initialization methods are proposed 
-depending on different search strategies. 
-The first strategy corresponds to standard ramped half-and-half initialization, 
-the second one corresponds to exhaustive systematic search, and the third one 
-represents a novel geometric strategy.
-Experiments on three well known benchmark problems show
-that the geometric strategy outperforms
-the standard generating method in success rate, best fitness value,
-time consumption and average individual size.
-The fact that a search strategy can be easily described by a rather simple filtration algorithm opens opportunity to easy exploration of more 
-population initialization methods, possibly in a kind of meta-evolution way.
+Tree based genetic programming (GP) traditionally uses simple S-expressions to represent programs, however more expressive representations, such as lambda calculus, can exhibit better results while being better suited for typed GP.
+In this paper we present population initialization methods within a framework of GP over simply typed lambda calculus that can be also used in the standard GP approach.
+Initializations can be parameterized by different search strategies, leading to wide spectrum of methods corresponding to standard ramped half-and-half initialization on one hand, or exhaustive systematic search on the other.
+A novel geometric strategy is proposed that balances those two approaches.
+Experiments on well known benchmark problems show that the geometric strategy outperforms the standard generating method in success rate, best fitness value, time consumption and average individual size.
 \end{abstract}
@@
 %The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation. 
-\red{In this paper we present a tree generating method suitable both for the standard GP\cite{koza92} and  typed GP over lambda calculus.
+In this paper we present a tree generating method suitable both for the standard GP\cite{koza92} and  typed GP over lambda calculus.
 We chose to describe the algorithm in the means of the typed lambda calculus since the standard S-expression representation can be understood as a special case of this representation.
@@
 We chose this evaluation approach because the result of the GP system depends profoundly on the crossover operation, which remains an open problem for the typed GP over lambda calculus.
-  }
-
-During the expansion phase \red{of the term generating procedure} the \red{space} 
-%set 
-of unfinished terms can be browsed with respect to various search strategies.
+
+During the expansion phase of the term generating procedure the space of unfinished terms can be browsed with respect to various search strategies.
 Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
@@
 %al 
-search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
+search strategy}\footnote{Not to be confused with geometric  crossover introduced by \cite{moraglio2010}.} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
 Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
@@
 polymorphic higher-order functions\footnote{Higher-order 
-function is a function taking another function as 
-input parameter.} and lambda abstractions  \cite{yu01}.
-Important point of interest in this work is use of
-\texttt{foldr} function as a tool for \textit{implicit recursion},
+function takes another function as an input parameter.
+} and lambda abstractions  \cite{yu01}.
+Important point of interest in this work is use of \texttt{foldr}\footnote{ \texttt{foldr f z [] $~~~$  = z\\ 
+$~~~\;$foldr f z (x:xs) = f x (foldr f z xs) }} function as a tool for \textit{implicit recursion},
 i.e. recursion without explicit recursive calls. 
@@
 \begin{align*}
-x   \in V \cup C  &\then x     \in \Lambda \\
+x   \in V \cup C  &\then x     \in \Lambda 
+\textit{~~~~~~~~~~~~~(Variable or constant)}\\
 M,N \in \Lambda   &\then (M~N) \in \Lambda 
@@
 well known from common programming languages.
-For example in JavaScript 
+
+For example in JavaScript term
 $(M~N)$ translates to expression \texttt{$M$($N$)} and
@@
 $\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
-$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
-The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
-combining the previously defined \textit{\lterms} and \textit{types} together. \\
+$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$. Such "chain of arrows" types simulate types of functions with multiple (here $n-1$) inputs. 
+This technique is called \textit{currying}. One can grasp the trick by considering $(f~M_1~\dots~M_n)$ as shorthand for $f(M_1,\dots,M_n)$
+and by observing that functions with such types have terms of the form $\lam{x_1 x_2 \dots x_n }{M}$.
+
+The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by combining the previously defined \textit{\lterms} and \textit{types} together.
 
@@
 \textit{~~~~~~~~(Unfinished leaf)}\\
-x   \in V \cup C  &\then x     \in \Lpr \\
+x   \in V \cup C  &\then x     \in \Lpr
+\textit{~~~~~~~~~~~~~(Variable or constant)}\\
 M,N \in \Lpr   &\then (M~N) \in \Lpr 
@@
 
+%\red{na vhodna mista podotknout, jak to je v std GP}
+
 Our approach to \lterm generating is based on technique 
@@
 well-typed \lterms in their long normal form. 
-We use this technique to perform systematic exhaustive enumeration
-of \lterms in their long normal form in order from smallest to largest.
+We use this technique to perform a systematic exhaustive enumeration
+of \lterms in their long normal form in order from the smallest to the largest.
 We use well known \textit{A* algorithm} \cite{AIMA} for this task.
@@
 
-The inputs for the term generating algorithm are following.
+Let us look on the proposed generating procedure in a grater detail sufficient as guide for implementing it.
+The inputs for the term generating algorithm are following:
 \begin{enumerate}
  \item Desired type $\tau$ of generated terms.
- \item Context $\Gamma$ representing set of building symbols.
+ \item Context $\Gamma$ representing the set of building symbols.
  \item Number $n$ of terms to be generated.
@@
 trivial estimate $\vert\ul{\tau}{\Gamma}\vert = 1$ which is trivially admissible.
-This heuristic is not as silly as it might seem since it is
+This heuristic is not as naive as it might seem since it is
 quite usual to have $x$ such that $x : \tau \in \Gamma$.
@@
 According to the number of unfinished leafs in $M$ one of
-the following actions is performed.
+the following actions is performed:
 \begin{enumerate}
@@
 only second rule is applied, but if it results in creation of some unatomic 
-$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }.
+$\ul{\tau_i}{\Gamma}$, then first rules are applied, but during the same successor creation This step eliminates all unatomic unfinished leafs. }:
 \begin{enumerate}
- \item If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
+ \item \textit{Function rule:}
+
+ If $L = \ul{\rho_1 \ar \dots \ar \rho_n \ar \alpha}{\Gamma}$,
  	   where $\alpha$ is atomic type and $n \geq 1$, 
@@
        Thus this expansion results in exactly one successor.
- \item If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type,
-       then for each 
-       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}
+
+       In other words, a \textit{function} type is replaced by a new anonymous function, and for each of its inputs is added a fresh variable into the context of its body -- which results in possibility of using those new variables inside the body of the anonymous function.
+Notice that this rule is never used in problems satisfying the closure requirement.
+ \item \textit{Atomic rule:}
+
+ If $L = \ul{\alpha}{\Gamma}$ where $\alpha$ is \textit{atomic} type, then for each 
+       \mbox{$f : (\tau_1 \ar \dots \ar \tau_m \ar \alpha) \in \Gamma$}, $m \geq 0$
        the unfinished leaf $L$ is replaced by 
@@
        Thus this expansion results in many (possibly zero or one) successors.
+
+       In other words, an \textit{atomic} type is replaced by a constant symbol (for $m = 0$) or function symbol (for $m>0$) from the local context.
+Unfinished leafs $\ul{\tau_i}{\Gamma}$ correspond to the argument subtrees yet to be generated -- precisely as in S-expression generation.
+Notice that for problems satisfying closure requirement only this rule is applicable; therefore standard S-expressions are generated for such a problem.
 \end{enumerate}
+
+\textit{Algorithm 2} summarizes this key phase of generating procedure by pseudocode.
+
+\Pseudokod{Individual generating procedure.}
+{\textbf{generate}(Type $\tau$, Context $\Gamma$, Int $n$, Strategy $S$)}{
+
+	%star je TODO
+	%isGoal je TODO
+	%heur je TODO
+
+	$results \la$ empty set\;
+
+	\While{ $\vert results \vert < n$ }{ 		
+		$S.init()$\;
+		$open    \la$ empty priority queue \;
+		$start  \la \ul{\tau}{\Gamma}$\;   
+		$start$.G $\la 0$ \;
+		$start$.F $\la 1$ \;  % heur$( $start$ )
+		$open$.insert($start$)\;
+		\While{$\neg$ $open$.isEmpty() }{
+			$term \la$ $open$.popTermWithLowestF()\;
+			\eIf{ numUnfinLeafs($term$) = 0 }{
+
+				$results$.insert($term$)\;
+
+				\If{ $\vert results \vert = n$ }{
+					\Return $results$ \;
+				}
+			}{
+				$nexts \la $\textbf{expand}($term$)\;
+				$depth \la term$.depthOfExpandedLeaf()\; 
+				$nexts' \la S.$\textbf{filter}($nexts$, $depth$)\;			
+				\For{$next \in nexts'$ }{
+					$heur     \la $numUnfinLeafs($next$)\;
+					$next$.G $\la term.$G$ + 1$ \;
+					$next$.F $\la next.$G$ + heur$  \;
+					$open$.insert($next$)\;			
+				}
+			}
+		}
+	}
+
+	\Return results
+}
+
+\Pseudokod{Expansion (successor) procedure.}
+{\textbf{expand}(UnfinishedTerm $term$)}{
+	$nexts \la$ empty set\;	
+	$\ul{\sigma}{\Gamma} \la $ \textit{DFS-first} unfinished leaf of $term$ \;
+	\Switch{$\sigma$}{
+	\Case{$\rho_1 \ar \dots \ar \rho_n \ar \alpha: n \geq 1, \alpha$ atomic}{
+		$t \la \lamb{x_1 \dots x_n}{\ul{\alpha}{\Gamma,x_1 : \rho_1\dots x_n :\rho_n}}$\;
+		$next \la $ replace $\ul{\sigma}{\Gamma}$ in $term$ with $t$\;  		
+		$nexts$.insert($next$)
+	}\Case{$\alpha$: $\alpha $ atomic}{
+		\For{$f : \tau_1 \ar \dots \ar \tau_m \ar \alpha \in \Gamma$, $m \geq 0$  
+		}{
+			$t \la (~f~\ul{\tau_1}{\Gamma}~\dots~\ul{\tau_m}{\Gamma}~)$\; 
+			$next \la $ replace $\ul{\sigma}{\Gamma}$ in $term$ with $t$\;  		
+			$nexts$.insert($next$)
+		}
+	}
+	\Return $nexts$
+}
+}
+
+\Pseudokod{Geometric strategy.}
+{Geometric.\textbf{filter}(UnfinTerms $nexts$, Int $depth$)}{
+	$nexts' \la$ empty set\;	
+	\For{$next \in nexts$}{
+		\textbf{with probability} $q^{depth}$ \textbf{do} $nexts'$.insert($next$) 	
+	}
+	\Return $nexts'$
+}
 
@@
 again initializes its internal state and the process continues.
-
-\Lets now discuss three such search strategies.\\
+The whole generating procedure is summarized by \textit{Algorithm 1}.
+
+\Lets now discuss three such search strategies.
 
@@
 first $n$ smallest lambda terms in their
-\textit{long normal form}.\\
+\textit{long normal form}.
 
@@
 If $isFull = false$, then the subset to select from is 
-the whole input set.\\
+the whole input set.
 
@@
 $$ p = q^{depth} $$
+
 This formula is motivated by idea that it is important to
@@
 "too big trees should be killed". 
-
-~\\
+\textit{Algorithm 3} summarizes the filtration procedure performed by the geometric strategy.
+
 Furthermore, our system utilizes generalization of the standard 
@@
 Another performance enhancing transformation is option of using "applicative" tree representation (coming directly from inductive definition of \lterms) instead of more traditional S-expression representation.
+Favorable properties of applicative tree representation are also reported in \cite{yu1998polygp}.
 
@@
 \textit{geometric strategy} (with default parameter $q=0.75$).
+
+Despite the fact that the geometric strategy was designed for typed GP over lambda trees it works well for standard GP problems satisfying the closure requirement.
+Furthermore, for such problems our approach generates standard S-expressions only.
+Another reason for including standard GP benchmarks in our experiments is the fact that the results are seriously affected by a crossover
+operator.
+Design of a good crossover for typed GP remains an open problem.
+Our crossover is quite complicated and beyond the scope of this paper, but for problems satisfying the closure requirement it behaves exactly the same as the standard GP crossover.
+
+Three well known benchmark problems
+(Simple symbolic regression, Artificial ant and
+Even parity problem) were chosen for our experiments.
+First two satisfy the closure requirement, 
+whereas the third one does not.
+
+In order to ensure that the experiments are replicable we chose to use the standard GP system described in \cite{koza92}, with all the parameters set to their default values (which are sumarized in table~I), only with the following exceptions:
+
+\begin{enumerate}
+  \item In the second and third experiment the preservation of the best individual into the next generation is performed.
+  \item In the third experiment we use our crossover operation briefly mentioned above together with use of "applicative" tree representation.
+\end{enumerate}
+
+For problems satisfying closure requirement our generating method generates S-expressions. Therefore it can be directly plugged into the standard GP system.
 Each experiment consisted of 50 independent runs 
@@
 
-Despite the fact that the presented system is designed
-for typed genetic programming it works well for
-standard GP problems satisfying the closure requirement. For such problems our approach 
-generates standard S-expressions only. 
-Another reason for including standard GP
-benchmarks in our experiments is that fact that
-the results are seriously affected by a crossover 
-operator. Design of a good crossover for typed GP
-remains an open problem. 
-Our crossover is quite complicated and beyond
-the scope of this paper, but for problems satisfying
-the closure requirement it behaves exactly the same
-as the standard GP crossover. 
-Three well known benchmark problems
-(Simple symbolic regression, Artificial ant and
-Even parity problem) were chosen for our experiments.
-First two satisfy the closure requirement, 
-whereas the third one does not.
-
-\subsection{Simple symbolic regression}
+\begin{table}[!t]
+\caption{Standard GP parameters.}
+\centering
+\begin{tabular}{|l|c|}
+\hline
+~&~\\
+Population size             & 500\\
+Number of generations       & 51\\
+Probability of crossover    & 90\%\\
+Probability of reproduction & 10\%\\
+Probability of mutation     & 0\%\\
+Max depth (initial)         & 6\\
+Max depth (crossover)       & 17\\
+Representation              & S-expressions\\
+Crossover                   & Standard tree swapping (90-10)\\
+Selection                   & Fitness-proportionate\\
+~&~\\
+\hline
+Generating method           & Ramped half-and-half / geometric s.\\
+\hline
+\end{tabular}
+\end{table}
+
+\subsection{Simple Symbolic Regression}
+
+\begin{figure}[!ht]
+  \centering
+  \caption{Graphs for Simple symbolic regression.}
+  \includegraphics[scale=0.12]{imgs/SSR.eps}
+\end{figure}
+
 \textit{Simple Symbolic Regression} is a problem described
@@
 \noindent
-\begin{minipage}{.5\linewidth}
+\begin{minipage}{.9\linewidth}
 \begin{align*}
@@
 
-\begin{minipage}{.5\linewidth}
+\begin{minipage}{.9\linewidth}
 \begin{align*}
@@
 
-~\\
-
 Fitness function is computed as follows.
@@
 An individual $f$ such that $\vert f(x_i)-y_i \vert < 0.01 $ for all data samples is 
-considered as a correct individual.\\
-
-\begin{figure}[h!]
-  \centering
-  \caption{Graphs for Simple symbolic regression.}
-  \includegraphics[scale=0.241]{imgs/cec/SSRb.PNG}
-\end{figure}
-
-\begin{table}[t]
+considered as a correct individual.
+
+\begin{table}[!t]
+\caption{Statistical analysis - Simple symbolic regression.}
 \centering
@@
 \end{tabular}
-\caption{Statistical analysis - Simple symbolic regression.}
 \end{table}
@@
 Figure~1 shows results of this experiment. 
-Table~I summarizes statistical analysis
+Table~II summarizes statistical analysis
 of the fitness values of the best individuals of the run 
@@
 
-\subsection{Artificial ant}
+\subsection{Artificial Ant}
 \textit{Artificial Ant} is another problem described
@@
 In \cite{koza92} this limit is said to be 400 time steps.
-But there is also mentioned following solution, 
-which is described as correct solution:
-\texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
-This program needs 545
-time steps; if it is given only 400 time steps, then it eats only 79 pieces
-of food. Thus we use 600 time steps. 
+But there is also mentioned one solution, which is described as correct solution. This program needs 545 time steps; if it is given only 400 time steps, then it eats only 79 pieces of food. 
+%But there is also mentioned following solution, which is described as correct solution: \texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
+%This program needs 545 time steps; if it is given only 400 time steps, then it eats only 79 pieces of food. 
+Thus we use 600 time steps. 
 }.
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
   \centering
   \caption{Graphs for Artificial ant problem.}
-  \includegraphics[scale=0.26]{imgs/cec/ANTb.PNG}
+  \includegraphics[scale=0.13]{imgs/ANT.eps}
 \end{figure}
 
-\begin{table}[t]
+\begin{table}[!t]
+\caption{Statistical analysis - Artificial ant problem.}
 \centering
@@
 t-value &  6.3479           & $\alpha = 0.05$\\
-p-value &  \textless 0.0001 &  \textbf{Extremely statistically significant.}\\
+p-value &  \textless 0.0001 &  \textbf{Extremely stat. significant.}\\
 \hline
 \end{tabular}
-\caption{Statistical analysis - Artificial ant problem.}
 \end{table}
@@
 Figure~2 shows results of this experiment. 
-Table~II summarizes statistical analysis
+Table~III summarizes statistical analysis
 of the fitness values of the best individuals of the run 
@@
 
-\subsection{Even parity problem}
+\subsection{Even Parity Problem}
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
   \centering
   \caption{Graphs for Even parity problem.}
-  \includegraphics[scale=0.26]{imgs/cec/EPb.PNG}
+  \includegraphics[scale=0.12]{imgs/EP.eps}
 \end{figure}
 
-\begin{table}[t]
+\begin{table}[!t]
+\caption{Statistical analysis - Even parity problem.}
 \centering
@@
 \end{tabular}
-\caption{Statistical analysis - Even parity problem.}
 \end{table}
@@
 Figure~3 shows results of this experiment. 
-Table~III summarizes statistical analysis
+Table~IV summarizes statistical analysis
 of the fitness values of the best individuals of the run 
@@
 17,500 in generation 4. But by observation of typical solution which is often
-some modification of \texttt{foldr xor False inputList} one sees that important 
+some modification of \texttt{foldr xor True inputList} one sees that important 
 task for GP here is to create \texttt{xor} function. This is advantage for the 
@@
 
-%\red{Ještě nějak rozšířit.}
-
-%\red{ [ještě něco navíc nebo podrobněji, říct poly že neni ale že se na něm pracuje a de hezky zandat do již systemu] }
+\section*{Acknowledgements}
+Tom\'{a}\v{s} K\v{r}en has been partially supported by the project GA \v{C}R P202/10/1333 and by the SVV project number 260~104.
+Roman Neruda has been partially supported by the The Ministry of Education of the Czech Republic project COST LD 13002.
 

### 2014-04-15 tex/gecco14/gecco14.tex
cla

--- 2014-01-30 tex/gecco14/gecco14.tex
@@
 \documentclass{sig-alternate}
+
+\newfont{\mycrnotice}{ptmr8t at 7pt}
+\newfont{\myconfname}{ptmri8t at 7pt}
+\let\crnotice\mycrnotice%
+\let\confname\myconfname%
+
+\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
+Copyrights for components of this work owned by others than the author(s) must be honored.
+Abstracting with credit is permitted.
+To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
+Request permissions from permissions@acm.org.}
+\conferenceinfo{GECCO'14,}{July 12--16, 2014, Vancouver, BC, Canada. \\
+{\mycrnotice{Copyright is held by the owner/author(s). Publication rights licensed to ACM.}}}
+\copyrightetc{ACM \the\acmcopyr}
+\crdata{978-1-4503-2662-9/14/07\ ...\$15.00.\\
+http://dx.doi.org/10.1145/2576768.2598361}
+
+\clubpenalty=10000 
+\widowpenalty = 10000
+
+%\begin{document}
+
+%.....
 
@@
 \newcommand{\lnf}{\textit{lnf}\xspace}
-\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+%\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+\newcommand{\Ae}{\mathop{\mathrm{AE}}}
 \newcommand{\Bcomb }{\mathbf{B}}   
@@
 
-\conferenceinfo{GECCO'14,} {July 12-16, 2014, Vancouver, BC, Canada.}
-\CopyrightYear{2014}
-\crdata{TBA}
-\clubpenalty=10000
-\widowpenalty = 10000
+\setlength{\pdfpagewidth}{8.5in}
+\setlength{\pdfpageheight}{11in}
+
+%\conferenceinfo{GECCO'14,} {July 12-16, 2014, Vancouver, BC, Canada.}
+%\CopyrightYear{2014}
+%\crdata{TBA}
+%\clubpenalty=10000
+%\widowpenalty = 10000
 
@@
 
-% \numberofauthors{2} 
-%
-%  \author{
-%  % 1st. author
-%  \alignauthor
-%  Tom\'{a}\v{s} K\v{r}en\\
-%    \affaddr{Faculty of Mathematics and Physics}\\
-%    \affaddr{Charles University in Prague}\\
-%    \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
-%    \affaddr{Prague, Czech Republic}\\
+ \numberofauthors{2} 
+  \author{
+  % 1st. author
+  \alignauthor
+  Tom\'{a}\v{s} K\v{r}en\wu\wg\\
+    \email{tomkren@gmail.com}\\
+    ~\\
+    \affaddr{{\wu}Faculty of Mathematics and Physics}\\
+    \affaddr{Charles University in Prague}\\
+    \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
+    \affaddr{Prague, Czech Republic}
+  % 2nd. author
+  \alignauthor
+  Roman Neruda\wg\\
+    \email{roman@cs.cas.cz}\\
+    ~\\
+    \affaddr{{\wg}Institute of Computer Science}\\
+    \affaddr{Academy of Sciences of the Czech Republic}\\
+    \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
+    \affaddr{Prague, Czech Republic}    
+  }
+
+%\numberofauthors{2}
+%\author{
+%  \alignauthor \ Tom\'{a}\v{s} K\v{r}en\wu\wg\\
 %  \email{tomkren@gmail.com}
-%  % 2nd. author
-%  \alignauthor
-%  Roman Neruda\\
-%    \affaddr{Institute of Computer Science}\\
-%    \affaddr{Academy of Sciences of the Czech Republic}\\
-%    \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
-%    \affaddr{Prague, Czech Republic}\\
+%  %
+%  \alignauthor \ Roman Neruda\wu\\
 %  \email{roman@cs.cas.cz}
-%  }
+%  %
+%  \sharedaffiliation
+%  \begin{tabular}{ccc}
+%    \affaddr{{\wu}Institute of Computer Science{\ }}      & & \affaddr{{\wg}Faculty of Mathematics and Physics{\ }} \\
+%    \affaddr{Academy of Sciences of the Czech Republic}   & & \affaddr{Charles University in Prague} \\
+%    \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,} & & \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,} \\
+%    \affaddr{Prague, Czech Republic}                      & & \affaddr{Prague, Czech Republic} \\
+%  \end{tabular}
+%}
 
@@
 In this paper we discuss advantages of this representation coming from the use of reductions (beta and eta) and a way to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm.
-We discuss the role of those reductions in the process of generating initial population and propose several crossover approaches, two of which are based on abstraction elimination and compared using the even parity benchmark problem.
+We discuss the role of those reductions in the process of generating initial population and propose two novel crossover operations based on abstraction elimination capable of handling general form of typed lambda term while being a straight generalization of the standard crossover operation.
+We compare their performances using the even parity benchmark problem.
+%\red{Přidat víc o kříženích tak aby davál insight to the project. Možná i o tom že je to zobecnení std GP.}
 %including novel approach to crossover operator based both on reductions and abstraction elimination. 
@@
 
-\keywords{genetic programming, lambda calculus}
+\keywords{genetic programming; lambda calculus}
 
@@
 the traditional GP representation of individuals, with concepts from
-lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts have shown to be successful \cite{yu01}.
+lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory~\cite{yu01}.
 
@@
 
-In our system we use simply typed lambda calculus.
-Since abstraction elimination and reductions do not involve many type related issues it seems to us as a sufficient choice of type system.
+In our system we use a simply typed lambda calculus.
+Since abstraction elimination and reductions do not involve many type related issues it seems to us as a sufficient choice of a type system.
 Techniques described here would be analogous for a more sophisticated type system and can be directly applied to such systems as the Hindley–Milner type system.
 
+In this paper we present an approach to typed GP over lambda calculus capable of operating with the full array of notions given by a simply typed lambda calculus while being a direct generalization of standard GP~\cite{koza92}.
+We discuss the advantages of generating lambda terms in their long normal form and present two novel crossover operations for simply typed lambda terms.
+Compared to the approach to typed GP over lambda calculus  proposed by
+Yu~\cite{yu01}, it uses more general term representation which is not restricting the use of outer local variables in the body of an anonymous function.
+Compared to the purely combinator approach by Briggs and
+O’Neill~\cite{kes}, it has the advantage of reducing the size of the search space in the term generation phase.
+
 %\red{Možná někam poznamenat že přínos toho článku je taky v tom že zpřístupňuje pojmy z teorie typovaného lc komunitě typovaného GP esli to neni blbý..}
 
-%Pro inspiraci :
-%\nehodi{The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
-During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
-Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
-Thus, the natural idea is to employ an exhaustive systematic search.
-On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.
-These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
-This design goal also differentiate our approach from the three state of the art proposals for typed GP known to us that are discussed in the following section.
-Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
-Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.}
-
 The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
-The body of the paper is presented in sections \ref{generating} and \ref{crossover}; the first discusses the term generating 
+The body of the paper is presented in sections \ref{generating} and \ref{crossover}; the former discusses the term generating 
 procedure and its relations
-with reductions, the second describes several approaches to lambda term
-crossover and how they solve the variable problem. Section~\ref{experiments} examines those ideas on the even parity benchmark problem, while the paper is concluded by section~\ref{conclusions}.
+with reductions, the latter describes several approaches to lambda term
+crossover, and how they solve the variable problem. Section~\ref{experiments} examines those ideas on the even parity benchmark problem, while the paper is concluded by section~\ref{conclusions}.
+
+\vfill
 
@@
 Yu presents a GP system utilizing
-polymorphic higher-order functions\footnote{Higher-order 
-function is a function taking another function as 
-input parameter.} and lambda abstractions  \cite{yu01}.
-An important point of interest in the work is use of
-\texttt{foldr} function as a tool for \textit{implicit recursion}, i.e.
-recursion without explicit recursive calls.
+polymorphic higher-order functions
+%\footnote{Higher-order function is a function taking another function as input parameter.} 
+and lambda abstractions~\cite{yu01}.
+An important point of interest in the work is the use of
+\texttt{foldr} function as a tool for \textit{implicit recursion}, i.e., recursion without explicit recursive calls.
 The terminal set for constructing lambda abstraction subtrees  is limited to use only constants and variables of that particular lambda abstraction, i.e., outer variables are not allowed to be used as terminals in this work.
@@
 utilizing typed GP with combinators \cite{kes}.
-The difference between approach presented in their work
+The difference between the approach presented in their work
 and our approach is that they generate terms in a straightforward way directly from the library of combinators, without any use of lambda abstractions.
 They use the Hindley–Milner type system.
-They also present interesting concept of \textit{Generalized genetic operator} based on term generation.
-The combinator based genetic algorithm has the advantage of
-avoiding the need of dealing with variables.
-
-Binard and Felty use even stronger type system (\textit{System F}) \cite{binard2008genetic}.
+They also present an interesting concept of \textit{Generalized genetic operator} based on term generation.
+The combinator based genetic algorithm has the advantage of avoiding the need of dealing with variables.
+
+Binard and Felty use an even stronger type system, the (\textit{System F}) \cite{binard2008genetic}.
 But with the increasing power of the type system also comes an increasing difficulty of term generation.
 For this reason, the evolution in this work takes interesting and nonstandard shape (fitness is associated with \textit{genes} which are evolved together with \textit{species} which together participate in creation of individuals).
-This differs from our approach, which tries to be a generalization of the standard GP\cite{koza92}.
-
-In contrast with the above mentioned works, our approach uses very simple type system (simply typed lambda calculus) and concentrates on process of generation  
-able to generate all possible well-typed normalized lambda terms. In order to do so we use technique based on \textit{inhabitation machines} described by Barendregt \cite{barendregt10}.
+This differs from our approach which tries to be a straightforward generalization of the standard GP\cite{koza92}.
+
+In contrast to the above mentioned works, our approach uses a very simple type system (the simply typed lambda calculus) and concentrates on process of generation  
+able to generate all possible well-typed normalized lambda terms. In order to do so we use technique based on \textit{inhabitation machines} described by Barendregt\cite{barendregt10}.
 
@@
 \label{preliminaries}
+
+%\red{Buď přidat neformální komentář a/nebo odebrat nadbytečný věci.}
 
@@
 
-For better readability, 
+To ensure better readability, 
 $M_1~M_2~M_3~\dots~M_n$ will be an abbreviation for 
-$(\dots((M_1~M_2)~M_3)~\dots~M_n)$,
-and $\lam{x_1 x_2 \dots x_n }{M}$ will be an abbreviation for 
+$(\dots((M_1~M_2)~M_3)~\dots~M_n)$.
+And $\lam{x_1 x_2 \dots x_n }{M}$  for
 $\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
@@
 mechanism for term evaluation. In $\lambda$-calculus there
-is a \mbox{$\beta$-reduction} procedure for this reason.\\
+is a \mbox{$\beta$-reduction} procedure for this reason.
 
@@
 of insertion of arguments supplied to a function into 
-its body. \\
+its body.
 
@@
 
-Relation $\bearr\;=\;\barr \cup \earr$. 
-(Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)
+Relation\footnote{Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.} $\bearr\;=\;\barr \cup \earr$.
 Similarly as for $\bbarr$ and $\beq$ we can define relations 
-$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.\\
-
-$\eta^{-1}$-reduction (also called $\eta$-expansion) is 
+$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.
+
+The $\eta^{-1}$-reduction (also called $\eta$-expansion) is 
 the reduction converse to $\eta$-reduction.
@@
 & A \lterm M \textit{has} a \bnf if $M \beq N$
-and $N$ is a \bnf.\\
+and $N$ is a \bnf.
 \end{easylist}
 \end{definition}
+
 A normal form may be thought of as a result of a term evaluation. 
 Normalization of the term $M$ is the process of finding normal form of $M$. 
-
 Similarly we can define \enf and \benf.
@@
 A statement $M\colon\sigma$ is \textit{derivable from}
-a context $\Gamma$ (notation 
+context $\Gamma$ (notation 
 \mbox{$\Gamma\vdash{}M\colon\sigma$}) 
@@
 %which according to Theorem 2B.4 is always the case, 
-then it also has a unique \lnf, 
+then it also has an unique \lnf, 
 which is also its unique \beenf.
@@
 \subsection{Abstraction elimination}
-
 \textit{Abstraction elimination} is a process of transforming 
@@
 free symbols from former \lterm and some new symbols standing for 
-combinators $\Scomb$, $\Kcomb$ and $\Icomb$. \\
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$.
 
@@
 \begin{align*}
-\Ae[x]           &= x &\\[0.4em]
-\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
-\Ae[\lam{x}{x}]  &= \Icomb &\\
-\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
-\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
-&\textbf{if } x \in \FV(M)\\
-\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
-&\textbf{if } x \in \FV(M)\\
-&&\vee~ x \in \FV(N)\\
+\Ae[x]           &= x\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N])\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) \textbf{ if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  \textbf{ if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])\\
+&\textbf{ if } x \in \FV(M) \vee~ x \in \FV(N)\\
 \end{align*}
@@
 The algorithm presented here is a simple version of this process. 
-More optimized version (used in our \textit{hybrid} crossover described below), by means of the size of resulting term and its time performance is presented in \cite{jones87}.
-
-\section{Term generating and reductions}
+More optimized version (used in our \textit{hybrid} cross\-over described below), by means of the size of resulting term and its time performance is presented in \cite{jones87}.
+
+\section{Generating and reductions}
 \label{generating}
 
-This section is focused on the individual generating method producing
+This section is focused on the individual generating meth\-od producing
 terms in their long normal form, which can be understood as a straight 
@@
 $\tau_1 , \dots , \tau_n$ (where $n > 0$) 
-we have a rule of the second form:\\
+we have a rule of the second form:
 
@@
 ( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
-$~\\
+$
 
@@
 \lterms $M$,$N$ such that $M \not= N$, but $M =_{\beta\eta} N$.
-In other words, we avoid generating two programs with different 
-source codes, but performing the same computation.
+This means that $M$,$N$ correspond to two programs with different source codes, but both perform the same computation.
 
@@
 is not omitted, because it is the same computation
-as the computation performed by \lterm $L$. \\
+as the computation performed by \lterm $L$.
 
@@
 \lterms $M$,$N$ such that $M \not= N$ and $M =_{\eta} N$; 
-but generating in \bnf does not have such a property..\\
+but generating in \bnf does not have such a property.
 
@@
 (a \lterm representation derived from the definition of \lterms 
-where @ stands for \textit{function application}).\\
+where @ stands for \textit{function application}).
+
 \textit{\ered}: \Tree [.$\lh{x}$ [.@ $M$ $x$ ] ] 
@@
 ~~~~~~~~~~~
-\textit{\bredex}: \Tree [.@ [.$\lh{x}$ $M$ ] $N$ ] \\
+\textit{\bredex}: \Tree [.@ [.$\lh{x}$ $M$ ] $N$ ]
 
@@
 In previous proposition is shown that \ered of \lnf
-leads to a term in \benf.\\
+leads to a term in \benf.
 
 In order to show that a function is bijection it is sufficient to show that it is
-both \textit{injection} and \textit{surjection}.\\
+both \textit{injection} and \textit{surjection}.
 
@@
 so $M_1 =_{\beta\eta^{-1}} M_2$.
-This contradicts with $M_1,M_2$ being distinct \lnf{}s.\\
+This contradicts with $M_1,M_2$ being distinct \lnf{}s.
 
@@
 by swapping randomly selected subtrees in each parent 
-S-expression.
+S-expre\-ssion.
 For typed lambda terms two difficulties arise: Types and variables.
@@
 The following example illustrates the problem. \Lets have these two
-parent trees with selected nodes in bold.\\
+parent trees with selected nodes in bold.
+
+\vfill
 
@@
 
-~\\The swap of subtrees results in following trees:\\
+\vfill
+
+The swap of subtrees results in following trees:
+
+\vfill
 
@@
 
-~\\The problem is that variable $x_2$ in second tree
+\vfill
+
+The problem is that variable $x_2$ in second tree
 is not bound by any $\lambda$-head and since
@@
 
-Generally speaking, the variable problem is caused by the fact the local variable is not defined in the new place, or the variable is defined in the new place, but has
-some a different type. Let us list some possible approaches to solving this problem.
+Generally speaking, the variable problem is caused by the fact that the local variable is not defined in the new place, or the variable is defined in the new place, but has a different type.
+Let us list some possible approaches to solving this problem.
 
@@
 
+We also use a little trick; if the type of the individual is a function type it is not necessary to abstract out the input variables (and to them associated lambda heads) of the function since all the individuals share them.
+(This is illustrated by the figure~2 bellow, where the root lambda head $\lambda x$ is still on the trees and so is the variable $x$ in the leafs.)
+
 We identified two approaches that utilize abstraction elimination.
@@
 lambda calculus representation and the purely combinator representation
-used in \cite{kes}. The advantage of generating terms as \lnf lambda terms
-instead of generating them directly as combinator terms is that it reduces
-the search space, because there is more than one way to represent a \lnf 
-lambda term as a combinator term (this can be seen considering that 
-\lnf term $\lamb{x}{x}$ is equivalent to both $\Icomb$ and 
-$\Scomb \Kcomb \Kcomb$), but every combinator term has only 
-one unique \lnf form. Unfortunately, one disadvantage is also important;
+used in \cite{kes}.
+The advantage of generating terms as \lnf lambda terms instead of generating them directly as combinator terms is that it reduces the search space, because there is more than one way to represent a \lnf lambda term as a combinator term (this can be seen considering that \lnf term $\lamb{x}{x}$ is equivalent to both $\Icomb$ and $\Scomb \Kcomb \Kcomb$), but every combinator term has only one unique \lnf form.
+Unfortunately, one disadvantage is also important;
 the abstraction elimination can cause up to a quadratic 
@@
 
-The first thing to do in a standard subtree swapping crossover is to select random node in the first parent.
+The first thing to do in a standard subtree swapping cross\-over is to select random node in the first parent.
 
@@
 
-After selecting a node in the first parent we select node in the
-second parent such that the type of that node must correspond to the type of
-of the first node. Again, this may eliminate the "90-10" step of
-first deciding whether the selected node will be internal node 
-or leaf node.
-
-When both nodes are selected we may swap the trees. \\
+After selecting a node in the first parent we select node in the second parent such that the type of that node must correspond to the type of the first node.
+Again, this may eliminate the "90-10" step of first deciding whether the selected node will be internal node or leaf node.
+
+When both nodes are selected we may swap the trees.
 
 If the abstraction elimination was performed, then 
-since the trees are of the same type, and there are no variables to be 
-moved from their scope, the offspring trees are well typed.\\
+since the trees are of the same type, and there are no variables to be moved from their scope, the offspring trees are well typed.
 
@@
 corresponds to a subtree of \atTree, but there are subtrees
-of \atTree that do not correspond to a subtree of a \sexprTree.\\
-
-The following example should clarify this.
+of \atTree that do not correspond to a subtree of a \sexprTree.
+
+The following example should clarify this:
+
+\vfill
 
@@
 ]
-\Tree[.\textbf{f} x y ]~\\
+\Tree[.\textbf{f} x y ]
+
+\vfill
 
@@
 whereas in \sexprTree it is an internal node and thus not a subtree.
-Those favorable properties of \atTree{} are also reported in \cite{yu1998polygp}.\\
+Those favorable properties of \atTree{} are also reported in \cite{yu1998polygp}.
 
@@
 is that they are the same representation as S-expressions
-used by standard GP.\\
+used by standard GP.
 
@@
 
-~\\
-
 \section{Experiments}
@@
 
-We have tested our approach on the well-known even parity problem.
-
+In this section we demonstrate our approach on the well-known even parity problem. More extensive experimental evaluation will be a subject of future work. 
 We made two experiments, in the first we have used the \textit{hybrid} crossover, while in the second we have used the \textit{unpacking} crossover.
@@
 Each experiment consisted of 50 independent runs 
-of GP algorithm. Each run had maximally 51 generations and population size
+of GP algorithm. Each run had a limit of 51 generations, and the population size
 was 500 individuals.
@@
 the average size of an individual,
-the ability of the system to produce a correct solution 
+the ability of the system to produce a correct solution, 
 and computational cost estimates throughout generations. 
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
   \centering
   \caption{Graphs for Even parity problem.}
-  \includegraphics[scale=0.251]{grafy/EP.PNG}
+  \includegraphics[scale=0.2]{grafy/EP.PNG} %0.251
 \end{figure}
@@
 \caption{Statistical analysis.}
-\end{table}~\\
+\end{table}
 
@@
 The average individual size for generation 50 was 83.1.
-The experiment took 388 minutes.
 
@@
 The average individual size for generation 50 was 74.3.
-The experiment took 33 minutes.
-
-Comparison with other results is summarized in table 2.
+
+Comparison with other results taken from literature is summarized in table 2.
+Concerning the time complexity, both experiments were run under same conditions on a fairly standard machine with 4 core 3.3 GHz processor and 4 GB RAM.
+The first experiment took 388 minutes while the second one took
+33 minutes.
 
@@
 \caption{Results comparison.}
-\end{table}~\\
+\end{table}
+
+We have also analyzed ratio of failed crossovers due to exceeding the maximal permissible depth.
+Only around 0.5\% of offspring created by the unpacking crossover fail due to exceeding the limit and 0.2\% for the hybrid crossover.
+Thus we conclude that the case of crossover failure does not have a significant effect on the performance of the operators.
+
+%\red{ukazat jedince krizeni a okomentovat}
+
+In order to illustrate how the individuals actually look like and how the crossover works, figure~2 depicts creation of the correct solution by the hybrid crossover in one particular run of the system.
+
+The found correct solution can be rewritten as: 
+\begin{align*}
+\lambda~x~.~&foldr~(\Scomb(\Scomb (\Kcomb~\Scomb) (\Scomb (\Kcomb (\Scomb  (\Kcomb~nor) ) ) and) ) nor )  \\
+ & (nor~(head'~x)~(head'~x))~ ~(tail'~x)\\
+ \beeq & \\
+\lambda~x~.~&foldr~(\lambda~y~z~.~nor~(and~y~z)~(nor~y~z)  )\\
+ & (nor~(head'~x)~(head'~x))~ ~(tail'~x)\\
+ \textit{Which~} & \textit{is equivalent to:}\\
+\lambda~x~.~&foldr~xor~(not~(head'~x))~ ~(tail'~x) 
+\end{align*}
+
+\begin{figure}[!ht]
+  \centering
+  \caption{Hybrid crossover example.}
+  \includegraphics[scale=0.59]{crossover.png} %0.251
+\end{figure}
 
@@
 
-We have presented two approaches to crossover for typed GP over lambda terms.
-The first one (\textit{hybrid}) transforms all generated individuals into combinator terms at the end of generation phase.
-The second crossover approach (\textit{unpacking}) uses the lambda term representation during the whole run, except for the temporary phases during the crossover.
-We have compared their relative performance on even parity problem.
-It shows that the first one is performing the same or better in all observed metrics.
-This was surprising result for us.
+In this paper we presented a generalization of standard GP based on simply typed lambda calculus trees.
+We discussed the advantages of generating lambda terms in their long normal form and presented two novel crossover operations for this system.
+
+The first one --- the \textit{hybrid} crossover --- transforms all generated individuals into combinator terms at the end of generation phase.
+The second crossover approach --- the \textit{unpacking} crossover --- uses the lambda term representation during the whole run, except for the temporary phases during the crossover.
+We have compared their relative performance on even parity problem; the first one is performing the same or better in all observed metrics.
+Since the use of the \textit{unpacking} crossover brings a significant computational overhead while its performance is the same or worse than the performance of the \textit{hybrid} crossover we conclude that
+the \textit{hybrid} crossover is better suited for the job.
+%\red{This was surprising result for us.}
 
 According to comparison of "individuals to be processed" metric our GP approach is performing better than the GP system presented in \cite{kes} and worse than the one presented in \cite{yu01}.
-By observation of typical solution which is often some modification of \texttt{foldr xor False inputList} one sees that an important task for GP here is to create \texttt{xor} function.
-This is the advantage for the more specialized term representation used in \cite{yu01}, which
-restricts the use of outer variables. 
+By observation of typical solution which is often some modification of \texttt{foldr xor True inputList} one sees that an important task for GP here is to create \texttt{xor} function.
+This is the advantage for the more specialized term representation used in \cite{yu01}, which restricts the use of outer variables.
+
 It is also fair to say, that this popular metric (in which the comparison data are available) is somehow problematic, since it is prone to have large variation across various instances of the same experiment \cite{luke2002perfect}.
-As is also visible from our results; both our approaches scored best in generation 0 thus the result depended only on the generation method which those two approaches shared, yet the individuals to be processed estimates are quite different.
-
-In the future work, we would like to accompany our theoretical results concerning relations of generating procedure and reductions described in this paper with experimental results.
-
-In this paper we touched varied aspects of the two lambda calculus concepts -- reductions and abstraction elimination -- in the hope that it will shed some more
-light on the promising intersection of lambda calculus and genetic programming.
+As is also visible from our results; both our approaches scored best in generation 0 thus the result shown in table 2 depended only on the generation method which those two approaches shared, yet the individuals to be processed estimates are quite different.
+
+In the future work, we would like to support our theoretical results concerning relations of generating procedure and reductions described in this paper with substantially extended experimental results.
+
+In this paper we touched varied aspects of the two lambda calculus concepts -- reductions and abstraction elimination -- in the hope that it will shed some more light on the promising intersection of lambda calculus and genetic programming.
 
@@
 %ACKNOWLEDGMENTS are optional
-%\section{Acknowledgments}
-%This section is optional; it is a location for you to acknowledge grants, funding, editing assistance and what have you.
-In the present case, for example, the authors would like to thank Gerald Murray of ACM for his help in codifying this \textit{Author's Guide} and the \textbf{.cls} and \textbf{.tex} files that it describes.
-
-%\section{Osnova}
-%\begin{lizt}
-%& Jak reprezentovat stromy programu pro GP?
-%  && Reprezentace v klasickym kozovy je S-expression.
-%  && Nebo mužem používat kombinátory jako 
-%     Briggs a O’Neil (to si myslim je jakoby 
-%     hlavní konkurence, vuči který by to chtělo obhájit)
-%  && Lambda termy a jejich awesomeness
-%& \textbf{Povídaní o redukcích}
-%& Povídání o lnf
-%  && že lnf je přirozený rozšíření s-exprešnu do lambda 
-%     kalkulu vlastnosti termu v lnf 
-%     &&& proč je eta redukovat (eta redukcí lnf dostanem 
-%         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
-%& Generování
-%  && Generování gramatikou
-%  && Gramatika pro lnf
-%  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
-%& Problémy s proměnýma a jak je řešit.
-%  && Křížit lambda termy i s proměnejma a abstrakcema, 
-%     problémy řešit když nastanou (pomluvit a odsoudit)
-%  && Zmenšit prostor termu (tim že někerý nejsme schopný 
-%     vygenerovat) s kterým operujeme tak aby křížení už 
-%     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
-%     jen použití proměnných z její hlavy)
-%  && Převod hned po vygenerování
-%  && Převod až při křížení
-%     &&& eliminace abstrakcí
-%     &&& skřížim
-%     &&& vložený kombinátory nahradim odpovídajícím termem
-%     &&& celý to redukuju
-% & Pokusy (?!)
-% & Závěr
-%\end{lizt}
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
+\section{Acknowledgments}
+
+Tom\'{a}\v{s} K\v{r}en has been partially supported by the project GA \v{C}R P202/10/1333 and by the SVV project number 260~104.
+Roman Neruda has been partially supported by The Ministry of Education of the Czech Republic project COST LD 13002.
+
 \bibliographystyle{abbrv}

### 2014-04-15 tex/gecco14_old/gecco14.tex
cla

--- 2014-04-15 tex/gecco14/gecco14.tex
@@
 
-\keywords{genetic programming; lambda calculus}
+\keywords{genetic programming, lambda calculus}
 
@@
 the traditional GP representation of individuals, with concepts from
-lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory~\cite{yu01}.
+lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory. Such attempts have shown to be successful \cite{yu01}.
 
@@
 
-In our system we use a simply typed lambda calculus.
-Since abstraction elimination and reductions do not involve many type related issues it seems to us as a sufficient choice of a type system.
+In our system we use simply typed lambda calculus.
+Since abstraction elimination and reductions do not involve many type related issues it seems to us as a sufficient choice of type system.
 Techniques described here would be analogous for a more sophisticated type system and can be directly applied to such systems as the Hindley–Milner type system.
 
-In this paper we present an approach to typed GP over lambda calculus capable of operating with the full array of notions given by a simply typed lambda calculus while being a direct generalization of standard GP~\cite{koza92}.
+In this paper we present an approach to typed GP over lambda calculus capable of operating with the full arsenal given by simply typed lambda calculus while being a direct generalization of standard GP\cite{koza92}.
 We discuss the advantages of generating lambda terms in their long normal form and present two novel crossover operations for simply typed lambda terms.
-Compared to the approach to typed GP over lambda calculus  proposed by
-Yu~\cite{yu01}, it uses more general term representation which is not restricting the use of outer local variables in the body of an anonymous function.
-Compared to the purely combinator approach by Briggs and
-O’Neill~\cite{kes}, it has the advantage of reducing the size of the search space in the term generation phase.
+Compared to the approach to typed GP over lambda calculus  proposed by Yu\cite{yu01} it uses more general term representation which is not restricting the use of outer local variables in the body of an anonymous function.
+Compared to the purely combinator approach by Briggs and O’Neill\cite{kes} it has the advantage of reducing the size of the search space in the term generation phase.
 
@@
 The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
-The body of the paper is presented in sections \ref{generating} and \ref{crossover}; the former discusses the term generating 
+The body of the paper is presented in sections \ref{generating} and \ref{crossover}; the first discusses the term generating 
 procedure and its relations
-with reductions, the latter describes several approaches to lambda term
-crossover, and how they solve the variable problem. Section~\ref{experiments} examines those ideas on the even parity benchmark problem, while the paper is concluded by section~\ref{conclusions}.
+with reductions, the second describes several approaches to lambda term
+crossover and how they solve the variable problem. Section~\ref{experiments} examines those ideas on the even parity benchmark problem, while the paper is concluded by section~\ref{conclusions}.
 
@@
 %\footnote{Higher-order function is a function taking another function as input parameter.} 
-and lambda abstractions~\cite{yu01}.
-An important point of interest in the work is the use of
-\texttt{foldr} function as a tool for \textit{implicit recursion}, i.e., recursion without explicit recursive calls.
+and lambda abstractions  \cite{yu01}.
+An important point of interest in the work is use of
+\texttt{foldr} function as a tool for \textit{implicit recursion}, i.e.
+recursion without explicit recursive calls.
 The terminal set for constructing lambda abstraction subtrees  is limited to use only constants and variables of that particular lambda abstraction, i.e., outer variables are not allowed to be used as terminals in this work.
@@
 utilizing typed GP with combinators \cite{kes}.
-The difference between the approach presented in their work
+The difference between approach presented in their work
 and our approach is that they generate terms in a straightforward way directly from the library of combinators, without any use of lambda abstractions.
 They use the Hindley–Milner type system.
-They also present an interesting concept of \textit{Generalized genetic operator} based on term generation.
+They also present interesting concept of \textit{Generalized genetic operator} based on term generation.
 The combinator based genetic algorithm has the advantage of avoiding the need of dealing with variables.
 
-Binard and Felty use an even stronger type system, the (\textit{System F}) \cite{binard2008genetic}.
+Binard and Felty use even stronger type system (\textit{System F}) \cite{binard2008genetic}.
 But with the increasing power of the type system also comes an increasing difficulty of term generation.
 For this reason, the evolution in this work takes interesting and nonstandard shape (fitness is associated with \textit{genes} which are evolved together with \textit{species} which together participate in creation of individuals).
-This differs from our approach which tries to be a straightforward generalization of the standard GP\cite{koza92}.
-
-In contrast to the above mentioned works, our approach uses a very simple type system (the simply typed lambda calculus) and concentrates on process of generation  
+This differs from our approach, which tries to be a generalization of the standard GP\cite{koza92}.
+
+In contrast with the above mentioned works, our approach uses very simple type system (simply typed lambda calculus) and concentrates on process of generation  
 able to generate all possible well-typed normalized lambda terms. In order to do so we use technique based on \textit{inhabitation machines} described by Barendregt\cite{barendregt10}.
@@
 Each experiment consisted of 50 independent runs 
-of GP algorithm. Each run had a limit of 51 generations, and the population size
+of GP algorithm. Each run had maximally 51 generations and population size
 was 500 individuals.
@@
 the average size of an individual,
-the ability of the system to produce a correct solution, 
+the ability of the system to produce a correct solution 
 and computational cost estimates throughout generations. 
@@
 The average individual size for generation 50 was 83.1.
+The experiment took 388 minutes.
 
@@
 The average individual size for generation 50 was 74.3.
-
-Comparison with other results taken from literature is summarized in table 2.
-Concerning the time complexity, both experiments were run under same conditions on a fairly standard machine with 4 core 3.3 GHz processor and 4 GB RAM.
-The first experiment took 388 minutes while the second one took
-33 minutes.
+The experiment took 33 minutes.
+
+Comparison with other results is summarized in table 2.
 
@@
 
-We have also analyzed ratio of failed crossovers due to exceeding the maximal permissible depth.
+We also analyzed ratio of failed crossovers due to exceeding the maximal permissible depth.
 Only around 0.5\% of offspring created by the unpacking crossover fail due to exceeding the limit and 0.2\% for the hybrid crossover.
-Thus we conclude that the case of crossover failure does not have a significant effect on the performance of the operators.
+Thus we conclude that it does not have a significant effect on the performance of the operators.
 
@@
   \caption{Hybrid crossover example.}
-  \includegraphics[scale=0.59]{crossover.png} %0.251
+  \includegraphics[scale=0.6]{crossover.png} %0.251
 \end{figure}
@@
 
-The first one --- the \textit{hybrid} crossover --- transforms all generated individuals into combinator terms at the end of generation phase.
-The second crossover approach --- the \textit{unpacking} crossover --- uses the lambda term representation during the whole run, except for the temporary phases during the crossover.
+The first one (\textit{hybrid}) transforms all generated individuals into combinator terms at the end of generation phase.
+The second crossover approach (\textit{unpacking}) uses the lambda term representation during the whole run, except for the temporary phases during the crossover.
 We have compared their relative performance on even parity problem; the first one is performing the same or better in all observed metrics.
@@
 According to comparison of "individuals to be processed" metric our GP approach is performing better than the GP system presented in \cite{kes} and worse than the one presented in \cite{yu01}.
-By observation of typical solution which is often some modification of \texttt{foldr xor True inputList} one sees that an important task for GP here is to create \texttt{xor} function.
+By observation of typical solution which is often some modification of \texttt{foldr xor False inputList} one sees that an important task for GP here is to create \texttt{xor} function.
 This is the advantage for the more specialized term representation used in \cite{yu01}, which restricts the use of outer variables.
@@
 
+\vfill
+
 %ACKNOWLEDGMENTS are optional
@@
 
-Tom\'{a}\v{s} K\v{r}en has been partially supported by the project GA \v{C}R P202/10/1333 and by the SVV project number 260~104.
+Tomáš Křen has been partially supported by the project GA ČR P202/10/1333 and by the SVV project number 260~104.
 Roman Neruda has been partially supported by The Ministry of Education of the Czech Republic project COST LD 13002.

### 2014-04-15 tex/cec14/article.tex
cec14 hotovo

@@
 Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
-Use of types naturally opens door to enriching S-expressions,
+The use of types naturally leads to enriching S-expressions,
 the traditional GP representation of individuals, with concepts from
-lambda calculus, which is simple yet powerful functional mathematical and programming 
+lambda calculus, which is simple yet powerful functional, mathematical, and programming 
 language extensively used in type theory. Such attempts has shown to be 
@@
 %The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation. 
-In this paper we present a tree generating method suitable both for the standard GP\cite{koza92} and  typed GP over lambda calculus.
-We chose to describe the algorithm in the means of the typed lambda calculus since the standard S-expression representation can be understood as a special case of this representation.
-However, for evaluation of our approach we focus more on problems from standard GP.
-We chose this evaluation approach because the result of the GP system depends profoundly on the crossover operation, which remains an open problem for the typed GP over lambda calculus.
+In this paper we present a tree generating method suitable both for the standard GP~\cite{koza92} and for the typed GP over lambda calculus.
+We chose to describe the algorithm by means of the typed lambda calculus, since the standard S-expression representation can be understood as a special case of this representation.
+However, for evaluation of our approach, we focus more on problems from the standard GP domain.
+The main reason for this is that the result of GP system depends on the crossover operation, which still remains an open problem for the typed GP over lambda calculus.
 
@@
 search strategy}\footnote{Not to be confused with geometric  crossover introduced by \cite{moraglio2010}.} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
-Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
+Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows -- the bloat problem.
 
@@
 
-We made three experiments comparing performance of standard 
+We made three experiments comparing the performance of standard 
 \textit{ramped half-and-half strategy} with our
-\textit{geometric strategy} (with default parameter $q=0.75$).
-
-Despite the fact that the geometric strategy was designed for typed GP over lambda trees it works well for standard GP problems satisfying the closure requirement.
-Furthermore, for such problems our approach generates standard S-expressions only.
-Another reason for including standard GP benchmarks in our experiments is the fact that the results are seriously affected by a crossover
+\textit{geometric strategy} (with default parameter value $q=0.75$).
+
+Despite the fact that the geometric strategy was designed for typed GP over lambda trees, it works well also for standard GP problems satisfying the closure requirement.
+Furthermore, for such problems our approach generates the standard S-expressions only.
+Another reason for including standard GP benchmarks in our experiments is the fact that the results are seriously affected by the choice of a crossover
 operator.
 Design of a good crossover for typed GP remains an open problem.
-Our crossover is quite complicated and beyond the scope of this paper, but for problems satisfying the closure requirement it behaves exactly the same as the standard GP crossover.
+Our crossover design for typed GP is quite technical and beyond the scope of this paper, but for problems satisfying the closure requirement it behaves exactly the same as the standard GP crossover.
 
@@
 Even parity problem) were chosen for our experiments.
-First two satisfy the closure requirement, 
+The first two satisfy the closure requirement, 
 whereas the third one does not.
-
-In order to ensure that the experiments are replicable we chose to use the standard GP system described in \cite{koza92}, with all the parameters set to their default values (which are sumarized in table~I), only with the following exceptions:
+In order to ensure that the experiments are replicable we chose to use the standard GP system described in \cite{koza92}, with all the parameters set to their default values (which are summarized in table~I), with the following exceptions:
 
@@
   \item In the second and third experiment the preservation of the best individual into the next generation is performed.
-  \item In the third experiment we use our crossover operation briefly mentioned above together with use of "applicative" tree representation.
+  \item In the third experiment we use our crossover operation briefly mentioned above together with the appropriate tree representation.
 \end{enumerate}
@@
 Each experiment consisted of 50 independent runs 
-of GP algorithm. Each run had maximally 51 generations and population size
+of a GP algorithm. Each run had a limit of 51 generations, and the population size
 was 500 individuals.
-In those experiments we analyze the fitness of the best individual,
+In the experiments we analyze the fitness of the best individual,
 the average size of an individual,
 the ability of the system to produce a correct solution 
-and computational cost estimates throughout generations. 
+and the computational cost estimates throughout generations. 
 For the last two metrics we use the popular measurement  
-methods within GP field --- the \textit{performance curves}
+methods within the GP field --- the \textit{performance curves}
 described in \cite{koza92} --- $P(M,i)$ (cumulative probability of success) 
@@
 to yield a correct solution with probability $z =99\%$).
-Difference of those two generating methods 
+The difference of those two generating methods 
 is statistically analyzed by the Welch t-test
 by comparing the fitness values of the best individuals of the run. 
-Graphs showing mean values of the fitness of the best individual and
-the average size of an individual throughout generations
+The graphs showing mean values of the fitness of the best individual and
+the average size of individuals throughout generations
 contain error bars representing the standard error of the mean (SEM).
+
+The experiments were performed under identical conditions on a fairly standard
+desktop computer with 4 core 3.3 GHz processor and 4 GB RAM. The running times noted for each experiment should be considered only a rough estimate of the time complexity of particular methods.
 
@@
   \caption{Graphs for Simple symbolic regression.}
-  \includegraphics[scale=0.12]{imgs/SSR.eps}
+  \includegraphics[scale=0.115]{imgs/SSR.eps}
 \end{figure}
@@
 \textit{Simple Symbolic Regression} is a problem described
-in \cite{koza92}. Objective of this problem is to 
+in \cite{koza92}. The objective of this problem is to 
 find a function $f(x)$ that fits a sample
 of twenty given points. The target function is 
-function $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
-Desired type of generated programs $\tau$ and 
-building blocks context $\Gamma$ are following.
+a function defined as $f_{t}(x) = x^4 + x^3 + x^2 + x$.  
+The desired type of generated programs $\tau$ and 
+the building blocks context $\Gamma$ are the following.
 \begin{align*}
@@
 
-Fitness function is computed as follows.
+The fitness function is computed as follows.
 
@@
 
-Figure~1 shows results of this experiment. 
+Figure~1 shows the results of this experiment. 
 Table~II summarizes statistical analysis
@@
 Standard \textit{ramped half-and-half strategy} scored 17/50 (34\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 23 with 192,000 individuals to be processed.
+The minimal $I(M,i,z)$ value was in the generation 23 with 192,000 individuals to be processed.
 The average individual size for generation 50 was 68.8.
@@
 Our \textit{geometric strategy} scored 21/50 (42\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 29 with 150,000 individuals to be processed.
+The minimal $I(M,i,z)$ value was in generation 29 with 150,000 individuals to be processed.
 The average individual size for generation 50 was 44.1.
 The experiment took 26 minutes.
-So for the \textit{Simple symbolic regression}
+Thus, for the \textit{Simple symbolic regression}
 our geometric strategy is slightly more successful than 
-ramped half-and-half strategy in these observed metrics.
+the ramped half-and-half strategy in these observed metrics.
 
@@
 \textit{Artificial Ant} is another problem described
-in \cite{koza92}. Objective of this problem is to 
+in \cite{koza92}. The objective of this problem is to 
 find a control program for an artificial ant so
-that it can find all food located on "Santa Fe" trail.
-The Santa Fe trail lies on toroidal square grid.
+that it can find all food items located on the "Santa Fe" trail.
+The Santa Fe trail lies on a toroidal square grid.
 The ant is in the upper left corner, facing right.
@@
 
-Actions $l$ and $r$ turn the ant left and right, respectively. 
-Action $m$ moves the ant forward.
-Action $ifa~x~y$ \mbox{(\textit{if-food-ahead})} 
-performs action $x$ if a food piece is ahead of the ant,
-otherwise it performs action $y$.
-Actions $p2$ and $p3$ perform two and three consecutive actions, respectively.
-Actions $l, r$ and $m$ each take one time step to execute.
-Ants action is performed over and over again until it reaches predefined
+The actions $l$ and $r$ turn the ant left and right, respectively. 
+The action $m$ moves the ant forward.
+The action $ifa~x~y$ \mbox{(\textit{if-food-ahead})} 
+performs the action $x$ if a food piece is ahead of the ant,
+otherwise it performs the action $y$.
+The actions $p2$ and $p3$ perform two and three consecutive actions, respectively.
+The actions $l, r$ and $m$ each take one time step to execute.
+An ant action is performed over and over again until it reaches the predefined
 maximal number of steps. 
-Fitness value is equal to number of eaten food pieces.
+The fitness value is equal to number of eaten food pieces.
 An individual such that eats all 89 pieces of food is 
-considered as a correct solution.
+considered a correct solution.
 This limit is set to be 600 time steps\footnote{
 In \cite{koza92} this limit is said to be 400 time steps.
-But there is also mentioned one solution, which is described as correct solution. This program needs 545 time steps; if it is given only 400 time steps, then it eats only 79 pieces of food. 
+But there is also one solution mentioned as correct but needing 545 time steps. (for 400 time steps it eats only 79 pieces of food). 
 %But there is also mentioned following solution, which is described as correct solution: \texttt{(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))}.
@@
   \caption{Graphs for Artificial ant problem.}
-  \includegraphics[scale=0.13]{imgs/ANT.eps}
+  \includegraphics[scale=0.112]{imgs/ANT.eps}
 \end{figure}
@@
 by conventional criteria, this difference is considered to be extremely statistically significant.
-Standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 28 with 449,500 individuals to be processed.
+The standard \textit{ramped half-and-half strategy} scored 7/50 (14\%) success rate. 
+The minimal $I(M,i,z)$ value was in the generation 28 with 449,500 individuals to be processed.
 The average individual size for generation 50 was 279.5.
@@
 Our \textit{geometric strategy} scored 19/50 (38\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 10 with 115,500 individuals to be processed.
+The minimal $I(M,i,z)$ value was in generation 10 with 115,500 individuals to be processed.
 The average individual size for generation 50 was 94.1.
@@
 
-Previous two problems were instances of standard GP, i.e. their building 
+The previous two problems were instances of standard GP, i.e. their building 
 symbols obeyed the closure requirement. This third experiment will 
-break the closure requirement, thus ideal for testing typed GP techniques. 
+break the closure requirement, thus it is an ideal candidate for testing typed GP techniques. 
 The even-parity function is a boolean function taking as inputs $N$
@@
 that obtained by Yu in \cite{yu01},
-where is presented approach to evolve recursive and modular programs
-by use of higher-order functions and $\lambda$-abstractions.
+where an approach to evolve recursive and modular programs
+by use of higher-order functions and $\lambda$-abstractions is presented.
 We use very similar set of \textit{building symbols} as in \cite{yu01}. 
@@
 function \texttt{foldr}. 
-And modifications of functions \texttt{head} 
+Modifications of functions \texttt{head} 
 (returning the first element of the list) 
@@
   \caption{Graphs for Even parity problem.}
-  \includegraphics[scale=0.12]{imgs/EP.eps}
+  \includegraphics[scale=0.111]{imgs/EP.eps}
 \end{figure}
@@
 
-Standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 17 with 333,000 individuals to be processed.
+The standard \textit{ramped half-and-half strategy} scored 9/50 (18\%) success rate. 
+The minimal $I(M,i,z)$ value was in generation 17 with 333,000 individuals to be processed.
 The average individual size for generation 50 was 47.7.
@@
 Our \textit{geometric strategy} scored 32/50 (64\%) success rate. 
-Minimal $I(M,i,z)$ was in generation 0 with 28,000 individuals to be processed.
+The minimal $I(M,i,z)$ value was in generation 0 with 28,000 individuals to be processed.
 The average individual size for generation 50 was 74.3.
@@
 
-Once again this is a big improvement against standard method.
+Once again, this represents a big improvement over the standard GP method.
 
@@
 that term size around 75 is similar as for the previous two problems, 
-it seems to us that ramped half-and-half maybe has difficulties with 
-constructing more complex terms, and therefore is "staying small".
-
-Another remarkable observation is that 4 times out of 50 (8\% of all runs) 
-the 100\% correct solution was present in the generation 0. This is pretty good
+it seems to us that the ramped half-and-half strategy may have difficulties with 
+constructing more complex terms, and therefore it is "staying small".
+
+Another itnteresting observation is that 4 times out of 50 (8\% of all runs) 
+the 100\% correct solution was present in the generation 0. This is good
 result for uninformed search. However, our results are less successful 
@@
 building symbols); they scored 40/50 (80\%) success rate and $I(M,i,z)$ of 
-17,500 in generation 4. But by observation of typical solution which is often
-some modification of \texttt{foldr xor True inputList} one sees that important 
-task for GP here is to create \texttt{xor} function. This is advantage for the 
+17,500 in generation 4. But by an observation of the typical solution which is often
+some modification of \texttt{foldr xor True inputList} one sees that the important 
+task for GP here is to create the \texttt{xor} function. This is the advantage for the 
 more specialized term representation used in \cite{yu01} 
-(restricting use of outer variables). 
+(restricting the use of outer variables). 
 The difference in crossover operators may also be involved in the difference, 
-but such analysis is beyond the scope of this paper. Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
+but such analysis is beyond the scope of this paper. 
+Finally, our result at least outperforms other results mentioned in \cite{yu01}: \textit{Generic genetic programming} 
 scored 17/60 (28\%); min $I(M,i,z) = 220,000$. \textit{GP with ADFs} 
@@
 In this work, we have proposed a generalization of genetic programming 
-for simply typed lambda calculus. The main focus of this paper is on individual generation algorithms. 
+for simply typed lambda calculus. The main focus of this paper is on the individual generation algorithms. 
 The main idea is to explore the spectrum of available approaches bounded on one end by the traditional
@@
 the second one corresponds to exhaustive search, and the third one 
-represents a novel geometric strategy which outperforms standard genetic 
-programming in success rate, resource consumption and average individual size 
-on the presented experiments.
-
+represents a novel geometric strategy.
 %\red{ [větu dvě o experimentech]}
-
 Experiments on three well 
@@
 time consumption and average individual size.
-
 The relevance of these algorithms does not effect only initialization, but it can be used during mutation

### 2014-04-29 tex/wds14/2/wds2e_ex.tex
wds14 zaklad

%to be compiled with LATEX 2e

%use this setting with 11pt basic font:
\documentclass[11pt]{article}
\usepackage{wds11,epsf}

%use this setting with 10pt basic font:
%\documentclass{article}
%\usepackage{wds10,epsf}

%\tighten

% Preamble Information

\lefthead{SAFRANKOVA ET AL.}
\righthead{MAGNETOSHEATH RESPONSE}

\setcounter{secnumdepth}{0}

\begin{document}

\title{Example of WDS style article: IMF tangential discontinuity}

\author{J. \v Safr\'ankov\'a, L. P\v rech, and Z. N\v eme\v cek}

\affil{Charles University, Faculty  of  Mathematics  and  Physics,
     Prague, Czech Republic.}

\author{D. G. Sibeck}

\affil{JHU, APL, Laurel, Maryland.}

\author{T. Mukai}

\affil{Institute of Space and Astronautical Science, Sagamihara, Japan.}

\begin{abstract}
This is an example of the WDS style article in LATEX. The
original article has been heavily truncated.
We present a multipoint observational study of the magnetosheath
response to the interplanetary magnetic field tangential
discontinuities which form  hot flow anomaly-like structures.
We identify these structures (HFAs) in the vicinity of the bow shock
as well as deeper in the magnetosheath. Two or
more points of simultaneous observations allow us to describe the
gradual evolution and propagation of these HFAs through the
magnetosheath. From tens of events recorded by INTERBALL-1, we present
two cases. In the first, GEOTAIL identified HFAs in the solar  wind
near the bow shock, and INTERBALL-1 and MAGION-4
observed related events in the magnetosheath. During the second
interval, all three spacecraft observed the HFA
features in the magnetosheath. Our analysis of reported events
suggests the negligible evolution of these structures in the
magnetosheath. A survey of the INTERBALL-1 data has shown that
magnetosheath HFAs are observed predominantly during periods of
fast solar wind.
\end{abstract}

\begin{article}

\section{Introduction}

The interaction of the solar wind with the Earth's magnetosphere
generates a population of backstreaming ions directed from the
bow shock into the solar wind. This high-energy population was
invoked to explain Hot Flow Anomalies (HFAs) [\markcite{{\it
e.g., Schwartz et al.}, 1985; {\it Thomsen et al.}, 1986}]
identified as heated regions of solar wind plasma
flowing nearly perpendicular to the Earth-Sun line.
The main observational features of HFAs include \markcite{{\it
Schwartz} [1995]}: (1) Central regions with hot plasma flowing
significantly slower than that in the ambient solar wind in a
direction highly deflected (nearly $90^o$) from the Sun-Earth line.
The flow velocities are often roughly tangential to the nominal
bow shock shape [\markcite{{\it Schwartz et al.}, 1988}].
(2) HFAs are bounded by regions of enhanced
magnetic field strength, density, and temperature. The outer
edges of these enhancements are fast shocks generated by
pressure enhancements within the core region. The inner
edges of the enhancements are probably  tangential
discontinuities [\markcite{{\it Paschmann et al.}, 1988}].
Published examples indicated that  many HFAs are bounded by only
one enhancement.
(3) HFAs occur in conjunction with significant changes in the IMF
direction. The angle between pre- and post event orientations is
typically $\sim 70^o$.

\markcite{{\it Thomsen et al.} [1986, 1988]} have shown that many
HFAs occur when the geometry of the bow shock is changing and large fraction
of specularly reflected ions are present. These and others
observations led to many theoretical/numerical studies.
One-dimensional simulations with a finite length backstreaming
ion beam suggested
that the interaction of the beam and background plasma can produce hot,
low-density regions from which the solar wind plasma is largely
excluded [\markcite{{\it Onsager et al.}, 1990b}]. However, an
examination of the ion temperature inside HFAs indicated that
complete thermalization of the ion beam and subsequent adiabatic
expansion of the heated plasma leads to final ion temperatures that
are generally below those observed [\markcite{{\it Onsager et al.},
1990a}]. \markcite{{\it Thomas and Brecht} [1988]} presented
a 2-D hybrid simulation for a beam of backstreaming ions
relative to the ambient solar wind. The authors demonstrated that
the thermalized backstreaming ions create a diamagnetic cavity of
depressed magnetic field  strengths and densities.
According to numerical simulations of \markcite{{\it  Thomas et
al.} [1991]} and \markcite{{\it
Lin} [1997]}, kinetic effects at the intersection of the magnetic
discontinuities with the bow shock can create very deflected
flows of heated plasma surrounded by enhanced densities and
magnetic field strengths.

\section{Observations}
To demonstrate HFA properties, we have chosen two representative
examples with favourable positions for all mentioned spacecraft.
The first case study shows an interval when two of three IMF tangential
discontinuities identified in WIND solar wind data resulted in HFAs.
These HFAs are observed in the bow shock region by GEOTAIL,
and simultaneously by
INTERBALL-1 in the magnetosheath. The second case is related to the
propagation of HFAs through the magnetosheath. We have found an
interval when the same HFA is observed by GEOTAIL in the dayside
magnetosheath and by INTERBALL-1 and MAGION-4 in the dusk
magnetosheath flank, about $\sim 20\> R_E$ downstream.

In order to
show that the solar wind velocity can be an important factor for the
HFA creation, we have reanalyzed data published in
[\markcite{{\it Onsager et al.}, 1990a}] and
complemented their survey of ISEE and AMPTE observations by our
analysis of INTERBALL-1 measurements.

\subsection{Case 1}
On August 31, 1996, the INTERBALL-1 satellite
registered a series of HFA-like events in the dawn magnetosheath
at (-0.2; -17.2; -5.1)$_{GSE}\> R_E$. Figure~1 shows observations during two of
these events.  This figure presents 15~s
time resolution VDP ion flux [\markcite{{\it Safrankova et
al.}, 1997}] and 1~s time resolution MIF-M magnetic field
[\markcite{{\it Klimov et al.}, 1997}] observations for the interval
from 0740 to 0820 UT.
%The events at $\sim 0745$ and 0805~UT
The first HFA at 0740-0750 UT as well as the second one at
0805-0815 UT can be identified by rapid decreases of the ion flux
bounded by transient enhancements (top panel in Figure~1). The time
resolution of the measurements does not allow us to see the full
amplitude of the enhancements in this case but our analysis of
similar events has shown that they can exceed the mean magnetosheath flux
by a factor of 3.
The cone angle (second panel in Figure~1) of the ion flux shows
that the flow is highly deflected. Whereas the undisturbed
magnetosheath flow is deflected by $\sim 20^o$ from the
Sun--Earth's line, consistent with INTERBALL-1's location near
the terminator, the ions inside the structures are highly
deflected. They flow nearly perpendicularly to this line during our
second event.

%\begin{figure}[htb]
%\begin{center}
%\begin{tabular}{c}
%  \epsfxsize=102mm
%  \epsfysize=100mm
%%\epsfbox{iii60831.ps}
%\epsfbox{iii.ps}
%\end{tabular}
%\end{center}
%\caption{An observation of HFAs in the magnetosheath:
%INTERBALL-1 data. From top to bottom: $f_{\rm VDP}$ (ion flux), alpha
%(cone angle between ion bulk velocity and the $X_{\rm GSE}$
%coordinate), $f_{\rm DOK}$ (flux of electron in the range
%28-31~keV), $B$ (magnetic field magnitude), $B_X, B_Y, B_Z$ (components of
%the magnetic field), $E_{\rm e0}$ (electron energy spectrogram).}
%\end{figure}

We can compute its normal from $({\bf \vec B_1\times \vec
B_2})/({\bf |\vec B_1| \cdot |\vec B_2|})$ where the subscripts 1
and 2 refer to the region upstream and downstream of a
discontinuity, respectively. We tested the normal direction by
minimum variance analysis of the WIND high-resolution (3~s) data.
Both approaches yield the same direction of the discontinuity
normal (${\bf n} = 0.22, 0.67, 0.70$).
The normal defines the discontinuity plane and we can compute
the time lag for observations of the discontinuity at
WIND, IMP-8 and GEOTAIL (GEOTAIL observations will be discussed
in detail later).

Thus, we can conclude that both necessary conditions ($B_n = 0$ and
$u_n = 0$ where $B_n$ and $u_n$ are normal components of the magnetic field
and velocity, respectively) are fulfilled and that all
discontinuities in this time interval are tangential
discontinuities. Such discontinuities can create HFAs, when the
motional electric field ${\bf \vec v_{sw} \times \vec B}$ is oriented
toward the discontinuity
on one or both sides of the current sheet [\markcite{{\it Lin et
al.}, 1997}; \markcite{{\it Thomsen et al.}, 1988}].
We tested this criterion and found that it is obeyed for the first
and third discontinuities but not for the second one.
%has oppositely oriented electric field and
Thus only the first and third discontinuities can create HFAs
when interacting with the bow shock.
This is consistent with magnetosheath observations presented in
Figure~1 where only two HFA-like events are seen during the time
interval under question.

\subsection{Case 2}
A similar situation occurred on January 31, 1997. The IMF
tangential discontinuity observed by WIND at $\sim 1953$~UT resulted
in the creation of an HFA observed in the dusk
magnetosheath by three spacecraft: GEOTAIL, INTERBALL-1, and
MAGION-4. For the sake of simplicity, Figure~2 shows only the ion
flux and magnetic field magnitudes observed by these
spacecraft but we have carefully examined all parameters to ensure
that the magnetosheath event has  HFA characteristics. It should be
noted that the characteristics of the event differ a little at
different spacecraft positions.

%\begin{figure}[htb]
%\begin{center}
%\begin{tabular}{c}
%  \epsfxsize=102mm
%  \epsfysize=100mm
%  \epsfbox{gim70131.ps}
%\end{tabular}
%\end{center}
%\caption{Propagation and evolution of HFA through the
%magnetosheath. $B$ and $f$ stand for the magnetic field magnitude
%and ion flux, respectively, and subscripts are abbreviations of
%the spacecraft names.}
%\end{figure}

The locations of the spacecraft projected onto the ecliptic plane are
schematically shown in the left part of Figure~3. GEOTAIL was
located in the magnetosheath near the magnetopause and registered
the leading enhancement of HFA at 2034~UT. Due to the decreased
density in the core region of HFA, GEOTAIL crossed magnetopause
and entered the plasma sheet. It exited into HFA at 2041~UT to
observe the trailing enhancement of the ion flux.

%\begin{figure}[htb]
%\begin{center}
%\begin{tabular}{c}
%  \epsfxsize=82mm
%%  \epsfysize=30mm
%  \epsfbox{dvepara2.eps}
%\end{tabular}
%\end{center}
%\caption{ A projection of the satellite positions into ecliptic
%plane during discussed events. The right part of the figure
%belongs to August 31, 1996 event, the left part refers to January
%31, 1996 event. The intersection of discontinuities with the
%ecliptic plane is shown by double lines.}
%\end{figure}

\subsection{A survey of INTERBALL HFA observations}
Both cases analyzed above occurred during a period of relatively
high solar wind speed (505$\>km/s$ for the first case and
495$\>km/s$ for the
second case). In order to determine whether or not this is a random
occurrence, we surveyed  3 months of the INTERBALL-1
observations (January - March, 1997).
For our survey, we have taken into account only those events
which had all basic HFA characteristics, i.e., a region of tenuous
highly deflected flow and low magnetic field bounded by short
regions of significantly enhanced ion density and magnetic field
magnitude. We exclude events connected with bow shock or
magnetopause crossings because the plasma and magnetic field
behaviour are more complicated in such cases and a complex
analysis of all parameters is needed to decide whether the particular
event is connected with the HFA.

\section{Discussion}
We have presented two multipoint case studies of the
magnetosheath response to the arrival of  IMF tangential
discontinuity at the bow shock.
In both of them, an HFA is created at the bow shock and then
observed in the magnetosheath. The INTERBALL-1 observation of HFA
in the magnetosheath is supported by simultaneous GEOTAIL
observation at the bow shock region in our first case. The analysis
of this event shows that the characteristics of HFA in the
solar wind and in the magnetosheath are basically the same. HFAs
can be distinguished as regions of the hot tenuous plasma bounded
by density enhancements in both regions.

In the second case, a single HFA was observed in the dusk magnetosheath
near the magnetopause by three spacecraft.
We suggested that the differing profiles of the
events at different points of the magnetosheath could be attributed
to the shape of a HFA cavity rather than to a temporal evolution of
its dimensions on a time scale of several minutes.
We suggest that the durations of HFAs depend on the locations
where IMF discontinuities intersect the bow shock.
Thus, the different durations of the events observed by
INTERBALL-1 and GEOTAIL  in our first case  may result from
GEOTAIL's separation from the subsolar region where the plasma
observed by INTERBALL-1 enter the magnetosheath.
Simulations indicate that HFAs need about 10
ion gyroperiods to develop into notable disturbances [\markcite{{\it
Thomas et al.}, 1991]}.

\section{Conclusion}
HFAs are an important magnetosheath phenomenon.
They can be encountered throughout the
magnetosheath from the bow shock to the magnetopause. Our
observations are limited by the INTERBALL-1 satellite's orbit to the
range $|X_{GSE}| < 10\> R_E$ but the negligible evolution of reported
events suggests that HFAs should be observed further tailward, if the
duration of the event ($\sim$ several minutes) and the temporal
resolution of observations allows their identification.

We have pointed out a {\it double structure} of magnetosheath
HFAs. A complete description of this feature requires a
multidevice analysis of high-time resolution data.

The magnetosheath HFAs exhibit a clear tendency to occur
predominantly during periods of enhanced solar wind speed.
This needs more careful examination because high speed streams
have different characteristics than slow speed streams,
% one and probably only a good
perhaps computer simulation can determine which factors are
most important in HFA formation.

\acknowledgments %\footnotesize
{The authors thank the WIND, IMP-8, GEOTAIL, and INTERBALL working
teams for the
magnetic field and plasma data. The  present  work  was  supported by
the Czech  Grant  Agency  under  Contracts  205/99/1712, and
205/00/1686 and by the Charles University Grant Agency under
Contract 181.}

\begin{references}

%\footnotesize
\reference
Klimov, S. {\it et al.}, ASPI experiment: Measurements of fields and
waves onboard the Interball-1 spacecraft, {\it Ann. Geophys.}, 15,
514-527, 1997.

\reference
Kudela, K., M. Slivka, J. Rojko, and V. N. Lutsenko, The
apparatus DOK-2 (project INTERBALL): Output data structure and
modes of operation, preprint of Inst. Exp. Phys. UEF-01-95,
Kosice, 20, 1995.

\reference
Lin Y., Generation of anomalous flows near the bow shock by
its interaction with interplanetary discontinuities, {\it J. Geophys.
Res.}, 102, 24265-24281, 1997.

\reference
Lutsenko, V. N., J. Rojko, K. Kudela, T. V. Gretchko, J. Balaz,
J. Matisin, E. T. Sarris, K. Kalaitzides, and N. Paschalidis,
Energetic particle experiment DOK-2 (Interball project), in:
{it INTERBALL Mission and Payload,} ed. by Yu. Galperin, IKI-CNES,
249, 1995.

%\reference
%Nemecek, Z., J. Safrankova, L. Prech, S. Kokubun, T. Mukai,
%and D. G. Sibeck, Transient flux events in the magnetosheath,
%{\it Geophys. Res. Lett.}, 25, 1273-1276, 1998.

\reference
Onsager, T. G., M. F. Thomsen, J. G. Gosling, and S. J.
Bame, Observational test of a hot flow anomaly formation
mechanism, {\it J. Geohys. Res.}, 95, 11.967-11.974, 1990a.

\reference
Onsager, T. G., M. F. Thomsen, and D. Winske,
Hot flow anomaly formation by magnetic deflection,
{\it Geophys. Res. Lett.}, 17, 1621-1624, 1990b.

\reference
Schwartz, S. J., G. E. Paschmann, N. Sckopke, T. M. Bauer, M. W.
Dunlop, A. N. Fazakerley, and M. F. Thomsen,  Hot flow anomalies
revisited, {\it Int. J. Geomag. and Aeronomy}, 1, 1999, in print.

\end{references}

\end{article}

\end{document}

### 2014-04-29 tex/wds14/wds14.tex
wds14 zaklad

--- 2014-01-26 tex/gecco14/gecco14.tex
@@
-\documentclass{sig-alternate}
+%to be compiled with LATEX 2e
 
-\usepackage[utf8]{inputenc}
+%use this setting with 11pt basic font:
+\documentclass[11pt]{article}
+\usepackage{wds11,epsf}
+
+%use this setting with 10pt basic font:
+%\documentclass{article}
+%\usepackage{wds10,epsf}
+
+%use to get author-year citations with BibTeX
+\usepackage[square]{natbib}
+
 \usepackage{xspace}
-\usepackage[ampersand]{easylist}
-
-\newtheorem{definition}{Definition}
-\newtheorem{proposition}{Proposition}
-
-\newenvironment{lizt}
-{\begin{easylist}[itemize]}
-{\end{easylist}}
 
@@
 \newcommand{\lheads}{$\lambda$-heads\xspace}
-\newcommand{\la}{\leftarrow\xspace}
+\newcommand{\myla}{\leftarrow\xspace}
 \newcommand{\Lp}  {\Lambda^{\prime}\xspace}
@@
 \newcommand{\Real}{\mathbb{R}}
-\newcommand{\gar}{\longmapsto}
 
@@
 
+%\tighten
+
+% Preamble Information
+
+%%%%%%%%%%% nevim co to tu bylo za omg, tak jsem to zakomentoval
+%%\lefthead{SAFRANKOVA ET AL.}
+%%\righthead{MAGNETOSHEATH RESPONSE}
+
+\setcounter{secnumdepth}{0}
+
 \begin{document}
 
-\conferenceinfo{GECCO'14,} {July 12-16, 2014, Vancouver, BC, Canada.}
-\CopyrightYear{2014}
-\crdata{TBA}
-\clubpenalty=10000
-\widowpenalty = 10000
+\title{Typed Genetic Programming over Lambda Calculus}
 
-\title{Utilization of Reductions and Abstraction Elimination in Typed Genetic Programming}
+\author{T. K\v ren}
 
-\numberofauthors{2}
+\affil{Charles University, Faculty  of  Mathematics  and  Physics,
+     Prague, Czech Republic.}
 
-\author{
-% 1st. author
-\alignauthor
-Tom\'{a}\v{s} K\v{r}en\\
-  \affaddr{Faculty of Mathematics and Physics}\\
-  \affaddr{Charles University in Prague}\\
-  \affaddr{Malostransk\'e n\'am\v{e}st\'\i~25, 11000,}\\
-  \affaddr{Prague, Czech Republic}\\
-  \email{tomkren@gmail.com}
-% 2nd. author
-\alignauthor
-Roman Neruda\\
-  \affaddr{Institute of Computer Science}\\
-  \affaddr{Academy of Sciences of the Czech Republic}\\
-  \affaddr{Pod Vod\'arenskou v\v{e}\v{z}\'\i~2, 18207,}\\
-  \affaddr{Prague, Czech Republic}\\
-  \email{roman@cs.cas.cz}
-}
-
-%\date{24 January 2014}
-
-\maketitle
 \begin{abstract}
-Lambda calculus representation of programs offers a more expressive alternative to traditional S-expressions.
-In this paper we discuss advantages of this representation coming from use of reductions (beta and eta) and how to overcome disadvantages caused by variables occurring in the programs by use of the abstraction elimination algorithm.
-We discuss the role of those reductions in the process of generating initial population and compare several crossover approaches including novel approach to crossover operator based both on reductions and abstraction elimination.
-The design goal of this operator is to turn the disadvantage of abstraction elimination  - possibly quadratic increase of program size - into a virtue; our approach leads to more crossover points.
-At the same time, utilization of reductions provides offspring of small sizes.
+Todo.
 \end{abstract}
 
-% A category with the (minimum) three required fields
-\category{H.4}{TODO}{Miscellaneous}
-%A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{TODO}
-
-\keywords{TODO, ToDo, todo}
-
-\section{Plán boje}
-
-\subsection{Jak to celý pojmout?}
-
-Pokud chceme dělat GP nad stromama větší vyjadřovací síly než maj klasický S-výrazy, tak máme v záse dvě přirozený možnosti pro obecný řešení: buď to dělat celý v (polymorfních) kombinátorech od začátku (a vyhnout se tak proměnejm a lambda abstrakcím), nebo to skusit v lambda kalkulu a nejak se vypořádat s proměnejma a lambdama.
-Motivace pro práci s lambda termama je, že za nima košatá teorie, která například popisuje redukce lambda termů (ty zajišťujou jak zmenšení samotných stromu, tak to, že prohledávací prostor se zmenší, díky tomu, že se různé stromy redukují na ten samej).
-
-Jak se vypořádat s proměnejma a lambdama? (aka jak křížit)
-
-\begin{itemize}
-\item Po vy generování převod do kombinátorů pomocí eliminace abstrakcí (tak jak se to dělá v diplomce).
-To ale působí v něčem neohrabaně, když to porovnáme s generovánim přímo v kombinatorech, když vezmeme v potaz to že eliminace abstrakcí má za důsledek až kvadratickej nárůst stromu - čili to co sme nahnali na redukcích stratíme eliminací.
-
-\item Jedince držíme jako redukovane malinké-kompaktní-a-elegantní lambda termy.
-Ve chvíly kdy křížíme provedeme na obou rodičích eliminaci abstrakcí (která ubere proměny a lambdy a namísto toho tam dá kombinátory S,K,I případně i další při fikanějších eliminacích), tím nám sice narostou, ale my z toho máme jedine radost, protože tím se nám zvýšil počet míst ke křížení (což se ukazuje jako dobráv věc, viz s-expr reprezentace vs @-tree reprezentace lambda termů).
-Po skřížení se vložený kombinátory nahradí odpovídajícim lambda termem (tzn např všude kde je K dám (\ x y .
-x) atd) výslednej term zredukuju a dostávam zase malinké-kompaktní-a-elegantní dítě.
-Nevýhoda toho zahrnout do článku i tohle je v tom, že k tomu nemám ještě žádný pokusy - ale k tomu zbytku mám upřímě v zato taky dost ubohý pokysy, takže toho bych se asi nebál.
-Většinu potřebnýho kodu bych k tomu ale měl už mít víceméně hotovou, takže pokud by se to na něčem nezaseklo, tak myslim že je realný udělat i pokus do toho dvacátýho.
-Zvlášť přitažlivý mi tohle křížení příde i kvůli tomu, že v přírodě se taky rozbalujou a zabaloujou chromozomi při meioze/mitoze.
-\end{itemize}
-
-\subsection{Osnova}
-
-\begin{lizt}
-& Jak reprezentovat stromy programu pro GP?
-  && Reprezentace v klasickym kozovy je S-expression.
-  && Nebo mužem používat kombinátory jako 
-     Briggs a O’Neil (to si myslim je jakoby 
-     hlavní konkurence, vuči který by to chtělo obhájit)
-  && Lambda termy a jejich awesomeness
-& \textbf{Povídaní o redukcích}
-& Povídání o lnf
-  && že lnf je přirozený rozšíření s-exprešnu do lambda 
-     kalkulu vlastnosti termu v lnf 
-     &&& proč je eta redukovat (eta redukcí lnf dostanem 
-         beta-eta-nf a že nemusíme beta redukovat pač se to tim nerozbyje)
-& Generování
-  && Generování gramatikou
-  && Gramatika pro lnf
-  && (?) Inhabitation trees jako intuitivní model takovýhodle lnf generování
-& Problémy s proměnýma a jak je řešit.
-  && Křížit lambda termy i s proměnejma a abstrakcema, 
-     problémy řešit když nastanou (pomluvit a odsoudit)
-  && Zmenšit prostor termu (tim že někerý nejsme schopný 
-     vygenerovat) s kterým operujeme tak aby křížení už 
-     nebyl problém (to dělá Yu - v těle lambda fce dovoluje 
-     jen použití proměnných z její hlavy)
-  && Převod hned po vygenerování
-  && Převod až při křížení
-     &&& eliminace abstrakcí
-     &&& skřížim
-     &&& vložený kombinátory nahradim odpovídajícím termem
-     &&& celý to redukuju
- & Pokusy (?!)
- & Závěr
-\end{lizt}
+\begin{article}
 
 \section{Introduction}
-
-Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\cite{koza92,koza03}.
-Early attempts to enhance the GP approach with the concept of types include the seminal work~\cite{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
-Use of types naturally opens door to enriching S-expressions,
-the traditional GP representation of individuals, with concepts from
-lambda calculus, which is simple yet powerful functional mathematical and programming 
-language extensively used in type theory. Such attempts has shown to be 
-successful \cite{yu01}.
-
-The key issue in the lambda calculus approach to enrich GP with types is the method of individual generation.
-During the expansion phase the set of unfinished terms can be browsed with respect to various search strategies.
-Our approach to this problem aims to utilize the full arsenal given by the simply typed lambda calculus.
-Thus, the natural idea is to employ an exhaustive systematic search.
-On the other hand, if we were to mimic the standard GP approach, a quite arbitrary yet common and successful ramped half-and-half generating heuristic~\cite{fg} should probably be used.
-These two search methods in fact represent boundaries between which we will try to position our parameterized solution that allows us to take advantage of both strategies.
-This design goal also differentiate our approach from
-the three state of the art proposals for typed GP known to us that are discussed in the following section. 
-Our proposed \emph{geometrical search strategy} described in this paper is such a successful hybrid mixture of random and systematic exhaustive search.
-Experiments show that it is also very efficient dealing with one of the traditional GP scarecrows - the bloat problem.
-
-The rest of the paper is organized as follows: The next section briefly discusses related work in the field of typed GP, while section~\ref{preliminaries} introduces necessary notions.
-Main original results about search strategies in individual generating are described in section~\ref{approach}.
-Section~\ref{experiments} presents results of our method on three well-known tasks, and the paper is concluded by section~\ref{conclusions}.
 
@@
 
-Yu presents a GP system utilizing
+\cite{yu01} presents a GP system utilizing
 polymorphic higher-order functions\footnote{Higher-order 
-function is a function taking another function as 
-input parameter.} and lambda abstractions  \cite{yu01}.
-Important point of interest in this work is use of
-\texttt{foldr} function as a tool for \textit{implicit recursion},
+function takes another function as an input parameter.
+} and lambda abstractions.
+Important point of interest in this work is use of \texttt{foldr}\footnote{ In the functional programming language Haskell \texttt{foldr} can be defined as:\\ \texttt{foldr f z [] $~~~$  = z\\ 
+foldr f z (x:xs) = f x (foldr f z xs) }} function as a tool for \textit{implicit recursion},
 i.e. recursion without explicit recursive calls. 
@@
 
-Briggs and O’Neill present technique 
-utilizing typed GP with combinators \cite{kes}.
+\cite{kes} present technique 
+utilizing typed GP with combinators.
 The difference between approach presented in this work
@@
 
-Binard and Felty use even 
-stronger type system (\textit{System F}) \cite{binard2008genetic}.  
+\cite{binard2008genetic} use even 
+stronger type system (\textit{System F}).  
 But with increasing power of the type system comes increasing difficulty of term generation.
@@
 
-\section{Preliminaries}
-\label{preliminaries}
+%%% Abych věděl jak citovat tak sem si tu kousek nechal
+%The interaction of the solar wind with the Earth's magnetosphere
+%generates a population of backstreaming ions directed from the
+%bow shock into the solar wind. This high-energy population was
+%invoked to explain Hot Flow Anomalies (HFAs)
+%\citep[e.g.][]{sw}. \citet{th} identified as heated regions of solar wind plasma
+%flowing nearly perpendicular to the Earth-Sun line.
+%The main observational features of HFAs include \citep[and others]{sw}: (1) Central regions with hot plasma flowing
+%significantly slower than that in the ambient solar wind in a
+%direction highly deflected (nearly $90^o$) from the Sun-Earth line.
+%The flow velocities are often roughly tangential to the nominal
+%bow shock shape \citep{sw}.
+%(2) HFAs are bounded by regions of enhanced
+%magnetic field strength, density, and temperature. The outer
+%edges of these enhancements are fast shocks generated by
+%pressure enhancements within the core region. The inner
+%edges of the enhancements are probably  tangential
+%discontinuities.
+%Published examples indicated that  many HFAs are bounded by only
+%one enhancement.
+%(3) HFAs occur in conjunction with significant changes in the IMF
+%direction. The angle between pre- and post event orientations is
+%typically $\sim 70^o$.
 
-In this section, several notions necessary to build a typed GP based on lambda calculus are introduced. 
-First, \lets describe a programming language, 
-in which the GP algorithm generates individual programs --- the so called \lterms.
+Todo.
 
-\begin{definition}
-Let $V$ be infinite countable set of {\it 
-variable names}. Let $C$ be set of {\it constant names}, 
-$V \cap C = \emptyset$.	 	
-Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
-\begin{align*}
-x   \in V \cup C  &\then x     \in \Lambda \\
-M,N \in \Lambda   &\then (M~N) \in \Lambda 
-\textit{~~~~~~(Function application)} \\
-x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
-\textit{~~~~($\lambda$-abstraction)} 
-\end{align*}
-\end{definition}
+\section{Conclusion}
+Todo.
 
-\textit{Function application} and 
-\textit{$\lambda$-abstraction} are concepts
-well known from common programming languages. 
-For example in JavaScript 
-$(M~N)$ translates to expression \texttt{$M$($N$)} and
-$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
-In other words, the function application 
-corresponds to the act of supplying a function 
-with an argument, and
-the $\lambda$-abstraction is equivalent to 
-\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
-they were recently introduced to C++, and they are expected to be supported in Java 8.}.
+\acknowledgments %\footnotesize
+{???}
 
-For better readability, 
-$M_1~M_2~M_3~\dots~M_n$ is an abbreviation for
-$(\dots((M_1~M_2)~M_3)~\dots~M_n)$
-and $\lam{x_1 x_2 \dots x_n }{M}$ for 
-$\lamb{x_1}{\lamb{x_2}{\dots\lamb{x_n}{M}\dots}}$.
+%% TWO METHODS FOR INCLUDING THE BIBLIOGRAPHY (LIST OF REFERENCES)
+%% EITHER TYPE IN THE ENTRIES YOURSELF AS SHOWN HERE IN
+%% `thebibliography' ENVIRONMENT,
+%%        OR
+%% USE THE FOLLOWING TWO COMMANDS SO THAT BIBTEX WILL GENERATE
+%% `thebibliography' TEXT FOR YOU AND READ IT IN.
+%%
+\bibliographystyle{egs}%<-- LIST OF REFERENCES TO BE IN "EGS" STYLE (FILE "EGS.STY")
+\bibliography{my}       %<-- REFERENCES ARE IN FILE "SAMPLE.BIB"
+%%
+%%  IF THE ABOVE TWO COMMANDS ARE USED, THEN thebibliography ENVIRONMENT
+%%  MUST BE REMOVED.
 
-\subsection{$\beta$-reduction}
+%\begin{thebibliography}{}
+%
+%\itemsep=0.4ex
+%%\footnotesize
+%\bibitem[\it Klimov and Friend(1997)]{kl}
+%Klimov, S., Friend A., ASPI experiment: Measurements of fields and
+%waves onboard the Interball-1 spacecraft, {\it Ann. Geophys.}, 15,
+%514-527, 1997.
+%
+%\bibitem[\it Kudela et al.(1995)]{ku}
+%Kudela, K., M. Slivka, J. Rojko, and V. N. Lutsenko, The
+%apparatus DOK-2 (project INTERBALL): Output data structure and
+%modes of operation, preprint of Inst. Exp. Phys. UEF-01-95,
+%Kosice, 20, 1995.
+%
+%\bibitem[\it Lin(1997)]{li}
+%Lin Y., Generation of anomalous flows near the bow shock by
+%its interaction with interplanetary discontinuities, {\it J. Geophys.
+%Res.}, 102, 24265-24281, 1997.
+%
+%\bibitem[\it Lutsenko et al.(1995)]{lu}
+%Lutsenko, V. N., J. Rojko, K. Kudela, T. V. Gretchko, J. Balaz,
+%J. Matisin, E. T. Sarris, K. Kalaitzides, and N. Paschalidis,
+%Energetic particle experiment DOK-2 (Interball project), in:
+%{it INTERBALL Mission and Payload,} ed. by Yu. Galperin, IKI-CNES,
+%249, 1995.
+%
+%
+%\bibitem[\it Onsager et al.(1990a)]{ona}
+%Onsager, T. G., M. F. Thomsen, J. G. Gosling, and S. J.
+%Bame, Observational test of a hot flow anomaly formation
+%mechanism, {\it J. Geophys. Res.}, 95, 11.967-11.974, 1990a.
+%
+%\bibitem[\it Onsager et al.(1990b)]{onb}
+%Onsager, T. G., M. F. Thomsen, and D. Winske,
+%Hot flow anomaly formation by magnetic deflection,
+%{\it Geophys. Res. Lett.}, 17, 1621-1624, 1990b.
+%
+%\bibitem[\it Schwartz et al.(1999d)]{sw}
+%Schwartz, S. J., G. E. Paschmann, N. Sckopke, T. M. Bauer, M. W.
+%Dunlop, A. N. Fazakerley, and M. F. Thomsen,  Hot flow anomalies
+%revisited, {\it Int. J. Geomag. and Aeronomy}, 1, 1999, in print.
+%
+%\bibitem[\it Thomsen et al.(1992)]{th}
+%Thomsen M. F., J. G. Gosling, and S. J.
+%Bame, Observational test of a hot flow anomaly formation
+%mechanism, {\it J. Geophys. Res.}, 97, 1967-1974, 1992.
+%
+%
+%\end{thebibliography}
 
-In order to perform computation there must be some
-mechanism for term evaluation. In $\lambda$-calculus there
-is $\beta$-reduction for this reason.\\
-
-\newcommand{\bRedex}{$\beta$-redex\xspace}
-\newcommand{\bRedexes}{$\beta$-redexes\xspace}
-\newcommand{\bArrow}{\rightarrow_\beta\xspace}
-\newcommand{\eArrow}{\rightarrow_\eta\xspace}
-\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
-
-A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
-A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
-This fact is written as \textit{relation} $\bArrow$ 
-of those two terms:
-\begin{equation} \label{eq:bRed}
-\lamb{x}{M}N \bArrow M[x:=N]
-\end{equation}
-It is also possible to reduce \textit{subterm \bRedexes} 
-which can be formally stated as:
-\begin{align*}
-P \bArrow Q &\then (R~P)      \bArrow (R~Q) \\
-P \bArrow Q &\then (P~R)      \bArrow (Q~R) \\
-P \bArrow Q &\then \lam{x}{P} \bArrow \lam{x}{Q}  
-\end{align*}
-
-In other words, $\beta$-reduction is the process 
-of insertion of arguments supplied to a function into 
-its body. \\
-
-Another useful relations are $\bbarr$ and $\beq$ defined as follows.
-
-\begin{enumerate}
- \item \begin{enumerate}
- 	\item $M \bbarr M$
- 	\item $M \barr N \then M \bbarr N$
- 	\item $M \bbarr N , N \bbarr L \then M \bbarr L$ 	
- \end{enumerate}
- \item \begin{enumerate}
- 	\item $M \bbarr N \then M \beq N$
- 	\item $M \beq N \then N \beq M$
- 	\item $M \beq N , N \beq L \then M \beq L$
- \end{enumerate}
-
-\end{enumerate}
-
-We read those relations as follows.
-\begin{enumerate}
- 	\item $M \bbarr N$ --- "$M$ $\beta$-reduces to $N$."  
- 	\item $M \barr N$  --- "$M$ $\beta$-reduces to $N$
- 	      in one step."
- 	\item $M \beq N$ --- "$M$ is $\beta$-convertible to $N$."	
- \end{enumerate}
-
-\subsection{$\eta$-reduction}
-
-Similarly as for $\beta$-reduction we can define $\eta$-reduction 
-except that instead of \ref{eq:bRed} we use:
-
-$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$~
-
-Analogically, a term of a form $\lamb{x}{(M~x)}$ is called 
-\textit{$\eta$-redex}.
-
-Relation $\bearr\;=\;\barr \cup \earr$. 
-(Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.)
-
-Similarly as for $\bbarr$ and $\beq$ we can define relations 
-$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.
-
-\subsection{$\eta^{-1}$-reduction}
-
-$\eta^{-1}$-reduction (also called $\eta$-expansion) is 
-the reduction converse to $\eta$-reduction.
-Again it may be obtained by replacing \ref{eq:bRed}, now with:
-
-$$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$
-
-\subsection{Normal forms}
-
-~\begin{easylist}[enumerate]
-& A \lterm is a \textit{$\beta$-normal form} (\bnf) 
-if it does not have a $\beta$-redex as subterm.
-& A \lterm M \textit{has} a \bnf if $M \beq N$
-and $N$ is a \bnf.\\
-\end{easylist}
-A normal form may be thought of as a result of a term evaluation.
-
-Similarly we can define \enf and \benf.
-
-\subsection{Types etc}
-
-A \lterm as described above
-corresponds to a program expression with no type information
-included. Now we will describe \textit{types} (or \textit{type terms}).
-
-\begin{definition}
-Let $A$ be set of {\it atomic type names}. 
-Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
-\begin{align*}
-\alpha      \in A  &\then   \alpha \in \T \\
-\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
-\end{align*}
-\end{definition}
-
-Type $\sigma \ar \tau$ is type for functions taking as input
-something of a type $\sigma$ and returning 
-as output something of a type $\tau$. 
-$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
-$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
-The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
-combining the previously defined \textit{\lterms} and \textit{types} together.
-
-\begin{definition}~
-
-\begin{enumerate}
- \item 	Let $\Lambda$ be set of {\it \lterms}. 
-	Let $\mathbb{T}$ be set of {\it types}.       
-	A {\it statement} $M : \sigma$ is a pair 
-	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
-	Statement $M : \sigma$ is vocalized as 
-	{\it "$M$ has type $\sigma$"}.
-	The term $M$ is called the {\it subject} of the 
-	statement $M : \sigma$.
- \item A \textit{declaration} is a statement 
- $x : \sigma$ where $x \in V \cup C$.
-
- \item A \textit{context} 
- is set of declarations with distinct variables as subjects.
-\end{enumerate}
-\end{definition}
-%~\\
-
-Context is a basic type theoretic concept suitable as a typed alternative
-for terminal and function set in standard GP. 
-Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
-such that $\Gamma$ does not contain any declaration with $x$ as subject.
-We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
-
-\begin{definition}
-A statement $M\colon\sigma$ is \textit{derivable from}
-a context $\Gamma$ (notation 
-\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
-if it can be produced by the following rules.
-\begin{align*}
-x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
-\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
-&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
-\tur{\Gamma,x:\sigma}{M}{\tau}
-&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
-\end{align*}
-\end{definition}
-
-Přeformulovat aby se to hodilo do tohodle kontextu....
-Our goal in term generation is to produce terms $M$
-for a given pair $\ul{\tau}{\Gamma}$
-such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
-
-\subsection{Long normal form}
-\label{lnf}
-
-%v barendrechtovi je to v definici lnf
-%$\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo $M_m$ 
-%ale to musí bejt určitě překlep...
-
-\begin{definition}
-Let \GMS where 
-$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha, n \geq 0$.
-	\begin{enumerate}
-	  \item	
-		Then $M$ is in \textit{long normal form} (\lnf) if following 
-		conditions are satisfied.
-		\begin{enumerate}
-		 \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
-		  (specially for $n = 0$, $M$ is term of the form $f$).
-		 \item Each $M_i$ is in \lnf.
-		\end{enumerate}	
-	  \item 
-	    $M$ has a \lnf if $M =_{\beta\eta} N$ and $N$ is in \lnf.
-	\end{enumerate}
-\end{definition}~
-
-As is shown in \cite{barendregt10}, \lnf has following nice properties.
-
-\begin{proposition}
-If $M$ has a \bnf, 
-%which according to Theorem 2B.4 is always the case, 
-then it also has a unique \lnf, 
-which is also its unique \beenf.
-\end{proposition}
-
-\begin{proposition}
-Every $B$ in \bnf has a \lnf 
-$L$ such that $L \twoheadrightarrow_{\eta} B$.
-\end{proposition}
-
-\subsection{Grammar producing \lterms in \lnf}
-
-In \cite{barendregt10} is shown term generating grammar with 
-following rules (our notation is used, but we will not 
-highlighted terminals anymore).
-
-%\footnote{
-%I was using term generating grammars for term generation
-%on my own before I encountered with \cite{barendregt10},
-%where i happily discovered that Barendregt is using
-%almost identical notation. 
-%But his \lnf grammar was more clever then my grammar
-%that was using grammar rules corresponding to the three basic
-%inference rules. 
-%}
-
-\begin{align*}
-( \alpha , \Gamma )  
-&\gar
-(~f~( \rho_1 , \Gamma )~\dots~( \rho_m , \Gamma )~)
-& \textbf{if } \alpha \in A,
-(f : \rho_1 \ar \dots \ar \rho_m \ar \alpha) \in \Gamma
-\\ 
-( \sigma \rightarrow \tau , \Gamma )  
-&\gar
-(~\lambda~x~.~( \tau ; \Gamma,x:\sigma )~)
-&   
-\end{align*}
-
-The second rule can be replaced by more effective one.
-\[ 
-( \tau_1 \ar \dots \ar \tau_n \ar \alpha , \Gamma )  
-\gar
-(~\lambda~x_1~\dots~x_n~.~
-( \alpha ; \Gamma , x_1:\tau_1 , \dots , x_n:\tau_n  )~)
-~~~~ \textbf{if } n > 0
-\]
-
-This rule packs consecutive uses of the second rule into one use.
-This is valid since the use of the second rule is deterministic;
-it is used if and only if the non-terminal's type is not atomic.
-
-\section{Our approach}
-\label{approach}
-
-\section{Experiments}
-\label{experiments}
-
-\section{Conclusions}
-\label{conclusions}
-
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance and
-what have you.  In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
-
-%
-% The following two commands are all you need in the
-% initial runs of your .tex file to
-% produce the bibliography for the citations in your paper.
-\bibliographystyle{abbrv}
-\bibliography{evogp}  % sigproc.bib is the name of the Bibliography in this case
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-
-% That's all folks!
+\end{article}
 \end{document}

### 2014-05-11 tex/gecco14_oprava/ACM example/sig-alternate.tex
gecco nepouzita oprava

% This is "sig-alternate.tex" V2.0 May 2012
% This file should be compiled with V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls'
% V2.5 LaTeX2e document class file. It is for those submitting
% articles to ACM Conference Proceedings WHO DO NOT WISH TO
% STRICTLY ADHERE TO THE SIGS (PUBS-BOARD-ENDORSED) STYLE.
% The 'sig-alternate.cls' file will produce a similar-looking,
% albeit, 'tighter' paper resulting in, invariably, fewer pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which
% DOES NOT produce 1) thru' 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within
% the source .tex file, over both the CopyrightYear
% (defaulted to 200X) and the ACM Copyright Data
% (defaulted to X-XXXXX-XX-X/XX/XX).
% e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use
% the .bib file (from which the .bbl file % is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission, you *NEED* to 'insert'
% your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% ================= IF YOU HAVE QUESTIONS =======================
% Questions regarding the SIGS styles, SIGS policies and
% procedures, Conferences etc. should be sent to
% Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to
% Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate}

\begin{document}
%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Produces the permission block, and
copyright information). For use with
SIG-ALTERNATE.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{8} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
\alignauthor
G.K.M. Tobin\titlenote{The secretary disavows
any knowledge of this author's actions.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{P.O. Box 1212}\\
       \affaddr{Dublin, Ohio 43017-6221}\\
       \email{webmaster@marysville-ohio.com}
% 3rd. author
\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
one who did all the really hard work.}\\
       \affaddr{The Th{\o}rv{\"a}ld Group}\\
       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       \affaddr{Hekla, Iceland}\\
       \email{larst@affiliation.org}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Lawrence P. Leipuner\\
       \affaddr{Brookhaven Laboratories}\\
       \affaddr{Brookhaven National Lab}\\
       \affaddr{P.O. Box 5000}\\
       \email{lleipuner@researchlabs.org}
% 5th. author
\alignauthor Sean Fogarty\\
       \affaddr{NASA Ames Research Center}\\
       \affaddr{Moffett Field}\\
       \affaddr{California 94035}\\
       \email{fogartys@amesres.org}
% 6th. author
\alignauthor Charles Palmer\\
       \affaddr{Palmer Research Laboratories}\\
       \affaddr{8600 Datapoint Drive}\\
       \affaddr{San Antonio, Texas 78229}\\
       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
TThis paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings. It is an {\em alternate} style which produces
a {\em tighter-looking} paper and was designed in response to
concerns expressed, by authors, over page-budgets.
It complements the document \textit{Author's (Alternate) Guide to
Preparing ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}.
This source file has been written with the intention of being
compiled under \LaTeX$2_\epsilon$\ and BibTeX.

The developers have tried to include every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.

To make best use of this sample document, run it through \LaTeX\
and BibTeX, and compare this source code with the printed
output produced by the dvi file. A compiled PDF version
is available on the web page to help you with the
`look and feel'.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference.
ACM seeks to give these conference by-products a uniform,
high-quality appearance.  To do this, ACM has some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (1.9 cm [0.75"]) top, (2.54 cm [1"]) bottom
and (1.9 cm [.75"]) left and right; specified column width
(8.45 cm [3.33"]) and gutter size (.83 cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{Two of these, the {\texttt{\char'134 numberofauthors}}
and {\texttt{\char'134 alignauthor}} commands, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands.

\section{The {\secit Body} of The Paper}
Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}
We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}\lim_{n\rightarrow \infty}x=0\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles \cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings \cite{clark:pct} or
books \cite{salas:calculus, Lamport:LaTeX} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}\cite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of \textbf{.eps}
and \textbf{.ps} files to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\epsfig{file=fly.eps}
\caption{A sample black and white graphic (.eps format).}
\end{figure}

\begin{figure}
\centering
\epsfig{file=fly.eps, height=1in, width=1in}
\caption{A sample black and white graphic (.eps format)
that has been resized with the \texttt{epsfig} command.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
and don't forget to end the environment with
{figure*}, not {figure}!

\begin{figure*}
\centering
\epsfig{file=flies.eps}
\caption{A sample black and white graphic (.eps format)
that needs to span two columns of text.}
\end{figure*}

Note that either {\textbf{.ps}} or {\textbf{.eps}} formats are
used; use
the \texttt{{\char'134}epsfig} or \texttt{{\char'134}psfig}
commands as appropriate for the different file types.

\begin{figure}
\centering
\psfig{file=rosette.ps, height=1in, width=1in,}
\caption{A sample black and white graphic (.ps format) that has
been resized with the \texttt{psfig} command.}
\vskip -6pt
\end{figure}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.  In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e. the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{Acknowledgments}
\subsection{Additional Authors}
This section is inserted by \LaTeX; you do not insert it.
You just add the names and information in the
\texttt{{\char'134}additionalauthors} command at the start
of the document.
\subsection{References}
Generated by bibtex from your ~.bib file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the ~.bbl file.  Insert that ~.bbl file into
the .tex source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}
The sig-alternate.cls file itself is chock-full of succinct
and helpful comments.  If you consider yourself a moderately
experienced to expert user of \LaTeX, you may find reading
it useful but please remember not to change it.
%\balancecolumns % GM June 2007
% That's all folks!
\end{document}

### 2014-05-14 tex/wds14/wds14.tex
wds14 pokracovani

@@
 %to be compiled with LATEX 2e
-
 %use this setting with 11pt basic font:
 \documentclass[11pt]{article}
+
+\usepackage[utf8]{inputenc}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\usepackage{xspace}
+\usepackage[ampersand]{easylist}
+\usepackage{qtree}
+\usepackage{color}
+
+%
+
 \usepackage{wds11,epsf}
-
 %use this setting with 10pt basic font:
@@
 %\usepackage{wds10,epsf}
-
 %use to get author-year citations with BibTeX
@@
 
-\usepackage{xspace}
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
 
@@
 \newcommand{\lnf}{\textit{lnf}\xspace}
-\newcommand{\Ae}{\mathop{\mathrm{\AE}}}
+\newcommand{\Ae}{\mathop{\mathrm{AE}}} %{\mathop{\mathrm{\AE}}}
 \newcommand{\Bcomb }{\mathbf{B}}   
@@
 \newcommand{\Real}{\mathbb{R}}
+
+\newcommand{\gar}{\longmapsto}
+\newcommand{\bRedex}{$\beta$-redex\xspace}
+\newcommand{\bRedexes}{$\beta$-redexes\xspace}
+\newcommand{\bArrow}{\rightarrow_\beta\xspace}
+\newcommand{\eArrow}{\rightarrow_\eta\xspace}
+\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
 
@@
 
-\title{Typed Genetic Programming over Lambda Calculus}
+\setlength{\pdfpagewidth}{210mm}
+\setlength{\pdfpageheight}{297mm}
+
+\title{Typed Genetic Programming in Lambda Calculus}
 
@@
 \begin{abstract}
-Todo.
+\red{TODO}
 \end{abstract}
@@
 \section{Introduction}
+
+Genetic programming (GP) is an AI technique, which falls into broader category of evolutionary algorithms  ---  mataheuristic  search algorithms inspired  by biological evolution by natural selection.
+GP was developed by John Koza in 1992 \cite{koza92}.
+I dare say many people perceive its beauty in the fact that GP is a computer program  constructing other computer programs with desired properties by breeding them.
+It is also pretty successful technique in a number of areas \citep{koza2003genetic}.
+And perhaps that is why GP has become very popular.
+
+Early attempts to enhance the GP approach with the concept of types include the seminal work~\citep{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
+Usage of types naturally opens door to enriching S-expressions,
+the traditional GP representation of individuals, with concepts from
+lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory~\citep{yu01}.
+
+One of the motivations for using lambda calculus is that it is backed by a considerable theoretical background.
+From this background we can utilize several useful concepts such as reductions, normalization or abstraction elimination, and use them to enhance the standard GP algorithm.
+
+\red{(- ještě něco k tomu že LC je cool pro TGP (taky z článků) ?)}
+
+\red{- že důležitý je navrhnout generování a gene ops. že gener de přímoč do mutace.}
+
+The rest of the paper is organized as follows: \red{TODO}
+
+\section{Preliminaries}
+
+In this section, several notions necessary to build a typed GP based on lambda calculus are introduced.
+
+\subsection{Genetic programming}
+
+A problem to be solved is given to GP in a form of \textit{fitness function}.
+Fitness function is a function which takes computer program as its input and  returns numerical value called \textit{fitness} as output.
+The bigger fitness of a computer program is, the better solution of a problem.
+GP maintains a collection of computer programs called \textit{population}.  A member of population is called \textit{individual}. By running GP algorithm evolution of those individuals is performed.
+
+Individuals are computer program \textit{expressions} kept as \textit{syntactic trees}.
+Basically those trees are rooted trees with a function symbol in each internal node and with constant symbol or variable symbol in each leaf node.
+Number of child nodes for each internal node corresponds to the number of arguments of a function whose symbol is in that node.
+
+Another crucial input besides fitness function is a collection of \textit{building symbols}.
+It is a collection of symbols (accompanied with an information about number of arguments).
+Those symbols are used to construct trees representing individuals.
+
+\Lets describe GP algorithm briefly.
+At the beginning, initial population is generated from building blocks randomly.
+A step of GP algorithm is stochastic transformation of the current population into the next population.
+
+This step consists of two sub steps:
+
+\begin{itemize} 
+\item Selection of \textit{parents} for individuals of the next population based on the fitness.
+The bigger fitness of an individual of the current population is, the better chance of success being selected as parent it has.
+\item Application of genetic operators (such as \textit{crossover},\textit{reproduction} and \textit{mutation}) on parent individuals producing new individuals of the next population.  
+\end{itemize}
+
+This transformation is repeatedly applied for a predefined number of steps (which is called number of \textit{generations}) or until some predefined criterion is met.
+
+\subsection{Lambda calculus}
+
+\Lets describe a programming language, 
+in which the GP algorithm generates individual programs --- the so called \lterms.
+
+\begin{definition}
+Let $V$ be infinite countable set of {\it 
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	 	
+Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.	
+\begin{align*}
+x   \in V \cup C  &\then x     \in \Lambda \\
+M,N \in \Lambda   &\then (M~N) \in \Lambda 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
+\textit{~~~~($\lambda$-abstraction)} 
+\end{align*}
+\end{definition}
+
+\textit{Function application} and 
+\textit{$\lambda$-abstraction} are concepts
+well known from common programming languages. 
+For example, in JavaScript, 
+$(M~N)$ translates to expression \texttt{$M$($N$)} and
+$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
+In other words, the function application 
+corresponds to the act of supplying a function 
+with an argument, and
+the $\lambda$-abstraction is equivalent to 
+\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
+they were recently introduced to C++, and they are expected to be supported in Java 8.}.
+To ensure better readability, 
+$M_1~M_2~M_3~\dots~M_n$ will be an abbreviation for 
+$(\dots((M_1~M_2)~M_3)~\dots~M_n)$.
+And $\lam{x_1 \dots x_n }{M}$  for
+$\lamb{x_1}{\dots\lamb{x_n}{M}\dots}$.
+
+\subsection{Reductions}
+
+In order to perform computation, there must be some
+mechanism for term evaluation. In $\lambda$-calculus there
+is a \mbox{$\beta$-reduction} procedure for this reason.
+
+A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
+A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
+This fact is written as \textit{relation} $\bArrow$ 
+of those two terms:
+\begin{equation} \label{eq:bRed}
+\lamb{x}{M}N \bArrow M[x:=N]
+\end{equation}
+It is also possible to reduce \textit{subterm \bRedexes} 
+which can be formally stated as:
+\begin{align*}
+P \bArrow Q &\then (R~P)      \bArrow (R~Q) \\
+P \bArrow Q &\then (P~R)      \bArrow (Q~R) \\
+P \bArrow Q &\then \lam{x}{P} \bArrow \lam{x}{Q}  
+\end{align*}
+
+In other words, $\beta$-reduction is the process 
+of insertion of arguments supplied to a function into 
+its body.
+
+Another useful relations are $\bbarr$ and $\beq$ defined as follows.
+
+\begin{enumerate}
+ \item \begin{enumerate}
+ 	\item $M \bbarr M$
+ 	\item $M \barr N \then M \bbarr N$
+ 	\item $M \bbarr N , N \bbarr L \then M \bbarr L$ 	
+ \end{enumerate}
+ \item \begin{enumerate}
+ 	\item $M \bbarr N \then M \beq N$
+ 	\item $M \beq N \then N \beq M$
+ 	\item $M \beq N , N \beq L \then M \beq L$
+ \end{enumerate}
+
+\end{enumerate}
+
+We read those relations as follows.
+\begin{enumerate}
+ 	\item $M \bbarr N$ --- "$M$ $\beta$-reduces to $N$."  
+ 	\item $M \barr N$  --- "$M$ $\beta$-reduces to $N$
+ 	      in one step."
+ 	\item $M \beq N$ --- "$M$ is $\beta$-convertible to $N$."	
+ \end{enumerate}
+
+Similarly as for $\beta$-reduction we can define $\eta$-reduction except that now it is defined as follows.  
+$$\lamb{x}{(M~x)} \eArrow M \textbf{ ~~~~if } x \not\in FV(M) $$
+
+Analogically, a term of a form $\lamb{x}{(M~x)}$ is called 
+\textit{$\eta$-redex}.
+
+Relation\footnote{Relation $R = \{\;(a,b)\;|\;a\;R\;b\;\}$.} $\bearr\;=\;\barr \cup \earr$.
+Similarly as for $\bbarr$ and $\beq$ we can define relations 
+$\eearr$, $\eeq$,$\bbeearr$ and $\beeq$.
+
+The $\eta^{-1}$-reduction (also called $\eta$-expansion) is 
+the reduction converse to $\eta$-reduction.
+It is defined as follows.  
+$$M \eeArrow \lamb{x}{(M~x)} \textbf{ ~~~~if } x \not\in FV(M) $$
+
+\subsection{Normal forms}
+\begin{definition}~
+
+\begin{easylist}[enumerate]
+& A \lterm is a \textit{$\beta$-normal form} (\bnf) 
+if it does not have a $\beta$-redex as subterm.
+& A \lterm M \textit{has} a \bnf if $M \beq N$
+and $N$ is a \bnf.
+\end{easylist}
+\end{definition}
+
+A normal form may be thought of as a result of a term evaluation. 
+Normalization of the term $M$ is the process of finding normal form of $M$. 
+Similarly we can define \enf and \benf.
+
+\subsection{Types}
+
+A \lterm as described above
+corresponds to a program expression with no type information
+included. Now we will describe \textit{types} (or \textit{type terms}).
+
+\begin{definition}
+Let $A$ be set of {\it atomic type names}. 
+Then $\mathbb{T}$ is set of {\it types} inductively defined as follows.
+\begin{align*}
+\alpha      \in A  &\then   \alpha \in \T \\
+\sigma,\tau \in \T &\then ( \sigma \ar  \tau ) \in \T 
+\end{align*}
+\end{definition}
+
+Type $\sigma \ar \tau$ is type for functions taking as input
+something of a type $\sigma$ and returning 
+as output something of a type $\tau$. 
+$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
+$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$.
+The system called \textit{simply typed $\lambda$-calculus} is now easily obtained by
+combining the previously defined \textit{\lterms} and \textit{types} together.
+
+\begin{definition}~
+
+\begin{enumerate}
+ \item 	Let $\Lambda$ be set of {\it \lterms}. 
+	Let $\mathbb{T}$ be set of {\it types}.       
+	A {\it statement} $M : \sigma$ is a pair 
+	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
+	Statement $M : \sigma$ is vocalized as 
+	{\it "$M$ has type $\sigma$"}.
+	The term $M$ is called the {\it subject} of the 
+	statement $M : \sigma$.
+ \item A \textit{declaration} is a statement 
+ $x : \sigma$ where $x \in V \cup C$.
+
+ \item A \textit{context} 
+ is set of declarations with distinct variables as subjects.
+\end{enumerate}
+\end{definition}
+%~\\
+
+Context is a basic type theoretic concept suitable as a typed alternative
+for terminal, and function set in standard GP. 
+Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
+such that $\Gamma$ does not contain any declaration with $x$ as subject.
+We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
+
+\begin{definition}
+A statement $M\colon\sigma$ is \textit{derivable from}
+context $\Gamma$ (notation 
+\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
+if it can be produced by the following rules.
+\begin{align*}
+x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
+\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
+&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
+\tur{\Gamma,x:\sigma}{M}{\tau}
+&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
+\end{align*}
+\end{definition}
+
+Our goal in term generation is to produce terms $M$
+for a given pair $\ul{\tau}{\Gamma}$
+such that for each $M$ is $\tur{\Gamma}{M}{\tau}$.
+
+\subsection{Long normal form}
+\label{lnf}
+
+%v barendrechtovi je to v definici lnf
+%$\lam{x_1 \dots x_n}{f~M_1~\dots~M_n}$ místo $M_m$ 
+%ale to musí bejt určitě překlep...
+
+\begin{definition}
+Let \GMS where 
+$\sigma = \tau_1 \ar \dots \ar \tau_n \ar \alpha, n \geq 0$.
+	\begin{enumerate}
+	  \item	
+		Then $M$ is in \textit{long normal form} (\lnf) if following 
+		conditions are satisfied.
+		\begin{enumerate}
+		 \item $M$ is term of the form $\lam{x_1 \dots x_n}{f~M_1~\dots~M_m}$\\
+		  (specially for $n = 0$, $M$ is term of the form $f$).
+		 \item Each $M_i$ is in \lnf.
+		\end{enumerate}	
+	  \item 
+	    $M$ has a \lnf if $M =_{\beta\eta} N$ and $N$ is in \lnf.
+	\end{enumerate}
+\end{definition}~
+
+As is shown in \cite{barendregt10}, \lnf has following nice properties.
+
+\begin{proposition}
+If $M$ has a \bnf, 
+%which according to Theorem 2B.4 is always the case, 
+then it also has an unique \lnf, 
+which is also its unique \beenf.
+\end{proposition}
+
+\begin{proposition}
+Every $B$ in \bnf has a \lnf 
+$L$ such that $L \twoheadrightarrow_{\eta} B$.
+\end{proposition}
+
+\subsection{Abstraction elimination}
+\textit{Abstraction elimination} is a process of transforming 
+an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables.
+The newly produced \lterm may contain function applications, 
+free symbols from former \lterm and some new symbols standing for 
+combinators $\Scomb$, $\Kcomb$ and $\Icomb$.
+
+Those combinators are defined as:
+\begin{align*}
+\Scomb &= \lam{f\,g\,x}{f\,x\,(g\,x)} \\
+\Kcomb &= \lam{x\,y}{x} \\
+\Icomb &= \lam{x}{x} 
+\end{align*}
+
+\Lets describe transformation $\Ae$ performing this 
+process.
+%\footnote{In our implementation we must also deal with types, because our individuals are annotated with types.
+But since this process is fairly straightforward but cumbersome to describe, we will skip explaining it here.
+Reader interested in seeing it can see the source code of our implementation.
+}
+\begin{align*}
+\Ae[x]           &= x\\[0.4em]
+\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N])\\[0.4em]
+\Ae[\lam{x}{x}]  &= \Icomb\\
+\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) \textbf{ if } x \not\in \FV(M)\\
+\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  \textbf{ if } x \in \FV(M)\\
+\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])\\
+&\textbf{ if } x \in \FV(M) \vee~ x \in \FV(N)\\
+\end{align*}
+
+As is stated in \citep{jones87},
+the biggest disadvantage of this technique is that the translated
+term is often much larger than in its lambda form --- the size of
+the translated term can be proportional to the square of the size 
+of the original term. But the advantage is also tempting --- no need to deal with variables and lambda abstractions.
+
+The algorithm presented here is a simple version of this process. 
+More optimized version (used in our \textit{hybrid} cross\-over described below), by means of the size of resulting term and its time performance is presented in \citep{jones87}.
+
+\red{- HM - z papírů, wiki}
+
+\red{- Typová unifikace}
+
+\red{- Type classes - z papírů (asi ne tak formální)}
+
+\red{- naznačit celou hierarchii, řict že na vršku je jakoby DTT.}
 
@@
 
-Todo.
+\red{Todo, eště neco?}
+
+\section{Our current approach}
+
+\subsection{Generating}
+
+Our approach to \lterm generating is based on technique 
+briefly described in \citep{barendregt10}, which generates
+well-typed \lterms in their long normal form. 
+We use this technique to perform a systematic exhaustive enumeration
+of \lterms in their long normal form in order from the smallest to the largest.
+We use well known \textit{A* algorithm} \citep{AIMA} for this task.
+A* is used to search in a given state space for a goal state. 
+It finds the optimal solution (in our case the smallest term)
+and uses "advising" heuristic function.
+It maintains a priority queue to organize states yet to be explored.
+Initially this queue contains only the initial state.
+
+Our state space to search in is the space of unfinished \lterms. 
+The initial state is the unfinished term $\ul{\tau}{\Gamma}$, 
+where $\tau$ is the desired type of
+terms to be generated and $\Gamma$ is the context
+representing the set of building symbols to be used in construction of
+terms (it corresponds to the set $T \cup F$ in
+standard GP enriched with types). The process of determining 
+successors of a state described below is designed so it constructs well-typed 
+\lterms and omits no \lterm in its long normal form. 
+A state is considered a goal state if it contains no unfinished
+leaf, i.e., it is a finished \lterm.
+
+Our generating method is based on simple modification of the
+standard A*, which we call \textit{forgetful A*}. This modification consist in 
+additional parameter for the A* algorithm -- the \textit{search strategy}. 
+It is a simple filtration function (along with initialization procedure)
+that is given the set of all successors of the state that is being examined
+and returns a subset of this input. This subset is added to the priority queue 
+to be further explored. In this way the search space may be reduced as 
+the filtration function may \textit{forget} some successors.
+If the queue becomes empty before the desired number of \lterms
+is generated, then the initial state is inserted to the queue
+and the process continues. For the standard A* this would be meaningless,
+but since our A* is forgetful this kind of restart makes sense.
+
+A* keeps a priority queue of states during the generation process,
+on the other hand the \textit{ramped half-and-half method}, 
+the standard GP algorithm for generating individuals, 
+keeps only one individual which is gradually constructed. This 
+behavior is easily achieved by use of suitable search strategy 
+that returns subset consisting of only one successor.
+The systematic search is obtained by search strategy that 
+returns whole input set.      
+Our novel \textit{geometric strategy} can be understood as
+point somewhere between those two extremes.
+
+%.... metaevoluce remark
+
+A remarkable benefit of parameterizing the generating method 
+by a search strategy taking the form of a simple filtration function
+is that such a function can be expressed by functional language in a very 
+economical way. Take for instance a search strategy that behaves same
+as the \textit{geometric} except that when it comes to 
+situation where all the elements are filtered out it
+acts as the \textit{ramped half-and-half}. 
+Such \textit{strategy composition}
+is easily describable as a higher order function. Thus comes 
+to mind the possibility of using our system to come up with new
+search strategies on its own.
+
+\subsection{Crossover}
+
+Our system utilizes generalization of the standard 
+tree-swapping crossover operator. Two main concerns with swapping typed 
+subtrees are types and free variables. Well-typed offspring is obtained 
+by swapping only subtrees of the same type. Only subtrees with 
+corresponding counterpart in the second parent are randomly chosen from.
+More interesting problem lies in free variables, which may cause trouble
+if swapped somewhere where it is suddenly not bounded. In order to circumvent this
+difficulty we utilize technique called \textit{abstraction elimination}\citep{jones87}
+that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions
+and no bound variables. After the initial population is generated,  it is transformed
+by abstraction elimination. Another possible transformation taking place
+after initialization is $\eta$-normalization shortening 
+rather long long normal form into shorter $\beta\eta$-normal form.  
+Another performance enhancing transformation is option of using "applicative" tree representation (coming directly from inductive definition of \lterms) instead of more traditional S-expression representation.
+Favorable properties of applicative tree representation are also reported in \citep{yu1998polygp}.
+
+\section{Future work}
+\red{
+Jak pojmout podstatu článku?
+
+Řekl bych, že vtip je v použití typovejch tříd na hendlování typovýho systému TGP.
+
+Tahle pasáž je imho klíčová a tak bych jí napsal jako první,
+podle noušnů v ní obsaženejch bych pak upravil preliminaries
+a akordingly to zmínil v related work na konci (něco jako, na rozdíl od 
+kanaďanů my se chceme skusit vydat na cestu k zobecnění HM tak, že 
+využijem typeclasses : výhody videíme v tom, že se prakticky využívaj 
+haskellu a jejich použití výrazně neztěžuje inferenční proces, naopak do něj
+přirozeně pasuje)
+}
+\red{nekde dyštak zmínit že to jsou fikaný javovský interfacy}
+
+Zásadním rozhodnutím při budování systému pro typované GP je volba underlying typového sytému.
+Současné přístupy k typovanému GP nad lambda calculem nejčastěji používají hidley-milnerův typový system \red{citovat yu, kombinatory} nebo experimentují s Systemem F, kterýžto je zobecněním HM.
+Alternativní přístup k obohacení HM would be enriching the HM type system with the type classes.
+
+A type class is a type system construct that supports ad-hoc polymorphism in a mathematically elegant way.
+The concept of the type class first appeared in the Haskell programming language \citep{morris2013type} and was originally designed as a way of implementing overloaded arithmetic and equality operators in a systematic fashion \citep{wadler1989make}.
+
+Basically a type class is a predicate over types. Let us demonstrate this notion on the simple example of a type class, the \texttt{Eq} type class for handling the equality operator.
+
+Funkci \texttt{intMember} zjištující zda číslo x je v seznamu čísel xs můžeme zapsat jako:
+
+\texttt{~\\
+intMember :: Int -> [Int] -> Bool\\
+intMember y [] = False\\
+intMember y (x:xs) = (x == y) || intMember y xs\\}
+
+Abychom nemusel mít funkci pro každý typ zvlášť, hodilo by se mít zobecněnou funkci \texttt{member :: a -> [a] -> Bool}, navíc bychom chtěli abychom její definici mohli převzít z intMember.
+Tomu brání pouze použití equality operatoru (==).
+Abychom tuto definici mohli použít, potřebujeme aby nad typem a byla definována rovnost.
+Tento fakt můžeme vyjádřit jako predikát \texttt{Eq a}.
+V typu funkce member se tato dodatečná podmínka projeví jako \texttt{member :: (Eq a) => a -> [a] -> Bool}.
+Zbívá však dořešit jakým způsobem se typová třída Eq definuje.
+Nejjednoduším způsobem definice by bylo(footnote doopravdy je složitější ale to je jedno):
+
+\texttt{~\\
+class Eq a where\\
+ (==) :: a -> a -> Bool\\}
+
+Ještě zbývá poslední věc a to jak říct systému, že nějaký typ splňuje predikát Eq.
+Řekněme, že to chceme udělat pro typ Int a že máme k dispozici funkci eqInt :: Int -> Int -> Bool která implemetuje operator rovnosti nad cisly.
+Pak stačí deklarovat následující:
+
+\texttt{~\\
+instance Eq Int where\\
+ (==) = eqInt\\}
+
+Co kdybychom ale chtěli definovat Eq nad parametrickým typem, řekněme nad stromem.
+Jak se vyhneme tomu abychom nemuseli deklarovat instanci pro každý typ seznamu zvlášť?
+K tomu nám poslouží následující notace:
+
+\texttt{~\\
+instance (Eq a) => Eq (Tree a) where\\ 
+  Leaf a         == Leaf b          =  a == b\\
+  (Branch l1 r1) == (Branch l2 r2)  =  (l1==l2) \&\& (r1==r2)\\
+  \_              == \_               =  False\\}
+
+\red{v rychlosti okomentovat a zmínit že obecně těch předpokladů tam může bejt víc a jakou to má notaci.
+řict že pro nás je z typovýho hlediska podstatná ta deklarace a to co následuje po where už je otázka implementace (možná to dohrotit tak, že tu implementaci ...-ovat.) }
+
+Jak to převést do GP? Konkrétněji jak generovat programy v HM rozšířeném o TC?
+Pro generování bez TC je potřeba zadat požadovaný typ a mn. stavebních symbolu (s informací o jejich typech) tj. context. Pro HM s TC musíme navíc zahrnout deklarace tříd a instancí.
+
+Deklarace tříd udává jednak podmínky které musí splnit deklarace instancí.
+Navíc pokud používáme TC Eq, je patrně nasnadě aby součástí contextu byly funkce, které daná třída definuje.
+E.g., pro smysluplné použití třídy Eq je potřeba aby součástí vstupniho contextu bylo \texttt{(==) :: (Eq a) => a -> a -> Bool}.
+
+Hlavní novinku do vstupu generujícího algoritmu zajišťují deklarace instancí.  
+Množina instancí nám  v záseadě přidává informaci o struktuře typů, nad kterými operují funkce z kontextu.
+
+Pro účely generování nás zajíma jen "info z prvního řádku", v zasadě máme dva typy instancí jednoduchou a s předpoklady.
+Zásadním faktem je, že tyto informace o instancích můžeme přímočaře vyjádřit jako logické programy \red{citovat něco kde se to dělá}.
+
+\texttt{instance $P$ $T$ where ...} odpovídá faktu $P(T)$.
+
+\texttt{instance ($P_1$ $T_1$,$\dots$,$P_n$ $T_n$) => $P$ $T$ where ...} odpovídá hornovské klauzuli $P(T) \leftarrow P_1(T_1),\dots,P_n(T_n)$.
+
+Pro přehlednost jsme uvedli predikátové symboli jen s jedním argumentem, přímočaře to však můžeme zobecnit pro predikáty libovolné arity.
+
+Odpověď na otázku zda daný typ je instancí nějaké TC zodpovíme odpovídajícím dotazem pro logický program odpovídající naší množině instancí.
+
+\red{
+- dodat tuto LP složku je na programátarovy, GP se pak stará o generování. 
+  TC jsou ale obecné konstrukty které jde jednoduše používat napříč problémy
+  (užití Systému F naopak nechává vše na GP, minimum na programátorovi)
+
+- neboť LP tur completní tak v zasadě můžeme naprogramovat libovolnej predikát
+
+- blackboxing - můžeme se vysrat na to pracně definovat predikáty v LP, stačí když se budou zvenku chovat tak jako by byli a vevnitř mohou bejt napsaný třeba v javě
+
+- praktický příklad takovýho blackboxingu - např plánování.
+ukazuje, že tyto dvě problem solving approaches are uzce connected - využití např že se muže GP inspirovat technikama z Planování, naopak zas že type theory může dát plánování expresivní jazyk / novej pohled na pro škatulkizaci různejch pojmů co odpovídaj nějakejm věcem co v GP jsou přirozeně (fce víc argumentů..., lamb abstr...)
+
+- nakonec (/na začatek) zmínit že máme i další future worky ale že vybíráme tento konkrétní
+}
+
+\section{Starší poznámky ...}
+
+- HM .. místo = dá unifikaci, a bez letu.
+
+(možná tam dát dk, spiš ne)
+
+- uvod k hm+tc 1-3 věty.
+
+\subsection{hinley milner + type classes}
+
+příklady
+
+(možná i ten s planovanim, dle času a místa)
+
+(metaevoluce)
+
+\subsection{Problémy}
+
+- sada benchmarku - citovat ten šit co sem našel
+
+- trhy
+
+. - vyhody: propojenej uzel uloh (slechteni manazera, modelu, konstruktera stroju, ), je to vypocetni model, na trh muzeme koukat jako na 
+. - meta model - výpočetní model co má to co optimalizuje zakodovany jako komoditu to co 
+. - navazuje na spolupráci mezi plánovánim a
 
@@
 
+\red{v referencích změnit toho barendrechta na toho z roku 2013}
+
 \acknowledgments %\footnotesize
@@
 
-%% TWO METHODS FOR INCLUDING THE BIBLIOGRAPHY (LIST OF REFERENCES)
-%% EITHER TYPE IN THE ENTRIES YOURSELF AS SHOWN HERE IN
-%% `thebibliography' ENVIRONMENT,
-%%        OR
-%% USE THE FOLLOWING TWO COMMANDS SO THAT BIBTEX WILL GENERATE
-%% `thebibliography' TEXT FOR YOU AND READ IT IN.
-%%
 \bibliographystyle{egs}%<-- LIST OF REFERENCES TO BE IN "EGS" STYLE (FILE "EGS.STY")
 \bibliography{my}       %<-- REFERENCES ARE IN FILE "SAMPLE.BIB"
-%%
-%%  IF THE ABOVE TWO COMMANDS ARE USED, THEN thebibliography ENVIRONMENT
-%%  MUST BE REMOVED.
-
-%\begin{thebibliography}{}
-%
-%\itemsep=0.4ex
-%%\footnotesize
-%\bibitem[\it Klimov and Friend(1997)]{kl}
-%Klimov, S., Friend A., ASPI experiment: Measurements of fields and
-%waves onboard the Interball-1 spacecraft, {\it Ann. Geophys.}, 15,
-%514-527, 1997.
-%
-%\bibitem[\it Kudela et al.(1995)]{ku}
-%Kudela, K., M. Slivka, J. Rojko, and V. N. Lutsenko, The
-%apparatus DOK-2 (project INTERBALL): Output data structure and
-%modes of operation, preprint of Inst. Exp. Phys. UEF-01-95,
-%Kosice, 20, 1995.
-%
-%\bibitem[\it Lin(1997)]{li}
-%Lin Y., Generation of anomalous flows near the bow shock by
-%its interaction with interplanetary discontinuities, {\it J. Geophys.
-%Res.}, 102, 24265-24281, 1997.
-%
-%\bibitem[\it Lutsenko et al.(1995)]{lu}
-%Lutsenko, V. N., J. Rojko, K. Kudela, T. V. Gretchko, J. Balaz,
-%J. Matisin, E. T. Sarris, K. Kalaitzides, and N. Paschalidis,
-%Energetic particle experiment DOK-2 (Interball project), in:
-%{it INTERBALL Mission and Payload,} ed. by Yu. Galperin, IKI-CNES,
-%249, 1995.
-%
-%
-%\bibitem[\it Onsager et al.(1990a)]{ona}
-%Onsager, T. G., M. F. Thomsen, J. G. Gosling, and S. J.
-%Bame, Observational test of a hot flow anomaly formation
-%mechanism, {\it J. Geophys. Res.}, 95, 11.967-11.974, 1990a.
-%
-%\bibitem[\it Onsager et al.(1990b)]{onb}
-%Onsager, T. G., M. F. Thomsen, and D. Winske,
-%Hot flow anomaly formation by magnetic deflection,
-%{\it Geophys. Res. Lett.}, 17, 1621-1624, 1990b.
-%
-%\bibitem[\it Schwartz et al.(1999d)]{sw}
-%Schwartz, S. J., G. E. Paschmann, N. Sckopke, T. M. Bauer, M. W.
-%Dunlop, A. N. Fazakerley, and M. F. Thomsen,  Hot flow anomalies
-%revisited, {\it Int. J. Geomag. and Aeronomy}, 1, 1999, in print.
-%
-%\bibitem[\it Thomsen et al.(1992)]{th}
-%Thomsen M. F., J. G. Gosling, and S. J.
-%Bame, Observational test of a hot flow anomaly formation
-%mechanism, {\it J. Geophys. Res.}, 97, 1967-1974, 1992.
-%
-%
-%\end{thebibliography}
 

### 2014-05-23 tex/wds14/wds14.tex
wds...

--- 2014-04-29 tex/wds14/wds14.tex
@@
 %to be compiled with LATEX 2e
-
 %use this setting with 11pt basic font:
 \documentclass[11pt]{article}
+
+% >>>>>>>>>>>>>>>>>>    nezapomen projet ze nezbyli zadny TODO      <<<<<<<
+
+\usepackage[utf8]{inputenc}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\usepackage{xspace}
+\usepackage[ampersand]{easylist}
+\usepackage{qtree}
+\usepackage{color}
+
+%
+
 \usepackage{wds11,epsf}
-
 %use this setting with 10pt basic font:
@@
 \newcommand{\Real}{\mathbb{R}}
+
+\newcommand{\gar}{\longmapsto}
+\newcommand{\bRedex}{$\beta$-redex\xspace}
+\newcommand{\bRedexes}{$\beta$-redexes\xspace}
+\newcommand{\bArrow}{\rightarrow_\beta\xspace}
+\newcommand{\eArrow}{\rightarrow_\eta\xspace}
+\newcommand{\eeArrow}{\rightarrow_{\eta^{-1}}\xspace}
+
+\newcommand{\HM}{Hindley–Milner\xspace}
 
@@
 %%%%%%%%%%% nevim co to tu bylo za omg, tak jsem to zakomentoval
-%%\lefthead{SAFRANKOVA ET AL.}
-%%\righthead{MAGNETOSHEATH RESPONSE}
+\lefthead{K\v{R}EN}
+\righthead{Typed Genetic Programming in Lambda Calculus}
 
@@
 \begin{abstract}
-Todo.
+\red{TODO\\~[Článek zatím obsahuje i české poznámky, které jsou hodně hala bala a neformální.
+A celkově je to ještě dost neučesané.
+Dost jsem zápasil s nedostatkem místem a hodně toho dal pryč, protože jsem dal radši přednost podrobnějšímu popisu klíčových pojmů..
+]}
 \end{abstract}
@@
 
+Genetic programming (GP) represents an efficient method for automatic generating of programs by means of evolutionary techniques~\citep{koza92,koza2003genetic}.
+Early attempts to enhance the GP approach with the concept of types include the seminal work~\citep{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
+The use of types naturally leads to enriching S-expressions,
+the traditional GP representation of individuals, with concepts from
+\textit{lambda calculus}, which is simple yet powerful functional, mathematical, and programming 
+language extensively used in type theory. Such attempts has shown to be 
+successful \cite{yu01}. 
+%\red{TODO}\footnote{\red{možná přesunout do preliminaries k oduvodnění proč na to dem juknout z formal hlediska a říct že $\Gamma$ je dk že to hezky zvoní}}
+One of the motivations for using lambda calculus is that it is backed by a considerable theoretical background.
+From this background we can utilize several useful concepts such as reductions, normalization or abstraction elimination, and use them to enhance the typed GP (TGP).
+
+%Genetic programming (GP) is an AI technique, which falls into broader category of evolutionary algorithms  ---  mataheuristic  search algorithms inspired  by biological evolution by natural selection.
+GP was developed by John  \cite{koza92}.
+I dare say many people perceive its beauty in the fact that GP is a computer program  constructing other computer programs with desired properties by breeding them.
+It is also pretty successful technique in a number of areas \citep{koza2003genetic}.
+And perhaps that is why GP has become very popular.
+
+%Early attempts to enhance the GP approach with the concept of types include the seminal work~\citep{montana95} where the ideas from Ada programming language were used to define a so-called strongly typed GP.
+Usage of types naturally opens door to enriching S-expressions, the traditional GP representation of individuals, with concepts from lambda calculus, which is a simple yet powerful functional, mathematical and programming language extensively used in type theory~\citep{yu01}.
+
+%%  >>>>>>>>>>>>
+
+%\red{(- ještě něco k tomu že LC je cool pro TGP (taky z článků - nebo diplomky tam to bylo rozebraný ale jen když bude místo) ?)}
+
+In this paper we compare typed lambda calculus approaches to GP known to us with approach we are suggesting as possible alternative.
+The crucial decision behind the design of a TGP system is the choice of the underlying \textit{type system}.
+The majority of such TGP systems uses \textit{\HM type system (HM)}.
+There are also experiments with more powerful type systems such as System F.
+An alternative approach we proposse is to use \textit{HM} enriching with the \textit{type classes}.
+
+The rest of the paper is organized as follows: \red{TODO}
+
+\section{Preliminaries}
+
+In this section several useful notions 
+%\textit{(GP, LC, STLC, HM, SF, TC)} 
+are described.
+
+\subsection{Genetic programming}
+
+A problem to be solved is given to GP in the form of a \textit{fitness function}.
+A fitness function is a function which takes a computer program as its input and  returns a numerical value called \textit{fitness} as an output.
+The bigger fitness of a computer program is, the better solution of a problem.
+GP maintains a collection of computer programs called \textit{population}.
+A member of a population is called \textit{individual}.
+By running GP algorithm evolution of those individuals is performed.
+Individuals are computer program \textit{expressions} kept as \textit{syntactic trees}.
+Another crucial input besides fitness function is a collection of \textit{building symbols}.
+It is a collection of symbols (accompanied with an information about number of arguments; or in the case of the typed GP, this information is the type of the symbol).
+Those symbols are used to construct trees representing individuals.
+
+%\Lets describe GP algorithm briefly.
+At the beginning, initial population is generated randomly from the set of building symbols.
+A step of GP algorithm is stochastic transformation of the current population into the next population.
+This transformation is given by the choice of genetic operators (e.g.
+crossover, mutation) and selection mechanism based on individual fitness.
+The transformation is repeatedly applied for a predefined number of steps (which is called number of \textit{generations}) or until some predefined criterion is met.
+
+\subsection{Lambda calculus}
+\Lets describe a programming language, 
+in which the GP algorithm generates individual programs --- the so called \lterms.
+
+\begin{definition}
+Let $V$ be infinite countable set of {\it 
+variable names}. Let $C$ be set of {\it constant names}, 
+$V \cap C = \emptyset$.	 	
+Then $\Lambda$ is set of {\it \lterms} defined inductively as follows.
+\begin{align*}
+x   \in V \cup C  &\then x     \in \Lambda \\
+M,N \in \Lambda   &\then (M~N) \in \Lambda 
+\textit{~~~~~~(Function application)} \\
+x   \in V , M \in \Lambda &\then \lamb{x}{M} \in \Lambda
+\textit{~~~~($\lambda$-abstraction)} 
+\end{align*}
+\end{definition}
+
+\textit{Function application} and 
+\textit{$\lambda$-abstraction} are concepts
+well known from common programming languages. 
+For example, in JavaScript, 
+$(M~N)$ translates to expression \texttt{$M$($N$)} and
+$\lamb{x}{M}$ translates to expression \texttt{function($x$)\{return $M$;\}}.
+In other words, the function application 
+corresponds to the act of supplying a function 
+with an argument, and
+the $\lambda$-abstraction is equivalent to 
+\textit{anonymous function}\footnote{Apart from JavaScript, anonymous functions are common e.g. in Python and Ruby, 
+they were recently introduced to C++, and they are expected to be supported in Java 8.}.
+To ensure better readability, 
+$M_1~M_2~M_3~\dots~M_n$ will be an abbreviation for 
+$(\dots((M_1~M_2)~M_3)~\dots~M_n)$.
+And $\lam{x_1 \dots x_n }{M}$  for
+$\lamb{x_1}{\dots\lamb{x_n}{M}\dots}$.
+
+In order to perform computation, there must be some
+mechanism for term evaluation. In $\lambda$-calculus there
+is a \mbox{$\beta$-reduction} procedure for this reason.
+A term of a form $\lamb{x}{M}N$ is called \textit{\bRedex}.
+A \bRedex can be $\beta$-reduced to term $M[x:=N]$. 
+It is also possible to reduce \textit{subterm \bRedexes}.
+In other words, $\beta$-reduction is the process 
+of insertion of arguments supplied to a function into 
+its body.
+
+\subsection{Simply typed lambda calculus}
+
+A \lterm as described above corresponds to a program expression with no type information included. Now we will describe \textit{simple types}.
+
+\begin{definition}
+Let $A$ be set of {\it atomic type names}. 
+Then $\mathbb{T}$ is set of {\it types} inductively defined by the following two rules:
+(1) $\alpha      \in A  \then   \alpha \in \T$, and
+(2) $\sigma,\tau \in \T \then ( \sigma \ar  \tau ) \in \T$.
+\end{definition}
+
+Type $\sigma \ar \tau$ is type for functions taking as input
+something of a type $\sigma$ and returning 
+as output something of a type $\tau$. 
+$\tau_1 \ar \tau_2 \ar \dots \ar \tau_n$ is an abbreviation for 
+$\tau_1 \ar (\tau_2 \ar (\dots \ar (\tau_{n-1} \ar \tau_n)\dots))$. Such "chain of arrows" types simulate types of functions with multiple (here $n-1$) inputs. 
+This technique is called \textit{currying}. One can grasp the trick by considering $(f~M_1~\dots~M_n)$ as shorthand for $f(M_1,\dots,M_n)$
+and by observing that functions with such types have terms of the form $\lam{x_1 x_2 \dots x_n }{M}$. 
+The \textit{type system} called \textit{simply typed $\lambda$-calculus (STLC)} is now easily obtained by combining the previously defined \textit{\lterms} and \textit{types} together.
+
+\begin{definition} Let $\Lambda$ be set of {\it \lterms}. 
+	Let $\mathbb{T}$ be set of {\it types}.       
+	A {\it statement} $M : \sigma$ is a pair 
+	$(M,\sigma) \in \Lambda \times \mathbb{T}$.
+	Statement $M : \sigma$ is vocalized as 
+	{\it "$M$ has type $\sigma$"}.
+	The term $M$ is called the {\it subject} of the 
+	statement $M : \sigma$.
+A \textit{declaration} is a statement 
+ $x : \sigma$ where $x \in V \cup C$.
+A \textit{context} 
+ is set of declarations with distinct variables as subjects.
+\end{definition}
+%~\\
+
+Context is a basic type theoretic concept suitable as a typed alternative
+for terminal and function set in standard GP. 
+Notation $\Gamma,x:\sigma $ denotes $ \Gamma\cup\{(x:\sigma)\}$ 
+such that $\Gamma$ does not contain any declaration with $x$ as subject.
+We also write $x:\sigma \in \Gamma$ instead of $(x,\sigma) \in \Gamma$.
+
+\begin{definition}
+A statement $M\colon\sigma$ is \textit{derivable from}
+a context $\Gamma$ (notation 
+\mbox{$\Gamma\vdash{}M\colon\sigma$}) 
+if it can be produced by the following rules.
+\begin{align*}
+x : \sigma \in \Gamma &~\then~ \tur{\Gamma}{x}{\sigma}\\
+\tur{\Gamma}{M}{\sigma \ar \tau}~,~\tur{\Gamma}{N}{\sigma} 
+&~\then~ \tur{\Gamma}{(M~N)}{\tau}\\  
+\tur{\Gamma,x:\sigma}{M}{\tau}
+&~\then~ \tur{\Gamma}{\lamb{x}{M}}{\sigma \ar \tau} 
+\end{align*}
+\end{definition}
+
+Those rules are called \textit{inference rules} (see the similarity with the \textit{definition 1}).
+A type system is defined by a collection of inference rules.
+When generating terms (e.g., during population initialization) our goal is to produce terms $M$ for a given pair $\ul{\tau}{\Gamma}$ such that for each $M$ is $\tur{\Gamma}{M}{\tau}$ (where $\tau$ is the desired type of generated individuals and $\Gamma$ is the set of \textit{building blocks}).
+More complex type systems (e.g.
+\HM or System F) can be defined similarly by specifying their set of inference rules, but (for the sake of brevity) we prefer to continue the overview in a less formal fashion.
+
+\subsection{\HM type system}
+
+\HM type system \citep{hindley1969principal} is a generalization of the simply typed lambda calculus extended by two concepts: parametric types and type variables (and with the \texttt{let} expression, but we omit its discussion since it is less important for the type related matter).
+
+Parametric types are types that take other type(s) as parameter. E.g., a list (or a tree) of values of the same type. 
+A type variable is a type that may stand for an arbitrary type. 
+Those two concepts play nicely together, we can see this by observing following simple example.
+The function \texttt{head} which returns the first element of a list.
+This function has type \texttt{a -> List a}, where \texttt{a} is a type variable and \texttt{List} is a parametric type (taking another type as parameter, in this case it is a type variable \texttt{a}).
+In STLC we must define the head function for every concrete instance of the List, whereas in the \HM we can define them all at once.
+
+%možná líp na ... replicate :: Int -> a -> [a]
+
+\subsection{System F}
+
+System F \red{\citep{TODO}} can be seen as a further generalization of both STLC and \HM. 
+The concept of type variables brings the question of variable quantification. 
+\HM uses implicit general quantification. System F is about making things explicit.
+This holds both for placing variable quantification symbols ($\forall$) explicitly inside the type terms, and for need to explicitly place the type term into the polymorphic function in order to cast it to the specific instance with no type variables.
+This for example enables possibility of generating "anonymous" type definitions\footnote{E.g., the standard \texttt{List a = Cons a (List a) | Nil} can be expressed "anonymously" as $\forall \alpha \forall \beta : (\alpha \ar \beta \ar \beta) \ar \beta \ar \beta  $.}.
+This rise in power brings the disadvantage of the substantially more computationally complex term generating procedure.
+
+\subsection{Type classes}
+
+A type class is a type system construct that supports ad-hoc polymorphism in a mathematically elegant way.
+The concept of the type class first appeared in the Haskell programming language \citep{morris2013type} and was originally designed as a way of implementing overloaded arithmetic and equality operators in a systematic fashion \citep{wadler1989make}.
+
+Basically a type class is a predicate over types. Let us demonstrate this notion on the simple example of a type class, the \texttt{Eq} type class for handling the equality operator:
+
+Let \texttt{isMemberInt : Int -> [Int] -> Bool} be a function performing check whether an integer (the first argument) is a meber of a list of integers (the second argument).
+And let us assume that this function in its definition uses the equality operator:
+\texttt{~\\
+%isMemberInt :: Int -> [Int] -> Bool\\
+isMemberInt y [] = False\\
+isMemberInt y (x:xs) = (x == y) || isMemberInt y xs}
+
+In order to overcame the need to have for each type (such as \texttt{Int}) a specific instance of the function it would be nice to have a way to specific 
+a function such as \texttt{member :: a -> [a] -> Bool}, but there is a trouble with
+the equality operator used in the definition. The \texttt{==} is defined only on some types. 
+The fact that an equality operator is defined for type \texttt{T} is denoted by use of the \textit{predicate} \texttt{Eq} by writing \texttt{(Eq T)}.  
+We can specify this additional requirement in the type as \texttt{member :: (Eq a) => a -> [a] -> Bool}.
+
+Such a predicate is called \textit{type class} and is defined in the following way:
+
+\texttt{class Eq a where (==) :: a -> a -> Bool}.
+
+In order to state the fact that a certain concrete type (e.g.
+\texttt{Int}) is a type for which the predicate holds we write the following declaration (and suppose that \texttt{eqInt :: Int -> Int -> Bool} is our implementation of integer equality):
+
+\texttt{instance Eq Int where (==) = eqInt}
+
+\red{Co kdybychom ale chtěli definovat Eq nad parametrickým typem, řekněme nad stromem.
+Jak se vyhneme tomu abychom nemuseli deklarovat instanci pro každý typ seznamu zvlášť?
+K tomu nám poslouží následující notace (where we omit the implementation part since we are interested in the signature):}
+
+\texttt{instance (Eq a) => Eq (Tree a) where ...}
+
+%\texttt{~\\
+%instance (Eq a) => Eq (Tree a) where\\ 
+%  Leaf a         == Leaf b          =  a == b\\
+%  (Branch l1 r1) == (Branch l2 r2)  =  (l1==l2) \&\& (r1==r2)\\
+%  \_              == \_               =  False\\}
+
+\red{v rychlosti okomentovat a zmínit že obecně těch předpokladů tam může bejt víc a jakou to má notaci.
+řict že pro nás je z typovýho hlediska podstatná ta deklarace a to co následuje po where už je otázka implementace }
+
+%\red{- naznačit celou hierarchii, řict že na vršku je jakoby DTT.}
+
 \section{Related work}
 \label{related}
+
+\red{Todo, přepsat aby to pasovalo na porovnání s type classama; připadně to porovnání udělat celý až v následující sekci o our approach}
 
@@
 This differs from our approach, which tries to be generalization of
-the standard GP\cite{koza92}.
+the standard GP\citep{koza92}.
 
@@
 
-Todo.
+\section{Our current approach}
+
+\red{
+In our previous work jsme se rozhodli pro přístup: nejdřív vzít nejjednoduší typovej systém a v něm navrhnout generování a gen operace tak aby jednak byly zobecnění std GP, a jednak aby byly schopný generovat libovolnej term (modulo NF), ne jen nějakou jednodušší podmnožinu.
+
+Pokud se to tam vejde tak zmínit, že HM lze z STLC získat tak že (zjednodušeně řečeno) na vhodných mistech změníme porovnání rovnosti typových termů za unifikaci typových termů.
+(V tý nasledující sekci pak k tomu navíc přidat, že se k unifikaci přidává běh logickýho programu, který je ale nutno pouštět po krokách tak, aby byly na stejný urovni kroky generujícího algoritmu a kroky běhu toho LP (tzn že rozpracovaný term muže být několik kroků A* v tom samém termu ale s jiným stádiem dopočítanosti logického programu pracujícího na doplnění dalšího symbolu).
+)
+
+Konkretní popis mojí generující (a křížící) metody zatim zakomentován protože to zabírá moc místa, nejdřív dopsat zbytek a pak vzít co se vejde, pokud vůbec něco..}
+
+%\subsection{Generating}
+
+%Our approach to \lterm generating is based on technique briefly described in \citep{barendregt10}, which generates well-typed \lterms in their \textit{long normal form}.
+
+%Our generating method is based on simple modification of the standard A* \citep{AIMA}, which we call \textit{forgetful A*}.
+This modification consist in additional parameter for the A* algorithm -- the \textit{search strategy}.
+It is a simple filtration function that is given the set of all successors of the state that is being examined and returns a subset of this input.
+This subset is added to the priority queue to be further explored.
+In this way the search space may be reduced as the filtration function may \textit{forget} some successors.
+
+%A* keeps a priority queue of states during the generation process, on the other hand the \textit{ramped half-and-half method}, the standard GP algorithm for generating individuals, keeps only one individual which is gradually constructed.
+This behavior is easily achieved by use of suitable search strategy that returns subset consisting of only one successor.
+The systematic exhaustive search is obtained by search strategy that returns whole input set.
+Our novel \textit{geometric strategy} can be understood as point somewhere between those two extremes.
+
+%%.... metaevoluce remark
+%%A remarkable benefit of parameterizing the generating method by a search strategy taking the form of a simple filtration function is that such a function can be expressed by functional language in a very economical way.
+Take for instance a search strategy that behaves same as the \textit{geometric} except that when it comes to situation where all the elements are filtered out it acts as the \textit{ramped half-and-half}.
+Such \textit{strategy composition} is easily describable as a higher order function.
+Thus comes to mind the possibility of using our system to come up with new search strategies on its own.
+
+%\subsection{Crossover}
+
+%Our system utilizes generalization of the standard tree-swapping crossover operator.
+Two main concerns with swapping typed subtrees are types and free variables.
+Well-typed offspring is obtained by swapping only subtrees of the same type.
+Only subtrees with corresponding counterpart in the second parent are randomly chosen from.
+More interesting problem lies in free variables, which may cause trouble if swapped somewhere where it is suddenly not bounded.
+In order to circumvent this difficulty we utilize technique called \textit{abstraction elimination}\citep{jones87} that transforms an arbitrary \lterm into \lterm that contains no lambda abstractions and no bound variables.
+After the initial population is generated,  it is transformed by abstraction elimination.
+Another possible transformation taking place after initialization is $\eta$-normalization shortening rather long long normal form into shorter $\beta\eta$-normal form.
+
+%Another performance enhancing transformation is option of using "applicative" tree representation (coming directly from inductive definition of \lterms) instead of more traditional S-expression representation.
+Favorable properties of applicative tree representation are also reported in \citep{yu1998polygp}.
+
+\section{Future work}
+
+%\red{Jak pojmout podstatu článku?
+Řekl bych, že vtip je v použití typovejch tříd na hendlování typovýho systému TGP.
+Tahle pasáž je imho klíčová a tak bych jí napsal jako první, podle noušnů v ní obsaženejch bych pak upravil preliminaries a akordingly to zmínil v related work na konci (něco jako, na rozdíl od kanaďanů my se chceme skusit vydat na cestu k zobecnění HM tak, že využijem typeclasses : výhody videíme v tom, že se prakticky využívaj haskellu a jejich použití výrazně neztěžuje inferenční proces, naopak do něj přirozeně pasuje)}
+
+%\red{nekde dyštak zmínit že to jsou fikaný javovský interfacy}
+
+Jak TC převést do GP? Konkrétněji jak generovat programy v HM rozšířeném o TC?
+Pro generování bez TC je potřeba zadat požadovaný typ a mn. stavebních symbolu (s informací o jejich typech) tj. context. Pro HM s TC musíme navíc zahrnout deklarace tříd a instancí.
+
+Deklarace tříd udává jednak podmínky které musí splnit deklarace instancí.
+Navíc pokud používáme TC Eq, je patrně nasnadě aby součástí contextu byly funkce, které daná třída definuje.
+E.g., pro smysluplné použití třídy Eq je potřeba aby součástí vstupniho contextu bylo \texttt{(==) :: (Eq a) => a -> a -> Bool}.
+
+Hlavní novinku do vstupu generujícího algoritmu zajišťují deklarace instancí.  
+Množina instancí nám  v záseadě přidává informaci o struktuře typů, nad kterými operují funkce z kontextu.
+
+Pro účely generování nás zajíma jen "info z prvního řádku", v zasadě máme dva typy instancí jednoduchou a s předpoklady.
+Zásadním faktem je, že tyto informace o instancích můžeme přímočaře vyjádřit jako logické programy \red{citovat něco kde se to dělá}.
+
+\texttt{instance $P$ $T$ where ...} odpovídá faktu $P(T)$.
+
+\texttt{instance ($P_1$ $T_1$,$\dots$,$P_n$ $T_n$) => $P$ $T$ where ...} odpovídá hornovské klauzuli $P(T) \leftarrow P_1(T_1),\dots,P_n(T_n)$.
+
+Pro přehlednost jsme uvedli predikátové symboli jen s jedním argumentem, přímočaře to však můžeme zobecnit pro predikáty libovolné arity.
+
+Odpověď na otázku zda daný typ je instancí nějaké TC zodpovíme odpovídajícím dotazem pro logický program odpovídající naší množině instancí.
+
+\red{
+- říct že sou 2 použití - jednak v původnim smyslu pro ad-hoc polymorphism - ale taky exploitačnim způsobem kde nám vůbec nejde o implementační složku (která může být prázdná), ale jde nám jen o LP věci.
+Názornym příkladem takovýho použití je zavedení čísel (pozor neplest s termovými čísly) do typovýho systému.
+
+- taková čísla se nám mohou hodit např při práci se seznamy pevné délky.
+
+- dodat tuto LP složku je na programátarovy, GP se pak stará o generování. 
+  TC jsou ale obecné konstrukty které jde jednoduše používat napříč problémy
+  (užití Systému F naopak nechává vše na GP, minimum na programátorovi)
+
+- neboť LP tur completní tak v zasadě můžeme naprogramovat libovolnej predikát
+
+- blackboxing - nemusíme se trápit s tím pracně definovat predikáty v LP, stačí když se budou zvenku chovat tak jako by byli a vevnitř mohou bejt napsaný třeba v javě
+
+- když zbude místo (což se asi nestane) tak tmínit příklad blackboxingu - plánování vyjádřený vjako specialní případ typovanýho GP.
+Spíš asi než takle komplikovanej příklad dát ale na víc míst par drobnejch příkladů využití
+
+- určitě shrnout výhody HM+TC approache...
+
+%- praktický příklad takovýho blackboxingu - např plánování.
+ukazuje, že tyto dvě problem solving approaches are uzce connected - využití např že se muže GP inspirovat technikama z Planování, naopak zas že type theory může dát plánování expresivní jazyk / novej pohled na pro škatulkizaci různejch pojmů co odpovídaj nějakejm věcem co v GP jsou přirozeně (fce víc argumentů..., lamb abstr...)
+
+%- nakonec (/na začatek) zmínit že máme i další future worky ale že vybíráme tento konkrétní
+}
+
+%\section{Starší poznámky ...}
+%
+%- HM .. místo = dá unifikaci, a bez letu.
+%
+%(možná tam dát dk, spiš ne)
+%
+%- uvod k hm+tc 1-3 věty.
+
+%\subsection{hinley milner + type classes}
+%
+%příklady
+%
+%(možná i ten s planovanim, dle času a místa)
+%
+%(metaevoluce)
+
+%\subsection{Problémy}
+%
+%- sada benchmarku - citovat ten šit co sem našel
+%
+%- trhy
+%
+%. - vyhody: propojenej uzel uloh (slechteni manazera, modelu, konstruktera stroju, ), je to vypocetni model, na trh muzeme koukat jako na 
+%. - meta model - výpočetní model co má to co optimalizuje zakodovany jako komoditu to co 
+%. - navazuje na spolupráci mezi plánovánim a
 
 \section{Conclusion}
-Todo.
+Todo nebo spojit s future work do jedný subsekce.
 
-\acknowledgments %\footnotesize
-{???}
+\red{v referencích změnit toho barendrechta na toho z roku 2013}
 
-%% TWO METHODS FOR INCLUDING THE BIBLIOGRAPHY (LIST OF REFERENCES)
-%% EITHER TYPE IN THE ENTRIES YOURSELF AS SHOWN HERE IN
-%% `thebibliography' ENVIRONMENT,
-%%        OR
-%% USE THE FOLLOWING TWO COMMANDS SO THAT BIBTEX WILL GENERATE
-%% `thebibliography' TEXT FOR YOU AND READ IT IN.
-%%
+% TODO nemaj se tam dát TODO
+%\acknowledgments %\footnotesize
+%{???}
+
 \bibliographystyle{egs}%<-- LIST OF REFERENCES TO BE IN "EGS" STYLE (FILE "EGS.STY")
 \bibliography{my}       %<-- REFERENCES ARE IN FILE "SAMPLE.BIB"
-%%
-%%  IF THE ABOVE TWO COMMANDS ARE USED, THEN thebibliography ENVIRONMENT
-%%  MUST BE REMOVED.
-
-%\begin{thebibliography}{}
-%
-%\itemsep=0.4ex
-%%\footnotesize
-%\bibitem[\it Klimov and Friend(1997)]{kl}
-%Klimov, S., Friend A., ASPI experiment: Measurements of fields and
-%waves onboard the Interball-1 spacecraft, {\it Ann. Geophys.}, 15,
-%514-527, 1997.
-%
-%\bibitem[\it Kudela et al.(1995)]{ku}
-%Kudela, K., M. Slivka, J. Rojko, and V. N. Lutsenko, The
-%apparatus DOK-2 (project INTERBALL): Output data structure and
-%modes of operation, preprint of Inst. Exp. Phys. UEF-01-95,
-%Kosice, 20, 1995.
-%
-%\bibitem[\it Lin(1997)]{li}
-%Lin Y., Generation of anomalous flows near the bow shock by
-%its interaction with interplanetary discontinuities, {\it J. Geophys.
-%Res.}, 102, 24265-24281, 1997.
-%
-%\bibitem[\it Lutsenko et al.(1995)]{lu}
-%Lutsenko, V. N., J. Rojko, K. Kudela, T. V. Gretchko, J. Balaz,
-%J. Matisin, E. T. Sarris, K. Kalaitzides, and N. Paschalidis,
-%Energetic particle experiment DOK-2 (Interball project), in:
-%{it INTERBALL Mission and Payload,} ed. by Yu. Galperin, IKI-CNES,
-%249, 1995.
-%
-%
-%\bibitem[\it Onsager et al.(1990a)]{ona}
-%Onsager, T. G., M. F. Thomsen, J. G. Gosling, and S. J.
-%Bame, Observational test of a hot flow anomaly formation
-%mechanism, {\it J. Geophys. Res.}, 95, 11.967-11.974, 1990a.
-%
-%\bibitem[\it Onsager et al.(1990b)]{onb}
-%Onsager, T. G., M. F. Thomsen, and D. Winske,
-%Hot flow anomaly formation by magnetic deflection,
-%{\it Geophys. Res. Lett.}, 17, 1621-1624, 1990b.
-%
-%\bibitem[\it Schwartz et al.(1999d)]{sw}
-%Schwartz, S. J., G. E. Paschmann, N. Sckopke, T. M. Bauer, M. W.
-%Dunlop, A. N. Fazakerley, and M. F. Thomsen,  Hot flow anomalies
-%revisited, {\it Int. J. Geomag. and Aeronomy}, 1, 1999, in print.
-%
-%\bibitem[\it Thomsen et al.(1992)]{th}
-%Thomsen M. F., J. G. Gosling, and S. J.
-%Bame, Observational test of a hot flow anomaly formation
-%mechanism, {\it J. Geophys. Res.}, 97, 1967-1974, 1992.
-%
-%
-%\end{thebibliography}
 

### 2014-05-31 tex/wds14/wds14.tex
wds......

@@
 \lefthead{K\v{R}EN}
-\righthead{Typed Genetic Programming in Lambda Calculus}
+\righthead{Typed Genetic Programming in Lambda Calculus with Type Classes}
 
@@
 
-\title{Typed Genetic Programming in Lambda Calculus}
+\title{Typed Genetic Programming in Lambda Calculus with Type Classes}
 
@@
 \begin{abstract}
-\red{TODO\\~[Článek zatím obsahuje i české poznámky, které jsou hodně hala bala a neformální.
+Genetic programming (GP) traditionally uses simple S-expressions to represent programs.
+Lambda calculus representation of programs offers a more expressive alternative, which can exhibit better results.
+Moreover, lambda calculus representation can benefit from a considerable theoretical background.
+The crucial decision behind the design of a typed GP system is the choice of the underlying type system.
+The majority of such systems uses \HM (HM) type system.
+And there are also experiments with System F, which is a generalization of HM.
+In this paper we outline a novel typed GP system  based on HM enriched with the concept of type classes -- an alternative way to generalizing HM.
+We discuss its benefits and connection to logic programming.
+\end{abstract}
+
+%\red{TODO\\~[Článek zatím obsahuje i české poznámky, které jsou hodně hala bala a neformální.
 A celkově je to ještě dost neučesané.
@@
 ]}
-\end{abstract}
 
 \begin{article}
+
+\red{\textbf{TODO: Přidat názorné příklady ...a par dalších poznámek}
+
+Základní příklad, který by se mohl táhnout napříč textem jsou seznamy.
+Můžeme ukázat, jak v simply typed LC je potřeba pro každý typ seznamu (seznam čísel, znaků, boolů) přidat zvlášť atomický typ, zatímco v HM to jde udělat díky polymorfizmu najednou.
+Ale pro specializovanější datovou strukturu \textbf{seznam pevné délky} je už HM slabý a každá delka se musí definovat zvlášť (problém je že potřebuju typ stojící za konkrétní číslo, a že typ seznamového konstruktoru musí v sobe obsahovat omezení, že vzniklý list je o jedna delší než ten z kterého vznikl přidáním prvku, což jde v HM těžko vyjádřit, protože máme jen "funkční symboly" ale žádný "predikáty".) ...a to je podobnost s předchozím omezením (v LC musim dělat zvlášť typy seznamů, tady zas zvlášť dýlky).
+
+Další příklad, který myslím hezky ilustruje moc TC a přitom neni moc komplikovaný je udělat dědičnost pomocí TC.
+
+Další příklad (navíc i příklad na "blackboxing") je "plánování jako special case typovanýho GP s TC" ale to je dost komplikovaný vysvětlit srozumitelně a zároveň stručně, takže to asi ne.
+
+A samozřejmě přichází výhoda v použití toho k čemu se používaj TC v Haskellu prakticky tzn. různý monády, arrows, funktory, atd. 
+%Např monáda je definovaná dvěma funkcema (>>= a return), ale pro pro každou svou instanci dělá něco komplet zajímavýho a jinýho (na listech "dělá paralelizmus",  ).
+
+Důležitý je zdůraznit, že oproti Systemu F se výrazně nezesložiťuje generující procedura, ale že to tam naopak celkem přirozeně zapadá jako rozšíření (výpočet logického porgramu a generování typovaného stromu v HM jsou skoro stejné procesy), což kontrastuje s tím co dělaj System F lidi, který kvůli komplexitě inference v SF používaj dost nestandardní GP proceduru (i když teda zajímavou).
+
+Myslim, že je také dobré jasně vymezit rozdíl ve filozofii použití SF a TC: system F jde cestou \textit{vše nechám na GP systému a to včetně definice typů samotných}, zatímco mnou navrhovaný přístup se nese více v duchu: \textit{rozdělme pole působnosti GP systému a programátora ve stylu programování vs meta-programování, GP-systém programuje (skládá funkce do programů) zatímco programátor nachystá vhodnou "knihovnu" a její datové struktury, ty však mezi sebou mohou mít poměrně komplexní vztahy - "programátor meta-programuje logický program (vztahů datových struktur) pomocí typových tříd"}.
+}
 
@@
 
-The rest of the paper is organized as follows: \red{TODO}
+%The rest of the paper is organized as follows: \red{TODO}
 
-\section{Preliminaries}
+\section{Basic notions} %\section{Preliminaries}
 
@@
 
-\subsection{Genetic programming}
+\subsection{\textit{Genetic programming}}
 
@@
 
-\subsection{Lambda calculus}
+\subsection{\textit{Lambda calculus}}
+\red{\textit{tuto subsekci zkracovat v případě překročení limitu....}}
+
 \Lets describe a programming language, 
@@
 
-\subsection{Simply typed lambda calculus}
+\subsection{\textit{Simply typed lambda calculus}}
 
@@
 
-\subsection{\HM type system}
+\subsection{\textit{\HM type system}}
 
@@
 
-\subsection{System F}
+\subsection{\textit{System F}}
 
@@
 
-\subsection{Type classes}
+\subsection{\textit{Type classes}}
 
@@
 
-\red{Todo, přepsat aby to pasovalo na porovnání s type classama; připadně to porovnání udělat celý až v následující sekci o our approach}
+\red{Todo, přepsat aby to pasovalo na porovnání s type classama; připadně to porovnání udělat celý až v následující sekci o our approach, ocitovat krátce kdo ukázal na propojení TC a LP a možnost programovat v typovym systému}
 
@@
 
-\section{Our current approach}
+\section{Our approach}
 
-\red{
-In our previous work jsme se rozhodli pro přístup: nejdřív vzít nejjednoduší typovej systém a v něm navrhnout generování a gen operace tak aby jednak byly zobecnění std GP, a jednak aby byly schopný generovat libovolnej term (modulo NF), ne jen nějakou jednodušší podmnožinu.
+\red{následující dvě subsekce (generování, křížení) pojmout tak, aby řekli: oporti obecnému trendu se my  zatim soustředíme výrazně víc na generující proceduru (než na operátory atd jak bejvá zvykem) a pojímame jí jako kontinuum od klasického GP přístupu až k systematické enumeraci bližšší automatickému dokazování.
+Uvidíme že tento přístup nám napomáhá hladce integrovat výpočet logického porgramu stojícího za soustavou typovejch tříd do generující procedury.
+}
 
-Pokud se to tam vejde tak zmínit, že HM lze z STLC získat tak že (zjednodušeně řečeno) na vhodných mistech změníme porovnání rovnosti typových termů za unifikaci typových termů.
+%\red{In our previous work jsme se rozhodli pro přístup: nejdřív vzít nejjednoduší typovej systém a v něm navrhnout generování a gen operace tak aby jednak byly zobecnění std GP, a jednak aby byly schopný generovat libovolnej term (modulo NF), ne jen nějakou jednodušší podmnožinu.
+
+%Pokud se to tam vejde tak zmínit, že HM lze z STLC získat tak že (zjednodušeně řečeno) na vhodných mistech změníme porovnání rovnosti typových termů za unifikaci typových termů.
 (V tý nasledující sekci pak k tomu navíc přidat, že se k unifikaci přidává běh logickýho programu, který je ale nutno pouštět po krokách tak, aby byly na stejný urovni kroky generujícího algoritmu a kroky běhu toho LP (tzn že rozpracovaný term muže být několik kroků A* v tom samém termu ale s jiným stádiem dopočítanosti logického programu pracujícího na doplnění dalšího symbolu).
@@
 
-Konkretní popis mojí generující (a křížící) metody zatim zakomentován protože to zabírá moc místa, nejdřív dopsat zbytek a pak vzít co se vejde, pokud vůbec něco..}
+%Konkretní popis mojí generující (a křížící) metody zatim zakomentován protože to zabírá moc místa, nejdřív dopsat zbytek a pak vzít co se vejde, pokud vůbec něco..}
 
-%\subsection{Generating}
+\subsection{\textit{Generating}}
 
-%Our approach to \lterm generating is based on technique briefly described in \citep{barendregt10}, which generates well-typed \lterms in their \textit{long normal form}.
+Our approach to \lterm generating is based on technique briefly described in \citep{barendregt10}, which generates well-typed \lterms in their \textit{long normal form}.
 
-%Our generating method is based on simple modification of the standard A* \citep{AIMA}, which we call \textit{forgetful A*}.
+Our generating method is based on simple modification of the standard A* \citep{AIMA}, which we call \textit{forgetful A*}.
 This modification consist in additional parameter for the A* algorithm -- the \textit{search strategy}.
@@
 
-%A* keeps a priority queue of states during the generation process, on the other hand the \textit{ramped half-and-half method}, the standard GP algorithm for generating individuals, keeps only one individual which is gradually constructed.
+A* keeps a priority queue of states during the generation process, on the other hand the \textit{ramped half-and-half method}, the standard GP algorithm for generating individuals, keeps only one individual which is gradually constructed.
 This behavior is easily achieved by use of suitable search strategy that returns subset consisting of only one successor.
@@
 
-%\subsection{Crossover}
+\subsection{\textit{Crossover}}
 
-%Our system utilizes generalization of the standard tree-swapping crossover operator.
+Our system utilizes generalization of the standard tree-swapping crossover operator.
 Two main concerns with swapping typed subtrees are types and free variables.
@@
 
-%Another performance enhancing transformation is option of using "applicative" tree representation (coming directly from inductive definition of \lterms) instead of more traditional S-expression representation.
+Another performance enhancing transformation is option of using "applicative" tree representation (coming directly from inductive definition of \lterms) instead of more traditional S-expression representation.
 Favorable properties of applicative tree representation are also reported in \citep{yu1998polygp}.
 
-\section{Future work}
+\section{\textit{Using \HM with type classes}}
 
@@
 
-\section{Conclusion}
+\section{Conclusions}
 Todo nebo spojit s future work do jedný subsekce.
